### 2018-04-05 paper/appendix.tex
add papers

\section{Generating a Minimal CDD Expansion}
\label{app:cons-cdd}
% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{For $\mathcal{R} \in \relpreds$, %
    $\mathcal{S} \in \chc{\mathcal{R}}$.}
  % Output: inductive invariants.
  \Output{A minimal CDD expansion of $\mathcal{S}$ and
    correspondence to $\mathcal{S}$.} %
  \myproc{$\expand{ \mathcal{R} }(\mathcal{S})$ %
    \label{line:expand-begin} }{ %
    % auxiliary procedure:
    \myproc{$\expandaux( \mathcal{S}' )$
      \label{line:expand-aux-begin} }{ %
      % case split: optional sharing clause:
      \Switch{$\sharingclause( \mathcal{S}' )$}{ %
        % case: no siblings with overlapping descendants
        \lCase{$\none$:}{ \Return{$\mathcal{ S }'$} %
          \label{line:expand-ret} } %
        % case: siblings with 
        \lCase{$\mathcal{C} \in \mathcal{S}', % 
          R \in \ctxs{ \mathcal{R} }$: }{ %
          % expand the return:
          \Return{$\expandaux( %
            \copyrel{ \mathcal{R} }( %
            \mathcal{S}', \mathcal{C}, R))$ } %
          \label{line:expand-recurse}
        } %
      } %
    \label{line:expand-aux-end} } %
    % base call: 
    \Return{$( \expandaux(\mathcal{S}), \corr{ \mathcal{R} } )$ %
      \label{line:expand-base-call}}
  } %
  %
  \caption{$\expand{ \mathcal{R} }$: for $\mathcal{R} \in \relpreds$
    given $\mathcal{S} \in \chcs{ \mathcal{R} }$, returns a minimal
    CDD expansion of $\mathcal{S}$. }
  \label{alg:expand}
\end{algorithm}
% walk through top-level procedure:
\autoref{alg:expand}, given $\mathcal{S} \in \chcs{ \mathcal{R} }$
(\autoref{line:expand-begin}), returns an expansion $\mathcal{S}'$ of
$\mathcal{S}$ over relational predicates $\ctxs{ \mathcal{R} }$
(defined below) accompanied by a correspondence from $\mathcal{S}'$ to
$\mathcal{S}$.
%
$\expand{ \mathcal{R} }$ defines a procedure $\expandaux$
(\autoref{line:expand-aux-begin}---\autoref{line:expand-aux-end}) that
takes a CHC system $\mathcal{S}$ and returns an expansion of
$\mathcal{S}$.
%
$\expand{ \mathcal{R} }$ runs $\expandaux$ on $\mathcal{S}$ and
returns the result, paired with the map $\corr{ \mathcal{R} } : \ctxs{
  \mathcal{R} } \to \mathcal{R}$ (\autoref{line:expand-base-call};
$\corr{\mathcal{R}}$ is defined below).

% walk through aux procedure:
$\expandaux$, given $\mathcal{S}' \in \chcs{ \ctxs{ \mathcal{R} } }$,
runs a procedure $\sharingclause$ on $\mathcal{S}'$, which tries to
find $\mathcal{C} \in \mathcal{S}'$ and $R \in \ctxs{ \mathcal{R} }$
applied in the body of $\mathcal{C}$ that is a transitive dependences of
clause siblings in $\mathcal{S}'$.
%
I.e., $\expandaux$ tries to find $\mathcal{C}$ and $R$ such that for
some siblings $R_0, R_1 \in \ctxs{ \mathcal{R} }$, $R \in \depsof{
  \mathcal{S}' }(R_0) \intersection \depsof{ \mathcal{S}' }(R_1)$ and
$\relof{ \headof{ \mathcal{C} }} \in \depsof{ \mathcal{S}' }(R_0)
\setminus \depsof{ \mathcal{S}' }(R_1)$.
%
In such a case, we say that $(\mathcal{C}, R)$ is a
\emph{sibling-shared dependence}.
% case: no sibling-shares
If $\sharingclause$ determines that no sibling-shared dependence
exists, then $\expandaux$ returns $\mathcal{S}'$
(\autoref{line:expand-ret}).

% case: there is a sibling-shared dependence:
Otherwise, for sibling-shared dependence $(\mathcal{C}, R)$ returned
by $\sharingclause$, $\expandaux$ runs $\copyrel{ \mathcal{R} }$ on
$\mathcal{S}'$, $\mathcal{C}$, and $R$, which returns an expansion of
$\mathcal{S}'$ in which the dependents shared by $R_0$ and $R_1$ are
those shared in $\mathcal{S}'$, without $(\mathcal{C}, R)$.
%
$\expandaux$ recurses on the CHC system returned by $\copyrel{
  \mathcal{R} }$ and returns the result
(\autoref{line:expand-recurse}).

% give the representation of ctxs
\subsubsection{Representation of $\ctxs{ \mathcal{R} }$}
%
$\ctxs{ \mathcal{R} }$ are relational predicates that store the
clauses in which they occur.
%
\begin{defn}
  \label{defn:ctxs}
  $\ctxs{ \mathcal{R} }$ is the smallest space such that $\mathcal{R}
  \subseteq \ctxs{ \mathcal{R} }$ and for $\mathcal{C} \in \clauses{
    \mathcal{ \ctxs{R} } }$ and $A \in \appsof{ \bodyof{ \mathcal{C} }
  }$, $(\mathcal{C}, \relof{A}) \in \ctxs{ \mathcal{R} }$.
\end{defn}
% give the correspondence between relational predicates:
$\corr{ \mathcal{R} } : \ctxs{R} \to \mathcal{R}$ is such that for $R
\in \mathcal{R}$, $\corr{ \mathcal{R} }(R) = R$ and for $\mathcal{C}
\in \clauses{ \ctxs{R} }$ and $R \in \ctxs{R}$ such that
$(\mathcal{C}, R) \in \ctxs{R}$, $\corr{\mathcal{R}}(\mathcal{C}, R) =
\corr{ \mathcal{R} }(R)$.

% give the function that copies instances of a relational predicate
\subsubsection{Implementation of $\copyrel{ \mathcal{R} }$}
%
$\copyrel{\mathcal{R}}$, given $\mathcal{S} \in \chcs{ \mathcal{R} }$,
$\mathcal{C} \in \mathcal{S}$, and $R \in \ctxs{ \mathcal{R} }$,
returns the following CHC system.
% clause: C, but the application now points to (C, A)
Let $\mathcal{C}' \in \clauses{ \ctxs{ \mathcal{R} } }$ be
$\mathcal{C}$, with $R$ replaced with $(\mathcal{C}, R)$ in the body
of $\mathcal{C}$.
%
Let $\mathcal{S}_{ \mathcal{C} } \subseteq \clauses{ \ctxs{
    \mathcal{R} } }$ be the clauses in $\mathcal{S}$ for which $R$ is
the relational predicate of the clause head, and let
$\mathcal{S}_{\mathcal{C}}'$ be each clause in $\mathcal{S}_{
  \mathcal{C} }$, with $R$ in the clause head replaced with
$(\mathcal{C}, R)$.
%
Then $\copyrel{ \mathcal{R} }$ returns $\add{ (\remove{ \mathcal{S} }{
    \add{ \mathcal{S}_{\mathcal{C}} }{\mathcal{C} } }) }{ \mathcal{C}'
} \union \mathcal{S}_{\mathcal{C}}'$.
%

% CDD systems:
$\expand{\mathcal{R}}$ always returns a CDD expansion of its input
(see \autoref{app:corr}, \autoref{lem:expand-corr}) that is minimal.
% discussion: there are variations:
$\expand{ \mathcal{R} }$ is certainly not unique as an algorithm for
generating a minimal CDD expansion.
%
In particular, feasible variations of $\expand{ \mathcal{R} }$ can be
generated from different implementations of $\sharingclause$, each of
which chooses clause-relations pairs to return based on different
heuristics.
%
We expect that other expansion algorithms can also be developed by
generalizing algorithms introduced in previous work on generating
compact verification conditions of hierarchical
programs~\cite{lal-qadeer15}.

\section{Proof of characterization of CDD systems}
\label{app:char}
The following is a proof of \autoref{thm:cdd-contains}.
%
\begin{proof}
  % proof structure:
  To prove that the union of the classes of linear systems and
  body-disjoint systems are strictly contained by the class of CDD
  systems, we prove %
  \textbf{(1)} the class of linear systems are contained by the class
  of CDD systems, %
  \textbf{(2)} the class of body-disjoint systems are contained by the
  class of CDD systems, and %
  \textbf{(3)} there is some CDD system that is not linear or
  body-disjoint.

  % contains linear systems:
  For goal \textbf{(1)}, let $\mathcal{S}$ be an arbitrary linear
  system.
  %
  $\mathcal{S}$ is CDD if for each clause $\mathcal{C}$ in
  $\mathcal{S}$, for each pair of distinct relational predicates in
  the body of $\mathcal{C}$, the transitive dependence are disjoint
  (\autoref{defn:cdds}).
  %
  Let $\mathcal{C}$ be an arbitrary clause in $\mathcal{S}$.
  %
  $\mathcal{C}$ has no pairs of distinct clauses, by the fact that
  $\mathcal{S}$ is linear.
  %
  Therefore, $\mathcal{S}$ is CDD.

  % contains body-disjoint systems:
  For goal \textbf{(2)}, let $\mathcal{S}$ be an arbitrary
  body-disjoint system.
  %
  The dependence relation of $\mathcal{S}$ is a tree $T$, by the
  definition of a body-disjoint system.
  %
  Let $\mathcal{C}$ be an arbitrary clause in $\mathcal{S}$, with
  distinct relational predicates $R_0$ and $R_1$ in its body.
  %
  All dependences of $R_0$ and $R_1$ are in subtrees of $T$, which
  are disjoint by the definition of a tree.
  %
  Thus, $\mathcal{S}$ is CDD, by \autoref{defn:cdds}.

  % distinguishing example:
  For goal \textbf{(3)}, the system $\mcchc$ is CDD, but is not linear
  or body-disjoint.
\end{proof}

\section{Proof of Correctness}
\label{app:corr}
%
In this section, we give a proof that \sys is a correct solver for
recursion-free CHC systems.
%
We first establish lemmas for the correctness of each procedure used
by \sys, namely \textsc{Collapse} (\autoref{lem:expansion-sound} and
\autoref{lem:expansion-complete}), $\expand{\mathcal{R}}$
(\autoref{lem:expand-corr}), and $\solvecdd{ \mathcal{R} }$
(\autoref{lem:vc}, \autoref{lem:cdd-soln-sound}, and
\autoref{lem:cdd-soln-complete}).
%
We combine the lemmas to prove correctness of \sys
(\autoref{thm:corr}).

% lem: result connecting expansions to constructing solutions
For CHC systems $\mathcal{S}$ and $\mathcal{S}'$, if $\mathcal{S}'$ is
an expansion of $\mathcal{S}$, then the result of collapsing a
solution of $\mathcal{S}'$ is a solution of $\mathcal{S}$.
%
\begin{lem}
  \label{lem:expansion-sound}
  For $\mathcal{R}, \mathcal{R}' \in \relpreds$, $\mathcal{S} \in
  \chcs{ \mathcal{R} }$, $\mathcal{S}' \in \chcs{ \mathcal{R}' }$,
  $\sigma' \in \interps{ \mathcal{R}' }$, and $\eta : \mathcal{R}' \to
  \mathcal{R}$ a correspondence from $\mathcal{S}'$ to $\mathcal{S}$,
  $\collapse{\eta}{\sigma'}$ is a solution of $\mathcal{S}$.
\end{lem}
%
\begin{proof}
  % introduce 
  Let $\mathcal{B} \in \bodies{\mathcal{R}}$, $\varphi \in \formulas$,
  and $R \in \mathcal{R}$ be such that $((\mathcal{B}, \varphi),
  R(\vars{ \mathcal{S} }(R))) \in \mathcal{S}$.
  %
  For each $R' \in \mathcal{R}'$ such that $\eta(R') = R$, there is
  some $\mathcal{B}' \in \bodies{ \mathcal{R}' }$ such that
  $((\mathcal{B}', \varphi), R'(\vars{ \mathcal{S} }(R))) \in
  \mathcal{S}'$, by the fact that $\mathcal{S}'$ is an expansion of
  $\mathcal{S}$ and the definition of an expansion.
  %
  $\sigma'(\mathcal{B}'), \varphi \entails \sigma'(\mathcal{R}')$ by
  the fact that $\sigma'$ is a solution of $\mathcal{S}'$.
  %
  $\collapse{ \eta }{ \sigma' }(\mathcal{B}), \varphi \entails
  \sigma'(\mathcal{R}')$, by the definition of
  $\collapse{\eta}{\sigma'}$.
  %
  Therefore,
  %
  \[ \collapse{ \eta }{ \sigma' }( \mathcal{B} ), \varphi \entails %
  \left( \bigland_{ \substack{R' \in \mathcal{R}' \\ \eta(R') = R} }
  \sigma(R') \right) = %
  \collapse{\eta}{\sigma'}(\mathcal{R}') 
  \]
  % 
  Therefore, $\collapse{\eta}{\sigma'}$ is a solution of clause
  $((\mathcal{B}, \varphi), R(\varsof{ \mathcal{S} }(R)))$.
  %
  Therefore, $\collapse{\eta}{\sigma'}$ is a solution of
  $\mathcal{S}$.
\end{proof}

% completeness of expansions:
For CHC solvable system $\mathcal{S}$, each expansion of $\mathcal{S}$
is solvable.
\begin{lem}
  \label{lem:expansion-complete}
  For $\mathcal{R}, \mathcal{R}' \in \relpreds$, $\mathcal{S} \in
  \chcs{ \mathcal{R} }$ such that $\mathcal{S}$ is solvable, and
  $\mathcal{S}' \in \chcs{ \mathcal{R}' }$ such that $\mathcal{S'}$ is
  an expansion of $\mathcal{S}$, $\mathcal{S}'$ is solvable.
\end{lem}
%
\begin{proof}
  Let $\sigma : \mathcal{R} \to \formulas$ be a solution of
  $\mathcal{S}$, and let $\eta : \mathcal{R}' \to \mathcal{R}$ be a
  correspondence from $\mathcal{R}'$ to $\mathcal{R}$.
  %
  Let $\sigma' : \mathcal{R}' \to \formulas$ be such that for each $R'
  \in \mathcal{R}'$, $\sigma'(R') = \sigma(\eta(R'))$.
  %
  Then $\sigma$ is a solution of $\mathcal{S}'$.
\end{proof}

% CDD systems:
$\expand{\mathcal{R}}$ always returns a CDD expansion of its input.
%
\begin{lem}
  \label{lem:expand-corr}
  For $\mathcal{R} \in \relpreds$ and $\mathcal{S} \in \chcs{
    \mathcal{R} }$, $\mathcal{S} \in \chcs{ \ctxs{ \mathcal{R} } }$,
  and $\eta : \ctxs{ \mathcal{R} } \to \mathcal{R}$ such that
  $(\mathcal{S}', \eta) = \expand{\mathcal{R}}(\mathcal{S})$, $\eta$
  is a correspondence from $\mathcal{S}'$ to $\mathcal{S}$.
\end{lem}
%
\begin{proof}
  Proof by induction on the evaluation of $\expand{ \mathcal{R} }$ on
  an arbitrary $\mathcal{S} \in \chcs{ \mathcal{R} }$.
  % inductive fact:
  The inductive fact is that in each step of evaluation \expandaux,
  $\corr{ \mathcal{R} }$ is a correspondence from argument
  $\mathcal{S}'$ to $\mathcal{S}$.
  % base case:
  For the base case, \expandaux is called initially on $\mathcal{S}$,
  by \autoref{alg:expand}.
  % 
  $\corr{ \mathcal{R} }$ is a correspondence from $\mathcal{S}$ to
  itself, by the definition of $\corr{ \mathcal{R} }$
  (\autoref{app:cons-cdd}).

  % inductive case:
  For the inductive case, for $\mathcal{C} \in \mathcal{S}$ and $R \in
  \ctxs{ \mathcal{R} }$, $\expandaux$ calls itself on $\copyrel{
    \mathcal{R} }(\mathcal{S}, \mathcal{C}, R)$.
  %
  For each $\mathcal{S}' \in \chcs{ \ctxs{ \mathcal{R} } }$ such that
  $\corr{ \mathcal{R} }$ is a correspondence from $\mathcal{S}'$ to
  $\mathcal{S}$, $\corr{ \mathcal{R} }$ is a correspondence from
  $\copyrel{ \mathcal{R} }( \mathcal{S}', \mathcal{C}, R)$ to
  $\mathcal{S}$, by definition of $\copyrel{ \mathcal{R} }$
  (\autoref{app:cons-cdd}).
  %
  By this fact and the inductive hypothesis, $\corr{ \mathcal{R}}$ is
  a correspondence from $\copyrel{ \mathcal{R} }( \mathcal{S}',
  \mathcal{C}, R)$ to $\mathcal{S}$.

  % final claim: expand returns an expansion:
  $\expandaux{ \mathcal{R} }$ returns its parameter at some step, by
  \autoref{alg:expand}.
  %
  Therefore, $\expandaux$ returns an expansion of $\mathcal{S}$.

  % prove that returned system is CDD:
  For $\mathcal{S}' \in \chcs{ \ctxs{ \mathcal{R} } }$ and $\eta :
  \ctxs{\mathcal{R}} \to \mathcal{R}$, if $(\mathcal{S}', \eta) =
  \expand{ \mathcal{R} }( \mathcal{S} )$, then $\sharingclause(
  \mathcal{S}') = \none$, by the definition of $\expand{ \mathcal{R}
  }$.
  %
  If $\sharingclause(\mathcal{S}') = \none$, then $\mathcal{S}'$ is
  CDD, by the definition of $\sharingclause$ and CDD systems
  (\autoref{defn:cdds}).
  %
  Therefore, $\mathcal{S}'$ is CDD.
\end{proof}
%
Furthermore, $\expand{ \mathcal{R} }$ returns a \emph{minimal} CDD
expansion of its input.
%
This fact is not required to prove \autoref{thm:corr}, and thus a
complete proof is withheld.

% lem: VC's characterize system sat:
For each CHC system $\mathcal{S}$, $\mathcal{S}$ has a solution if and
only if the verification condition of $\mathcal{S}$ is unsatisfiable.
%
\begin{lem}
  \label{lem:vc}
  For $\mathcal{R} \in \relpreds$ and $\mathcal{S} \in \chcs{
    \mathcal{R} }$ that is CDD and solvable, $\vc{ \mathcal{R} }(
  \mathcal{S} )$ is unsatisfiable.
\end{lem}
%
\begin{proof}
  %
  Assume that $\mathcal{S}$ has solution $\sigma$ and $m$ is a model
  of $\vc{ \mathcal{R} }( \mathcal{S})$.
  %
  There are some relational predicates $\mathcal{R}' \subseteq
  \mathcal{R}$ containing $\queryof{ \mathcal{S} }$ such that for each
  $R' \in \mathcal{R}'$, $m(\used{R})$ holds, by the definition of
  $\vc{ \mathcal{R} }$ (\autoref{sec:solve-cdd}).
  %
  For each $R'$, the interpretation of $m$ over $\varsof{ \mathcal{R}'
  }$ must satisfy $\sigma(R')$, by well-founded induction on $\deps_S$
  restricted to $\mathcal{R}'$.
  %
  But $\sigma(\queryof{ \mathcal{S} }) = \false$, by the definition of
  a solution of a CHC system (\autoref{defn:chc-soln}).
  %
  Therefore, there can be no such model $m$.
  %
  Therefore, $\vc{ \mathcal{R} }( \mathcal{S} )$ is unsatisfiable.
\end{proof}

\begin{lem}
  \label{lem:solve-aux}
  For $\mathcal{R} \in \relpreds$ and $\mathcal{S} \in \chcs{
    \mathcal{R} }$ such that $\vc{ \mathcal{R} }(\mathcal{S})$ is
  unsatisfiable, $\solvecdd{ \mathcal{R} }(\mathcal{S})$ is a solution
  of $\mathcal{S}$.
\end{lem}
%
\begin{proof}
  % overview of structure:
  Proof by induction on the evaluation of $\solveaux$ on
  interpretation $\sigma$.
  %
  The inductive fact is that at each step of evaluation, $\sigma$ is a
  partial solution of $\mathcal{S}$ such that $\sigma(\mathcal{S})$
  % base case:
  For the base case, \solveaux is initially called on the empty
  interpretation.
  %
  Therefore, $\vc{\mathcal{R}}(\sigma(\mathcal{S})) = \vc{
    \mathcal{R}}(\mathcal{S})$, which is unsatisfiable.

  % inductive case:
  For the inductive case, in each step, \solveaux calls itself on an
  interpretation $\sigma'$, which is its argument $\sigma$, extended
  to map relational predicate $R$ to an interpolant of $\prectr{
    \mathcal{R} }( \mathcal{S}, \sigma, R)$ and $\postctr{ \mathcal{R}
  }( \mathcal{S}, \sigma, R)$.
  %
  $\sigma'$ is a partial solution of $\mathcal{S}$ by a combination of
  the inductive hypothesis that $\sigma$ is a partial solution, the
  fact that any extension of $\sigma$ that maps $R$ to a formula
  implied by $\prectr{ \mathcal{R} }( \mathcal{S}, \sigma, R)$ is a
  partial solution, and the definition of an interpolant
  (\autoref{defn:itps}).
  %
  $\vc{ \mathcal{R} }(\sigma'(\mathcal{S}))$ is unsatisfiable by a
  combination of the inductive hypothesis that $\vc{ \mathcal{R}
  }(\sigma(\mathcal{ S }))$ is unsatisfiable, that for any extension
  $\sigma''$ of $\sigma$ that maps $R$ to a formula inconsistent with
  $\postctr{ \mathcal{R} }( \mathcal{S}, \sigma, R)$, $\vc{
    \mathcal{R} }(\sigma''(\mathcal{S}))$ is unsatisfiable, and the
  definition of an interpolant (\autoref{defn:itps}).

  % prove goal:
  $\mathcal{R}'$ (defined at \autoref{alg:solve-cdd},
  \autoref{line:cons-nodeps}) is empty only if $\mathcal{R} \setminus
  \domainof{ \sigma }$ is empty, by induction on evaluation of
  \solveaux.
  %
  Therefore, \solveaux only returns its parameter $\sigma$ if $\sigma$
  is in fact a complete interpretation.
  %
  This fact, combined with the inductive fact, imply that \solveaux
  returns a solution of $\mathcal{S}$.
\end{proof}

% SolveCDD: establish soundness
For $\mathcal{R} \in \relpreds$, $\solvecdd{ \mathcal{R} }$ is a
correct CHC solver for CDD systems over $\mathcal{R}$.
%
\begin{lem}
  \label{lem:cdd-soln-sound}
  For $\mathcal{R} \in \relpreds$, %
  $\mathcal{S} \in \chcs{ \mathcal{R} }$ such that $\mathcal{S}$ is
  CDD, and %
  $\sigma \in \interps{ \mathcal{R} }$ such that %
  $\sigma = \solvecdd{ \mathcal{R} }(\mathcal{S})$, %
  $\sigma$ is a solution of $\mathcal{S}$.
\end{lem}
%
\begin{proof}
  %
  $\vc{ \mathcal{R} }(\mathcal{S})$ is unsatisfiable, by the fact that
  $\sigma \in \interps{ \mathcal{R} }$ \autoref{alg:solve-cdd},
  \autoref{line:check-sat}.
  %
  Therefore, $\solvecdd{ \mathcal{R} }(\mathcal{S})$ is the result of
  calling \solveaux, by \autoref{alg:solve-cdd},
  \autoref{line:check-sat}.
  %
  Therefore, \autoref{lem:solve-aux} and the fact that $\vc{
    \mathcal{R} }(\mathcal{S})$ is unsatisfiable imply that \solveaux
  returns a complete solution of $\mathcal{S}$.
\end{proof}

% SolveCDD: establish completeness
\begin{lem}
  \label{lem:cdd-soln-complete}
  For $\mathcal{R} \in \relpreds$ and %
  $\mathcal{S} \in \chcs{ \mathcal{R} }$ such that $\mathcal{S}$ is
  solvable and CDD, %
  there is some $\sigma \in \interps{ \mathcal{R} }$ such that %
  $\sigma = \solvecdd{ \mathcal{R} }(\mathcal{S})$.
\end{lem}
%
\begin{proof}
  % final claim:
  $\vc{ \mathcal{R} }(\mathcal{S})$ is not satisfiable, by
  \autoref{lem:vc} and the fact that $\mathcal{S}$ is solvable.
  %
  Therefore, $\solvecdd{ \mathcal{R} }$ is the result of running
  \solveaux, by \autoref{alg:solve-cdd}.
  %
  Therefore, by \autoref{lem:solve-aux} and the fact that
  $\mathcal{S}$ is solvable, $\solvecdd{ \mathcal{R} }$ returns a
  solution of $\mathcal{S}$.
\end{proof}

% proof of main theorem:
The following is a proof of correctness of \sys
(\autoref{sec:approach}, \autoref{thm:corr}).
%
\begin{proof}
  % if returns a map, then S is a solution
  Assume that for $\sigma \in \interps{ \mathcal{R} }$, $\sigma =
  \shara{ \mathcal{R} }( \mathcal{S} ) \in \interps{ \mathcal{R} }$.
  %
  Let $\mathcal{S'} \in \chcs{ \ctxs{ \mathcal{R} } }$ and $\eta :
  \ctxs{ \mathcal{R} } \to \mathcal{R}$ be such that $(\mathcal{S}',
  \eta) = \expand{ \mathcal{R} }( \mathcal{S} )$;
  %
  $\eta$ is a correspondence from $\mathcal{S'}$ to $\mathcal{S}$, by
  \autoref{lem:expand-corr}.
  %
  There is some $\sigma' \in \interps{ \mathcal{R}' }$ such that
  $\sigma' = \solvecdd{ \ctxs{ \mathcal{R} } }( \mathcal{S}' )$, by
  the definition of $\shara{ \mathcal{R} }$.
  %
  $\sigma'$ is a solution of $\mathcal{S}'$, by
  \autoref{lem:cdd-soln-sound}.
  %
  $\collapse{ \eta }{ \sigma' }$ is a solution of $\mathcal{S}$, by
  \autoref{lem:expansion-sound}.

  % if S is solvable, then shara returns a solution
  Assume that $\mathcal{S}$ is solvable.
  %
  Let $\mathcal{S'} \in \chcs{ \ctxs{ \mathcal{R} } }$ and $\eta :
  \ctxs{ \mathcal{R} } \to \mathcal{R}$ be such that $(\mathcal{S}',
  \eta) = \expand{ \mathcal{R} }( \mathcal{S} )$.
  %
  $\mathcal{S}'$ is a CDD expansion of $\mathcal{S}$, by
  \autoref{lem:expand-corr}.
  %
  $\mathcal{S}'$ is solvable, by \autoref{lem:expansion-complete}.
  %
  $\solvecdd{ \ctxs{ \mathcal{R} } }( \mathcal{S'}) \in \interps{
    \mathcal{R'} } $, by \autoref{lem:cdd-soln-complete}.
  %
  $\shara{\mathcal{R}} \in \interps{ \mathcal{R} }$, by
  \autoref{alg:shara}.
\end{proof}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End:

### 2018-04-05 paper/approach.tex
add papers

% walk through the approach in technical detail
\section{Technical Approach}
\label{sec:approach}
%
This section presents the technical details of our approach.
%
\autoref{sec:CDD-defn} presents the class of Clause-Dependence Disjoint
systems and its key properties.
%
\autoref{sec:solve-cdd} describes how \sys solves CDD systems
directly.
%
\autoref{sec:core-solver} describes how \sys solves a given
recursion-free system by solving an equivalent CDD system.
%
Proofs of all theorems stated in this section are contained in appendix
\autoref{app:char} and \autoref{app:corr}.

\subsection{Clause-Dependence Disjoint Systems}
\label{sec:CDD-defn}
%
The key contribution of our work is the introduction of the class of
Clause-Dependence Disjoint (CDD) systems.
%
Solving an CHC system in CDD is coNP time.
%
CDD strictly contains
the union of classes of recursion-free systems that can be solved
in coNP time that have been introduced in previous work.
%
\begin{defn}
  \label{defn:cdds}
  For a given recursion-free CHC system $\cc{S}$,
  %
  if for all pair of uninterpreted predicates $P, Q \in \predof{S}$ that are siblings(\autoref{sec:chcs}), the 
  transitive dependences of $P$ and $Q$ are disjoint $\tdepsof{P} \cap \tdepsof{Q} = \emptyset$, and there is 
  no uninterpreted predicate shows more than once in a body of clause, then $\cc{S}$ is
  \emph{Clause-Dependence-Disjoint (CDD)}.
\end{defn}
%
CDD systems model hierarchical programs with branching and
procedure calls that in each execution path that each statments be
exected at most once.
%
\QZ{rework example}
\begin{ex}
  The CHC system $\mcchc$ (given in \autoref{sec:solve-ex}) is a CDD
  system.
  %
  The only clauses in $\mcchc$ that contain multiple uninterpreted
  predicates are \autoref{eqn:then} and \autoref{eqn:else}, which both
  contain $\cc{L4}$ and $\cc{dbl}$.
  %
  The dependences of $\cc{L4}$ and $\cc{dbl}$ are disjoint.
\end{ex}

% talk about body disjoint
The class of CDD systems contains body-disjoint and linear
systems introduced in previous work.
%
For recursion-free system $\cc{S}$, if for each uninterpreted predicate
$Q \in \predof{S}$, there is at most one caluse $\cc{C}$ that $Q \in \bodyof{C}$,
and there is no uninterpreted predicate shows more than once in a body of clause, then
$\cc{S}$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
%
If the body of each clause in $\cc{S}$ contains at most
relational predicate, then $\cc{S}$ is
\emph{linear}~\cite{albarghouthi12a}.

% talk about linear system
\begin{thm}
	\label{thm:cdd-contains}
  The class of CDD systems strictly contains the union of the classes
  of body-disjoint and linear systems.
\end{thm}
%
The CHC system $\mcchc$ (\autoref{sec:solve-ex}) is not body-disjoint
or linear, but is CDD.
%
A proof that both the classes of body-disjoint and linear systems are
contained in the class of CDD systems is given in \autoref{app:char}.

% describe how to solve a CDD system
\subsection{Solving a CDD system}
\label{sec:solve-cdd}

% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A CDD System $\cc{S}$.}
  % Output: inductive invariants.
  \Output{If $\cc{S}$ is solvable, then a solution of
    $\cc{S}$; %
    otherwise, the value $\nosoln$.}
  % 
  \myproc{$\solvecdd{ \mathcal{R} }(\mathcal{S})$ %
    \label{line:solve-begin}}{ %
    % recurse on a partial solution to the CHC system:
    \myproc{$\solveaux( \mathcal{ \sigma } )$ %
      \label{line:aux-begin}}{
      % construct predicates that:
      $\mathcal{R}' \assign \setformer{ R }{ %
        % are not defined by sigma,
        R \notin \domainof{ \sigma }, %
        \bigland_{ (R, R') \in \depsof{ \mathcal{S} } } 
        R' \in \domainof{ \sigma } }$ \label{line:cons-nodeps} \;
      % case split: on D:
      \Switch{$\mathcal{R}'$ \label{line:case-deps}}{ %
        \lCase{$\emptyset$:}{ %
          % D is empty: return sigma as solution
          \Return{$\sigma$} \label{line:ret-soln} } %
        \Case{$\elts{ R } \disjunion \mathcal{R}''$: %
          \label{line:case-frontier}}{ %
          % D contains an element R. Update sigma to map R to result
          % of interpolation query on suitable formulas.
          \Return{$\solveaux( %
            \upd{\sigma}{R}{ %
              \solveitp(\prectr{ \mathcal{R} }( %
              \mathcal{S}, \sigma, R), %
              \postctr{ \mathcal{R} }(\mathcal{S}, \sigma, R))})$} %
          \label{line:ret-recurse} \;
        } %
      } %
    \label{line:solve-aux-end} } %
    % check if query is feasible:
    \lIf{$\issat(\vc{ \mathcal{R} }(\mathcal{S}))$}{ % 
      \Return{$\nosoln$} \label{line:check-sat} }
    % return result of calling aux on base case:
    \Return{$\solveaux(\emptyset)$} \label{line:aux-empty}
    \; }
  %
  \caption{$\solvecdd{ \mathcal{R} }$: for $\mathcal{R} \in \relpreds$
    and CDD system $\mathcal{S} \in \chcs{ \mathcal{R} }$, returns a
    solution to $\mathcal{S}$ or the value $\none$ to denote that
    $\mathcal{S}$ has no solution.}
  \label{alg:solve-cdd}
\end{algorithm}
%
For $\mathcal{R} \in \relpreds$, the procedure $\solvecdd{ \mathcal{R}
}$ (\autoref{alg:solve-cdd}) is a solver for CDD systems in $\chcs{
  \mathcal{R} }$.
%
$\solvecdd{ \mathcal{R} }$, given $\mathcal{S} \in \chcs{ \mathcal{R}
}$, defines a procedure $\solveaux$ that, given a partial solution
$\sigma$ of $\mathcal{S}$, returns an extension of $\sigma$ that is a
solution of $\mathcal{S}$
(\autoref{line:aux-begin}---\autoref{line:solve-aux-end}).
%
$\solvecdd{ \mathcal{R} }$ runs a procedure $\vc{ \mathcal{R} }$ that,
given a CDD system $\mathcal{S} \in \chcs{ \mathcal{R} }$, returns a
formula that is satisfiable if and only if $\mathcal{S}$ has no
solution, which we refer to as a \emph{counterexample
  characterization} of $\mathcal{S}$ (\autoref{line:check-sat};
%
the implementation of $\vc{ \mathcal{R} }$ is given below).
%
$\solvecdd{ \mathcal{R} }$ runs $\issat$ (\autoref{sec:itps}) on $\vc{
  \mathcal{R} }(\mathcal{S})$;
%
if $\issat$ returns that $\vc{ \mathcal{R} }(\mathcal{S})$ is
satisfiable, then $\solvecdd{ \mathcal{R} }$ returns that
$\mathcal{S}$ has no solution (\autoref{line:check-sat}).
% how the aux function is used:
Otherwise, $\solvecdd{ \mathcal{R} }$ runs $\solveaux$ on the empty
partial solution and returns the result (\autoref{line:aux-empty}).

% walk through the aux function:
$\solveaux$, given partial solution $\sigma$, constructs the set
$\mathcal{R}'$ of each relational predicate $R$ such that $R$ is not
in the domain of $\sigma$ and all dependents of $R$ in $\mathcal{S}$
are in the domain of $\sigma$ (\autoref{line:cons-nodeps}).
%
If $\mathcal{R}'$ is empty, then $\solveaux$ returns $\sigma$ as a
solution of $\mathcal{S}$ (\autoref{line:ret-soln}).
%
Otherwise, if there is some $R \in \mathcal{R}'$
(\autoref{line:case-frontier}), then $\solveaux$ extends $\sigma$ to
map $R$ to an interpolant of the formulas $\prectr{ \mathcal{R}
}(\mathcal{S}, \sigma, R)$ and $\postctr{ \mathcal{R} }(\mathcal{S},
\sigma, R)$, defined below.
%
The interpolant of the two formulas is synthesized by running the
procedure $\solveitp$ (\autoref{sec:itps}).
%
$\solveaux$ recurses on the extended partial solution and returns the
result (\autoref{line:ret-recurse}).
%
\begin{ex}
  $\solvecdd{ \mcpreds }$, given $\mcchc$, may generate
  interpretations of $\mcpreds$ in any topological ordering of the
  dependency relation of $\mcchc$, depending on the relational
  predicates that it chooses at \autoref{line:case-deps}.
  %
  One such ordering is \cc{L3}, \cc{L4}, \cc{dbl}, \cc{L9}, \cc{main},
  $\mathsf{Err}$.
\end{ex}

\subsubsection{Constructing the Counterexample Characterization of a
  CDD}
%
For $\mathcal{S'} \in \chcs{ \mathcal{R} }$, the formula $\vc{
  \mathcal{R} }(\mathcal{S'})$ is a compact representation of the
counterexample characterization of $\mathcal{S'}$, defined as follows.
% introduce the isused predicates:
For each $R \in \mathcal{R}$, let there be a Boolean predicate
$\used{R}$.
% for each relational predicate, define constraint:
The constraint of $R$, denoted $\ctrof{R} \in \formulas$, constrains
that if $R$ occurs in a derivation of $\mathcal{S'}$, then there is
some $\mathcal{C} \in \mathcal{S'}$ in which $R$ is the head such that
\textbf{(1)} each $R' \in \mathcal{R}$ in an application in the body
of $\mathcal{C}$ is used and \textbf{(2)} the values assigned to each
application of $R'$ are equal to the values assigned to the head
variables of $R'$ in $\mathcal{S'}$.
%
I.e., $\ctrof{R}$ is
\begin{align*}
% for each clause with R as a head:
\biglor_{ \substack{ \mathcal{B} \in \bodies{ \mathcal{R} }, %
    R \in \mathcal{R}, \\
  (\mathcal{B}, (R, \varsof{ \mathcal{S'} }(R))) \in
  \mathcal{S'} } } %
\left( \ctrof{ \mathcal{B} } \land %
\bigland_{ \substack{R' \in \mathcal{R}, Y \in X^{*}, \\
    (R', Y) \in \appsof{ \mathcal{B} } } } %
\used{R'} \land \varsof{ \mathcal{S'} }(R') = Y \right)
\end{align*}
%
The counterexample characterization of $\mathcal{S'}$ constrains that
\textbf{(1)} the query predicate of $\mathcal{S'}$ must be used in a
derivation and %
\textbf{(2)} for $R \in \mathcal{R}$, if $R$ is used, then its
constraint must hold.
%
I.e., $\vc{ \mathcal{R} }(\mathcal{S'})$ is $\used{
  \queryof{\mathcal{S'} } } \land %
\bigland_{R \in \mathcal{R} } \left( \used{R} \implies \ctrof{R}
\right)$.

\begin{ex}
  % 
  $\vc{ \mcpreds }(\mcchc)$ characterizes all runs of \cc{multiCall}
  that result in an error.
  %
  Although the relational predicates \cc{L4} and \cc{dbl} occur in the
  bodies of multiple clauses, they are each modeled with only a single
  instance of logical variables.
  % 
  Such a collection is sufficient because each run of \cc{multiCall}
  reaches \cc{L4} and \cc{dbl} at most once.
  %
  Correspondingly, each derivation of $\mcchc$ contains at most one
  occurrence of relational predicates \cc{L4} and \cc{dbl}.
\end{ex}

% pre-constraint:
\subsubsection{Constructing the pre-constraint of a relational predicate}
\label{sec:cons-pre}
%
For partial solution $\sigma : \mathcal{R} \to \formulas$ of
$\mathcal{S}$, and $R \in \mathcal{R}$ for which each dependency is in
the domain of $\sigma$, the pre-constraint $\prectr{ \mathcal{R}
}(\mathcal{S}, \sigma, R)$ is a formula that is entailed by the
interpretation of $R$ in each solution of $\mathcal{S}$ that extends
$\sigma$.
%
In particular, $\prectr{ \mathcal{R} }(\mathcal{S}, \sigma, R)$ is:
\[
\biglor_{ \substack{ \mathcal{B} \in \bodies{ \mathcal{R} } \\
  (\mathcal{B}, (R, \varsof{ \mathcal{S} }(R) ) ) \in \mathcal{S} } }
\left( \ctrof{ \mathcal{ B } } \land %
\bigland_{ A \in \appsof{ \mathcal{B} } } %
\sigma( \relof{A} )[ \argsof{ A } ] \right)
\]
%
\begin{ex}
  \label{ex:pre-ctr}
  Assume that $\mcsolve$, given $\mcchc$, synthesizes a partial
  solution $\sigma \in \interps{ \mcpreds }$ that interprets \cc{L3}
  as $\true$, and chooses to then find an interpretation of \cc{L4}.
  %
  The pre-constraint of \cc{L4} is $\true \land \cc{res}_{\cc{4}} =
  0$, where $\cc{res}_{\cc{4}}$ is a logical variable that models the
  value of \cc{res} at line \cc{4}.
\end{ex}

\subsubsection{Constructing the post-constraint of a relational
  predicate}
\label{sec:cons-ctx}
%
For $\sigma \in \interps{ \mathcal{R} }$ a partial solution of
$\mathcal{S}$ and $R \in \mathcal{R}$, the post-constraint $\postctr{
  \mathcal{R} }(\mathcal{S}, \sigma, R) \in \formulas$ is mutually
inconsistent with the interpretation of $R$ in each solution of
$\mathcal{S}$ that extends $\sigma$.
% dependents of R:
In particular, let $\mathcal{D}_0$ be the reflexive dependents of $R$
in $\mathcal{S}$,
% clause siblings of dependents:
let $\mathcal{D}_1$ be the siblings in $\mathcal{S}$ of
$\mathcal{D}_0$,
%
let $\mathcal{D}_2$ be all dependences of $\mathcal{D}_1$, and
% collect all predicates that we define a constraint over:
let $\mathcal{D} = \mathcal{D}_0 \union \mathcal{D}_1 \union
\mathcal{D}_2$.
% define constraint for each predicate
Let $\restrict{ \mathcal{S} }{ \mathcal{D} }$ be $\mathcal{S}$
restricted to clauses whose head's relational predicate is in
$\mathcal{D}$.
% define instantiation of C
The post constraint for $\mathcal{S}$ under $\sigma$ at $R$ is the
counterexample characterization of $\restrict{ \mathcal{S} }{
  \mathcal{D} }$ grounded on $\sigma$ (defined in \autoref{sec:chcs}).
%
I.e., $\postctr{ \mathcal{R} }(\mathcal{S}, \sigma, R) = \vc{
  \mathcal{R} }(\sigma( \restrict{ \mathcal{S} }{ \mathcal{D} }))$.
% example post constraint:
\begin{ex}
  \label{ex:ctx-ctr}
  Consider the case in which $\mcsolve$, given $\mcchc$, generates an
  interpretation of \cc{L4}, introduced in \autoref{ex:pre-ctr}.
  %
  $\mcsolve$ collects relational predicates $\mathcal{D}_0 = \elts{
    \cc{L4}, \cc{L9}, \cc{main}, \mathsf{Err} }$, $\mathcal{D}_1 =
  \elts{ \cc{dbl} }$, and $\mathcal{D}_2 = \emptyset$.
  %
  $\mcsolve$ constructs $\postctr{ \mcpreds }(\mcchc, \sigma,
  \cc{L4})$ to be the counterexample characterization of $\mcchc$
  restricted to clauses whose head relational predicate is in
  $\mathcal{D} = \elts{ \cc{L4}, \cc{L9}, \cc{main}, \mathsf{Err} }$.
\end{ex}

\subsection{Solving recursion-free systems using CDD systems}
\label{sec:core-solver}

% give the general procedure:
$\shara{ \mathcal{R} }$, given $\mathcal{S} \in \chcs{ \mathcal{R} }$,
constructs a CDD system $\mathcal{S'}$ equivalent to $\mathcal{S}$.
%
\sys then directly solves $\mathcal{S'}$ and from its solution,
constructs a solution of $\mathcal{S}$.
% define an expansion of a CHC system:
For $\mathcal{R}, \mathcal{R}' \in \relpreds$, $\mathcal{S} \in \chcs{
  \mathcal{R} }$ and $\mathcal{S}' \in \chcs{ \mathcal{R}' }$, if
there is a homomorphism from $\mathcal{R}'$ to $\mathcal{R}$ that
preserves the relationship between the clauses of $\mathcal{S}'$ in
the clauses of $\mathcal{S}$, then $\mathcal{S}'$ is an expansion of
$\mathcal{S}$ (all definitions in this section will be over fixed
$\mathcal{R}$, $\mathcal{R}'$, $\mathcal{S}$, and $\mathcal{S}'$).
% formal defn:
\begin{defn}
  \label{defn:expansion}
  Let $f : \mathcal{R}' \to \mathcal{R}$ be such that
  % preserves arity:
  \textbf{(1)} for all $R' \in \mathcal{R}'$, $\arityof{f(R')} =
  \arityof{R'}$;
  % preserves clause structure:
  \textbf{(2)} for all $H \in \apps{\mathcal{R}}$, $\mathcal{B}
  \subseteq \apps{ \mathcal{R} }$, and $\varphi \in \formulas$ such
  that $((\mathcal{B}, \varphi), H) \in \mathcal{S}'$, clause
  %
  \[ \left( \left( \bigunion_{ A \in \mathcal{B} } %
        \{ (f(\relof{A}), \argsof{A}) \}, \varphi \right), %
      ( f(\relof{H}), \argsof{H} ) \right)
  \]
  %
  is in $\mathcal{S}$.
  %
  Then $f$ is a \emph{correspondence} from $\mathcal{S}'$ to
  $\mathcal{S}$.
\end{defn}
%
If there is a correspondence from $\mathcal{S}'$ to $\mathcal{S}$,
$\mathcal{S}'$ is an \emph{expansion} of $\mathcal{S}$, denoted
$\mathcal{S} \expandsto \mathcal{S}'$.
%
\begin{defn}
  \label{defn:min-cdd-expansion}
  If $\mathcal{S}'$ is CDD, %
  $\mathcal{S} \expandsto \mathcal{S}'$, and %
  there is no CDD system $\mathcal{S}''$ such that $\mathcal{S}
  \expandsto \mathcal{S}'' \expandsto \mathcal{S}$ and $\mathcal{S}''
  \not= \mathcal{S}'$, then $\mathcal{S}'$ is a \emph{minimal} CDD
  expansion of $\mathcal{S}$.
\end{defn}

% core algorithm
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{For $\mathcal{R} \in \relpreds$, $\mathcal{S} \in \chcs{
      \mathcal{R} }$.}
  % Output: inductive invariants.
  \Output{A solution to $\mathcal{S}$ or $\none$.}
  % expand procedure:
  \myproc{$\sys[ \mathcal{R} ](\mathcal{S})$ %
    \label{line:shara-begin}}{ %
    % construct expansion of C:
    $(\mathcal{S}', \eta) \assign %
    \expand{\mathcal{R}}(\mathcal{S})$ \label{line:shara-expand} \; %
    % expand, solve, case split:
    \Switch{ $\solvecdd{ \ctxs{ \mathcal{R} } }( \mathcal{S}' )$ %
      \label{line:shara-case} }{
      % case: result is no solution:
      \lCase{ $\none$: }{ \Return{$\none$} \label{line:shara-ret-none} } %
      % case: result is a solution:
      \lCase{ $\sigma' \in \interps{ \ctxs{ \mathcal{R} } }$: }{ %
        \Return{ $\collapse{ \eta }{ \sigma' }$ } %
        \label{line:shara-ret-collapse}
      } %
    } %
  } %
  %
  \caption{\sys: a solver for recursion-free CHCs, which uses
    procedures $\expand{\mathcal{R}}$ (see \autoref{app:cons-cdd}) and
    $\solvecdd{ \ctxs{ \mathcal{R} } }$ (see
    \autoref{sec:solve-cdd}). }
  \label{alg:shara}
\end{algorithm}
% walk through the sys algorithm:
$\shara{ \mathcal{R} }$ (\autoref{alg:shara}), given $\mathcal{S} \in
\chcs{\mathcal{R}}$ (\autoref{line:shara-begin}), returns a solution
to $\mathcal{S}$ or the value $\none$ to denote that $\mathcal{S}$ is
unsolvable.
%
$\shara{ \mathcal{R} }$ first runs a procedure $\expand{ \mathcal{R}
}$ on $\mathcal{S}$ to obtain a minimal CDD expansion $\mathcal{S}'
\in \chcs{ \ctxs{ \mathcal{R} } }$ (\autoref{defn:min-cdd-expansion})
of $\mathcal{S}$ (where $\ctxs{ \mathcal{R} }$ is a space of
relational predicates defined by $\mathcal{R}$, defined in
\autoref{app:cons-cdd}), and a correspondence $\eta : \ctxs{
  \mathcal{R} } \to \mathcal{R}$ (\autoref{line:shara-expand}).
%
$\shara{ \mathcal{R} }$ then runs a procedure $\solvecdd{ \ctxs{
    \mathcal{R} } }$ on $\mathcal{S}'$, which either returns the value
$\none$ to denote that $\mathcal{S}'$ has no solution or a solution
$\sigma'$ (\autoref{line:shara-case}).
%
If $\solvecdd{ \ctxs{ \mathcal{R} } }$ returns that $\mathcal{S}'$ has
no solution, then \sys returns that $\mathcal{S}$ has no solution
(\autoref{line:shara-ret-none}).
%
Otherwise, if $\solvecdd{ \ctxs{ \mathcal{R} } }$ returns a solution
$\sigma'$, then $\sys$ returns, as a solution to $\mathcal{S}$,
$\collapse{ \eta }{ \sigma' }$, defined below
(\autoref{line:shara-ret-collapse}).

% define collapse:
For $\sigma \in \interps{ \ctxs{ \mathcal{R} } }$ and %
correspondence $\eta$ from $\mathcal{S}'$ to $\mathcal{S}$,
$\collapse{\eta}{\sigma} \in \interps{ \mathcal{R} }$ maps each $R \in
\mathcal{R}$ to
%
$\bigland_{ \substack{R' \in \mathcal{R}' \\ \eta(R') = R} }
\sigma(R')$.
%
$\expand{ \mathcal{R} }$ is given in \autoref{app:cons-cdd}.

% shara correctness
\begin{thm}
  \label{thm:corr}
  $\mathcal{S}$ is solvable if and only if $\shara{ \mathcal{R}
  }(\mathcal{S}) \in \interps{ \mathcal{R} }$.
\end{thm}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End:

### 2018-04-05 paper/background.tex
add papers

% technical background material:
\section{Background}
\label{sec:background}

% interpolants:
\subsection{Logical interpolation}
\label{sec:itps}
%
\sys solves CHC systems, logic-programming problems in which a
solution maps predicate symbols to first-order formulas.
%
All objects defined in this paper are defined over a fixed space of
first-order logical variables $X$.
%
In particular, for theory $\mathcal{T}$, the space of $\mathcal{T}$
formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.
%
For each formula $\varphi \in \tformulas{\mathcal{T}}$, the set of
variables that occur in $\varphi$ (i.e., the \emph{vocabulary} of
$\varphi$) is denoted $\vocab(\varphi)$.
% define models, satisfaction, entailment
For formulas $\varphi_0, \ldots, \varphi_n, \varphi \in
\tformulas{\mathcal{T}}$, the fact that $\varphi_0, \ldots, \varphi_n$
\emph{entail} $\varphi$ is denoted $\varphi_0, \ldots, \varphi_n
\entails \varphi$.

% Define replacement and substitution:
For sequences of variables $Y = [ y_0, \ldots, y_n ]$ and $Z = [ z_0,
\ldots, z_n ]$, the $\mathcal{T}$ formula constraining the equality of
each element in $Y$ with its corresponding element in $Z$, i.e., the
formula $\bigland_{0 \leq i \leq n} y_i = z_i$, is denoted $Y = Z$.
%
The repeated replacement of variables $\varphi[ z_0 / y_0 \ldots z_{n}
/ y_{n} ]$ is denoted $\replace{\varphi}{Y}{Z}$.
%
For each formula $\varphi$ defined over free variables $Y$,
$\replace{\varphi}{Y}{Z}$ is denoted alternatively as $\varphi[ Z ]$.
%
The number of free variables in formula $\varphi$ is denoted
$\degreeof{\varphi}$.

% introduce interpolation:
An interpolant of mutually inconsistent $\mathcal{T}$ formulas
$\varphi_0$ and $\varphi_1$ is a $\mathcal{T}$ formula in their common
vocabulary that explains their inconsistency.
%
\begin{defn}
  \label{defn:itps}
  % define:
  For $\varphi_0, \varphi_1 \in \tformulas{\mathcal{T}}$, $I \in
  \tformulas{\mathcal{T}}$ such that
  %
  \textbf{(1)} $\varphi_0 \entails I$, %
  \textbf{(2)} $I, \varphi_1 \entails \false$, and %
  \textbf{(3)} $\vocab(I) \subseteq \vocab(\varphi_0) \intersection
  \vocab(\varphi_1)$,
  %
  $I$ is an \emph{interpolant} of $\varphi_0$ and $\varphi_1$.
\end{defn}
%
All spaces of formulas in the remainder of this paper will be defined
for a fixed, arbitrary theory $\mathcal{T}$ that supports
interpolation, such as combinations of the theories of linear
arithmetic and the theory of uninterpreted functions with equality.
% introduce decision procedure:
Although determining the satisfiability of formulas in such theories
is NP-complete in general, decision procedures~\cite{moura08} and
interpolating theorem provers~\cite{mcmillan04} for such theories have
been proposed that operate on such formulas efficiently.
%
\sys is defined as using a decision procedure for $\mathcal{T}$ named
\issat, and an interpolating theorem prover for $\mathcal{T}$ named
$\solveitp$.

% define CHC's:
\subsection{Constrained Horn Clauses}
\label{sec:chcs}

\subsubsection{Structure}
% definition of CHC
A Constrained Horn Clause contains a body and a head.
%
A body consisting of a conjunctive set of uninterpreted relational predicates applied to varialbes and a formula
that over variables.
%
A head can either be an uninterpreted relational predicates applied to variables or a formula 
relational predicate.
%
A Clause where the head is a formula is called a query.
\begin{defn}
  A Constrained Horn Clause is definied as follows: \\
  $\cc{chc} \Coloneqq \cc{head} \gets \cc{body}$ \\
  $\cc{head} \Coloneqq \varphi \mid \cc{pred}$  \\
  $\cc{body} \Coloneqq \varphi \wedge \cc{preds}$ \\
  $\cc{preds} \Coloneqq \cc{True} \mid \cc{pred} \wedge \cc{preds}$ \\
  $\cc{pred} \Coloneqq$ an uninterpreted predicate applied to variables \\
  $\varphi \Coloneqq$ a formula that all variables are vocabulary \\
\end{defn}
%
% define a body function, head function, and recursion-free
For a given constrained horn clause $C$, $\bodyof{C}$ denotes
the vector of uninterpreted predicates in the body of $C$.
%
It is possible for a uninterpreted predicate shows multiple times in the
body of $C$.
%
For a given constrained horn clause $C$ that is not a query,
$\headof{c}$ denotes the uninterpreted predicates of the head of $C$.
%

%define a CHC system
A Constrained Horn Clauses system is a set of constrained horn clauses 
that it has one and only one query.
%
For a given Constrained Horn Clauses system $S$,$\predof{S}$ denotes 
the set of all uninterpreted predicates in this CHC system.

% define dependency predicates
\begin{defn}
  Given a CHC system $\cc{S}$ and two uninterpreted predicates $\cc{P}$ and
  $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$ such that $\cc{P} = \headof{C}$
  and $\cc{Q} \in \bodyof{C}$, then $\cc{Q}$ is a dependency predicate of $\cc{P}$.
\end{defn}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$, $\depsof{P}$
denotes the set of all dependency predicates of $\cc{P}$ in $\cc{S}$.
%

%define transitive dependency predicates
\begin{defn}
  Given a CHC system $\cc{S}$ and three uninterpreted predicates $\cc{P},
  \cc{Q}$ and $\cc{R} \in \predof{S}$, if $\cc{Q} \in \depsof{P}$ and $\cc{R} \in \depsof {Q}$
  , then $\cc{R}$ is a transitive dependency predicate of $\cc{P}$.
\end{defn}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$, $\tdepsof{P}$
denotes the set of all transitive dependency predicates of $\cc{P}$ in $\cc{S}$.
%

%define siblings
\begin{defn}
  Given a CHC system $\cc{S}$ and two uninterpreted predicates $\cc{P}$ and
  $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$ such that $\cc{P} \in \bodyof{C}$
  and $\cc{Q} \in \bodyof{C}$, then $\cc{Q}$ and $\cc{P}$ are siblings of each other.
\end{defn}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$, $\siblingof{P}$
denotes the set of all siblings predicates of $\cc{P}$ in $\cc{S}$.
%

% define system of CHC's
We will present \sys as a solver for recursion-free CHCs.
%
For a given CHC system $\cc{S}$, there is no uninterpreted predicates $\cc{P} \in \predof{S}$ 
such that $\cc{P} \in \tdepsof{P}$, then $\cc{S}$ is a recursion-free CHCs system.

\subsubsection{Solutions}
\label{sec:chc-solns}
%
A solution to a CHC system $\mathcal{S}$ is an interpretation of each
relational predicate $R$ of arity $n$ as a formula over $n$ free
variables such that for each clause $\mathcal{C} \in \mathcal{S}$, the
conjunction of interpretations of all relational predicates in the
body of $\mathcal{C}$ and the constraint of $\mathcal{C}$ entail the
interpretation of the head of $\mathcal{C}$.
%
Let a map from each $R \in \relsof{ \mathcal{R} }$ to a
$\mathcal{T}$-formula over an ordered vector of $\arityof{ \mathcal{R}
}(R)$ free variables be an \emph{interpretation} of $\mathcal{R}$;
%
let the space of interpretations of $\mathcal{R}$ be denoted
$\interps{ \mathcal{R} }$.
%
\begin{defn}
  \label{defn:chc-soln}
  For $\mathcal{B} \in \bodies{ \mathcal{R} }$ and $H \in \apps{
    \mathcal{R} }$, %
  let $\sigma \in \interps{ \mathcal{R} }$ be such that for each $R
  \in \mathcal{R}$, $\arityof{\mathcal{R}}(R) = \degreeof{\sigma(R)}$
  and %
  %
  \[ \elts{ \sigma(\relof{ A })[ \argsof{ A } ] }_{ A \in \appsof{
      \mathcal{B} } }, %
  \ctrof{ \mathcal{B} } \entails %
  \sigma( \relof{H} )[ \argsof{ H } ]
  \]
  Then $\sigma$ is a \emph{solution} of $(\mathcal{B}, H)$.
\end{defn}
% define partial solution 
For $\mathcal{S} \in \chcs{ \mathcal{R} }$, if
% solution respects dependence order:
\textbf{(a)} for each $R \in \domain(\sigma)$ (where $\domain(\sigma)$
denotes the domain of $\sigma$) and $R' \in \mathcal{R}$ a dependence
of $R$ in $\mathcal{S}$, it holds that $R' \in \domain(\sigma)$;
%
\textbf{(b)} for each $\mathcal{C} \in \mathcal{S}$ such that
$\relof{ \headof{ \mathcal{C} } } \in \domain(\sigma)$, $\sigma$ is a
solution of $\mathcal{C}$;
%
then $\sigma$ is a \emph{partial solution} of $\mathcal{S}$.
%
If, in addition, $\sigma(\queryof{ \mathcal{S} }) \entails \false$,
then $\sigma$ is a \emph{solution} of $\mathcal{S}$.
%
The problem addressed in this paper is, given a recursion-free CHC
system $\mathcal{S}$, to synthesize a solution for $\mathcal{S}$.

% define grounding of clause body:
Each partial solution $\sigma$ of $\mathcal{S}$ defines a simpler CHC
system in which each relational predicate in the domain of $\sigma$ is
replaced with its interpretation in $\sigma$.
%
For $\mathcal{A} \subseteq \apps{\mathcal{R}}$, $\varphi \in
\formulas$, and $\sigma \in \interps{ \mathcal{R} }$, the clause body
\begin{align*}
  \left( \bigunion_{ \substack{ %
      R \in \mathcal{R} \setminus \domain(\sigma), \\
      Y \in X^{*}, 
      (R, Y) \in \mathcal{A} } } \elts{(R, Y)}, 
  \varphi \land %
  \bigland_{ \substack{ R \in \domain(\sigma) \\ 
      Y \in X^{*},
      (R, Y) \in \mathcal{A} } } \sigma(R)[ Y ] \right)
\end{align*}
%
is the \emph{grounding} of $(\mathcal{A}, \varphi)$ on $\sigma$,
denoted $\sigma(\mathcal{A}, \varphi)$.

% groundings of systems:
For each $\mathcal{S} \in \chcs{ \mathcal{R} }$ and $\sigma \in
\interps{ \mathcal{R} }$ a partial solution of $\mathcal{S}$, the
grounding of $\mathcal{S}$ on $\sigma$ is a CHC system in which each
clause is a grounding of a clause in $\mathcal{S}$ with a head
relational predicate not in the domain of $\sigma$.
%
In particular, let $\mathcal{S}' \subseteq \chc{ \mathcal{R} \setminus
  \domain(\sigma) }$ be such that for each $\mathcal{B} \in \bodies{
  \mathcal{R} }$ and $H \in \apps{ \mathcal{R} }$ with $\relof{H}
\notin \domain(\sigma)$, the clause $(\sigma(\mathcal{B}), H)$ is in
$\mathcal{S}'$.
%
Then $\mathcal{S}'$ is the \emph{grounding} of $\mathcal{S}$ on
$\sigma$, denoted $\sigma(\mathcal{ S })$.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End:

### 2018-04-05 paper/evaluation.tex
add papers

\section{Evaluation}
\label{sec:evaluation}
% describe systems:
We performed an empirical evaluation of \sys to determine in what
cases \sys performs better or worse than existing solvers for
recursion-free CHC systems.
%
To do so, we implemented \sys as a modification of \duality CHC
solver, which is included in the \zthree theorem prover~\cite{z3}.
%
We modified \duality to use \sys as its solver for recursion-free CHC
systems.
%
We modified the algorithm used by \duality to generate recursion-free
unwindings of a given recursive system so that in each iteration, it
generates an unwinding with a maximal set of relational predicates and
clauses.
%
In the following context, ``\sys'' refers to \duality modified to
solve general CHC systems.

% describe benchmarks:
We evaluated \sys and \duality on 4,309 CHC systems generated from
programs in the SV-COMP 2015~\cite{svcomp15} verification benchmark
suite.
%
To generate CHC systems, we ran the \seahorn~\cite{gurfinkel15}
verification framework with its default settings, set to timeout at 90
seconds.
%
We used the benchmarks in SV-COMP 2015~\cite{svcomp15} because they
were used to evaluate \duality in previous work~\cite{mcmillan14}.
%
We plan to make \sys publicly available.

% experimental setups:
All experiments were run using a single thread 
on a machine with 16 1.4 GHz processors and
128 GB of RAM.
%
We ran the solvers on each benchmark, timing out each implementation
after 180 seconds.

% summarize the result
Out of 4,309 benchmarks, \sys solved or refuted 2,408, timed out on
762, and for 1,139 benchmarks, generated a constraint that caused
\zthree's interpolating theorem prover to fail, which caused \sys to
neither be able to solve or refute the CHC system.
%
\duality solved or refuted 2,321 benchmarks, timed out on 1,145, and
failed due to an error thrown by \zthree's interpolating theorem
prover on 843.
%
The two solvers can induce a failure in \zthree on different systems
because in attempting to solve a given system, they generate different
interpolation queries.

%
We also attempted to compare \sys to the \eldarica CHC solver, but
it could not parse the CHC systems generated by \seahorn.
%
We compared \sys and \eldarica on an alternative set of benchmarks
generated by the UFO model checker~\cite{albarghouthi12c}, and found
that \sys outperformed \eldarica by at least an order of magnitude on
an overwhelming number of cases.
%
As a result, we focus our discussion on a comparison of \sys and
\duality.

\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.48\textwidth] %
    {\caption{Solving times of \sys vs. \duality.
        %
        The $x$ and $y$ axes range over the solving times in seconds
        of \sys and \duality, respectively.
        %
        Each point depicts the performance of a benchmark.
        %
        The line $y = x$ is shown in red.} %
      \label{fig:complete-data} }
    { \includegraphics[width=\linewidth]{fig/complete.png} }
    %
    \ffigbox[.48\textwidth] %
    { \caption{Times of \sys and \duality vs. system size.
        %
        The $x$-axis ranges over the size of a given system, and the
        $y$-axis ranges over solvers' times.
        %
        Measurements of \sys and \duality are shown in blue and
        red, respectively. } %
      \label{fig:size} }
     { \includegraphics[width=\linewidth]{fig/size.png} }
  \end{floatrow}
\end{figure}
% introduce figure, summarize data not included:
The results of our evaluation are contained in
\autoref{fig:complete-data}.
%
Of the 4,040 benchmarks on which both solvers took a negligible amount
of time---less than five seconds---\sys solved the benchmarks in an
average of 0.51 seconds and \duality solved them in an average of 0.42
seconds.
%
\autoref{fig:complete-data} contains data on only the benchmarks that
were not solved in a negligible amount of time.
% give key numbers on non-negligible cases:
Out of the 269 non-negligible benchmarks, \sys solved 185 in less time
than \duality, and solved 159 in less than half the time of \duality.
%
\duality solved 84 in less time than \sys, and solved 53 in less than
half the time of \sys.
%
Of the 762 benchmarks on which \sys timed out, \duality solved or
found a counterexample to 185.
%
Of the 1,145 benchmarks on which \duality timed out, \sys solved or
found a counterexample to 470.

% analyze data on system size:
\autoref{fig:size} shows the relationship between the solving times of
\duality and \sys and the size of a given system, measured as lines of
code in the format generated by \seahorn.
%
Because the majority of files have size ranging between 1,000 and
100,000 lines, \autoref{fig:size} is restricted to this range to
clarify the presentation.
%
The data indicates that that performance improvement of \sys compared
to \duality is consistent across systems of all sizes available.

%result is good
The results indicate that \textbf{(1)} on a significant number of
different verification problems, \sys can perform significantly better than
\duality, but
%
\textbf{(2)} there are some cases in which the strengths of each
algorithm yield better results.
%
We collected the differences between sizes of a given system and its
minimal CDD expansion generated by \sys and found that they were
independent of \sys's performance compared to \duality.
%
Thus, while \duality may in the worst case enumerate exponentially
many derivation trees, it appears to enumerate far fewer
than the worst-case bound in some cases, causing it to perform better
than \sys.
%
Our results indicate that a third approach that combines the strengths
of both \duality and \sys, perhaps by lazily unwinding a given
recursion free system into a series of CDD systems instead of
derivation trees, could yield further improvements.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End:

### 2018-04-05 paper/introduction.tex
add papers

% intro:
\section{Introduction}
\label{sec:intro}
% context, problem defn:
Many critical problems in program verification, such as the
verification of safety properties of sequential recursive programs and
concurrent programs, can be reduced to solving systems of
\emph{Constrained Horn Clauses} (CHCs), a class of logic-programming
problems~\cite{bjorner13,flanagan03,rummer13a,rummer13b}.

% general introduce chc system
A CHC consists of a body of applications of uninterpreted relational
predicates, a constraint in a first-order theory, and a head
application of a relational-predicate.
%
A CHC system consists of a set of CHCs and a query relational
predicate.
% 
A solution of a system is an interpretation of each relational
predicate as a formula such that each clause body with predicate
symbols substituted with their interpretation entails the interpreted
clause head.

% introduce the problem of solving recursion-free systems:
Recursion-free CHC systems, in which each
relational-predicate does not occur in a derivation (i.e., is not
a \emph{dependency}) of itself, is an important subclass of general
CHC systems for two reasons.
%
First, many modern solvers solves general CHC systems by solving 
a series of recursion-free systems from bounded unwindings of given
recursive CHC system and combining their solutions~\cite{bjorner13}.
%
The performance of the solver for recursive systems critically depends
on the performance of the solver for recursion-free systems that it
uses. 
%
Second, solving recursion-free systems formulates proving safety of
hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12}, i.e.,
programs with control branches and procedure calls, but only bounded
iteration and recursion.

% current general techniques for solving recursion-free systems
Many efficient CHC solvers, solve a recursion-free CHC system 
by reducing it to one or a series of subclass of 
recursion-free CHC system.
%
Then solvers solve these subclass of recursion-free CHC systems by
issuing an interpolation query for each relational predicate.
%
We refer to such subclass of systems as \emph{directly solvable}.
%

% state the complexity of solving general problem,and explain the 
% why it is fast
In general, solving a recursion-free CHC system for the theory of
integer linear arithmetic is co-NEXPTIME-complete~\cite{rummer13b}.
%
In contrast, solving a directly solvalbe system for the theory of
integer linear arithmetic is co-NP~~\cite{rummer13b}.
%
Therefore, the performance of a CHC solver for a recursion-free CHC system 
typically is directly determined by the size(i.e the number of predicates) 
of the subclass systems it reduces to.

% current subclass
Previous work has introduced three classes of directly solvable systems,
% derivation trees:
\textbf{(1)} tree structure systems~\cite{heizmann10,bjorner13,mcmillan14}, %
% disjunctive trees:
\textbf{(2)} body-disjoint systems, which represent multiple
derivation trees with a single disjunctive tree
system~\cite{rummer13a,rummer13b}, and
% linear systems:
\textbf{(3)} linear CHC systems, which can compactly represent
multiple linear derivations with a single system structured as a
Directed Acyclic Graph (DAG)~\cite{albarghouthi12a}.
%
The class of body-disjoint systems strictly contains the class of
derivation trees, and is independent of the class of linear systems
(i.e., the classes overlap, and neither contains the
other)~\cite{rummer13a,rummer13b}.

% contribution of this paper: CDD systems:
The first contribution of this paper is the introduction of a novel
class of directly solvable systems that strictly
contains the union of the classes of body-disjoint and linear systems.
%
In particular, the class consists of all CHC systems such that, for
each clause $\mathcal{C}$, the dependences of all distinct relational
predicates in $\mathcal{C}$ are disjoint;
%
we thus refer to such systems as \emph{Clause-Dependent Disjoint
  (CDD)} systems.
%
We present a symbolic algorithm that demonstrates that CDD is directly solvable.
%

% a new solver
The second contribution of this paper is a novel solver for
CHC systems, named \sys.
%
Given recursion-free system $\mathcal{S}$, \sys reduces the problem
of solving $\mathcal{S}$ to solving an equivalent CDD system
$\mathcal{S}'$.
%
In the worst case, the size of $\mathcal{S}'$ may be exponential in the size of
$\mathcal{S}$.
%
In our experience, the size of $\mathcal{S}'$ is usually close enough
to the size of $\mathcal{S}$ for \sys to perform significantly better
than the best known CHC solvers.
%
Because CDD strictly contains the union of the classes 
of body-disjoint and linear systems, the size of an equivalent minimum 
CDD system is equal or less then the size of equivalent body-disjoint and
linear system.
%
The procedure implemented in \sys to construct $\mathcal{S}'$ is a
generalization of existing procedures for generating compact
verification conditions of hierarchical
programs~\cite{flanagan01,lal-qadeer15}.
%
Given a recursion system, \sys solves a series recursion-free systems 
from bounded unwinding the original system and combines the solutions
of recursion-free systems to generate the solution for the recursion system
as previous work proposed~\cite{rummer13b}.
%

% experience:
\QZ{need to change if need}
We implemented \sys within the \duality CHC solver~\cite{bjorner13},
which is implemented within the \zthree automatic theorem
prover~\cite{moura08}.
%
We evaluated the effectiveness of \sys on standard benchmarks drawn
from SVCOMP15~\cite{svcomp15}.
%
The results indicate that in a strong majority of practical cases,
\sys performs better than modern solvers.
%
The results indicate that solvers that perform better than both modern
solvers and \sys could be designed by combining the strengths of both
approaches (discussed in \autoref{sec:evaluation}).

% paper outline:
The rest of this paper is organized as follows.
%
\autoref{sec:overview} illustrates the operation of \sys using an
example hierarchical program and corresponding CHC system.
%
\autoref{sec:background} reviews technical work on which \sys is
based.
%
\autoref{sec:approach} describes \sys in technical detail.
%
\autoref{sec:evaluation} gives the results of our empirical evaluation
of \sys.
%
\autoref{sec:related-work} compares \sys to related work.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End:

### 2018-04-05 paper/overview.tex
add papers

\section{Overview}
\label{sec:overview}

%first section
In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
$\mcchc$, that formulates verifying the safety of the program
\cc{multiCall} (\autoref{fig:multicall-code}), and illustrate that
$\mcchc$ is not in the directly solvable classes introduced in
previous work.
%
In \autoref{sec:solve-ex}, we show that $\mcchc$ is a
\emph{Clause-Dependent Disjoint (CDD)} system and how \sys solves it.

% introduce running example program
\subsection{Verifying \cc{multiCall}: an example hierarchical program}
\label{sec:running-ex}

% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.26\textwidth] %
    { \caption{\cc{multiCall}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    { \input{code/multiCall.java} }
    %
    \ffigbox[.7\textwidth] %
    {\caption{Dependencies of CHC system $\mcchc$, depicted as a
        hypergraph.
        % 
        Each relational predicate is a node (circle) and each clause
        is a hyperedge (box). } %
      \label{fig:ex-deps} }
    { \includegraphics[width=\linewidth]{fig/ex-deps.pdf} }
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{multiCall} includes a procedure \cc{dbl} that, given an integer
variable \cc{n}, returns the value of \cc{n} doubled (line 1).
%
\cc{main}, the entry procedure of \cc{multiCall}, given integer \cc{n}
(line 3), first binds variable \cc{res} to $0$ (line 4).
%
\cc{main} then tests if \cc{n} is greater than or equal to $0$ (line
5).
%
If so, \cc{main} doubles the value in \cc{n} and binds the result to
\cc{res} (line 6).
%
If not, \cc{main} doubles the value in \cc{n}, negates the result, and
binds the final result to \cc{res} (line 8).
%
\cc{main} then returns the value stored in \cc{res} (line 9).

% state the safety problem:
The problem that we address in this work, in the context of
\cc{multiCall}, is to verify that for each input, \cc{multiCall}
returns a value greater than or equal to $0$.
% talk about solving Horn Clauses:
Verifying that \cc{multiCall} satisfies such a property can be reduced
to solving a recursion-free CHC system over a set of relational
predicates $\mcpreds$ that contains each control location and
procedure name in \cc{multiCall}.
%
A solution interprets each control location \cc{L} as a summary that
relates the arguments in each active call of the function containing
\cc{L} to the program state at \cc{L}, and interprets each procedure
name \cc{f} as a summary of \cc{f}.
%
For each step of \cc{multiCall} to \cc{L}, there is a corresponding
clause with head predicate \cc{L}.
%
For each procedure \cc{f}, there is a clause relating state at return
locations of \cc{multiCall} to a summary of their procedure.
%
In particular, one such system $\mcchc$ is
%
\begin{gather}
  % semantic constraint of dbl:
  \cc{dbl}(\cc{n}, 2 \cdot \cc{n}) \qquad
  % semantic constraint of line 3:
  \cc{L3}(\cc{n}) \implies 
  % semantic constraint of line 4:
  \cc{L4}(\cc{n}, 0) \notag \\
  % semantic constraint of line 5 (then branch)
  \cc{L4}(\cc{n}, \cc{res}), \cc{dbl}(\cc{n}, \cc{res}'), %
  \cc{n} \geq 0 \implies %
  \cc{L9}(\cc{n}, \cc{res}') \label{eqn:then} \\
  % semantic constraint of line 6 (else branch)
  \cc{L4}(\cc{n}, \cc{res}), \cc{dbl}(\cc{n}, \cc{res}'), %
  \cc{n} < 0 \implies %
  \cc{L9}(\cc{n}, -\cc{res}') \label{eqn:else} \\
  % semantic constraint of main procedure:
  \cc{L9}(\cc{n}, \cc{res}) \implies %
  \cc{main}(\cc{n}, \cc{res}) \qquad
  % errors:
  \cc{main}(\cc{n}, \cc{res}), \cc{res} < 0 \implies \mathsf{Err}()
  \notag
\end{gather}
%
While $\mcchc$ has been presented as the result
of a translation from hierarchical program \cc{multiCall}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

% walk through the zoo of CHC systems:
Competitive solvers of recursion-free CHC systems reduce the problem
of solving a given CHC system $\mathcal{S}$ to solving multiple
systems, constructed as copies of sub-systems of $\mathcal{S}$, each
of which is directly solvable (as described in \autoref{sec:intro}).
%
In particular, CHC solvers based on tree interpolation reduce solving
$\mathcal{S}$ to solving copies of subsystems of $\mathcal{S}$ in
which each relational predicate in $\mathcal{S}$ is the head of at
most one clause~\cite{bjorner13,heizmann10}.
%
Such systems do not include $\mcchc$, which contains multiple clauses
with the same head.

% consequences for solvers:
As a result, CHC solvers that synthesize tree interpolants may, in the
worst case, solve systems that correspond to each derivation of a
given CHC system.
%
Such solvers, given $\mcchc$, would solve it by generating two
tree-interpolation problems, one of which corresponds to the \cc{then}
branch of \cc{multiCall} and the other which corresponds to the
\cc{else} branch of \cc{multiCall}.
%
CHC solvers have been proposed which can in many cases avoid
enumerating and solving all derivation trees of a given recursion-free
system~\cite{mcmillan14}.
%
However, such solvers may enumerate all derivations of a given system
in the worst case.

% Body-Disjoint: 
CHC solvers that synthesize disjunctive interpolants reduce solving a
given recursion-free system to solving a series of CHC systems in
which each relational predicate occurs in the body of at most one
clause, and occurs in the body of such a clause at most once;
%
such systems are referred to as \emph{body-disjoint}
systems~\cite{rummer13a,rummer13b}.
%
However, such systems cannot compactly model a program with multiple
control paths that share a common subpath, typically modeled as a CHC
system with a relational predicate that occurs in multiple clauses
(i.e., a relational predicate that is \emph{shared}).
%
E.g., $\mcchc$ is not a body-disjoint system because the relational
predicates \cc{L4} and \cc{dbl} occur in both clauses
\autoref{eqn:then} and \autoref{eqn:else}.

% consequences for relational predicates:
CHC solvers that solve body-disjoint systems, given a general recursion-free CHC system,
generate multiple copies of sub-systems that body has shared
relational predicates.
%
Such a solver, given $\mcchc$, generates a body-disjoint system
consisting of multiple copies of the sub-systems with head relational
predicates \cc{L4} and \cc{dbl}.

% solving linear systems:
CHC solvers that synthesize DAG interpolants~\cite{albarghouthi12a}
reduce the problem of solving a given CHC system to solving a series
of systems in which each clause contains at most one relational
predicate (i.e., \emph{linear} systems).
%
However, such systems cannot directly model the control flow of a
program that contains procedure calls, such as \cc{multiCall}.
%
Calls can only be compactly modeled using non-linear clauses, such as
clauses \autoref{eqn:then} and \autoref{eqn:else}.

% solvers based on linear systems:
CHC solvers that solve linear systems, given an arbitrary
recursion-free system $\mathcal{S}$, effectively inline constraints on
relational predicates that occur in non-linear clauses of
$\mathcal{S}$~\cite{albarghouthi12b}.
%
Such approaches can, in general, generate an inlined system that is
exponentially larger than a given system.

% solving the running example
\subsection{\cc{multiCall} as a Clause-Dependent Disjoint system}
\label{sec:solve-ex}
% restate defn of CDD systems:
\sys solves $\mcchc$ directly by issuing a series of
interpolation queries, each of which generates an interpretation for a
relational predicate in $\mcchc$.
%
The key observation behind the design of \sys is that for CHC system
$\mathcal{S}$, if the dependences of each relational predicate in
each clause of $\mathcal{S}$ are disjoint (in which we case we say
that $\mathcal{S}$ is \emph{Clause-Dependent-Disjoint (CDD)}), then
$\mathcal{S}$ is directly solvable.
%
The formal definition of CDD and its key properties are given in
\autoref{sec:CDD-defn}.

% example system is CDD and can be easily solved:
E.g., $\mcchc$ is CDD because for each clause $\mathcal{C}$ of
$\mcchc$, the dependences of all distinct relational predicates in the
body of $\mathcal{C}$ are disjoint.
%
In particular, for clauses \autoref{eqn:then} and \autoref{eqn:else},
the dependences of relational predicates \cc{L4} and \cc{dbl} are
disjoint.
%
The dependences of \cc{L4} is \{\cc{L3}\}, and the dependneces of \cc{dbl}
is $\varnothing$.
%
\autoref{fig:ex-deps} contains a hypergraph $H_{\cc{MC}}$ that depicts
the dependences induced by the clauses of $\mcchc$.
%
The fact that $\mcchc$ is CDD corresponds to the fact that no node in
$H_{\cc{MC}}$ reaches distinct sources of a common hyperedge.

% walk through how a CDD system is solved directly:
A solution of an arbitrary CDD system $\mathcal{S}$ can be constructed
from the results of a series of interpolation queries, each one issued
to find an interpretation of a relational predicate of $\mathcal{S}$
in the topological order of dependences between relational predicates.
%
For relational predicate $R$ in $\mathcal{S}$, the interpretation of
$R$ is the interpolant of two formulas: %
\textbf{(1)} a Boolean combination of the interpretations found for
each dependence of $R$ in $\mathcal{S}$ and constraints in
$\mathcal{S}$ and %
\textbf{(2)} a Boolean combination of constraints that occur in
$\mathcal{S}$.

% talk performance:
The performance of the solver is thus largely determined by the
ability of the interpolating theorem prover that it uses \textbf{(1)}
to generate interpolants quickly and %
\textbf{(2)} to generate interpolants of size close to that of their
input formulas.
%
In the worst case, all interpolating theorem provers given in previous
work may use exponential time to generate interpolants of size
exponential in the size of their inputs.
%
However, when given formulas that model practical verification
problems, modern theorem provers generate interpolants efficiently,
which have size close to the size of the given formulas.

% walk through multiCall
E.g., when solving $\mcchc$, \sys generates interpretations of
relational predicates in the order \cc{L3}, \cc{L4}, \cc{dbl},
\cc{L9}, and \cc{main}.
%
It generates an interpretation of \cc{L9} in particular as the
interpolant of \textbf{(1)} a \emph{pre}-formula, constructed as a
Boolean combination of interpretations generated for \cc{L4}, \cc{dbl}
and the constraints in \autoref{eqn:then} and \autoref{eqn:else} and %
\textbf{(2)} a \emph{post}-formula constructed from constraints in
$\mcchc$ that model the steps of execution of \cc{multiCall} following
\cc{9}.
%
A procedure for solving a CDD system directly is given in
\autoref{sec:solve-cdd}.

% talk about expansion:
The class of CDD systems strictly contains all directly-solvable
classes introduced in previous work, but does not contain all
recursion-free systems.
%
E.g., a recursion-free system that models \cc{multiCall} mutated to
call \cc{dbl} in sequence, instead of in alternate branches, would not
be CDD.
%
\sys, given an arbitrary recursion-free CHC system $\mathcal{S}$,
solves it by generating an equivalent CDD system $\mathcal{S}'$ and
solving $\mathcal{S}'$ directly.
%
In general, $\mathcal{S}'$ may have size exponential in the size of
$\mathcal{S}$.
%
In practice, a suitable equivalent system can be generated using
heuristics analogous to those used to generate compact verification
conditions of hierarchical programs~\cite{flanagan01,lal-qadeer15}.
%
A general procedure for constructing a CDD expansion of a given CHC
system is given in \autoref{app:cons-cdd}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End:

### 2018-04-05 paper/related-work.tex
add papers

\section{Related Work}
\label{sec:related-work}
% different classes of CHC solvers:
A significant body of previous work has presented solvers for
different classes of Constrained Horn Clauses, or finding inductive
invariants of programs that correspond to solutions of CHCs.
% solving linear systems:
\impact attempts to verify a given sequential procedure, which
corresponds to solving a recursive linear CHC
system~\cite{mcmillan06}.
%
\impact attempts to verify a given procedure by iteratively selecting
paths and synthesizing invariants for each path.
%
Such an approach corresponds to iteratively and selecting and solving
derivations of a corresponding linear CHC system.

% interprocedural verification:
Previous work also proposed a verifier for recursive
programs~\cite{heizmann10}.
%
The proposed approach selects interprocedural paths of a program and
synthesizes invariants for each as nested interpolants.
%
Such an approach corresponds to attempting to solve a recursive CHC
system $\mathcal{S}$ by selecting derivation trees of $\mathcal{S}$
and solving each tree.

% solving recursive systems:
Previous work has proposed solvers for recursive systems that, given a
system $\mathcal{S}$, attempt to solve $\mathcal{S}$ by generating and
solving a series of recursion-free unwindings of $\mathcal{S}$.
%
In particular, \duality attempts to solve each unwinding
$\mathcal{S}'$ by selecting and solving derivation-trees of
$\mathcal{S}'$~\cite{bjorner13}.
%
Other optimizations selects derivation trees to solve using symbolic
symbolic analogs of Prolog evaluation with
tabling~\cite{jaffar09,mcmillan14}.
%
\eldarica attempts to solve each recursion-free CHC system by lazily
copying its subsystems to form a body-disjoint
over-approximation~\cite{rummer13a,rummer13b}.

% DAG interpolation:
\whale attempts to verify a sequential recursive program by generating
and solving hierarchical programs (i.e., programs that may contain
conditional branches and procedure calls, but do not contain loops or
recursion), which correspond to recursion-free CHC
systems~\cite{albarghouthi12b}.
%
To solve a particular recursion-free system $\mathcal{S}$, \whale
generates a linear inlining $\mathcal{S}'$ of $\mathcal{S}$ and solves
it using a procedure \vinta~\cite{albarghouthi12a}.
%
In general, $\mathcal{S}'$ may have size exponential in the size of
$\mathcal{S}$.

% compare sys to all of the previous work:
\sys is similar to all of the approaches given above for solving
recursion-free CHC systems in that it reduces the problem of solving a
given recursion-free CHC system $\mathcal{S}$ to solving a CHC system
in a directly-solvable class.
%
\sys is distinct from all of the approaches given above in that it
reduces solving a recursion-free CHC system to solving a
Clause-Dependent Disjoint (CDD) system.
%
CDD systems strictly contain the union of all classes of
directly-solvable CHC systems used by the above approaches, and can
themselves be solved directly.

% bullshit CADE papers:
Previous work has given solvers for non-linear Horn clauses over
particular theories.
%
In particular, verifiers have been proposed for recursion-free systems
over the theory of linear arithmetic~\cite{komuravelli14}.
%
Because the verifier relies on quantifier elimination, it is not clear
if it can be extended to richer theories that support interpolation,
such as the combination of linear arithmetic with uninterpreted
functions.
%
Other work gives a solver for the class of \emph{timed pushdown
  systems}, a subclass of CHC systems over the theory of linear real
arithmetic~\cite{hoder12}.
%
Unlike both approaches, \sys can solve systems over any theory that
supports interpolation.

% DAG inlining:
DAG inlining, given a hierarchical program $P$, attempts to generate a
compact verification condition for $P$~\cite{lal-qadeer15}.
%
\sys contains a procedure that, given a recursion-free Horn Clause
system $\mathcal{S}$, attempts to construct a compact CDD system
$\mathcal{S}'$ such that each solution of $\mathcal{S}'$ defines a
solution of $\mathcal{S}$.
%
Because hierarchical programs and recursion-free Horn Clauses
correspond closely to each other, algorithms that operate on
hierarchical programs directly correspond to algorithms that operate
on recursion-free Horn Clauses.
%
However, an algorithm for constructing a verification condition of
hierarchical programs cannot apparently be directly used to synthesize
a solution of a recursion-free CHC system.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End:

### 2018-04-05 paper/code/multiCall.java

@@ -1,0 +1,9 @@
+int dbl(int n) {
+  return 2 * n; }
+int main(int n) {
+  int res = 0;
+  if (n >= 0)
+    res = dbl(n);
+  else
+    res = -1 * dbl(n);
+  return res; }

### 2018-04-06 src/Data/Grammar.hs
started implementing formal grammar approach

@@ -1,0 +1,265 @@
+module Data.Grammar where
+
+import           Control.Lens hiding (Empty, elements, (:<))
+import           Control.Monad.State
+import           Control.Monad.Writer (Writer, runWriter, tell)
+import           Control.Applicative
+
+import           Data.Data (Data)
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+import qualified Data.Set as S
+import           Data.Semigroup
+import           Data.List (intercalate, nub)
+
+import           Test.QuickCheck hiding (NonEmpty)
+
+import           Prelude hiding (seq, sum, product)
+
+import Debug.Trace
+
+newtype NT = NT { getNT :: Int }
+  deriving (Show, Read, Eq, Ord, Data, Real, Bounded, Enum, Integral, Num)
+
+instance Arbitrary NT where
+  arbitrary = do
+    topNT <- growingElements [0..3]
+    NT <$> elements (0:[1..topNT])
+
+data Rule a
+  = Null
+  | Eps
+  | Alt (Rule a) (Rule a)
+  | Seq (Rule a) (Rule a)
+  | Nonterminal NT
+  | Terminal a
+  deriving (Show, Read, Eq, Ord, Data, Functor, Foldable, Traversable)
+
+isNull :: Rule a -> Bool
+isNull Null = True
+isNull _ = False
+
+isEps :: Rule a -> Bool
+isEps Eps = True
+isEps _ = False
+
+seq :: Rule a -> Rule a -> Rule a
+seq Null _      = Null
+seq _ Null      = Null
+seq Eps x       = x
+seq x Eps       = x
+seq (Seq x y) z = Seq x (seq y z)
+seq x y         = Seq x y
+
+alt :: Rule a -> Rule a -> Rule a
+alt Eps Eps     = Eps
+alt Eps x       = alt x Eps
+alt Null x      = x
+alt x Null      = x
+alt (Alt x y) z = Alt x (alt y z)
+alt x y         = Alt x y
+
+instance Arbitrary a => Arbitrary (Rule a) where
+  arbitrary =
+    oneof [ pure Null
+          , pure Eps
+          , alt <$> arbitrary <*> arbitrary
+          , seq <$> arbitrary <*> arbitrary
+          , Nonterminal <$> arbitrary
[…]

### 2018-04-06 src/Data/InfGrammar.hs

@@ -1,0 +1,74 @@
+module Data.InfGrammar where
+
+import           Control.Lens
+import           Control.Monad.State
+import           Control.Applicative
+
+import           Data.Semigroup
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+
+import           Data.Grammar (Grammar, Rule)
+import qualified Data.Grammar as G
+
+type InfGrammar m a = Grammar (Inf m a)
+type InfRule m a = Rule (Inf m a)
+
+data Inf m a
+  = Finite a
+  | Infinite (m (InfGrammar m a))
+
+finite :: InfGrammar m a -> Grammar a
+finite (G.Grammar st rs) = G.mkGrammar (finR st) (fmap finR rs)
+  where
+    finR :: InfRule m a -> Rule a
+    finR = \case
+      G.Null -> G.Null
+      G.Eps -> G.Eps
+      G.Alt x y -> finR x <|> finR y
+      G.Seq x y -> finR x <> finR y
+      G.Nonterminal nt -> G.Nonterminal nt
+      G.Terminal (Finite a) -> G.Terminal a
+      G.Terminal (Infinite _) -> G.Null
+
+-- TODO ignore nonterminals in the given set.
+unroll :: Monad m => Set G.NT -> InfGrammar m a -> m (InfGrammar m a)
+unroll noUnroll (G.Grammar st rs) =
+  uncurry G.mkGrammar <$> runStateT (M.traverseWithKey lunr rs >> unr st) M.empty
+  where
+    lunr :: Monad m => G.NT -> InfRule m a -> StateT (Map G.NT (InfRule m a)) m ()
+    lunr nt r = do
+      r' <- unr r
+      modify (M.unionWith (<|>) (M.singleton nt r'))
+
+    unr :: Monad m => InfRule m a -> StateT (Map G.NT (InfRule m a)) m (InfRule m a)
+    unr = \case
+      G.Null -> pure G.Null
+      G.Eps -> pure G.Eps
+      G.Seq x y -> (<>) <$> unr x <*> unr y
+      G.Alt x y -> (<|>) <$> unr x <*> unr y
+      G.Nonterminal nt -> pure (G.Nonterminal nt)
+      G.Terminal (Finite t) -> pure (G.Terminal $ Finite t)
+      G.Terminal (Infinite ac) -> do
+        G.Grammar st' rs' <- lift ac
[…]

### 2018-04-06 src/Grammar/Shara.hs

@@ -1,85 +1,85 @@
 {-# LANGUAGE QuasiQuotes #-}
 module Grammar.Shara where
 
-import           Control.Monad.IO.Class
+-- import           Control.Monad.IO.Class
 
-import           Data.Map (Map)
-import qualified Data.Map as M
-import           Data.Set (Set)
+-- import           Data.Map (Map)
+-- import qualified Data.Map as M
+-- import           Data.Set (Set)
 
-import           Formula hiding (Rule)
+-- import           Formula hiding (Rule)
 
-import           Grammar.Grammar
-import           Grammar.Shara.LicketySplit
-import           Grammar.Shara.TreeUnwind
+-- import           Grammar.Grammar
+-- import           Grammar.Shara.LicketySplit
+-- import           Grammar.Shara.TreeUnwind
 
-import Data.Text.Prettyprint.Doc
+-- import Data.Text.Prettyprint.Doc
 
-data SharaCfg = SharaCfg
-  { interpolator :: Interpolator
-  , unwindStrategy :: UnwindStrategy
-  } deriving (Show, Read, Eq, Ord)
+-- data SharaCfg = SharaCfg
+--   { interpolator :: Interpolator
+--   , unwindStrategy :: UnwindStrategy
+--   } deriving (Show, Read, Eq, Ord)
 
-data Interpolator
-  = LicketySplit InterpolationStrategy
-  | AnyOrder
-  deriving (Show, Read, Eq, Ord)
+-- data Interpolator
+--   = LicketySplit InterpolationStrategy
+--   | AnyOrder
+--   deriving (Show, Read, Eq, Ord)
 
-data UnwindStrategy
-  = TreeUnwind
-  | DisjointDependencyUnwind
-  deriving (Show, Read, Eq, Ord)
+-- data UnwindStrategy
+--   = TreeUnwind
+--   | DisjointDependencyUnwind
+--   deriving (Show, Read, Eq, Ord)
 
-shara :: Expr -> Grammar -> IO (Either Model (Map Symbol Expr))
-shara = shara' defaultCfg
-  where
-    defaultCfg =
-      SharaCfg
-      { interpolator = LicketySplit ConcurrentInterpolation
-      , unwindStrategy = TreeUnwind
-      -- TODO swap this
-      }
[…]

### 2018-04-06 src/Grammar/Shara/Test.hs

@@ -1,121 +1,121 @@
 {-# LANGUAGE TemplateHaskell #-}
 module Grammar.Shara.Test where
 
-import           Data.Set (Set)
-import qualified Data.Set as S
-import           Data.Map (Map)
-import qualified Data.Map as M
+-- import           Data.Set (Set)
+-- import qualified Data.Set as S
+-- import           Data.Map (Map)
+-- import qualified Data.Map as M
 
-import           Formula hiding (Rule)
-import           Formula.Expr
-import           Grammar.Grammar
-import           Grammar.Graph2
-import           Grammar.Shara.CDD
-import           Grammar.Shara.DAGUnwind
-import           Grammar.Shara.Pre
+-- import           Formula hiding (Rule)
+-- import           Formula.Expr
+-- import           Grammar.Grammar
+-- import           Grammar.Graph2
+-- import           Grammar.Shara.CDD
+-- import           Grammar.Shara.DAGUnwind
+-- import           Grammar.Shara.Pre
 
-data Vertex = Vertex Int [Var] [Edge]
+-- data Vertex = Vertex Int [Var] [Edge]
 
-data Edge = Edge [Vertex] Rule
+-- data Edge = Edge [Vertex] Rule
 
-test1 :: IO ()
-test1 = do
-  let vars1 = [(Var "x1" Int)]
-  let vars2 = [(Var "x2" Int)]
-  let vars3 = [(Var "x3" Int)]
-  let vars4 = [(Var "x4" Int)]
-  let vars4 = [(Var "x5" Int)]
-  let n1 = Nonterminal 0 vars1
-  let n2 = Nonterminal 1 vars2
-  let n3 = Nonterminal 2 vars3
-  let n4 = Nonterminal 3 vars4
-  let n5 = Nonterminal 4 vars4
-  let r1 = Rule n4 (LBool True) [n2]
-  let r2 = Rule n4 (LBool True) [n3] 
-  let r3 = Rule n2 (LBool True) [n1] 
-  let r4 = Rule n3 (LBool True) [n1] 
-  let r5 = Rule n5 (LBool True) [n4]
-  let g = mkGraph (Grammar 0 [r1,r2,r3,r4,r5])
[…]

### 2018-04-06 src/Grammar/Shara2.hs

@@ -1,4 +1,4 @@
-module Grammar.Shara where
+module Grammar.Shara2 where
 
 import           Control.Monad.IO.Class
 import           Control.Arrow (second)

### 2018-04-06 src/Solver/Chc.hs

@@ -1,0 +1,20 @@
+module Solver.Chc where
+
+import Data.Grammar
+import Data.Semigroup
+
+import Formula.Expr
+import Formula.Var
+
+-- | A fragment of a constrained Horn Clause rule.
+data Fragment
+  = Capture Var
+  | Uncapture
+  | Apply Var
+  | Fact Expr
+  deriving (Show, Read, Eq, Ord)
+
+type Chc = Grammar Fragment
+
+scope :: Var -> Rule Fragment -> Rule Fragment
+scope v r = Terminal (Capture v) <> r <> Terminal Uncapture

### 2018-04-06 src/Solver/Shara.hs

@@ -1,0 +1,111 @@
+module Solver.Shara where
+
+import           Control.Lens
+import           Control.Monad.State
+import           Control.Applicative
+
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+import qualified Data.Set as S
+import           Data.Maybe (isJust, fromJust)
+import           Data.Semigroup
+
+import           Formula.Expr
+
+import           Data.Grammar (Grammar, Rule, NT)
+import qualified Data.Grammar as G
+
+import           Data.InfGrammar (InfGrammar)
+import qualified Data.InfGrammar as IG
+
+import           Solver.Types
+
+-- | Encode the CHC as a (possibly infinite) unrollable grammar such that each
+-- unrolling is a directly solvable CHC system.
+directlySolvable :: MonadState DirectState m => Chc -> m (InfGrammar m Expr)
+directlySolvable = undefined
+
+-- | Apply the interpolator to the given directly solvable system and yield
+-- either a counterexample model or solutions for the nonterminals.
+solveDirect :: MonadIO m => Grammar Expr -> m (Either Model (Map NT Expr))
+solveDirect = undefined
+
+-- | Interpolate the given nonterminal in the context of the directly solvable grammar.
+interpolateNT :: MonadIO m => NT -> Grammar Expr -> Map NT Expr -> m (Either Model (Map NT Expr))
+interpolateNT = undefined
+
+-- | Solve the Relational Post Fixed-Point Problem as encoded by the given CHC
+-- system.
+solveChc :: (MonadState DirectState m, MonadIO m) => Chc -> m (Either Model (Map NT Expr))
+solveChc chc = solve =<< directlySolvable chc
+
+solve :: (MonadState DirectState m, MonadIO m) => InfGrammar m Expr -> m (Either Model (Map NT Expr))
+solve g = solveDirect (IG.finite g) >>= \case
+  Left m -> pure (Left m)
+  Right sol -> do
+    ps <- use proofStructure
+    (mproof, indSet) <- inductive ps sol
+    case mproof of
+      Just proof -> do
+        cs <- use clones
+        let sol' = M.filterWithKey (\k _ -> k `elem` G.nonterminals proof) sol
+        pure (Right (collapse cs sol'))
+      Nothing -> IG.unroll indSet g >>= solve
+
[…]

### 2018-04-06 src/Solver/TreeUnwind.hs

@@ -1,0 +1,149 @@
+{-# LANGUAGE QuasiQuotes #-}
+{-# LANGUAGE TemplateHaskell #-}
+module Solver.TreeUnwind where
+
+import           Control.Lens hiding (mapping)
+import           Control.Monad.State
+import           Control.Monad.Reader
+import           Control.Applicative
+
+import           Data.Data (Data)
+import           Data.Semigroup
+import           Data.Set (Set)
+import qualified Data.Set as S
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Sequence (Seq)
+import qualified Data.Sequence as Seq
+import           Data.Grammar (Grammar, NT, Rule)
+import qualified Data.Grammar as G
+import           Data.InfGrammar (InfGrammar, InfRule)
+import qualified Data.InfGrammar as IG
+
+import           Formula hiding (Rule, Chc)
+
+import           Solver.Types hiding (DirectState(..))
+import           Solver.Chc
+
+import Data.Text.Prettyprint.Doc
+
+data UnwindState = UnwindState
+  { _ntCounter :: Int
+  , _varCounter :: Int
+  , _clones :: Map NT (Set NT)
+  , _proofStructure :: ProofStructure
+  , _varQueue :: Seq Var
+  , _varMapping :: [(Var, Var)]
+  }
+makeLenses ''UnwindState
+
+emptyState :: UnwindState
+emptyState = UnwindState
+  { _ntCounter = 0
+  , _varCounter = 0
+  , _clones = M.empty
+  , _proofStructure = G.nullGrammar
+  , _varQueue = Seq.empty
+  , _varMapping = []
+  }
+
+-- Unwind the given CHC system to a tree. In addition, rename the variables in
+-- the query to be consistent with the new CHC vocabulary and generate a map
+-- from symbols in the original grammar to symbols in the new grammar.
+treeUnwind :: MonadState UnwindState m => Chc -> m (InfGrammar m Expr)
+treeUnwind chc = unwindFrom S.empty chc (G.start chc)
+
+unwindFrom :: MonadState UnwindState m
+           => Set NT -> Chc -> Rule Fragment -> m (InfGrammar m Expr)
+unwindFrom vis chc r = do
+  (st, rs) <- runStateT (runReaderT (unw r) vis) M.empty
[…]

### 2018-04-06 src/Solver/Types.hs

@@ -1,0 +1,17 @@
+{-# LANGUAGE TemplateHaskell #-}
+module Solver.Types where
+
+import           Control.Lens
+
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+import           Data.Grammar (Grammar, Rule, NT)
+import qualified Data.Grammar as G
+
+type ProofStructure = Grammar Proof
+
+data Proof = Continue | Entails NT NT
+  deriving (Show, Read, Eq, Ord)
+
+data DirectState = DirectState

### 2018-04-06 src/Data/Grammar.hs
added unwinding

@@ -5,6 +5,7 @@ import           Control.Monad.State
 import           Control.Monad.Writer (Writer, runWriter, tell)
 import           Control.Applicative
 
+import           Data.Foldable (asum)
 import           Data.Data (Data)
 import           Data.Map (Map)
 import qualified Data.Map as M
@@ -104,29 +105,8 @@ mkGrammar :: Rule a -> Map NT (Rule a) -> Grammar a
 mkGrammar st rs =
   let g = Grammar st rs
       r = reaching g
-      rs' = M.filterWithKey (\k _ -> k `elem` r) rs
-      g' = Grammar st rs'
-  in clean g'
-
-clean :: Grammar a -> Grammar a
-clean (Grammar st rs) =
-  let (g', Any change) = runWriter (Grammar <$> go rs st <*> traverse (go rs) rs)
-      g'' = case g' of
-        Grammar st' rs' ->
-          Grammar st' (M.filter (\r -> not (isNull r || isEps r)) rs')
-  in if change then clean g'' else g''
-  where
-    go :: Map NT (Rule a) -> Rule a -> Writer Any (Rule a)
-    go m = \case
-      Null -> pure Null
-      Eps -> pure Eps
-      Alt r1 r2 -> alt <$> go m r1 <*> go m r2
-      Seq r1 r2 -> seq <$> go m r1 <*> go m r2
-      Nonterminal nt -> case M.findWithDefault Null nt m of
-        Null -> tell (Any True) >> pure Null
-        Eps -> tell (Any True) >> pure Eps
-        _ -> pure (Nonterminal nt)
-      Terminal t -> pure (Terminal t)
+      rs' = M.filter (not . isNull) $ M.filterWithKey (\k _ -> k `elem` r) rs
+  in Grammar st rs'
 
 reaching :: Grammar a -> Set NT
 reaching (Grammar st rs) = execState (go S.empty) (S.fromList $ ruleNonterminals st)
@@ -182,11 +162,11 @@ instance Monad Grammar where
 
 product :: Grammar a -> Grammar a -> Grammar a
 product (Grammar st rs) (Grammar st' rs') =
-  Grammar (seq st st') (M.unionWith alt rs rs')
+  mkGrammar (seq st st') (M.unionWith alt rs rs')
 
 sum :: Grammar a -> Grammar a -> Grammar a
 sum (Grammar st rs) (Grammar st' rs') =
-  Grammar (alt st st') (M.unionWith alt rs rs')
[…]

### 2018-04-06 src/Solver/TreeUnwind.hs

@@ -27,81 +27,162 @@ import           Solver.Chc
 
 import Data.Text.Prettyprint.Doc
 
+import Debug.Trace
+
 data UnwindState = UnwindState
   { _ntCounter :: Int
   , _varCounter :: Int
   , _clones :: Map NT (Set NT)
-  , _proofStructure :: ProofStructure
+  , _proofStart :: Rule Proof
+  , _proofRules :: Map NT (Rule Proof)
   , _varQueue :: Seq Var
   , _varMapping :: [(Var, Var)]
-  }
+  , _descendants :: Map NT (Map NT (Set NT))
+  } deriving (Show, Read, Eq, Ord)
+
+data UnwindCtxt = UnwindCtxt
+  { _visited :: Set NT
+  , _branchClones :: Map NT (Set NT)
+  } deriving (Show, Read, Eq, Ord)
+
 makeLenses ''UnwindState
+makeLenses ''UnwindCtxt
 
 emptyState :: UnwindState
 emptyState = UnwindState
   { _ntCounter = 0
   , _varCounter = 0
   , _clones = M.empty
-  , _proofStructure = G.nullGrammar
+  , _proofStart = G.Null
+  , _proofRules = M.empty
   , _varQueue = Seq.empty
   , _varMapping = []
+  , _descendants = M.empty
   }
 
--- Unwind the given CHC system to a tree. In addition, rename the variables in
--- the query to be consistent with the new CHC vocabulary and generate a map
--- from symbols in the original grammar to symbols in the new grammar.
-treeUnwind :: MonadState UnwindState m => Chc -> m (InfGrammar m Expr)
-treeUnwind chc = unwindFrom S.empty chc (G.start chc)
+emptyCtxt :: UnwindCtxt
+emptyCtxt = UnwindCtxt
+  { _visited = S.empty
+  , _branchClones = M.empty
+  }
 
-unwindFrom :: MonadState UnwindState m
-           => Set NT -> Chc -> Rule Fragment -> m (InfGrammar m Expr)
-unwindFrom vis chc r = do
-  (st, rs) <- runStateT (runReaderT (unw r) vis) M.empty
-  pure (G.mkGrammar st rs)
+instance Monad m => Expandable (StateT UnwindState m) where
+  getProof = G.mkGrammar <$> use proofStart <*> use proofRules
+  getClones = use clones
+
+-- | Unwind the given CHC system to a tree.
+-- In addition, generate a proof structure which indicates what an inductive
[…]

### 2018-04-06 src/Solver/Types.hs

@@ -1,4 +1,3 @@
-{-# LANGUAGE TemplateHaskell #-}
 module Solver.Types where
 
 import           Control.Lens
@@ -14,4 +13,6 @@ type ProofStructure = Grammar Proof
 data Proof = Continue | Entails NT NT
   deriving (Show, Read, Eq, Ord)
 
-data DirectState = DirectState
+class Monad m => Expandable m where
+  getProof :: m ProofStructure
+  getClones :: m (Map NT (Set NT))

### 2018-04-06 test/GrammarTest.hs

@@ -1,0 +1,108 @@
+import           Control.Lens
+import           Control.Monad
+import           Control.Applicative
+import           Data.Grammar as G
+
+import           Data.Semigroup
+
+import Test.QuickCheck
+import Test.QuickCheck.Function
+import Test.Hspec
+
+main :: IO ()
+main = hspec $ do
+  describe "a grammar rule" $ do
+    describe "monoid" $ do
+      it "left identity" $ property rMonoidLeftId
+      it "right identity" $ property rMonoidRightId
+      it "associativity" $ property rMonoidAssoc
+    describe "alternative" $ do
+      it "left identity" $ property rAltLeftId
+      it "right identity" $ property rAltRightId
+      it "associativity" $ property rAltAssoc
+    describe "monad" $ do
+      it "left identity" $ property rMonadLeftId
+      it "right identity" $ property rMonadRightId
+      it "associativity" $ property rMonadAssoc
+  describe "a context free grammar" $ do
+    describe "monoid" $ do
+      it "left identity" $ property gMonoidLeftId
+      it "right identity" $ property gMonoidRightId
+      it "associativity" $ property gMonoidAssoc
+    describe "alternative" $ do
+      it "left identity" $ property gAltLeftId
+      it "right identity" $ property gAltRightId
+      it "associativity" $ property gAltAssoc
+    describe "monad" $ do
+      it "left identity" $ property gMonadLeftId
+      it "right identity" $ property gMonadRightId
+      it "associativity" $ property gMonadAssoc
+    describe "partition and connect" $
+      it "connect inverts partition" $ property connectPartition
+
+rMonoidLeftId :: Rule Int -> Bool
+rMonoidLeftId g = mempty <> g == g
+
+rMonoidRightId :: Rule Int -> Bool
+rMonoidRightId g = g <> mempty == g
+
+rMonoidAssoc :: Rule Int -> Rule Int -> Rule Int -> Bool
+rMonoidAssoc a b c = a <> (b <> c) == (a <> b) <> c
+
+rAltLeftId :: Rule Int -> Bool
[…]

### 2018-04-06 paper/approach.tex
editing approach, need to rework overview, and move some appendix into main paper

@@
   % 
-  \myproc{$\solvecdd{ \mathcal{R} }(\mathcal{S})$ %
+  \myproc{$\solvecdd(\cc{S})$ %
     \label{line:solve-begin}}{ %
-    % recurse on a partial solution to the CHC system:
-    \myproc{$\solveaux( \mathcal{ \sigma } )$ %
-      \label{line:aux-begin}}{
-      % construct predicates that:
-      $\mathcal{R}' \assign \setformer{ R }{ %
-        % are not defined by sigma,
-        R \notin \domainof{ \sigma }, %
-        \bigland_{ (R, R') \in \depsof{ \mathcal{S} } } 
-        R' \in \domainof{ \sigma } }$ \label{line:cons-nodeps} \;
-      % case split: on D:
-      \Switch{$\mathcal{R}'$ \label{line:case-deps}}{ %
-        \lCase{$\emptyset$:}{ %
-          % D is empty: return sigma as solution
-          \Return{$\sigma$} \label{line:ret-soln} } %
-        \Case{$\elts{ R } \disjunion \mathcal{R}''$: %
-          \label{line:case-frontier}}{ %
-          % D contains an element R. Update sigma to map R to result
-          % of interpolation query on suitable formulas.
-          \Return{$\solveaux( %
-            \upd{\sigma}{R}{ %
-              \solveitp(\prectr{ \mathcal{R} }( %
-              \mathcal{S}, \sigma, R), %
-              \postctr{ \mathcal{R} }(\mathcal{S}, \sigma, R))})$} %
-          \label{line:ret-recurse} \;
-        } %
-      } %
-    \label{line:solve-aux-end} } %
-    % check if query is feasible:
-    \lIf{$\issat(\vc{ \mathcal{R} }(\mathcal{S}))$}{ % 
-      \Return{$\nosoln$} \label{line:check-sat} }
-    % return result of calling aux on base case:
-    \Return{$\solveaux(\emptyset)$} \label{line:aux-empty}
-    \; }
+    $\sigma \assign \emptyset$ \\
+    $\cc{Preds} \assign \topSort(\predof{S})$ \\
+    \For {$\cc{P} \in \cc{Preds}$}
+    {$\cc{interpolant} \assign \solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$\\
+     \Switch{\cc{interpolant}}{
+       \lCase{\cc{SAT}:}{\Return{$\nosoln$}}
+       \lCase{$\cc{I}$:}{$\sigma$ [$\cc{P} \to \cc{I}$]}
+     }
+    }
+    \Return{$\sigma$}
+    }
   %
-  \caption{$\solvecdd{ \mathcal{R} }$: for $\mathcal{R} \in \relpreds$
-    and CDD system $\mathcal{S} \in \chcs{ \mathcal{R} }$, returns a
-    solution to $\mathcal{S}$ or the value $\none$ to denote that
-    $\mathcal{S}$ has no solution.}
+  \caption{$\solvecdd$: for a CDD system $\cc{S}$, returns a
+    solution to $\cc{S}$ or the value $\none$ to denote that
+    $\cc{S}$ has no solution.}
   \label{alg:solve-cdd}
 \end{algorithm}
+
+% introduce solveCDD
+the procedure $\solvecdd$ (\autoref{alg:solve-cdd}) is a solver for CDD systems.
 %
-For $\mathcal{R} \in \relpreds$, the procedure $\solvecdd{ \mathcal{R}
-}$ (\autoref{alg:solve-cdd}) is a solver for CDD systems in $\chcs{
-  \mathcal{R} }$.
+$\solvecdd$, given a CDD system $\cc{S}$, it first do a topological sort on all
+uninterpreted predicates in $\cc{S}$ based on their dependency relations.
 %
-$\solvecdd{ \mathcal{R} }$, given $\mathcal{S} \in \chcs{ \mathcal{R}
-}$, defines a procedure $\solveaux$ that, given a partial solution
-$\sigma$ of $\mathcal{S}$, returns an extension of $\sigma$ that is a
-solution of $\mathcal{S}$
-(\autoref{line:aux-begin}---\autoref{line:solve-aux-end}).
+In the order of topological sort, for each uninterpreted predicates $\cc{P}$ 
+ $\solvecdd$ calles $\solveitp$ to compute the binary interpolation of 
+ encoded pre and post formulas with the current solution.
  %
-$\solvecdd{ \mathcal{R} }$ runs a procedure $\vc{ \mathcal{R} }$ that,
-given a CDD system $\mathcal{S} \in \chcs{ \mathcal{R} }$, returns a
-formula that is satisfiable if and only if $\mathcal{S}$ has no
-solution, which we refer to as a \emph{counterexample
-  characterization} of $\mathcal{S}$ (\autoref{line:check-sat};
+ If the conjunction of pre and post formulas is satisifiable, then $\solvecdd$
+ returns $\nosoln$ to indicate this CDD system $\cc{S}$ is not solvable.
  %
-the implementation of $\vc{ \mathcal{R} }$ is given below).
+ Otherwise, $\solvecdd$ binds the uninterpreted predicate $\cc{P}$ to the
+ interpolation result $\cc{I}$ as part of the soltuion.
  %
-$\solvecdd{ \mathcal{R} }$ runs $\issat$ (\autoref{sec:itps}) on $\vc{
-  \mathcal{R} }(\mathcal{S})$;
+ The encoding procedure $\prectr$ and $\postctr$ are defined below.
  %
-if $\issat$ returns that $\vc{ \mathcal{R} }(\mathcal{S})$ is
-satisfiable, then $\solvecdd{ \mathcal{R} }$ returns that
-$\mathcal{S}$ has no solution (\autoref{line:check-sat}).
-% how the aux function is used:
-Otherwise, $\solvecdd{ \mathcal{R} }$ runs $\solveaux$ on the empty
-partial solution and returns the result (\autoref{line:aux-empty}).
+ $\solvecdd$ returns the soltuion $\sigma$.
 
-% walk through the aux function:
-$\solveaux$, given partial solution $\sigma$, constructs the set
-$\mathcal{R}'$ of each relational predicate $R$ such that $R$ is not
-in the domain of $\sigma$ and all dependents of $R$ in $\mathcal{S}$
-are in the domain of $\sigma$ (\autoref{line:cons-nodeps}).
-%
-If $\mathcal{R}'$ is empty, then $\solveaux$ returns $\sigma$ as a
-solution of $\mathcal{S}$ (\autoref{line:ret-soln}).
-%
-Otherwise, if there is some $R \in \mathcal{R}'$
-(\autoref{line:case-frontier}), then $\solveaux$ extends $\sigma$ to
-map $R$ to an interpolant of the formulas $\prectr{ \mathcal{R}
-}(\mathcal{S}, \sigma, R)$ and $\postctr{ \mathcal{R} }(\mathcal{S},
-\sigma, R)$, defined below.
-%
-The interpolant of the two formulas is synthesized by running the
-procedure $\solveitp$ (\autoref{sec:itps}).
-%
-$\solveaux$ recurses on the extended partial solution and returns the
-result (\autoref{line:ret-recurse}).
+\QZ{rework example}
 %
@@
 
-\subsubsection{Constructing the Counterexample Characterization of a
-  CDD}
+\subsubsection{Constructing constrains of uninterpreted 
+predicate}
 %
-For $\mathcal{S'} \in \chcs{ \mathcal{R} }$, the formula $\vc{
-  \mathcal{R} }(\mathcal{S'})$ is a compact representation of the
-counterexample characterization of $\mathcal{S'}$, defined as follows.
+For an uninterpreted predicate $\cc{P} \in \predof{S}$ and $\cc{S}$ is an CDD system  
+,the formula $\ctrof{Q}$ is a compact representation of constrains of $\cc{P}$, defined as follows.
 % introduce the isused predicates:
-For each $R \in \mathcal{R}$, let there be a Boolean predicate
-$\used{R}$.
+For each $\cc{Q} \in \predof{S}$, there is a corresponding boolean variable
+$b_Q$.
 % for each relational predicate, define constraint:
-The constraint of $R$, denoted $\ctrof{R} \in \formulas$, constrains
-that if $R$ occurs in a derivation of $\mathcal{S'}$, then there is
-some $\mathcal{C} \in \mathcal{S'}$ in which $R$ is the head such that
-\textbf{(1)} each $R' \in \mathcal{R}$ in an application in the body
-of $\mathcal{C}$ is used and \textbf{(2)} the values assigned to each
-application of $R'$ are equal to the values assigned to the head
-variables of $R'$ in $\mathcal{S'}$.
+The constraint of $P$, denoted $\ctrof{P}$, constrains
+that if $P$ occurs in a derivation of $S$, then there is
+some clauses $C \in S$ in which $P$ is the head of $\cc{C}$
+\textbf{(1)} each $Q \in \bodyof{C}$ is used and 
+\textbf{(2)} the constrained of the clause $\consof{C}$ is encoded.
 %
-I.e., $\ctrof{R}$ is
-\begin{align*}
-% for each clause with R as a head:
-\biglor_{ \substack{ \mathcal{B} \in \bodies{ \mathcal{R} }, %
-    R \in \mathcal{R}, \\
-  (\mathcal{B}, (R, \varsof{ \mathcal{S'} }(R))) \in
-  \mathcal{S'} } } %
-\left( \ctrof{ \mathcal{B} } \land %
-\bigland_{ \substack{R' \in \mathcal{R}, Y \in X^{*}, \\
-    (R', Y) \in \appsof{ \mathcal{B} } } } %
-\used{R'} \land \varsof{ \mathcal{S'} }(R') = Y \right)
-\end{align*}
+I.e., $\ctrof{P}$ is
+\[\lnot b_Q \land 
+\left(
+\biglor_{ (\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
+\left( \consof{C_i} \land %
+\bigland_{ \cc{Q} \in \bodyof{C_i}} %
+b_Q \right) \right)
+\]
 %
@@
 
+\QZ{rework example}
 \begin{ex}
@@
 %
-For partial solution $\sigma : \mathcal{R} \to \formulas$ of
-$\mathcal{S}$, and $R \in \mathcal{R}$ for which each dependency is in
-the domain of $\sigma$, the pre-constraint $\prectr{ \mathcal{R}
-}(\mathcal{S}, \sigma, R)$ is a formula that is entailed by the
-interpretation of $R$ in each solution of $\mathcal{S}$ that extends
-$\sigma$.
+For a CDD system $\cc{S}$, given a uninterpreted predicate $\cc{P} \in \predof{\cc{S}}$
+, because $\solvecdd$ solve each uninterpreted predicate in a toplogical order, 
+it has a partial solution $\sigma : \cc{Q} \to \formula$ for all uninterpreted predicates in
+$\depsof{P}$.
 %
-In particular, $\prectr{ \mathcal{R} }(\mathcal{S}, \sigma, R)$ is:
-\[
-\biglor_{ \substack{ \mathcal{B} \in \bodies{ \mathcal{R} } \\
-  (\mathcal{B}, (R, \varsof{ \mathcal{S} }(R) ) ) \in \mathcal{S} } }
-\left( \ctrof{ \mathcal{ B } } \land %
-\bigland_{ A \in \appsof{ \mathcal{B} } } %
-\sigma( \relof{A} )[ \argsof{ A } ] \right)
+the pre-constraint $\prectr(\cc{P},\sigma)$ is a formula that the conjunction of the constrains
+of $\cc{P}$ and solutions of all dependencies of $P$ with their boolean varialbe.
+%
+In particular, $\prectr(\cc{P},\sigma)$ is:
+\[\ctrof{P} \land
+\left(
+\bigland_{\cc{Q} \in \depsof{P}}
+\left(b_Q \land \sigma[ \cc{Q} ]  \right)
+\right)
 \]
 %
+\QZ{rework example}
 \begin{ex}
@@
 % give the general procedure:
-$\shara{ \mathcal{R} }$, given $\mathcal{S} \in \chcs{ \mathcal{R} }$,
-constructs a CDD system $\mathcal{S'}$ equivalent to $\mathcal{S}$.
+$\shara$, given an recursion-free CHC system $S$,
+constructs a CDD system $S'$ equivalent to $S$.
 %
-\sys then directly solves $\mathcal{S'}$ and from its solution,
-constructs a solution of $\mathcal{S}$.
+\sys then directly solves $S'$ and from its solution,
+constructs a solution of $S$.
 % define an expansion of a CHC system:
-For $\mathcal{R}, \mathcal{R}' \in \relpreds$, $\mathcal{S} \in \chcs{
-  \mathcal{R} }$ and $\mathcal{S}' \in \chcs{ \mathcal{R}' }$, if
-there is a homomorphism from $\mathcal{R}'$ to $\mathcal{R}$ that
-preserves the relationship between the clauses of $\mathcal{S}'$ in
-the clauses of $\mathcal{S}$, then $\mathcal{S}'$ is an expansion of
-$\mathcal{S}$ (all definitions in this section will be over fixed
-$\mathcal{R}$, $\mathcal{R}'$, $\mathcal{S}$, and $\mathcal{S}'$).
+For two given recursion-free CHC system $S$ and $S'$, if
+there is a homomorphism from $\predof{S'}$ to $\predof{S}$ that
+preserves the relationship between the clauses of $S'$ in
+the clauses of $S$, then $S'$ is an expansion of
+$S$ (all definitions in this section will be over fixed
+$S$, and $S'$).
 % formal defn:
@@
   \label{defn:expansion}
-  Let $f : \mathcal{R}' \to \mathcal{R}$ be such that
+  Let $f : \predof{S'} \to \predof{S}$ be such that
   % preserves arity:
-  \textbf{(1)} for all $R' \in \mathcal{R}'$, $\arityof{f(R')} =
-  \arityof{R'}$;
+  \textbf{(1)} for all $P' \in \predof{S'}$, $P'$ has the
+  same vector of parameters as $f(P')$;
   % preserves clause structure:
-  \textbf{(2)} for all $H \in \apps{\mathcal{R}}$, $\mathcal{B}
-  \subseteq \apps{ \mathcal{R} }$, and $\varphi \in \formulas$ such
-  that $((\mathcal{B}, \varphi), H) \in \mathcal{S}'$, clause
+  \textbf{(2)} for all clause $C' \in S'$, the new clause $C$
+  by substituting all predicates $P'$ by $f(P')$ is in $S$.
   %
-  \[ \left( \left( \bigunion_{ A \in \mathcal{B} } %
-        \{ (f(\relof{A}), \argsof{A}) \}, \varphi \right), %
-      ( f(\relof{H}), \argsof{H} ) \right)
-  \]
-  %
-  is in $\mathcal{S}$.
-  %
-  Then $f$ is a \emph{correspondence} from $\mathcal{S}'$ to
-  $\mathcal{S}$.
+  Then $f$ is a \emph{correspondence} from $S'$ to $S$.
 \end{defn}
 %
-If there is a correspondence from $\mathcal{S}'$ to $\mathcal{S}$,
-$\mathcal{S}'$ is an \emph{expansion} of $\mathcal{S}$, denoted
-$\mathcal{S} \expandsto \mathcal{S}'$.
+If there is a correspondence from $S'$ to $S$,
+$S'$ is an \emph{expansion} of $S$, denoted
+$S \expandsto S'$.
 %
-\begin{defn}
-  \label{defn:min-cdd-expansion}
-  If $\mathcal{S}'$ is CDD, %
-  $\mathcal{S} \expandsto \mathcal{S}'$, and %
-  there is no CDD system $\mathcal{S}''$ such that $\mathcal{S}
-  \expandsto \mathcal{S}'' \expandsto \mathcal{S}$ and $\mathcal{S}''
-  \not= \mathcal{S}'$, then $\mathcal{S}'$ is a \emph{minimal} CDD
-  expansion of $\mathcal{S}$.
-\end{defn}
 
@@
   % Inputs: a heap program and an error location.
-  \Input{For $\mathcal{R} \in \relpreds$, $\mathcal{S} \in \chcs{
-      \mathcal{R} }$.}
+  \Input{For a recursion-free CHC system $S$.}
   % Output: inductive invariants.
-  \Output{A solution to $\mathcal{S}$ or $\none$.}
+  \Output{A solution to $S$ or $\none$.}
   % expand procedure:
-  \myproc{$\sys[ \mathcal{R} ](\mathcal{S})$ %
+  \myproc{$\sys(S)$ %
     \label{line:shara-begin}}{ %
     % construct expansion of C:
-    $(\mathcal{S}', \eta) \assign %
-    \expand{\mathcal{R}}(\mathcal{S})$ \label{line:shara-expand} \; %
+    $(S', \eta) \assign %
+    \expand(S)$ \label{line:shara-expand} \; %
     % expand, solve, case split:
-    \Switch{ $\solvecdd{ \ctxs{ \mathcal{R} } }( \mathcal{S}' )$ %
+    \Switch{ $\solvecdd(S' )$ %
       \label{line:shara-case} }{
@@
       % case: result is a solution:
-      \lCase{ $\sigma' \in \interps{ \ctxs{ \mathcal{R} } }$: }{ %
+      \lCase{ $\sigma'$: }{ %
         \Return{ $\collapse{ \eta }{ \sigma' }$ } %
@@
   \caption{\sys: a solver for recursion-free CHCs, which uses
-    procedures $\expand{\mathcal{R}}$ (see \autoref{app:cons-cdd}) and
-    $\solvecdd{ \ctxs{ \mathcal{R} } }$ (see
+    procedures $\expand{}$ (see \autoref{app:cons-cdd}) and
+    $\solvecdd$ (see
     \autoref{sec:solve-cdd}). }
@@
 % walk through the sys algorithm:
-$\shara{ \mathcal{R} }$ (\autoref{alg:shara}), given $\mathcal{S} \in
-\chcs{\mathcal{R}}$ (\autoref{line:shara-begin}), returns a solution
-to $\mathcal{S}$ or the value $\none$ to denote that $\mathcal{S}$ is
+$\shara$ (\autoref{alg:shara}), given an recursion-free CHC system $S$ (\autoref{line:shara-begin}), 
+returns a solution
+to $S$ or the value $\none$ to denote that $S$ is
 unsolvable.
 %
-$\shara{ \mathcal{R} }$ first runs a procedure $\expand{ \mathcal{R}
-}$ on $\mathcal{S}$ to obtain a minimal CDD expansion $\mathcal{S}'
-\in \chcs{ \ctxs{ \mathcal{R} } }$ (\autoref{defn:min-cdd-expansion})
-of $\mathcal{S}$ (where $\ctxs{ \mathcal{R} }$ is a space of
-relational predicates defined by $\mathcal{R}$, defined in
-\autoref{app:cons-cdd}), and a correspondence $\eta : \ctxs{
-  \mathcal{R} } \to \mathcal{R}$ (\autoref{line:shara-expand}).
+$\shara$ first runs a procedure $\expand{}$ on $S$ to obtain a CDD expansion $S'$ (\autoref{defn:min-cdd-expansion})
+of $S$.
 %
-$\shara{ \mathcal{R} }$ then runs a procedure $\solvecdd{ \ctxs{
-    \mathcal{R} } }$ on $\mathcal{S}'$, which either returns the value
-$\none$ to denote that $\mathcal{S}'$ has no solution or a solution
+$\shara$ then runs a procedure $\solvecdd$ on $S'$, which either returns the value
+$\none$ to denote that $S'$ has no solution or a solution
 $\sigma'$ (\autoref{line:shara-case}).
 %
-If $\solvecdd{ \ctxs{ \mathcal{R} } }$ returns that $\mathcal{S}'$ has
-no solution, then \sys returns that $\mathcal{S}$ has no solution
+If $\solvecdd$ returns that $S'$ has
+no solution, then \sys returns that $S$ has no solution
 (\autoref{line:shara-ret-none}).
 %
-Otherwise, if $\solvecdd{ \ctxs{ \mathcal{R} } }$ returns a solution
-$\sigma'$, then $\sys$ returns, as a solution to $\mathcal{S}$,
+Otherwise, if $\solvecdd$ returns a solution
+$\sigma'$, then $\sys$ returns, as a solution to $S$,
 $\collapse{ \eta }{ \sigma' }$, defined below
@@
 % define collapse:
-For $\sigma \in \interps{ \ctxs{ \mathcal{R} } }$ and %
-correspondence $\eta$ from $\mathcal{S}'$ to $\mathcal{S}$,
-$\collapse{\eta}{\sigma} \in \interps{ \mathcal{R} }$ maps each $R \in
-\mathcal{R}$ to
-%
-$\bigland_{ \substack{R' \in \mathcal{R}' \\ \eta(R') = R} }
-\sigma(R')$.
+For $\sigma'$ and %
+correspondence $\eta$ from $P' \in \predof{S'}$ to $P \in \predof{S}$,
+$\collapse{\eta}{\sigma'}$ maps solution $\sigma'$ to
+$\sigma$[$P \to \bigland_{ \eta(P') = P}
+\sigma'(P')$].
 %
@@
   \label{thm:corr}
-  $\mathcal{S}$ is solvable if and only if $\shara{ \mathcal{R}
-  }(\mathcal{S}) \in \interps{ \mathcal{R} }$.
+  $S$ is solvable if and only if $\shara$ returns solution $\sigma$.
 \end{thm}

### 2018-04-06 paper/background.tex
editing approach, need to rework overview, and move some appendix into main paper

@@
 For a given constrained horn clause $C$, $\bodyof{C}$ denotes
-the vector of uninterpreted predicates in the body of $C$.
+the vector of uninterpreted predicates in the body of $C$, $\consof{C}$
+denotes the formula in the body.
 %
@@
 
-\subsubsection{Solutions}
-\label{sec:chc-solns}
 %
-A solution to a CHC system $\mathcal{S}$ is an interpretation of each
-relational predicate $R$ of arity $n$ as a formula over $n$ free
-variables such that for each clause $\mathcal{C} \in \mathcal{S}$, the
-conjunction of interpretations of all relational predicates in the
-body of $\mathcal{C}$ and the constraint of $\mathcal{C}$ entail the
-interpretation of the head of $\mathcal{C}$.
-%
-Let a map from each $R \in \relsof{ \mathcal{R} }$ to a
-$\mathcal{T}$-formula over an ordered vector of $\arityof{ \mathcal{R}
-}(R)$ free variables be an \emph{interpretation} of $\mathcal{R}$;
-%
-let the space of interpretations of $\mathcal{R}$ be denoted
-$\interps{ \mathcal{R} }$.
-%
-\begin{defn}
-  \label{defn:chc-soln}
-  For $\mathcal{B} \in \bodies{ \mathcal{R} }$ and $H \in \apps{
-    \mathcal{R} }$, %
-  let $\sigma \in \interps{ \mathcal{R} }$ be such that for each $R
-  \in \mathcal{R}$, $\arityof{\mathcal{R}}(R) = \degreeof{\sigma(R)}$
-  and %
-  %
-  \[ \elts{ \sigma(\relof{ A })[ \argsof{ A } ] }_{ A \in \appsof{
-      \mathcal{B} } }, %
-  \ctrof{ \mathcal{B} } \entails %
-  \sigma( \relof{H} )[ \argsof{ H } ]
-  \]
-  Then $\sigma$ is a \emph{solution} of $(\mathcal{B}, H)$.
-\end{defn}
-% define partial solution 
-For $\mathcal{S} \in \chcs{ \mathcal{R} }$, if
-% solution respects dependence order:
-\textbf{(a)} for each $R \in \domain(\sigma)$ (where $\domain(\sigma)$
-denotes the domain of $\sigma$) and $R' \in \mathcal{R}$ a dependence
-of $R$ in $\mathcal{S}$, it holds that $R' \in \domain(\sigma)$;
-%
-\textbf{(b)} for each $\mathcal{C} \in \mathcal{S}$ such that
-$\relof{ \headof{ \mathcal{C} } } \in \domain(\sigma)$, $\sigma$ is a
-solution of $\mathcal{C}$;
-%
-then $\sigma$ is a \emph{partial solution} of $\mathcal{S}$.
-%
-If, in addition, $\sigma(\queryof{ \mathcal{S} }) \entails \false$,
-then $\sigma$ is a \emph{solution} of $\mathcal{S}$.
-%
-The problem addressed in this paper is, given a recursion-free CHC
-system $\mathcal{S}$, to synthesize a solution for $\mathcal{S}$.
-
-% define grounding of clause body:
-Each partial solution $\sigma$ of $\mathcal{S}$ defines a simpler CHC
-system in which each relational predicate in the domain of $\sigma$ is
-replaced with its interpretation in $\sigma$.
-%
-For $\mathcal{A} \subseteq \apps{\mathcal{R}}$, $\varphi \in
-\formulas$, and $\sigma \in \interps{ \mathcal{R} }$, the clause body
-\begin{align*}
-  \left( \bigunion_{ \substack{ %
-      R \in \mathcal{R} \setminus \domain(\sigma), \\
-      Y \in X^{*}, 
-      (R, Y) \in \mathcal{A} } } \elts{(R, Y)}, 
-  \varphi \land %
-  \bigland_{ \substack{ R \in \domain(\sigma) \\ 
-      Y \in X^{*},
-      (R, Y) \in \mathcal{A} } } \sigma(R)[ Y ] \right)
-\end{align*}
-%
-is the \emph{grounding} of $(\mathcal{A}, \varphi)$ on $\sigma$,
-denoted $\sigma(\mathcal{A}, \varphi)$.
-
-% groundings of systems:
-For each $\mathcal{S} \in \chcs{ \mathcal{R} }$ and $\sigma \in
-\interps{ \mathcal{R} }$ a partial solution of $\mathcal{S}$, the
-grounding of $\mathcal{S}$ on $\sigma$ is a CHC system in which each
-clause is a grounding of a clause in $\mathcal{S}$ with a head
-relational predicate not in the domain of $\sigma$.
-%
-In particular, let $\mathcal{S}' \subseteq \chc{ \mathcal{R} \setminus
-  \domain(\sigma) }$ be such that for each $\mathcal{B} \in \bodies{
-  \mathcal{R} }$ and $H \in \apps{ \mathcal{R} }$ with $\relof{H}
-\notin \domain(\sigma)$, the clause $(\sigma(\mathcal{B}), H)$ is in
-$\mathcal{S}'$.
-%
-Then $\mathcal{S}'$ is the \emph{grounding} of $\mathcal{S}$ on
-$\sigma$, denoted $\sigma(\mathcal{ S })$.
+A solution to a CHC system $S$ is an interpretation of each
+uninterpreted predicate $\cc{Q} \in \predof{S}$ that each clause holds
+after substitute all predicates by its interpretation.
 

### 2018-04-07 src/Data/Grammar.hs
added interpolator for arbitrary nonterminals in a CDD grammar

@@ -1,8 +1,10 @@
-module Data.Grammar where
+module Data.Grammar
+  ( module Data.Grammar
+  , module Data.Rule
+  ) where
 
 import           Control.Lens hiding (Empty, elements, (:<))
 import           Control.Monad.State
-import           Control.Monad.Writer (Writer, runWriter, tell)
 import           Control.Applicative
 
 import           Data.Foldable (asum)
@@ -12,95 +14,27 @@ import qualified Data.Map as M
 import           Data.Set (Set)
 import qualified Data.Set as S
 import           Data.Semigroup
-import           Data.List (intercalate, nub)
+import           Data.List (nub)
 
-import           Test.QuickCheck hiding (NonEmpty)
+import           Data.Rule
+
+import           Test.QuickCheck
 
 import           Prelude hiding (seq, sum, product)
 
-import Debug.Trace
-
-newtype NT = NT { getNT :: Int }
-  deriving (Show, Read, Eq, Ord, Data, Real, Bounded, Enum, Integral, Num)
-
-instance Arbitrary NT where
-  arbitrary = do
-    topNT <- growingElements [0..3]
-    NT <$> elements (0:[1..topNT])
-
-data Rule a
-  = Null
-  | Eps
-  | Alt (Rule a) (Rule a)
-  | Seq (Rule a) (Rule a)
-  | Nonterminal NT
-  | Terminal a
-  deriving (Show, Read, Eq, Ord, Data, Functor, Foldable, Traversable)
-
-isNull :: Rule a -> Bool
-isNull Null = True
-isNull _ = False
-
-isEps :: Rule a -> Bool
-isEps Eps = True
-isEps _ = False
-
-seq :: Rule a -> Rule a -> Rule a
-seq Null _      = Null
-seq _ Null      = Null
-seq Eps x       = x
-seq x Eps       = x
-seq (Seq x y) z = Seq x (seq y z)
-seq x y         = Seq x y
-
-alt :: Rule a -> Rule a -> Rule a
-alt Eps Eps     = Eps
-alt Eps x       = alt x Eps
-alt Null x      = x
-alt x Null      = x
-alt (Alt x y) z = Alt x (alt y z)
-alt x y         = Alt x y
-
-instance Arbitrary a => Arbitrary (Rule a) where
-  arbitrary =
-    oneof [ pure Null
-          , pure Eps
-          , alt <$> arbitrary <*> arbitrary
[…]

### 2018-04-07 src/Data/Rule.hs

@@ -1,0 +1,145 @@
+module Data.Rule where
+
+import           Control.Lens hiding (Empty, elements, (:<))
+import           Control.Monad.State
+import           Control.Applicative
+
+import           Data.Data (Data)
+import           Data.Semigroup
+
+import           Test.QuickCheck hiding (NonEmpty)
+
+import           Prelude hiding (seq, sum, product)
+
+-- | A nonterminal exists to identify rules in the context free grammar.
+newtype NT = NT { getNT :: Int }
+  deriving (Show, Read, Eq, Ord, Data, Real, Bounded, Enum, Integral, Num)
+
+instance Arbitrary NT where
+  arbitrary = do
+    topNT <- growingElements [0..5]
+    NT <$> elements (0:[1..topNT])
+
+-- | The right hand side of a grammar production.
+data Rule a
+  -- The rule which recognizes nothing.
+  = Null
+  -- The rule which only recognizes an empty string.
+  | Eps
+  -- The rule which recognizes based on either subrule.
+  | Alt (Rule a) (Rule a)
+  -- The rule which recognizes based on the sequence of subrules.
+  | Seq (Rule a) (Rule a)
+  -- The rule which recognizes based on the right hand side for the given nonterminal.
+  | Nonterminal NT
+  -- The rule which recognizes the given symbol.
+  | Terminal a
+  deriving (Show, Read, Eq, Ord, Data, Functor, Foldable, Traversable)
+
+-- | Is the given rule exactly the rule `Null`?
+isNull :: Rule a -> Bool
+isNull Null = True
+isNull _ = False
+
+-- | Is the given rule exactly the rule `Eps`?
+isEps :: Rule a -> Bool
+isEps Eps = True
+isEps _ = False
+
+-- | `seq` is built over the `Seq` constructor. It respects rule sequence
+-- associativity and the fact that `Null` is a 0 element and `Eps` is a 1
+-- element.
+seq :: Rule a -> Rule a -> Rule a
+seq Null _      = Null
+seq _ Null      = Null
+seq Eps x       = x
+seq x Eps       = x
+seq (Seq x y) z = Seq x (seq y z)
+seq x y         = Seq x y
+
+-- | `alt` is built over the `Alt` constructor. It respects rule alternation
+-- associativity and the fact that `Null` is a 1 element and that more than one
+-- alternation of `Eps` is redundant.
[…]

### 2018-04-07 src/Solver/Interpolate.hs

@@ -1,0 +1,83 @@
+module Solver.Interpolate where
+
+import           Control.Lens
+import           Control.Monad.State
+import           Control.Monad.Reader
+
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+import qualified Data.Set as S
+import           Data.Grammar (Grammar, Rule, NT)
+import qualified Data.Grammar as G
+import           Data.Maybe (fromMaybe)
+
+import           Formula
+import qualified Formula.Z3 as Z3
+
+import Data.Text.Prettyprint.Doc
+
+topologicalInterpolation :: MonadIO m => Grammar Expr -> m (Either Model (Map NT Expr))
+topologicalInterpolation g =
+  let nts = G.topologicalOrder g
+  in runStateT (loop nts) M.empty >>= \case
+    (Left m, _) -> pure (Left m)
+    (_, m) -> pure (Right m)
+  where
+    loop :: MonadIO m => [NT] -> StateT (Map NT Expr) m (Either Model ())
+    loop [] = pure (Right ())
+    loop (nt:nts) = interpolateNT' nt g >>= \case
+      Left m -> pure (Left m)
+      _ -> loop nts
+
+interpolateNT' :: MonadIO m => NT -> Grammar Expr -> StateT (Map NT Expr) m (Either Model ())
+interpolateNT' target g = do
+  solns <- get
+  interpolateNT target solns g >>= \case
+    Left m -> pure (Left m)
+    Right phi -> modify (M.insert target phi) >> pure (Right ())
+
+-- | Interpolate the given nonterminal in the context of the directly solvable grammar.
+interpolateNT :: MonadIO m => NT -> Map NT Expr -> Grammar Expr -> m (Either Model Expr)
+interpolateNT target solns g =
+  let (phi1, phi2) = interpolationForms target solns g
+  in Z3.interpolate phi1 phi2
+
+-- | Find the two formulas on either side of the target nonterminal. The two
+-- formulas are found by partitioning the grammar at the location of the
+-- nonterminal.
+interpolationForms :: NT -> Map NT Expr -> Grammar Expr -> (Expr, Expr)
+interpolationForms target solns g =
+  let (reaching, reached) = G.partition [target] g
+  in ( grammarExpr target solns reaching
+     , grammarExpr target solns reached)
+
[…]

### 2018-04-07 src/Solver/Shara.hs

@@ -2,6 +2,7 @@ module Solver.Shara where
 
 import           Control.Lens
 import           Control.Monad.State
+import           Control.Monad.Reader
 import           Control.Applicative
 
 import           Data.Map (Map)
@@ -10,8 +11,10 @@ import           Data.Set (Set)
 import qualified Data.Set as S
 import           Data.Maybe (isJust, fromJust)
 import           Data.Semigroup
+import           Data.List.Split
 
-import           Formula.Expr
+import           Formula hiding (Chc)
+import qualified Formula.Z3 as Z3
 
 import           Data.Grammar (Grammar, Rule, NT)
 import qualified Data.Grammar as G
@@ -20,35 +23,37 @@ import           Data.InfGrammar (InfGrammar)
 import qualified Data.InfGrammar as IG
 
 import           Solver.Types
+import           Solver.Interpolate
+import           Solver.Chc
+import           Solver.TreeUnwind
 
 -- | Encode the CHC as a (possibly infinite) unrollable grammar such that each
 -- unrolling is a directly solvable CHC system.
-directlySolvable :: MonadState DirectState m => Chc -> m (InfGrammar m Expr)
-directlySolvable = undefined
+-- directlySolvable :: Expandable m => Chc -> m (InfGrammar m Expr)
+-- directlySolvable = treeUnwind
 
 -- | Apply the interpolator to the given directly solvable system and yield
 -- either a counterexample model or solutions for the nonterminals.
 solveDirect :: MonadIO m => Grammar Expr -> m (Either Model (Map NT Expr))
-solveDirect = undefined
-
--- | Interpolate the given nonterminal in the context of the directly solvable grammar.
-interpolateNT :: MonadIO m => NT -> Grammar Expr -> Map NT Expr -> m (Either Model (Map NT Expr))
-interpolateNT = undefined
+solveDirect = topologicalInterpolation
 
 -- | Solve the Relational Post Fixed-Point Problem as encoded by the given CHC
 -- system.
-solveChc :: (MonadState DirectState m, MonadIO m) => Chc -> m (Either Model (Map NT Expr))
-solveChc chc = solve =<< directlySolvable chc
+-- solveChc :: (Expandable m, MonadIO m) => Chc -> m (Either Model (Map NT Expr))
+solveChc chc = solve =<< treeUnwind chc
 
[…]

### 2018-04-07 src/Solver/TreeUnwind.hs

@@ -5,6 +5,7 @@ module Solver.TreeUnwind where
 import           Control.Lens hiding (mapping)
 import           Control.Monad.State
 import           Control.Monad.Reader
+import           Control.Arrow (second)
 import           Control.Applicative
 
 import           Data.Data (Data)
@@ -24,6 +25,7 @@ import           Formula hiding (Rule, Chc)
 
 import           Solver.Types hiding (DirectState(..))
 import           Solver.Chc
+import           Solver.Interpolate
 
 import Data.Text.Prettyprint.Doc
 
@@ -69,6 +71,10 @@ emptyCtxt = UnwindCtxt
 instance Monad m => Expandable (StateT UnwindState m) where
   getProof = G.mkGrammar <$> use proofStart <*> use proofRules
   getClones = use clones
+
+instance Expandable m => Expandable (ReaderT UnwindCtxt m) where
+  getProof = lift getProof
+  getClones = lift getClones
 
 -- | Unwind the given CHC system to a tree.
 -- In addition, generate a proof structure which indicates what an inductive
@@ -203,7 +209,7 @@ subst' vs = subst (foldr (uncurry M.insert) M.empty vs)
 test :: Chc
 test =
   G.mkGrammar
-  (  G.Terminal (Fact [expr|x = 41|])
+  (  G.Terminal (Fact [expr|x = 5|])
   <> G.Terminal (Apply x)
   <> G.Nonterminal 0
   ) $
@@ -225,6 +231,12 @@ experiment = runReaderT (evalStateT ( do
   g' <- IG.unroll S.empty g
   liftIO $ putStrLn "\n\n"
   liftIO (putStr $ G.draw $ (show . pretty) <$> IG.finite g')
+
+  -- itp <- topologicalInterpolation (IG.finite g')
+  -- case itp of
+  --   Left m -> liftIO $ print $ pretty (M.toList m)
+  --   Right m -> liftIO $ print (map (second pretty) (M.toList m))
+
   g'' <- IG.unroll S.empty g'
   liftIO $ putStrLn "\n\n"
   liftIO (putStr $ G.draw $ (show . pretty) <$> IG.finite g'')
@@ -233,8 +245,18 @@ experiment = runReaderT (evalStateT ( do
   liftIO $ putStrLn "\n\n"
   liftIO (putStr $ G.draw $ (show . pretty) <$> IG.finite g''')
 
[…]

### 2018-04-07 test/GrammarTest.hs

@@ -1,4 +1,3 @@
-import           Control.Lens
 import           Control.Monad
 import           Control.Applicative
 import           Data.Grammar as G

### 2018-04-09 paper/introduction.tex
basically done, need a little more work in overview and its graph

@@
 %
-In particular, the class consists of all CHC systems such that, for
-each clause $\mathcal{C}$, the dependences of all distinct relational
-predicates in $\mathcal{C}$ are disjoint;
+In particular, the class consists of all recursion-free CHC systems such that, 
+for each clause $C$, the dependences of all distinct relational
+predicates in the body of $C$ are disjoint;
 %
@@
 %
-Given recursion-free system $\mathcal{S}$, \sys reduces the problem
-of solving $\mathcal{S}$ to solving an equivalent CDD system
-$\mathcal{S}'$.
+Given recursion-free system $S$, \sys reduces the problem
+of solving $S$ to solving an equivalent CDD system
+$S'$.
 %
-In the worst case, the size of $\mathcal{S}'$ may be exponential in the size of
-$\mathcal{S}$.
+In the worst case, the size of $S'$ may be exponential in the size of
+$S$.
 %
-In our experience, the size of $\mathcal{S}'$ is usually close enough
-to the size of $\mathcal{S}$ for \sys to perform significantly better
+In our experience, the size of $S'$ is usually close enough
+to the size of $S$ for \sys to perform significantly better
 than the best known CHC solvers.
@@
 %
-The procedure implemented in \sys to construct $\mathcal{S}'$ is a
+The procedure implemented in \sys to construct $S'$ is a
 generalization of existing procedures for generating compact

### 2018-04-09 paper/overview.tex
basically done, need a little more work in overview and its graph

@@
 
-%first section
+%first introduce the question
 In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
 $\mcchc$, that formulates verifying the safety of the program
-\cc{multiCall} (\autoref{fig:multicall-code}), and illustrate that
+\cc{multiCall} (\autoref{fig:multicall-code}).
+%
+In \autoref{sec:solve-ex}, we show that $\mcchc$ is a
+\emph{Clause-Dependent Disjoint (CDD)} system and how \sys can solve it by 
+encoding it into binary interpolants.
+%
+In \autoref{sec:not-in}, we illustrate that
 $\mcchc$ is not in the directly solvable classes introduced in
@@
 %
-In \autoref{sec:solve-ex}, we show that $\mcchc$ is a
-\emph{Clause-Dependent Disjoint (CDD)} system and how \sys solves it.
 
@@
 
+\QZ{remake the graph}
 % include code of running example and some figure:
@@
 Verifying that \cc{multiCall} satisfies such a property can be reduced
-to solving a recursion-free CHC system over a set of relational
-predicates $\mcpreds$ that contains each control location and
+to solving a recursion-free CHC system over a set of uninterpreted predicates
+$\mcpreds$ that each represents a control location and
 procedure name in \cc{multiCall}.
@@
 %
+\QZ{left allign}
 \begin{gather}
   % semantic constraint of dbl:
-  \cc{dbl}(\cc{n}, 2 \cdot \cc{n}) \qquad
+  \cc{dbl}(\cc{n}, \cc{res}) \gets \cc{res} = 2 * \cc{n} \\
   % semantic constraint of line 3:
-  \cc{L3}(\cc{n}) \implies 
-  % semantic constraint of line 4:
-  \cc{L4}(\cc{n}, 0) \notag \\
+   \cc{L4}(\cc{n}, \cc{res}) \gets  \cc{L3}(\cc{n}) \land \cc{res} = 0 \\
   % semantic constraint of line 5 (then branch)
-  \cc{L4}(\cc{n}, \cc{res}), \cc{dbl}(\cc{n}, \cc{res}'), %
-  \cc{n} \geq 0 \implies %
-  \cc{L9}(\cc{n}, \cc{res}') \label{eqn:then} \\
+   \cc{L9}(\cc{n}, \cc{res'}) \gets \cc{L4}(\cc{n}, \cc{res}) \land \cc{dbl}(\cc{n}, \cc{res}') \land
+   \cc{n} \geq 0 \\
   % semantic constraint of line 6 (else branch)
-  \cc{L4}(\cc{n}, \cc{res}), \cc{dbl}(\cc{n}, \cc{res}'), %
-  \cc{n} < 0 \implies %
-  \cc{L9}(\cc{n}, -\cc{res}') \label{eqn:else} \\
+   \cc{L9}(\cc{n}, \cc{res_1}) \gets \cc{L4}(\cc{n}, \cc{res}) \land \cc{dbl}(\cc{n}, \cc{res_0}) \land
+   \cc{n} < 0  \land \cc{res_1} = -\cc{res_0} \\
   % semantic constraint of main procedure:
-  \cc{L9}(\cc{n}, \cc{res}) \implies %
-  \cc{main}(\cc{n}, \cc{res}) \qquad
-  % errors:
-  \cc{main}(\cc{n}, \cc{res}), \cc{res} < 0 \implies \mathsf{Err}()
-  \notag
+   False \gets \cc{L9}(\cc{n}, \cc{res}) \land \cc{res} < 0 
 \end{gather}
@@
 
+% solving the running example
+\subsection{\cc{multiCall} as a Clause-Dependent Disjoint system}
+\label{sec:solve-ex}
+% state it is CDD
+The recursion-free CHC system \cc{multiCall} is a \emph{Clause-Dependent-Disjoint (CDD)}) system.
+% state the defn of CDD:
+\emph{Clause-Dependent-Disjoint (CDD)}) system is an recursion-free CHC system such that
+for any puninterpreted predictes in the body of the same clause, their transitive dependencies
+are disjoint(i.e,they do not have common predicates).
+%
+The formal definition of CDD and its key properties are given in
+\autoref{sec:CDD-defn}.
+% give a example
+In particular, for clauses $(3)$ and $(5)$, uninterpreted predicates \cc{L4} and \cc{dbl} shows in
+the body of same clause.
+%
+The transitive dependency of \cc{L4} is \{\cc{L3}\}, and the transitive dependency of \cc{dbl}
+is $\varnothing$.
+%
+Their transitive dependencies are disjoint, because \{\cc{L3}\} $\cap \varnothing = \varnothing$.
+%
+The formal definition of transitive dependency is given in\QZ{reference}.
+%
+And the other clauses in \cc{multiCall} only has one uninterpreted predicates in the body, so
+\cc{multiCall} is a CDD system.
+%
+The fact that $\mcchc$ is CDD corresponds to the fact that there no statements executed more than
+once in one execution path.
+%
+
+% restate defn of CDD systems:
+\sys solves $\mcchc$ directly by issuing a binary 
+interpolation queriey for each uninterpreted predicates in $\mcchc$ in a topological order.
+% walk through how a CDD system is solved directly:
+A solution of an arbitrary CDD system $\mathcal{S}$ can be constructed
+from the results of a series of interpolation queries, each one issued
+to find an interpretation of a relational predicate of $\mathcal{S}$
+in the topological order of dependences between relational predicates.
+%
+For relational predicate $R$ in $\mathcal{S}$, the interpretation of
+$R$ is the interpolant of two formulas: %
+\textbf{(1)} a Boolean combination of the interpretations found for
+each dependence of $R$ in $\mathcal{S}$ and constraints in
+$\mathcal{S}$ and %
+\textbf{(2)} a Boolean combination of constraints that occur in
+$\mathcal{S}$.
+
+% talk performance:
+The performance of the solver is thus largely determined by the
+ability of the interpolating theorem prover that it uses \textbf{(1)}
+to generate interpolants quickly and %
+\textbf{(2)} to generate interpolants of size close to that of their
+input formulas.
+%
+In the worst case, all interpolating theorem provers given in previous
+work may use exponential time to generate interpolants of size
+exponential in the size of their inputs.
+%
+However, when given formulas that model practical verification
+problems, modern theorem provers generate interpolants efficiently,
+which have size close to the size of the given formulas.
+
+% walk through multiCall
+E.g., when solving $\mcchc$, \sys generates interpretations of
+relational predicates in the order \cc{L3}, \cc{L4}, \cc{dbl},
+\cc{L9}, and \cc{main}.
+%
+It generates an interpretation of \cc{L9} in particular as the
+interpolant of \textbf{(1)} a \emph{pre}-formula, constructed as a
+Boolean combination of interpretations generated for \cc{L4}, \cc{dbl}
+and the constraints in \autoref{eqn:then} and \autoref{eqn:else} and %
+\textbf{(2)} a \emph{post}-formula constructed from constraints in
+$\mcchc$ that model the steps of execution of \cc{multiCall} following
+\cc{9}.
+%
+A procedure for solving a CDD system directly is given in
+\autoref{sec:solve-cdd}.
+
+% talk about expansion:
+The class of CDD systems strictly contains all directly-solvable
+classes introduced in previous work, but does not contain all
+recursion-free systems.
+%
+E.g., a recursion-free system that models \cc{multiCall} mutated to
+call \cc{dbl} in sequence, instead of in alternate branches, would not
+be CDD.
+%
+\sys, given an arbitrary recursion-free CHC system $\mathcal{S}$,
+solves it by generating an equivalent CDD system $\mathcal{S}'$ and
+solving $\mathcal{S}'$ directly.
+%
+In general, $\mathcal{S}'$ may have size exponential in the size of
+$\mathcal{S}$.
+%
+In practice, a suitable equivalent system can be generated using
+heuristics analogous to those used to generate compact verification
+conditions of hierarchical programs~\cite{flanagan01,lal-qadeer15}.
+%
+A general procedure for constructing a CDD expansion of a given CHC
+system is given in \autoref{app:cons-cdd}.
+
+\subsection{\cc{multiCall} is not in other system}
+\label{sec:not-in}
 % walk through the zoo of CHC systems:
@@
 
-% solving the running example
-\subsection{\cc{multiCall} as a Clause-Dependent Disjoint system}
-\label{sec:solve-ex}
-% restate defn of CDD systems:
-\sys solves $\mcchc$ directly by issuing a series of
-interpolation queries, each of which generates an interpretation for a
-relational predicate in $\mcchc$.
-%
-The key observation behind the design of \sys is that for CHC system
-$\mathcal{S}$, if the dependences of each relational predicate in
-each clause of $\mathcal{S}$ are disjoint (in which we case we say
-that $\mathcal{S}$ is \emph{Clause-Dependent-Disjoint (CDD)}), then
-$\mathcal{S}$ is directly solvable.
-%
-The formal definition of CDD and its key properties are given in
-\autoref{sec:CDD-defn}.
-
-% example system is CDD and can be easily solved:
-E.g., $\mcchc$ is CDD because for each clause $\mathcal{C}$ of
-$\mcchc$, the dependences of all distinct relational predicates in the
-body of $\mathcal{C}$ are disjoint.
-%
-In particular, for clauses \autoref{eqn:then} and \autoref{eqn:else},
-the dependences of relational predicates \cc{L4} and \cc{dbl} are
-disjoint.
-%
-The dependences of \cc{L4} is \{\cc{L3}\}, and the dependneces of \cc{dbl}
-is $\varnothing$.
-%
-\autoref{fig:ex-deps} contains a hypergraph $H_{\cc{MC}}$ that depicts
-the dependences induced by the clauses of $\mcchc$.
-%
-The fact that $\mcchc$ is CDD corresponds to the fact that no node in
-$H_{\cc{MC}}$ reaches distinct sources of a common hyperedge.
-
-% walk through how a CDD system is solved directly:
-A solution of an arbitrary CDD system $\mathcal{S}$ can be constructed
-from the results of a series of interpolation queries, each one issued
-to find an interpretation of a relational predicate of $\mathcal{S}$
-in the topological order of dependences between relational predicates.
-%
-For relational predicate $R$ in $\mathcal{S}$, the interpretation of
-$R$ is the interpolant of two formulas: %
-\textbf{(1)} a Boolean combination of the interpretations found for
-each dependence of $R$ in $\mathcal{S}$ and constraints in
-$\mathcal{S}$ and %
-\textbf{(2)} a Boolean combination of constraints that occur in
-$\mathcal{S}$.
-
-% talk performance:
-The performance of the solver is thus largely determined by the
-ability of the interpolating theorem prover that it uses \textbf{(1)}
-to generate interpolants quickly and %
-\textbf{(2)} to generate interpolants of size close to that of their
-input formulas.
-%
-In the worst case, all interpolating theorem provers given in previous
-work may use exponential time to generate interpolants of size
-exponential in the size of their inputs.
-%
-However, when given formulas that model practical verification
-problems, modern theorem provers generate interpolants efficiently,
-which have size close to the size of the given formulas.
-
-% walk through multiCall
-E.g., when solving $\mcchc$, \sys generates interpretations of
-relational predicates in the order \cc{L3}, \cc{L4}, \cc{dbl},
-\cc{L9}, and \cc{main}.
-%
-It generates an interpretation of \cc{L9} in particular as the
-interpolant of \textbf{(1)} a \emph{pre}-formula, constructed as a
-Boolean combination of interpretations generated for \cc{L4}, \cc{dbl}
-and the constraints in \autoref{eqn:then} and \autoref{eqn:else} and %
-\textbf{(2)} a \emph{post}-formula constructed from constraints in
-$\mcchc$ that model the steps of execution of \cc{multiCall} following
-\cc{9}.
-%
-A procedure for solving a CDD system directly is given in
-\autoref{sec:solve-cdd}.
-
-% talk about expansion:
-The class of CDD systems strictly contains all directly-solvable
-classes introduced in previous work, but does not contain all
-recursion-free systems.
-%
-E.g., a recursion-free system that models \cc{multiCall} mutated to
-call \cc{dbl} in sequence, instead of in alternate branches, would not
-be CDD.
-%
-\sys, given an arbitrary recursion-free CHC system $\mathcal{S}$,
-solves it by generating an equivalent CDD system $\mathcal{S}'$ and
-solving $\mathcal{S}'$ directly.
-%
-In general, $\mathcal{S}'$ may have size exponential in the size of
-$\mathcal{S}$.
-%
-In practice, a suitable equivalent system can be generated using
-heuristics analogous to those used to generate compact verification
-conditions of hierarchical programs~\cite{flanagan01,lal-qadeer15}.
-%
-A general procedure for constructing a CDD expansion of a given CHC
-system is given in \autoref{app:cons-cdd}.
-
 %%% Local Variables: 

### 2018-04-09 app/Main.hs
attempt to run concurrent

@@ -1,0 +1,99 @@
+{-# LANGUAGE QuasiQuotes #-}
+module Main where
+
+import           Control.Monad.State
+import           Control.Monad.Reader
+import           Control.Applicative
+import           Control.Arrow (second)
+
+import           Data.Grammar (Grammar, NT, Rule)
+import qualified Data.Grammar as G
+import           Data.InfGrammar (InfGrammar, InfRule)
+import qualified Data.InfGrammar as IG
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+import qualified Data.Set as S
+
+import           Formula hiding (Rule, Chc)
+
+import           Solver.Shara
+import           Solver.Chc
+import           Solver.Types hiding (DirectState(..))
+import           Solver.Interpolate
+import           Solver.TreeUnwind
+
+import           System.Environment (getArgs)
+
+import           Data.Text.Prettyprint.Doc
+
+main :: IO ()
+main = do
+  as <- getArgs
+  let defStrat = LicketySplit ConcurrentInterpolation
+  let cfg = case as of
+        [] -> defStrat
+        (a:_) -> case a of
+          "topo" -> Topological
+          "seq" -> LicketySplit SequentialInterpolation
+          _ -> defStrat
+  print cfg
+  solve cfg test >>= \case
+    Left e -> print e
+    Right m -> print $ map (second pretty) $ M.toList m
+
+test :: Chc
+test =
+  G.mkGrammar
+  (  G.Terminal (Fact [expr|x = 41|])
+  <> G.Terminal (Apply x)
+  <> G.Nonterminal 0
+  ) $
+  M.fromList
+    [ (0, scope x (G.Terminal (Fact [expr|x = 0|]))
+      <|> scope x' (scope x $ G.Terminal (Fact [expr|x' = x+2|])
+                               <> G.Terminal (Apply x)
+                               <> G.Nonterminal 0))
+    ]
+  where
+    x = Var "x" Int
+    x' = Var "x'" Int
+
+
+experiment :: IO ()
+experiment = runReaderT (evalStateT ( do
+  g <- treeUnwind test
+  liftIO (putStr $ G.draw $ (show . pretty) <$> IG.finite g)
+
+  g' <- IG.unroll S.empty g
[…]

### 2018-04-09 src/Data/Grammar.hs

@@ -196,8 +196,8 @@ partition nts g = (before, after)
   where
     before = evalState (go (start g)) (S.fromList nts)
     after =
-      evalState (asum <$> mapM (go . (`ruleFor` g)) nts)
-        (nonterminals before S.\\ S.fromList nts)
+      evalState (asum <$> mapM (\nt -> fmap (abstract nt) $ go $ ruleFor nt g) nts)
+        (nonterminals before)
     go = \case
       Null -> pure empty
       Eps -> pure mempty

### 2018-04-09 src/Data/InfGrammar.hs

@@ -54,21 +54,3 @@ unroll noUnroll (G.Grammar st rs) =
         G.Grammar st' rs' <- lift ac
         modify (M.unionWith (<|>) rs')
         pure st'
-
-fin :: a -> InfRule m a
-fin = G.Terminal . Finite
-
-inf :: m (InfGrammar m a) -> InfRule m a
-inf = G.Terminal . Infinite
-
-testInf :: InfGrammar IO Int
-testInf =
-  G.mkGrammar
-    (fin 1 <> G.Nonterminal 1)
-    (M.fromList
-      [(1, inf readRs <|> fin 2) ])
-  where
-    readRs :: IO (InfGrammar IO Int)
-    readRs = do
-      n <- read <$> getLine
-      pure (pure (Finite n))

### 2018-04-09 src/Formula/Expr.hs

@@ -147,6 +147,9 @@ appMany = foldl (:@)
 mkImpl :: Expr -> Expr -> Expr
 mkImpl = app2 Impl
 
+mkIff :: Expr -> Expr -> Expr
+mkIff = app2 Iff
+
 mkAdd :: Type -> Expr -> Expr -> Expr
 mkAdd _ (LInt a) (LInt b) = LInt (a + b)
 mkAdd t e e' = app2 (Add t) e e'

### 2018-04-09 src/Solver/Chc.hs

@@ -1,10 +1,22 @@
+{-# LANGUAGE TemplateHaskell #-}
 module Solver.Chc where
 
-import Data.Grammar
-import Data.Semigroup
 
-import Formula.Expr
-import Formula.Var
+import           Control.Lens
+import           Control.Monad.State
+
+import           Data.Data (Data)
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+import           Data.Sequence (Seq)
+import qualified Data.Sequence as Seq
+import           Data.Grammar (Grammar, NT, Rule)
+import qualified Data.Grammar as G
+import           Data.Semigroup
+
+import           Formula.Expr
+import           Formula.Var
 
 -- | A fragment of a constrained Horn Clause rule.
 data Fragment
@@ -17,4 +29,72 @@ data Fragment
 type Chc = Grammar Fragment
 
 scope :: Var -> Rule Fragment -> Rule Fragment
-scope v r = Terminal (Capture v) <> r <> Terminal Uncapture
+scope v r = G.Terminal (Capture v) <> r <> G.Terminal Uncapture
+
+type ProofStructure = Grammar Proof
+
+data Proof
+  = Continue
+  | Entails (Map NT (Set NT)) NT (Rule Fragment)
+  deriving (Show, Read, Eq, Ord)
+
+class Monad m => Expandable m where
+  getProof :: m ProofStructure
+  getClones :: m (Map NT (Set NT))
+
+data ChcState = ChcState
+  { _varMapping :: [(Var, Var)]
+  , _varQueue :: Seq Var
+  , _varCounter :: Int
+  } deriving (Show, Read, Eq, Ord)
+makeLenses ''ChcState
+
+emptyChcState :: ChcState
+emptyChcState = ChcState
+  { _varMapping = []
+  , _varQueue = Seq.empty
+  , _varCounter = 0
+  }
+
+resolve :: Monad m => Fragment -> StateT ChcState m Expr
+resolve = \case
+  Apply v -> do
+    -- Applying a variable adds the variable to the vocabulary and has a
+    -- trivial proof.
+    m <- use varMapping
+    varQueue %= (Seq.<|) (subst' m v)
+    pure (LBool True)
+  Capture v -> do
+    -- Capturing a variable either yields fresh variables or pops the
+    -- vocabulary. The proof is trivial.
+    Seq.viewr <$> use varQueue >>= \case
+      Seq.EmptyR -> do
[…]

### 2018-04-09 src/Solver/Cut.hs

@@ -1,0 +1,101 @@
+module Solver.Cut where
+
+import           Control.Lens hiding (below)
+import           Control.Monad.State
+
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+import qualified Data.Set as S
+import           Data.List (partition)
+import           Data.Ratio
+
+import           Solver.Graph
+import qualified Solver.MinCut as MC
+
+cut :: Graph -> (Set Symbol, Set Symbol, Set Symbol)
+cut g =
+  let (mcg, caps) = mkCutGraph g
+  in MC.mincut caps mcg
+
+-- | Construct a simple, non-hyper graph and a map of costs from an input
+-- hypergraph.
+mkCutGraph :: Graph -> (MC.Graph, Map (Symbol, Symbol) Rational)
+mkCutGraph g =
+  -- Find each pairing where each pairing represents a hyperedge. The list of
+  -- symbols are the sources of the edge and the distinguished symbol is the
+  -- sink.
+  let pairs = concatMap (\s -> do
+        ps <- predecessors s g
+        pure (s, ps)) (symbols g) :: [(Symbol, [Symbol])]
+      -- The basic graph has the same initial and terminal symbols as the hypergraph.
+      is = initials g
+      ts = terminals g
+      mcg = MC.empty & MC.initials .~ is
+                     & MC.terminals .~ ts
+  in
+  -- To build up the graph, look at each hyperedge.
+  foldr handlePair (mcg, M.empty) pairs
+
+  where
+    -- Count up the number of vertices above and below each vertex.
+    (up, down) = count g
+
+    -- Modify the result due to the effects of a hyperedge.
+    handlePair :: (Symbol, [Symbol]) -> (MC.Graph, Map (Symbol, Symbol) Rational)
+                                     -> (MC.Graph, Map (Symbol, Symbol) Rational)
+    handlePair (snk, srcs) (mcg, cap) =
+      let below = (snk, M.findWithDefault 0 snk down)
+          aboves = map (\src -> (src, M.findWithDefault 0 src up)) srcs
+      in
+      -- To handle the hyperedge, consider the pairing of each source with the
+      -- sink individually.
+      foldr (singlePair below aboves) (mcg, cap) srcs
+
+    singlePair :: (Symbol, Integer)
+               -> [(Symbol, Integer)]
+               -> Symbol
[…]

### 2018-04-09 src/Solver/Graph.hs

@@ -1,0 +1,51 @@
+module Solver.Graph where
+
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+import qualified Data.Set as S
+import           Data.Grammar
+
+type Symbol = Int
+
+data Graph = Graph
+  { graphForward :: Map Symbol [Symbol]
+  , graphBackward :: Map Symbol [[Symbol]]
+  } deriving (Show, Read, Eq, Ord)
+
+successors :: Symbol -> Graph -> [Symbol]
+successors s = M.findWithDefault [] s . graphForward
+
+predecessors :: Symbol -> Graph -> [[Symbol]]
+predecessors s = M.findWithDefault [] s . graphBackward
+
+symbols :: Graph -> Set Symbol
+symbols g = M.keysSet (graphForward g) `S.union` M.keysSet (graphBackward g)
+
+initials :: Graph -> Set Symbol
+initials g =
+  S.filter (\s -> all null (predecessors s g)) (symbols g)
+
+terminals :: Graph -> Set Symbol
+terminals g = symbols g `S.difference` M.keysSet (graphForward g)
+
+grammarToGraph :: Grammar a -> Graph
+grammarToGraph (Grammar _ rs) = Graph forw back
+  where
+    back = M.mapKeys (\(NT nt) -> nt) (fmap ruleSymbols rs)
+    forw = foldr addForw M.empty (M.toList back)
+
+    addForw :: (Symbol, [[Symbol]]) -> Map Symbol [Symbol] -> Map Symbol [Symbol]
+    addForw (snk, srcs) m = foldr (\src -> M.insertWith (++) src [snk]) m (concat srcs)
+
+    ruleSymbols :: Rule a -> [[Symbol]]
+    ruleSymbols = \case
+      Null -> []
+      Eps -> []
+      Alt a b -> ruleSymbols a ++ ruleSymbols b
+      Seq a b ->
+        let as = ruleSymbols a
+            bs = ruleSymbols b
+        in if null as then bs else if null bs then as else [a' ++ b' | a' <- as, b' <- bs]
+      Nonterminal (NT nt) -> [[nt]]
+      Terminal _ -> []

### 2018-04-09 src/Solver/Interpolate.hs

@@ -3,6 +3,8 @@ module Solver.Interpolate where
 import           Control.Lens
 import           Control.Monad.State
 import           Control.Monad.Reader
+import           Control.Monad.Except
+import           Control.Concurrent.Async (concurrently)
 
 import           Data.Map (Map)
 import qualified Data.Map as M
@@ -11,31 +13,79 @@ import qualified Data.Set as S
 import           Data.Grammar (Grammar, Rule, NT)
 import qualified Data.Grammar as G
 import           Data.Maybe (fromMaybe)
+import           Data.Foldable
 
 import           Formula
 import qualified Formula.Z3 as Z3
 
+import           Solver.Graph
+import           Solver.Cut
+
+import Debug.Trace
 import Data.Text.Prettyprint.Doc
+
+data InterpolationStrategy
+  = SequentialInterpolation
+  | ConcurrentInterpolation
+  deriving (Show, Read, Eq, Ord)
+
+-- | `licketySplit` computes interpolants of a DAG. It attempts to do this
+-- maximally in parallel by cutting the DAG consideration at each iteration.
+licketySplit :: MonadIO m
+             => InterpolationStrategy
+             -> Grammar Expr -> m (Either Model (Map NT Expr))
+licketySplit strategy g = liftIO (loop strategy g M.empty)
+
+-- | The core loop for interpolating a DAG. The loop requires two versions of
+-- the DAG, one which is a full representation and another which is a
+-- restricted version that only contains a subset of the vertices.
+-- On each iteration, the loop cuts the restricted subset using a minimum cut
+-- algorithm. Those vertices along the cut are interpolated as a part of the
+-- current iteration. The two subgraphs on either side of the cut are the new
+-- restrictions which can be interpolated in parallel.
+loop :: InterpolationStrategy
+     -> Grammar Expr -> Map NT Expr -> IO (Either Model (Map NT Expr))
+loop strategy g@(G.Grammar st rs) m =
+  if null (M.keysSet rs S.\\ M.keysSet m)
+  then pure (Right m)
+  else do
+    let gr = grammarToGraph g
+    let (now, _, _) = cut gr
+    runStateT (runExceptT $ mapM_ (`interpolateNT'` g) (S.map G.NT now)) m >>= \case
[…]

### 2018-04-09 src/Solver/MinCut.hs

@@ -1,0 +1,232 @@
+{-# LANGUAGE TemplateHaskell #-}
+module Solver.MinCut where
+
+import           Control.Lens
+import           Control.Monad.State
+import           Control.Monad.Reader
+import           Control.Monad.Extra (concatMapM)
+
+import           Data.Map (Map)
+import qualified Data.Map as M
+import           Data.Set (Set)
+import qualified Data.Set as S
+
+type Symbol = Int
+
+data Graph = Graph
+  { _forward   :: Map Symbol (Set Symbol)
+  , _backward  :: Map Symbol (Set Symbol)
+  , _initials  :: Set Symbol
+  , _terminals :: Set Symbol
+  } deriving (Show, Read, Eq, Ord)
+makeLenses ''Graph
+
+addEdge :: Symbol -> Symbol -> Graph -> Graph
+addEdge s t g =
+  g & forward  %~ M.insertWith S.union s (S.singleton t)
+    & backward %~ M.insertWith S.union t (S.singleton s)
+
+deleteEdge :: Symbol -> Symbol -> Graph -> Graph
+deleteEdge s t g =
+  g & forward %~ M.adjust (S.delete t) s
+    & backward %~ M.adjust (S.delete s) t
+
+addInitial :: Symbol -> Graph -> Graph
+addInitial s = initials %~ S.insert s
+
+addTerminal :: Symbol -> Graph -> Graph
+addTerminal s = terminals %~ S.insert s
+
+empty :: Graph
+empty = Graph M.empty M.empty S.empty S.empty
+
+data CutState = CutState
+  { _capacities :: Map (Symbol, Symbol) Rational
+  , _weights    :: Map (Symbol, Symbol) Rational
+  } deriving (Show, Read, Eq, Ord)
+makeLenses ''CutState
+
+type Cut = State CutState
+
+data Direction = Forward | Backward
+  deriving (Show, Read, Eq, Ord)
+
+data ExploreCtxt = ExploreCtxt
+  { _direction :: Direction
+  , _visited :: Set Symbol
+  , _cutState :: CutState
+  , _theGraph :: Graph
+  } deriving (Show, Read, Eq, Ord)
+makeLenses ''ExploreCtxt
+
+type Edge = (Symbol, Symbol, Direction)
+
+edges :: Graph -> [(Symbol, Symbol)]
+edges g =
+  concatMap (\(s, ts) -> [(s, t) | t <- S.toList ts]) (M.toList (view forward g))
+
[…]

### 2018-04-09 src/Solver/Shara.hs

@@ -9,6 +9,8 @@ import           Data.Map (Map)
 import qualified Data.Map as M
 import           Data.Set (Set)
 import qualified Data.Set as S
+import           Data.Sequence (Seq)
+import qualified Data.Sequence as Seq
 import           Data.Maybe (isJust, fromJust)
 import           Data.Semigroup
 import           Data.List.Split
@@ -27,6 +29,13 @@ import           Solver.Interpolate
 import           Solver.Chc
 import           Solver.TreeUnwind
 
+import Data.Text.Prettyprint.Doc
+
+data SolveKind
+  = Topological
+  | LicketySplit InterpolationStrategy
+  deriving (Show, Read, Eq, Ord)
+
 -- | Encode the CHC as a (possibly infinite) unrollable grammar such that each
 -- unrolling is a directly solvable CHC system.
 -- directlySolvable :: Expandable m => Chc -> m (InfGrammar m Expr)
@@ -34,19 +43,24 @@ import           Solver.TreeUnwind
 
 -- | Apply the interpolator to the given directly solvable system and yield
 -- either a counterexample model or solutions for the nonterminals.
-solveDirect :: MonadIO m => Grammar Expr -> m (Either Model (Map NT Expr))
-solveDirect = topologicalInterpolation
+solveDirect :: MonadIO m => SolveKind -> Grammar Expr -> m (Either Model (Map NT Expr))
+solveDirect = \case
+  Topological -> topologicalInterpolation
+  LicketySplit strat -> licketySplit strat
 
 -- | Solve the Relational Post Fixed-Point Problem as encoded by the given CHC
 -- system.
 -- solveChc :: (Expandable m, MonadIO m) => Chc -> m (Either Model (Map NT Expr))
-solveChc chc = solve =<< treeUnwind chc
+-- solveChc chc = solveLoop =<< treeUnwind chc
 
-treeSolve :: Chc -> IO (Either Model (Map NT Expr))
-treeSolve chc = evalStateT (runReaderT (solve =<< treeUnwind chc) emptyCtxt) emptyState
+solve :: SolveKind -> Chc -> IO (Either Model (Map NT Expr))
+solve sk chc =
+  evalStateT (runReaderT (solveLoop sk =<< treeUnwind chc) emptyCtxt) emptyState
 
-solve :: (Expandable m, MonadIO m) => InfGrammar m Expr -> m (Either Model (Map NT Expr))
[…]

### 2018-04-09 src/Solver/TreeUnwind.hs

@@ -1,4 +1,3 @@
-{-# LANGUAGE QuasiQuotes #-}
 {-# LANGUAGE TemplateHaskell #-}
 module Solver.TreeUnwind where
 
@@ -27,24 +26,23 @@ import           Solver.Types hiding (DirectState(..))
 import           Solver.Chc
 import           Solver.Interpolate
 
-import Data.Text.Prettyprint.Doc
-
 import Debug.Trace
 
 data UnwindState = UnwindState
   { _ntCounter :: Int
-  , _varCounter :: Int
   , _clones :: Map NT (Set NT)
   , _proofStart :: Rule Proof
   , _proofRules :: Map NT (Rule Proof)
-  , _varQueue :: Seq Var
-  , _varMapping :: [(Var, Var)]
+  , _chcState :: ChcState
   , _descendants :: Map NT (Map NT (Set NT))
   } deriving (Show, Read, Eq, Ord)
 
 data UnwindCtxt = UnwindCtxt
   { _visited :: Set NT
   , _branchClones :: Map NT (Set NT)
+  , _left :: Maybe (NT, NT)
+  , _currentRule :: Rule Fragment
+  , _chcSystem :: Chc
   } deriving (Show, Read, Eq, Ord)
 
 makeLenses ''UnwindState
@@ -53,12 +51,10 @@ makeLenses ''UnwindCtxt
 emptyState :: UnwindState
 emptyState = UnwindState
   { _ntCounter = 0
-  , _varCounter = 0
   , _clones = M.empty
   , _proofStart = G.Null
   , _proofRules = M.empty
-  , _varQueue = Seq.empty
-  , _varMapping = []
+  , _chcState = emptyChcState
   , _descendants = M.empty
   }
 
@@ -66,6 +62,9 @@ emptyCtxt :: UnwindCtxt
 emptyCtxt = UnwindCtxt
   { _visited = S.empty
   , _branchClones = M.empty
+  , _left = Nothing
+  , _currentRule = mempty
+  , _chcSystem = mempty
   }
 
 instance Monad m => Expandable (StateT UnwindState m) where
@@ -82,15 +81,17 @@ instance Expandable m => Expandable (ReaderT UnwindCtxt m) where
 -- The proof structure is encoded in the state. As the user unwind
 treeUnwind :: (MonadReader UnwindCtxt m, MonadState UnwindState m)
            => Chc -> m (InfGrammar m Expr)
[…]

### 2018-04-09 src/Solver/Types.hs

@@ -7,12 +7,3 @@ import qualified Data.Map as M
 import           Data.Set (Set)
 import           Data.Grammar (Grammar, Rule, NT)
 import qualified Data.Grammar as G
-
-type ProofStructure = Grammar Proof
-
-data Proof = Continue | Entails NT NT
-  deriving (Show, Read, Eq, Ord)
-
-class Monad m => Expandable m where
-  getProof :: m ProofStructure
-  getClones :: m (Map NT (Set NT))

### 2018-04-09 paper/cade-2017-rebuttals/rebuttal.txt

We thank the reviewers for their detailed comments. Responses are
included below, where each summarized comment is labeled with the
reviews that contained it. Each of the points will be addressed in a
future version of the paper.

---------------------------------------------------------------------

Q1 (Reviews 1, 2, 3): This paper appears to only reproduce known
approaches that construct a body-disjoint CHC system and solve it
efficiently.

A: SHARA indeed resembles previous work in that it uses cloning to
generate a CHC system that can be solved efficiently. However, in this
work, we have introduced the class of Clause-Dependence Disjoint (CDD)
systems. To the best of our knowledge, the class of systems has not
been considered in previous work, and strictly contains the class of
body-disjoint systems.

In particular, a system is body disjoint if each relational predicate
R occurs in the body of at most one clause, and occurs in any clause
at most once ([16], Sec. 4.1, Defn. 3). A system is CDD if, for each
clause, the dependencies of each predicate that occur in the body of
the clause are disjoint (submission, Sec. 4, Defn. 7). The fact that
the class of body-disjoint systems is contained by the class of CDD
systems follows fairly directly from their definitions; we will
provide a detailed proof in a future version of this paper.

The following is a system that is not body disjoint but is CDD:

x = 0 => R1(x)
x > 0 => R2(x)
x < 0 => R4(x)
R1(x), R2(x) => R3(x)
R1(x), R4(x) => R3(x)

As a more practical example, the system S'[MC] (Sec. 2.2) is not body
disjoint, but is CDD.

The work on body-disjoint systems is indeed relevant, and we will
discuss it in greater detail in a future version of the paper.

--------------------------------------------------------------------

Q2 (Reviews 1, 2, 3): How did you choose the benchmarks on which you
evaluated SHARA? How did you convert SV-COMP benchmarks to CHC
systems?

A: We evaluated SHARA on all CHC systems in the SV-COMP13 and
SV-COMP15 benchmark suites. Both SV-COMP13 and SV-COMP15 contain the
exact CHC systems used, in directory "clauses". CHC systems in
SV-COMP13 were generated by the SV-COMP maintainers from C programs
using the UFO model checker. CHC systems in SV-COMP15 were generated
by the SV-COMP maintainers from the the NTS benchmark suite using
Eldarica.

---------------------------------------------------------------------

Q3 (Reviews 1, 3): How does the quality of solutions generated by
state-of-the-art techniques compare to those generated by SHARA?

A: Out of our 670 benchmark system, 651 are recursive. Our evaluation
thus indicates that SHARA provides solutions that have quality at
least as good as those of Duality for the purposes of finding
solutions to recursive systems. A more detailed evaluation, in which
we analyze contributions to improvements in performance due to the
time spent to solve each recursion-free system and the number of
recursion-free systems solved, would be interesting to perform, and
will be included in a future version of the paper.

--------------------------------------------------------------------

Q4 (Reviews 1, 3): Is the implementation of SHARA available in order to
reproduce the results?

A: The implementation and benchmarks are available at the following
anonymous URL.
https://www.dropbox.com/s/drh43cj9baf18ej/z3_modified.zip?dl=0

Instructions for running the solver are included within the zip in the
file README.txt.

--------------------------------------------------------------------

Q5 (Review 1, 3) The description of existing solvers is incorrect: most
solvers enumerate one derivation at a time and attempt to avoid
enumerating all derivations.

A: We completely agree with this description of existing solvers: this
was the intended reading of our discussion, particularly p4. para 2.

---------------------------------------------------------------------

Q6 (Review 1): The paper presents an algorithm that transforms a
non-linear CHC system to a linear CHC system.

A: SHARA indeed translates a given system S to one in a class that can
be solved tractably, but it does not necessarily translate S to a
system that is linear. SHARA translates a given system to an
equivalent CDD system, but the class of CDD systems strictly contains
the class of linear systems. E.g., the system given in the response to
Q1, and the example system S'[MC] (Sec. 2.2) are non-linear CDD
systems.

---------------------------------------------------------------------

Q7 (Review 1): SHARA is guaranteed to be intractable in an
overwhelming number of cases: to generate a CDD expansion, it performs
a step analogous to inlining each function per callsite. In
particular, in Alg. 3, the call to VC on line 10 causes exponential
blowup.

A: To generate a CDD expansion, SHARA indeed may, in some cases, clone
clauses that model a function, if the function is a dependency of
multiple predicates in the body of the same clause. However, SHARA
need not clone every predicate that models a function for each
occurrence in an arbitrary system (or in practice, in most
systems). For example, the program:

g(x) { return x + 1; }
f(x) { if (x >= 0) { res := g(0); } else { res := g(1); } return res;

Can be modeled by system:

res = x + 1 => G(x, res)
G(0, res), x >= 0 => F(x, res)
G(1, res), x < 0 => F(x, res)

The above system is CDD: thus, SHARA would solve it directly without
inlining the definition of G. Avoiding such an inlining can have a
significant consequence on performance if the definition of g() is
large.

In Alg. 3, VC, given a CDD system S, generates a formula of size
linear in S (see Sec. 4.3), so it is not the cause of exponential
blowup. Blowup, if it occurs, is the result of Alg. 2 constructing a
CDD system that is significantly larger than a given system.

---------------------------------------------------------------------

Q8 (Review 1): What is the contribution of the proposed approach,
given its high-worst case performance?

A: Previous work has established that solving recursion-free Horn
Claues over integer linear arithmetic is co-NEXPTIME ([14],
Sec. 5). Thus, finding a solution that performs well in general is
infeasible.

The main conceptual contribution of our work is to identify that the
class of CDD systems can be solved with a polynomial number of queries
to a theorem prover. The main practical contribution of our work is to
demonstrate that realistic benchmarks can be translated to CDD systems
that are often much smaller than their worst-case bound on size.

--------------------------------------------------------------------

Q9 (Review 1): SHARA should be compared to a previous approach [KGS14]
that solves a given recursion-free CHC system with a polynomial number
of queries. SHARA, given the example program and property in that
paper, would generate systems that are exponentially larger.

A: [KGS14] indeed addresses a relevant problem. However, The approach
proposed in [KGS14] only satisfies a polynomial bound when applied to
Boolean programs. [KGS14] also extends the proposed verifier from
Boolean programs to those that use linear arithmetic. Because the
verifier for linear arithmetic relies on quantifier elimination, it is
not clear if such an approach can be extended to richer theories that
support interpolation, such as the combination of linear arithmetic
with uninterpreted functions.

Consider the following encoding of the program and property from the
paper:

T(m, m'), D(m', m'''), D(m'', m''') => M(m, m''') (1)
t<=0 => T(t, t) (2)
T(t', t''), t > 0, t' = t - 2, t''' = t'' + 1 => T(t, t''') (3)
d' = d - 1 => D(d, d') (4)
m' < 2m + 4 => false (5)

Consider the family of recursion-free CHC systems in which the i-th
system is constructed from i occurences of clause (3). Each system in
the family is not CDD, because it contains a single instance of rule
(1), which contains two occurrences of relational predicate
D. However, the (unique) minimal CDD expansion of each system in the
family is identical except for the fact that it contains two copies of
relational predicate D. Thus the sizes of the CDD expansion of the
i-th system in the family is linear in terms of the size of the i-th
system.

--------------------------------------------------------------------

Q10 (Review 1): SHARA should be compared to a previous approach that
has been implemented for solving recursion-free CHC systems
[HB12]. SHARA, given the running example in that paper would generate
systems that are exponentially larger.

A: [HBS12] indeed addresses a relevant problem; the approach proposed
in [HB12] resembles SHARA, in that it uses interpolation to solve
CHC's over theories.

However, the class of systems of non-Boolean theories addressed in
[HB12] is the class of "timed pushdown systems" ([HB12, Sec. 5]), a
subclass of CHC systems with the theory of linear real
arithmetic. Thus, in principle, SHARA and [HB12] are applicable to
distinct classes of CHC systems. In practice, there were not common
benchmarks in SV-COMP on which we could compare SHARA and [HB12].

All families of unwindings constructed from n applications of the
recursive clauses in the example in HB12 are of size linear in n, and
are CDD. E.g., consider the family of systems:

R[0](0, 0, 0, 0)
...
T[i](2, 3, Y1, Y2, Y1)
R[i](L, M, Y1, Y2), T[i](L, L', Y1, Y2, Y1') => R[i + 1](L', M, Y1', Y2)
...
R[n](2, 2, Y1, Y2) => false

Each system in the family is CDD.

--------------------------------------------------------------------

Q11 (Review 1) There does not appear to be a significant difference
between SHARA and existing work that computes nested interpolants
using multiple interpolation queries [8].

A: SHARA is indeed similar to the approach proposed in [8] in that
both approaches construct a solution from the result of interpolation
queries. However, the solver proposed in [8] can only be applied to
CHC systems in which each relational predicate is the head of at most
one clause. SolveCDD, used in SHARA, can be applied to CDD systems;
the class of CDD systems strictly contains the class considered in
[8].

--------------------------------------------------------------------

Q12 (Review 1) What is the semantics of uninterpreted functions in
constraints? Should they be universally quantified?

A: Yes: the semantics matches that adopted by other solvers for CHC
systems over a theory that includes uninterpreted functions.

--------------------------------------------------------------------

Q13 (Review 2): What is the complexity of SHARA for solving a given CDD
system?

A: SHARA issues a number of theorem prover queries that is linear in
the number of clauses in the given system. The size of the formula in
each query is of size bounded linearly in the sum of the sizes of the
constraints in all clauses in the system.

The cost can be derived by a direct analysis of SolveCDD (Alg. 3); we
will provide a more detailed cost and complexity analysis in a future
version of the paper.

--------------------------------------------------------------------

Q14 (Review 2): Does the expansion algorithm change the complexity of
solving clauses?

A: This is an interesting question. While different algorithms can
generate expansions of different sizes, they cannot change the
structure of the first-order constraints in each clause, by the
definition of an expansion (Defn. 6). As a result, the complexity of
solving each expansion of a fixed system is determined by the number
of clauses in the system.

--------------------------------------------------------------------

Q15 (Review 2): is there a meaningful relationship between solving
times of minimal and non-minimal expansions?

A: This is an interesting question, although there does not appear to
be a relationship between the times required to solve them other than
what is implied by the relationship of their sizes.

--------------------------------------------------------------------

Q16 (Review 2): The paper should include additional information
regarding its evaluation.

Out of all benchmark systems found in the editions of SV-COMP, 99 have
a counter-derivation, 328 do not (i.e., have a higher-order solution),
nineteen caused SHARA to timeout, and 224 caused the solvers to return
the value "unknown." Such benchmarks were defined over a theory not
supported by the solvers, e.g. non-linear arithmetic.

Of the nineteen cases on which SHARA timed out, Duality determined
that nine had counter-derivations, three do not, timed out on six, and
returned unknown on one. Duality timed out on eleven cases total. Of
those, SHARA determined that three had a counter-derivation, two did
not, and timed out on the remaining six.

The number of relational predicates in the systems range from three to
293, with an average of eight per system. The number of clauses range
from three to 1,022, with an average of 17. Among the constraint
formulas that occur in each clause, the number of variables that occur
in a formula range from one to 4,819, with an average number of
variables of 4,476.

The ratios of the number of clauses in each original non-recursive
system solved to the number of clauses in its CDD expansion ranged
from one to 19, with an average of 1.73.

### 2018-04-09 paper/abstract.tex

\begin{abstract}
  % problem, context
  Recursion-free Constrained Horn Clauses (CHCs) are logic-programming
  problems that formulate verifying safety of programs with bounded
  iteration and recursion.
  %
  They both formulate bounded verification problems and are generated
  by solvers that attempt to verify safety of unbounded iterative or
  recursive systems.
  %
  Efficient solvers of recursion-free systems reduce the problem to
  solving a series of systems that can each be solved tractably.

  % contribution:
  In this paper, we define a novel tractable class of recursion-free
  systems, named \emph{Clause-Dependence Disjoint (CDD)} systems, that
  strictly generalizes tractable classes defined in previous work.
  %  
  We implemented a novel solver for recursion-free CHCs, named \sys,
  that reduces the problem of solving a given recursion-free system to
  solving an equivalent CDD system.
  %
  Our evaluation of \sys on standard benchmarks indicates that, in
  many practical cases, it performs significantly better than a
  state-of-the-art implementation.
\end{abstract}

### 2018-04-09 paper/fig/bench.tex

\begin{center} 
  \begin{table*} 
    \centering
    \begin{tabular}{| r| r | r | r | r |} 
      \hline
      \multicolumn{1}{|c }{ \textbf{Benchmarks}}
       & \multicolumn{2}{|c|}{\sys} & 
       \multicolumn{2}{c|}{\duality} \\
       \hline
       \multicolumn{1}{|c|}{\textbf{Number}} 
       &  \multicolumn{1}{|c|}{\textbf{2 X}} 
       &  \multicolumn{1}{c|}{\textbf{TO}} 
       &  \multicolumn{1}{|c|}{\textbf{2 X}} 
       &  \multicolumn{1}{c|}{\textbf{TO}} 
       \\
      \hline
       670 & 194 & 19 & 24 & 8\\
      \hline
    \end{tabular}
    % 
    \caption{The results of our evaluation of \sys.
      % 
      "Benchmarks'' contains features of the
      benchmarks, in particular the 
       number of CHC problems ("Number")
      % 
      "\sys" contains features of the
      performance of \sys, in particular  ("2 X") contains the number of 
      CHC problems that at least two times faster than \duality , ("TO") contains the 
      number of CHC problems that time out.
      %
       " \duality " contains features of the
      performance of a, in particular  ("2 X") contains the number of 
      CHC problems that at least two times faster than \sys, ("TO") contains the 
      number of CHC problems that time out.
      %
      When we compare the running time, we eliminate all cases if both solver can solve it under 0.1 second
      }
    \label{table:bench}
  \end{table*} 
\end{center}

### 2018-04-09 paper/background.tex
basic draft except background 3.1, evalutaion and related work

@@
 %
-A head can either be an uninterpreted relational predicates applied to variables or a formula 
-relational predicate.
+A head can either be an uninterpreted relational predicates applied to variables or \cc{False}.
 %
-A Clause where the head is a formula is called a query.
+A Clause where the head is \cc{False} is called query.
 \begin{defn}
@@
   $\cc{chc} \Coloneqq \cc{head} \gets \cc{body}$ \\
-  $\cc{head} \Coloneqq \varphi \mid \cc{pred}$  \\
+  $\cc{head} \Coloneqq \cc{False} \mid \cc{pred}$  \\
   $\cc{body} \Coloneqq \varphi \wedge \cc{preds}$ \\
@@
 %
+\begin{ex}
+  Because uninterpreted predicates \cc{L4} is in the body of clause (4) and uninterpreted predicates \cc{L8}
+  is the head of clause (4), \cc{L4} is a dependency predicate of \cc{L8}.\autoref{chc}
+\end{ex}
+%
 Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$, $\depsof{P}$
@@
   Given a CHC system $\cc{S}$ and three uninterpreted predicates $\cc{P},
-  \cc{Q}$ and $\cc{R} \in \predof{S}$, if $\cc{Q} \in \depsof{P}$ and $\cc{R} \in \depsof {Q}$
-  , then $\cc{R}$ is a transitive dependency predicate of $\cc{P}$.
+  \cc{Q}$ and $\cc{R} \in \predof{S}$, if $\cc{Q} \in \depsof{P}$ then \cc{Q} is a transitive dependency
+  predicate of \cc{P}. 
+  %
+  if \cc{Q} is a transitive dependency predicate of \cc{P}, and \cc{R} is a transitive dependency predicate
+  of \cc{Q},then $\cc{R}$ is a transitive dependency predicate of $\cc{P}$.
 \end{defn}
+%
+\begin{ex}
+  Because uninterpreted predicates \cc{L4} is a dependency predicate of \cc{L8}, 
+  \cc{L4} is a transitive dependency predicate of \cc{L8}.
+  %
+  And because \cc{L8} is an transitive dependency predicate of \cc{L9}, 
+  \cc{L4} is a transitive dependency predicate of \cc{L9}.\autoref{chc}
+\end{ex}
 %
@@
 \end{defn}
+%
+\begin{ex}
+  Because uninterpreted predicates \cc{L9} and \cc{db} both shows in the body of
+  clause (6), \cc{L9} and \cc{db} are siblings of each other.
+  %
+\end{ex}
 %

### 2018-04-09 paper/overview.tex
basic draft except background 3.1, evalutaion and related work

@@
 %
-\cc{multiCall} includes a procedure \cc{dbl} that, given an integer
-variable \cc{n}, returns the value of \cc{n} doubled (line 1).
+\cc{multiCall} includes a procedure \cc{db} that, given an integer
+variable \cc{m}, returns the value of \cc{m} doubled (line 1).
 %
 \cc{main}, the entry procedure of \cc{multiCall}, given integer \cc{n}
-(line 3), first binds variable \cc{res} to $0$ (line 4).
+(line 3), first binds variable \cc{abs} to $0$ (line 4).
 %
-\cc{main} then tests if \cc{n} is greater than or equal to $0$ (line
+\cc{main} tests if \cc{n} is greater than or equal to $0$ (line
 5).
 %
-If so, \cc{main} doubles the value in \cc{n} and binds the result to
-\cc{res} (line 6).
+If so, \cc{main} binds the value of \cc{n} to \cc{res} (line 6).
 %
-If not, \cc{main} doubles the value in \cc{n}, negates the result, and
-binds the final result to \cc{res} (line 8).
+If not, \cc{main} binds the negation value of \cc{n} to \cc{res} (line 8).
 %
-\cc{main} then returns the value stored in \cc{res} (line 9).
+\cc{main} then calles \cc{dbl} on \cc{abs} and binds the return value to \cc{res} (line 9)
+%
+\cc{main} then returns the value stored in \cc{res} (line 10).
 
 % state the safety problem:
-The problem that we address in this work, in the context of
-\cc{multiCall}, is to verify that for each input, \cc{multiCall}
+The property to be verified in \cc{multiCall} is, for each input, 
+\cc{multiCall}
 returns a value greater than or equal to $0$.
@@
 to solving a recursion-free CHC system over a set of uninterpreted predicates
-$\mcpreds$ that each represents a control location and
+that each represents a control location and
 procedure name in \cc{multiCall}.
-%
-A solution interprets each control location \cc{L} as a summary that
-relates the arguments in each active call of the function containing
-\cc{L} to the program state at \cc{L}, and interprets each procedure
-name \cc{f} as a summary of \cc{f}.
-%
-For each step of \cc{multiCall} to \cc{L}, there is a corresponding
-clause with head predicate \cc{L}.
-%
-For each procedure \cc{f}, there is a clause relating state at return
-locations of \cc{multiCall} to a summary of their procedure.
 %
@@
 \begin{gather}
+\label{chc}
   % semantic constraint of dbl:
-  \cc{dbl}(\cc{n}, \cc{res}) \gets \cc{res} = 2 * \cc{n} \\
+  \cc{db}(\cc{m}, \cc{d}) \gets \cc{d} = 2 * \cc{m} \\
   % semantic constraint of line 3:
-   \cc{L4}(\cc{n}, \cc{res}) \gets  \cc{L3}(\cc{n}) \land \cc{res} = 0 \\
+   \cc{L4}(\cc{n}, \cc{abs}) \gets  \cc{abs} = 0 \\
   % semantic constraint of line 5 (then branch)
-   \cc{L9}(\cc{n}, \cc{res'}) \gets \cc{L4}(\cc{n}, \cc{res}) \land \cc{dbl}(\cc{n}, \cc{res}') \land
-   \cc{n} \geq 0 \\
+   \cc{L6}(\cc{n}, \cc{abs}) \gets \cc{L4}(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
   % semantic constraint of line 6 (else branch)
-   \cc{L9}(\cc{n}, \cc{res_1}) \gets \cc{L4}(\cc{n}, \cc{res}) \land \cc{dbl}(\cc{n}, \cc{res_0}) \land
-   \cc{n} < 0  \land \cc{res_1} = -\cc{res_0} \\
+   \cc{L8}(\cc{n}, \cc{abs}) \gets \cc{L4}(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
+  % assign branch
+   \cc{L9}(\cc{n}, \cc{abs'}) \gets \cc{L6}(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
+   \cc{L9}(\cc{n}, \cc{abs'}) \gets \cc{L8}(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
   % semantic constraint of main procedure:
-   False \gets \cc{L9}(\cc{n}, \cc{res}) \land \cc{res} < 0 
+    \cc{main}(\cc{n},\cc{res}) \gets \cc{L9}(\cc{n}, \cc{abs'}) \land \cc{db}(\cc{m}, \cc{d})
+    \land \cc{abs'} = \cc{m} \land \cc{res} = \cc{d} \\
+  %
+   False \gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
 \end{gather}
@@
 \emph{Clause-Dependent-Disjoint (CDD)}) system is an recursion-free CHC system such that
-for any puninterpreted predictes in the body of the same clause, their transitive dependencies
+for any uninterpreted predictes in the body of the same clause, their transitive dependencies
 are disjoint(i.e,they do not have common predicates).
@@
 \autoref{sec:CDD-defn}.
+
 % give a example
-In particular, for clauses $(3)$ and $(5)$, uninterpreted predicates \cc{L4} and \cc{dbl} shows in
+In particular, for clause $(7)$, uninterpreted predicates \cc{L6} and \cc{db} shows in
 the body of same clause.
 %
-The transitive dependency of \cc{L4} is \{\cc{L3}\}, and the transitive dependency of \cc{dbl}
-is $\varnothing$.
+The transitive dependency of \cc{L6} is \{\cc{L6}\,\cc{L8}\,\cc{L4}\}, 
+and the transitive dependency of \cc{db} is $\varnothing$.
 %
-Their transitive dependencies are disjoint, because \{\cc{L3}\} $\cap \varnothing = \varnothing$.
+Their transitive dependencies are disjoint, because \{\cc{L6}\,\cc{L8}\,\cc{L4}\} $\cap \varnothing = \varnothing$.
 %
@@
 The fact that $\mcchc$ is CDD corresponds to the fact that there no statements executed more than
-once in one execution path.
+once within one execution path.
 %
@@
 interpolation queriey for each uninterpreted predicates in $\mcchc$ in a topological order.
-% walk through how a CDD system is solved directly:
-A solution of an arbitrary CDD system $\mathcal{S}$ can be constructed
-from the results of a series of interpolation queries, each one issued
-to find an interpretation of a relational predicate of $\mathcal{S}$
-in the topological order of dependences between relational predicates.
 %
-For relational predicate $R$ in $\mathcal{S}$, the interpretation of
-$R$ is the interpolant of two formulas: %
-\textbf{(1)} a Boolean combination of the interpretations found for
-each dependence of $R$ in $\mathcal{S}$ and constraints in
-$\mathcal{S}$ and %
-\textbf{(2)} a Boolean combination of constraints that occur in
-$\mathcal{S}$.
+For each uninterpreted predicate $P$ in $\mathcal{S}$, the solution of
+$P$ is the interpolant of two formulas: %
+\textbf{(1)} a \emph{pre}-formula, constructed as 
+the solution of each uninterpreted predicate in dependence of $P$, constraints
+in corresponding clauses and boolean indicator variable for 
+each uninterpreted predicates in $\mcchc$%
+\textbf{(2)} a \emph{post}-formula, constructed as  
+all constraints that occur after $P$ and boolean indicator variable for 
+each uninterpreted predicates in $\mcchc$.
+%
 
-% talk performance:
-The performance of the solver is thus largely determined by the
-ability of the interpolating theorem prover that it uses \textbf{(1)}
-to generate interpolants quickly and %
-\textbf{(2)} to generate interpolants of size close to that of their
-input formulas.
+% walk through the example
+E.g., when \sys solving $\mcchc$, \sys issues interpolation queries of 
+each uninterpreted predicates in the order \cc{L4}, \cc{L6}, \cc{L8}, 
+\cc{L9}, \cc{db} and \cc{main}.
 %
-In the worst case, all interpolating theorem provers given in previous
-work may use exponential time to generate interpolants of size
-exponential in the size of their inputs.
+In particular, when \sys issues interpolation query for uninterpreted predicates \cc{L9},
+\sys has generated solutions for \cc{L4}, \cc{L6}, \cc{L8} which shows in\QZ{refer to graph}.
 %
-However, when given formulas that model practical verification
-problems, modern theorem provers generate interpolants efficiently,
-which have size close to the size of the given formulas.
 
-% walk through multiCall
-E.g., when solving $\mcchc$, \sys generates interpretations of
-relational predicates in the order \cc{L3}, \cc{L4}, \cc{dbl},
-\cc{L9}, and \cc{main}.
+The solution of \cc{L9} is an interpolant of \textbf{(1)} 
+a \emph{pre}-formula, constructed as solutions generated for \cc{L4}, \cc{L6}
+the constraints in clauses (5) and (6) and boolean variables%
+\textbf{(2)} a \emph{post}-formula constructed from constraints in
+$\mcchc$ that all constrains after \cc{L9}, constrains for \cc{db}
+and boolean variables.
 %
-It generates an interpretation of \cc{L9} in particular as the
-interpolant of \textbf{(1)} a \emph{pre}-formula, constructed as a
-Boolean combination of interpretations generated for \cc{L4}, \cc{dbl}
-and the constraints in \autoref{eqn:then} and \autoref{eqn:else} and %
-\textbf{(2)} a \emph{post}-formula constructed from constraints in
-$\mcchc$ that model the steps of execution of \cc{multiCall} following
-\cc{9}.
+In particular, the \emph{pre}-formula for \cc{L9} is 
+\begin{gather}
+((b_{L8}\land abs'= -n) \lor (b_{L6} \land abs=n))\land (\neg b_{L6} \lor n \ge 0) \land (\neg b_{L8} \lor n<0)
+\end{gather}
 %
-A procedure for solving a CDD system directly is given in
+, $b_{L8}$ and $b_{L6}$ are boolean variables corresponding to predicates \cc{L8} and \cc{L6}.
+%
+The \emph{post}-formula for \cc{L9} is
+\begin{gather}
+(b_{main} \land res' < 0 ) \land (\neg b_{main} \lor (b_{db} \land abs'= m \land res =d )) 
+\land (\neg b_{db} \lor d = 2*m)
+\end{gather}
+, $b_{main}$ and $b_{db}$ are boolean variables corresponding to predicates \cc{main} and \cc{db}.
+%
+A detailed and formal description for the procedure of solving a CDD system is given in
 \autoref{sec:solve-cdd}.
 
-% talk about expansion:
-The class of CDD systems strictly contains all directly-solvable
-classes introduced in previous work, but does not contain all
+\subsection{\cc{multiCall} is not in other system}
+\label{sec:not-in}
+%
+There are three classes of directly solvable systems has been introduced
+in previous work(as described in \autoref{sec:intro}).
+%
+However, $\mcchc$ does not belong to any of these three classes.
+%
+
+% tree solver 
+In particular, previous work has introduced tree structure systems\cite{bjorner13,heizmann10},
+which each uninterpreted predicate in the system is the head of at most one clause.
+%
+Such systems do not include $\mcchc$, because predicate \cc{L9} is the head 
+of clause (5) and (6).
+%
+CHC solvers based on tree structure system has to reduce solving $\mcchc$ to
+solving two tree strucutre systems that contains by $\mcchc$, that one corresponds
+to the tree branch and the other corresponds to the false branch.
+% not that bad 
+There are works has been proposed which in some cases enumerating 
+and solving all derivation trees 
+of a given recursion-free
+system~\cite{mcmillan14}.
+%
+However, such solvers may enumerate all derivations tree of a given system
+in the worst case.
+
+% Body-Disjoint:
+Previous work has introduced \emph{body-disjoint}
+systems~\cite{rummer13a,rummer13b}, 
+which each uninterpreted predicate shows in the body of at most one
+clause, and shows in the body of such a clause at most once;
+%
+However, such systems cannot model a program with multiple
+control paths that share a common subpath, typically modeled as a CHC
+system with a uninterpreted predicate that occurs in the body of multiple clauses.
+%
+$\mcchc$ is not a body-disjoint system because the uninterpreted
+predicates \cc{L4} shows in the body of both clauses (3) and (4).
+%
+Such a solver, given $\mcchc$, generates a body-disjoint system
+consisting of two copies of the sub-systems with head relational
+predicates \cc{L4}.
+
+%linear systems:
+Previous work also introduced linear system~\cite{albarghouthi12a}, 
+which each clause's body has at most one uninterpreted predicate.
+%
+However, such systems cannot directly model the control flow of a
+program that contains procedure calls.
+%
+$\mcchc$ is not a linear system because the body of clause (7) has two
+predicates \cc{L9} and \cc{db}.
+%
+CHC solvers that solve linear systems, $\mcchc$, effectively inline constraints on
+relational predicates that occur in non-linear clauses~\cite{albarghouthi12b}.
+%
+Such approaches can, in general, generate an inlined system that is
+exponentially larger than a given system.
+
+%% talk about expansion:
+The class of CDD systems strictly contains these three classes, 
+but does not contain all
 recursion-free systems.
 %
-E.g., a recursion-free system that models \cc{multiCall} mutated to
-call \cc{dbl} in sequence, instead of in alternate branches, would not
+E.g., a recursion-free system that models \cc{multiCall} 
+call \cc{dbl} twice on \cc{abs} in sequence would not
 be CDD.
 %
-\sys, given an arbitrary recursion-free CHC system $\mathcal{S}$,
-solves it by generating an equivalent CDD system $\mathcal{S}'$ and
-solving $\mathcal{S}'$ directly.
+\sys, given an arbitrary recursion-free CHC system $S$,
+solves it by generating an equivalent CDD system $S'$ and
+solving $S'$ directly.
 %
-In general, $\mathcal{S}'$ may have size exponential in the size of
-$\mathcal{S}$.
+In general, $S'$ may have size exponential in the size of
+$S$.
 %
@@
 
-\subsection{\cc{multiCall} is not in other system}
-\label{sec:not-in}
-% walk through the zoo of CHC systems:
-Competitive solvers of recursion-free CHC systems reduce the problem
-of solving a given CHC system $\mathcal{S}$ to solving multiple
-systems, constructed as copies of sub-systems of $\mathcal{S}$, each
-of which is directly solvable (as described in \autoref{sec:intro}).
-%
-In particular, CHC solvers based on tree interpolation reduce solving
-$\mathcal{S}$ to solving copies of subsystems of $\mathcal{S}$ in
-which each relational predicate in $\mathcal{S}$ is the head of at
-most one clause~\cite{bjorner13,heizmann10}.
-%
-Such systems do not include $\mcchc$, which contains multiple clauses
-with the same head.
-
-% consequences for solvers:
-As a result, CHC solvers that synthesize tree interpolants may, in the
-worst case, solve systems that correspond to each derivation of a
-given CHC system.
-%
-Such solvers, given $\mcchc$, would solve it by generating two
-tree-interpolation problems, one of which corresponds to the \cc{then}
-branch of \cc{multiCall} and the other which corresponds to the
-\cc{else} branch of \cc{multiCall}.
-%
-CHC solvers have been proposed which can in many cases avoid
-enumerating and solving all derivation trees of a given recursion-free
-system~\cite{mcmillan14}.
-%
-However, such solvers may enumerate all derivations of a given system
-in the worst case.
-
-% Body-Disjoint: 
-CHC solvers that synthesize disjunctive interpolants reduce solving a
-given recursion-free system to solving a series of CHC systems in
-which each relational predicate occurs in the body of at most one
-clause, and occurs in the body of such a clause at most once;
-%
-such systems are referred to as \emph{body-disjoint}
-systems~\cite{rummer13a,rummer13b}.
-%
-However, such systems cannot compactly model a program with multiple
-control paths that share a common subpath, typically modeled as a CHC
-system with a relational predicate that occurs in multiple clauses
-(i.e., a relational predicate that is \emph{shared}).
-%
-E.g., $\mcchc$ is not a body-disjoint system because the relational
-predicates \cc{L4} and \cc{dbl} occur in both clauses
-\autoref{eqn:then} and \autoref{eqn:else}.
-
-% consequences for relational predicates:
-CHC solvers that solve body-disjoint systems, given a general recursion-free CHC system,
-generate multiple copies of sub-systems that body has shared
-relational predicates.
-%
-Such a solver, given $\mcchc$, generates a body-disjoint system
-consisting of multiple copies of the sub-systems with head relational
-predicates \cc{L4} and \cc{dbl}.
-
-% solving linear systems:
-CHC solvers that synthesize DAG interpolants~\cite{albarghouthi12a}
-reduce the problem of solving a given CHC system to solving a series
-of systems in which each clause contains at most one relational
-predicate (i.e., \emph{linear} systems).
-%
-However, such systems cannot directly model the control flow of a
-program that contains procedure calls, such as \cc{multiCall}.
-%
-Calls can only be compactly modeled using non-linear clauses, such as
-clauses \autoref{eqn:then} and \autoref{eqn:else}.
-
-% solvers based on linear systems:
-CHC solvers that solve linear systems, given an arbitrary
-recursion-free system $\mathcal{S}$, effectively inline constraints on
-relational predicates that occur in non-linear clauses of
-$\mathcal{S}$~\cite{albarghouthi12b}.
-%
-Such approaches can, in general, generate an inlined system that is
-exponentially larger than a given system.
-
 %%% Local Variables: 

### 2018-04-10 app/Main.hs
fixing inductiveness

@@ -30,7 +30,8 @@ import           Data.Text.Prettyprint.Doc
 main :: IO ()
 main = do
   as <- getArgs
-  let defStrat = LicketySplit ConcurrentInterpolation
+  let defStrat = Topological
+  -- let defStrat = LicketySplit ConcurrentInterpolation
   let cfg = case as of
         [] -> defStrat
         (a:_) -> case a of
@@ -45,7 +46,7 @@ main = do
 test :: Chc
 test =
   G.mkGrammar
-  (  G.Terminal (Fact [expr|x = 41|])
+  (  G.Terminal (Fact [expr|x = 5|])
   <> G.Terminal (Apply x)
   <> G.Nonterminal 0
   ) $

### 2018-04-10 src/Formula/Z3.hs

@@ -175,7 +175,7 @@ formToAst f =
 
     register v = do
       fd <- varDec v
-      fixedpointRegisterVariable fd
+      -- fixedpointRegisterVariable fd
       v' <- mkApp fd []
       envVars %= M.insert v v'
       return v'

### 2018-04-10 src/Solver/Chc.hs

@@ -56,8 +56,8 @@ emptyChcState = ChcState
   , _varCounter = 0
   }
 
-resolve :: Monad m => Fragment -> StateT ChcState m Expr
-resolve = \case
+generativeResolve :: Monad m => Fragment -> StateT ChcState m Expr
+generativeResolve = \case
   Apply v -> do
     -- Applying a variable adds the variable to the vocabulary and has a
     -- trivial proof.
@@ -79,6 +79,33 @@ resolve = \case
     -- Uncapture does nothing but change the vocabulary context. The proof
     -- is trivial.
     varMapping %= tail
+    pure (LBool True)
+  Fact e ->
+    -- A single fact unwinds by applying the current vocabulary to the
+    -- fact. The proof is trivial.
+    applyMapping e
+
+simpleResolve :: Monad m => Fragment -> StateT ChcState m Expr
+simpleResolve = \case
+  Apply v -> do
+    -- Applying a variable adds the variable to the vocabulary and has a
+    -- trivial proof.
+    m <- use varMapping
+    varQueue %= (Seq.<|) (subst' m v)
+    pure (LBool True)
+  Capture v -> do
+    -- Capturing a variable either yields fresh variables or pops the
+    -- vocabulary. The proof is trivial.
+    Seq.viewr <$> use varQueue >>= \case
+      Seq.EmptyR -> pure ()
+      (xs Seq.:> x) -> do
+        varQueue .= xs
+        varMapping %= (:) (v, x)
+    pure (LBool True)
+  Uncapture ->
+    -- Uncapture does nothing but change the vocabulary context. The proof
+    -- is trivial.
+    -- varMapping %= tail
     pure (LBool True)
   Fact e ->
     -- A single fact unwinds by applying the current vocabulary to the

### 2018-04-10 src/Solver/Shara.hs

@@ -65,6 +65,7 @@ solveLoop sk g = solveDirect sk (IG.finite g) >>= \case
   Right sol -> do
     ps <- getProof
     (mproof, indSet) <- inductive ps sol
+    liftIO $ putStrLn (G.draw ps)
     case mproof of
       Just proof -> do
         cs <- getClones
@@ -115,18 +116,28 @@ inductive ps sols = do
 
 entails :: MonadIO m => Map NT Expr -> Map NT (Set NT) -> NT -> Rule Fragment-> m Bool
 entails sols descs lhs rhs =
-  let lhs' = nonterm (LBool False) lhs
-      rhs' = evalState (go rhs) emptyChcState
-  in rhs' `Z3.entails` lhs'
+  let (rhs', lhs') = evalState (do
+        rhs' <- go rhs
+        lhs' <- applyMapping (nonterm (LBool False) lhs)
+        pure (rhs', lhs')) emptyChcState
+  in do
+    liftIO $ print $ pretty rhs'
+    liftIO $ print $ pretty lhs'
+    rhs' `Z3.entails` lhs'
   where
     go :: Rule Fragment -> State ChcState Expr
     go = \case
       G.Null -> pure (LBool False)
       G.Eps -> pure (LBool True)
-      G.Alt a b -> mkOr <$> go a <*> go b
+      G.Alt a b -> do
+        q <- get
+        a' <- go a
+        put q
+        b' <- go b
+        pure (mkOr a' b')
       G.Seq a b -> mkAnd <$> go a <*> go b
       G.Nonterminal nt -> applyMapping (nonterm (LBool True) nt)
-      G.Terminal x -> resolve x
+      G.Terminal x -> simpleResolve x
 
     nonterm :: Expr -> NT -> Expr
     nonterm def nt =
@@ -138,11 +149,6 @@ entails sols descs lhs rhs =
     unalias e = e & vars . varName %~ unal
     unal :: String -> String
     unal n = head (splitOn "%" n)
-
-    freshVar v = do
-      c <- use _3
-      _3 += 1
-      pure (v & varName <>~ ("%" ++ show c))
 
 collapse :: Map NT (Set NT) -> Map NT Expr -> Map NT Expr
 collapse _ = id -- TODO

### 2018-04-10 src/Solver/TreeUnwind.hs

@@ -110,14 +110,14 @@ unwindFrom r = do
       G.Alt x y -> do
         -- Unwinding an alternation requires unwinding both parts with the same
         -- vocabulary.  A proof must show both branches.
-        q <- use (chcState . varQueue)
+        q <- use chcState
         (x', px) <- unw x
-        chcState . varQueue .= q
+        chcState .= q
         (y', py) <- unw y
         pure (x' <|> y', px <> py)
       G.Terminal x -> do
         st <- use chcState
-        let (x', st') = runState (resolve x) st
+        let (x', st') = runState (generativeResolve x) st
         chcState .= st'
         pure (case x' of
           LBool True -> G.epsGrammar

### 2018-04-10 paper/approach.tex

@@
 %
-\QZ{rework example}
 \begin{ex}
@@
   %
-  The only clauses in $\mcchc$ that contain multiple uninterpreted
-  predicates are \autoref{eqn:then} and \autoref{eqn:else}, which both
-  contain $\cc{L4}$ and $\cc{dbl}$.
+  The only clauses in $\mcchc$ that its body has multiple uninterpreted
+  predicates are clause (7) that contains $\cc{L9}$ and $\cc{db}$.
   %
-  The dependences of $\cc{L4}$ and $\cc{dbl}$ are disjoint.
+  The transitive dependences of $\cc{L9}$ and $\cc{db}$ are \{\cc{L6}\,\cc{L8}\,\cc{L4}\} and $\varnothing$, which are disjoint.
 \end{ex}
@@
 % talk about body disjoint
-The class of CDD systems contains body-disjoint and linear
+The class of CDD systems contains tree structure system, and body-disjoint and linear
 systems introduced in previous work.
 %
-For recursion-free system $\cc{S}$, if for each uninterpreted predicate
-$Q \in \predof{S}$, there is at most one caluse $\cc{C}$ that $Q \in \bodyof{C}$,
+Tree structure system is contained by body-disjoint system. 
+%
+For recursion-free system $S$, if for each uninterpreted predicate
+$Q \in \predof{S}$, there is at most one caluse $C$ that $Q \in \bodyof{C}$,
 and there is no uninterpreted predicate shows more than once in a body of clause, then
-$\cc{S}$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
+$S$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
 %
@@
 %
+
 $\solvecdd$, given a CDD system $\cc{S}$, it first do a topological sort on all
@@
 
-\QZ{rework example}
 %
 \begin{ex}
-  $\solvecdd{ \mcpreds }$, given $\mcchc$, may generate
+  $\solvecdd$, given $\mcchc$, may generate
   interpretations of $\mcpreds$ in any topological ordering of the
-  dependency relation of $\mcchc$, depending on the relational
-  predicates that it chooses at \autoref{line:case-deps}.
+  dependency relation of $\mcchc$.
   %
-  One such ordering is \cc{L3}, \cc{L4}, \cc{dbl}, \cc{L9}, \cc{main},
-  $\mathsf{Err}$.
+  One such ordering is \cc{L4}, \cc{L6}, \cc{L8}, 
+   \cc{L9}, \cc{db} and \cc{main}.
 \end{ex}
@@
 %
-For an uninterpreted predicate $\cc{P} \in \predof{S}$ and $\cc{S}$ is an CDD system  
-,the formula $\ctrof{Q}$ is a compact representation of constrains of $\cc{P}$, defined as follows.
-% introduce the isused predicates:
-For each $\cc{Q} \in \predof{S}$, there is a corresponding boolean variable
-$b_Q$.
+For a CDD system $S$, given a uninterpreted predicate $\cc{P} \in \predof{S}$, and a partial 
+solution $\sigma$ that maps some uninterpreted predicate to its solution formula, 
+,the formula $\ctrof{P,\sigma}$ is a compact representation of constrains of $\cc{P}$ based on parital
+solution $\sigma$, defined as follows.
 % for each relational predicate, define constraint:
-The constraint of $P$, denoted $\ctrof{P}$, constrains
-that if $P$ occurs in a derivation of $S$, then there is
-some clauses $C \in S$ in which $P$ is the head of $\cc{C}$
-\textbf{(1)} each $Q \in \bodyof{C}$ is used and 
+If partial solution $\sigma$ does not contains $P$, 
+the constraint of $P$ under $\sigma$ that denoted as $\ctrof{P,\sigma}$ 
+is constructed as: for all clauses $C \in S$ in which $P$ is the head of $\cc{C}$
+\textbf{(1)} each predicate $Q \in \bodyof{C}$ is used and 
 \textbf{(2)} the constrained of the clause $\consof{C}$ is encoded.
 %
-I.e., $\ctrof{P}$ is
-\[\lnot b_Q \land 
-\left(
+I.e., $\ctrof{P,\sigma}$ is
+\[
 \biglor_{ (\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
@@
 \bigland_{ \cc{Q} \in \bodyof{C_i}} %
-b_Q \right) \right)
+b_Q \right)
 \]
 %
-The counterexample characterization of $\mathcal{S'}$ constrains that
-\textbf{(1)} the query predicate of $\mathcal{S'}$ must be used in a
-derivation and %
-\textbf{(2)} for $R \in \mathcal{R}$, if $R$ is used, then its
-constraint must hold.
+If partial solution $\sigma$ contains $P$, then $\ctrof{P,\sigma}$ is the solution formula $\sigma[P]$.
+
+% introduce the isused predicates:
+For each $\cc{P} \in \predof{S}$, there is a corresponding boolean variable
+$b_P$.
 %
-I.e., $\vc{ \mathcal{R} }(\mathcal{S'})$ is $\used{
-  \queryof{\mathcal{S'} } } \land %
-\bigland_{R \in \mathcal{R} } \left( \used{R} \implies \ctrof{R}
-\right)$.
+The counterexample characterization of $P$ constrains that
+if $P$ is used, which means $b_{P}$ is \cc{True}, then its constraint must hold.
+%
+I.e., $\vc{P,\sigma}$ is $\neg b_{P} \lor \ctrof{P,\sigma}$.
 
-\QZ{rework example}
 \begin{ex}
+  \label{ex:ctr}
+  In the CHC system $\mcchc$ (given in \autoref{sec:solve-ex}), 
+  when $\mcsolve$ solves the uninterpreted predicate \cc{L9}, it 
+  has a partial solution $\sigma$ that maps \cc{L6} to $n \ge 0$.
   %
-  $\vc{ \mcpreds }(\mcchc)$ characterizes all runs of \cc{multiCall}
-  that result in an error.
+  For predicate \cc{L6}, the the $\ctrof{L6,\sigma}$ is $n \ge 0$.
   %
-  Although the relational predicates \cc{L4} and \cc{dbl} occur in the
-  bodies of multiple clauses, they are each modeled with only a single
-  instance of logical variables.
+  For the 
+  uninterpreted \cc{L9}, the $\ctrof{L9,\sigma}$ is $(abs' = n \land b_{L6}) \lor (abs' = n \land b_{L8})$.
   %
-  Such a collection is sufficient because each run of \cc{multiCall}
-  reaches \cc{L4} and \cc{dbl} at most once.
-  %
-  Correspondingly, each derivation of $\mcchc$ contains at most one
-  occurrence of relational predicates \cc{L4} and \cc{dbl}.
+  The $\vc{L9,\sigma}$ is $\neg b_{L9} \lor \ctrof{L9,\sigma}$.
 \end{ex}
@@
 In particular, $\prectr(\cc{P},\sigma)$ is:
-\[\ctrof{P} \land
+\[\ctrof{P,\sigma} \land
 \left(
 \bigland_{\cc{Q} \in \depsof{P}}
-\left(b_Q \land \sigma[ \cc{Q} ]  \right)
+\left(\neg b_Q \lor \sigma[ \cc{Q} ]  \right)
 \right)
@@
 %
-\QZ{rework example}
 \begin{ex}
   \label{ex:pre-ctr}
-  Assume that $\mcsolve$, given $\mcchc$, synthesizes a partial
-  solution $\sigma \in \interps{ \mcpreds }$ that interprets \cc{L3}
-  as $\true$, and chooses to then find an interpretation of \cc{L4}.
+  In the CHC system $\mcchc$ (given in \autoref{sec:solve-ex}), 
+  when $\mcsolve$ solves the uninterpreted predicate \cc{L9}, because it solves
+  in topological order, partial solution $\sigma$ contains \cc{L6} and \cc{L8}.
   %
-  The pre-constraint of \cc{L4} is $\true \land \cc{res}_{\cc{4}} =
-  0$, where $\cc{res}_{\cc{4}}$ is a logical variable that models the
-  value of \cc{res} at line \cc{4}.
+  $\sigma$ maps
+  \cc{L6} to $n \ge 0$ and \cc{L8} to $n < 0$.
+  %
+  The pre-formula for \cc{L9} under $\sigma$ is $\ctrof{L9} \land (\neg b_{L6} \lor n \ge 0) \land (\neg b_{L8} \lor n < 0)$.
+  %
+  The forumula of $\ctrof{L9,\sigma}$ is given in~\autoref{ex:ctr}.
 \end{ex}
@@
 %
-For $\sigma \in \interps{ \mathcal{R} }$ a partial solution of
-$\mathcal{S}$ and $R \in \mathcal{R}$, the post-constraint $\postctr{
-  \mathcal{R} }(\mathcal{S}, \sigma, R) \in \formulas$ is mutually
-inconsistent with the interpretation of $R$ in each solution of
-$\mathcal{S}$ that extends $\sigma$.
+For $\sigma$ as a partial solution of
+$S$ and $P \in \predof{S}$, the post-constraint $\postctr(P, \sigma)$ is mutually
+inconsistent with the solution of $P$ in each solution of
+$S$ that extends $\sigma$.
 % dependents of R:
-In particular, let $\mathcal{D}_0$ be the reflexive dependents of $R$
-in $\mathcal{S}$,
+In particular, let $D_0$ be the reflexive transitive dependents of $P$
+in $S$,
 % clause siblings of dependents:
-let $\mathcal{D}_1$ be the siblings in $\mathcal{S}$ of
-$\mathcal{D}_0$,
+let $D_1$ be the siblings in $S$ of
+$(D_0 \union {P})$,
 %
-let $\mathcal{D}_2$ be all dependences of $\mathcal{D}_1$, and
+let $D_2$ be all transitive dependences of $D_1$, and
 % collect all predicates that we define a constraint over:
-let $\mathcal{D} = \mathcal{D}_0 \union \mathcal{D}_1 \union
-\mathcal{D}_2$.
-% define constraint for each predicate
-Let $\restrict{ \mathcal{S} }{ \mathcal{D} }$ be $\mathcal{S}$
-restricted to clauses whose head's relational predicate is in
-$\mathcal{D}$.
+let $D = D_0 \union D_1 \union D_2$.
 % define instantiation of C
-The post constraint for $\mathcal{S}$ under $\sigma$ at $R$ is the
-counterexample characterization of $\restrict{ \mathcal{S} }{
-  \mathcal{D} }$ grounded on $\sigma$ (defined in \autoref{sec:chcs}).
+The post constraint for $P$ under $\sigma$ is the conjunction of
+counterexample characterization of all predicates $Q \in D$ and the query clause.
 %
-I.e., $\postctr{ \mathcal{R} }(\mathcal{S}, \sigma, R) = \vc{
-  \mathcal{R} }(\sigma( \restrict{ \mathcal{S} }{ \mathcal{D} }))$.
+I.e., $\postctr(R, \sigma) =\cc{query} \land (\bigland_{\cc{Q} \in D} \vc{Q,\sigma})$.
 % example post constraint:
@@
   Consider the case in which $\mcsolve$, given $\mcchc$, generates an
-  interpretation of \cc{L4}, introduced in \autoref{ex:pre-ctr}.
+  interpretation for \cc{L9}, introduced in \autoref{ex:pre-ctr}.
   %
-  $\mcsolve$ collects relational predicates $\mathcal{D}_0 = \elts{
-    \cc{L4}, \cc{L9}, \cc{main}, \mathsf{Err} }$, $\mathcal{D}_1 =
-  \elts{ \cc{dbl} }$, and $\mathcal{D}_2 = \emptyset$.
+  The reflexive transitive dependents $D_0$ of \cc{L9} is \{\cc{main}\}.
   %
-  $\mcsolve$ constructs $\postctr{ \mcpreds }(\mcchc, \sigma,
-  \cc{L4})$ to be the counterexample characterization of $\mcchc$
-  restricted to clauses whose head relational predicate is in
-  $\mathcal{D} = \elts{ \cc{L4}, \cc{L9}, \cc{main}, \mathsf{Err} }$.
+  The siblings set $D_1$ is \{\cc{db}\}.
+  %
+  The transitive dependences of $D_1$ is $\varnothing$. 
+  %
+  Then the $D$ set is \{\cc{main},\cc{db}\}
+  %
+  The \cc{query} is $b_{main} \land res'<0$.
+  %
+  The post constraint for $L9$ under $\sigma$ is $\cc{query} \land \vc{main,\sigma} \land \vc{db,\sigma}$.
+  %
+  The whole formula is given in\QZ{add reference}. 
 \end{ex}

### 2018-04-10 paper/introduction.tex
pass on intro

@@
 Many critical problems in program verification, such as the
-verification of safety properties of sequential recursive programs and
+verification of safety properties of recursive programs and
 concurrent programs, can be reduced to solving systems of
@@
 clause head.
+\QZ{This seems really out of place -- DAH}
 
 % introduce the problem of solving recursion-free systems:
-Recursion-free CHC systems, in which each
-relational-predicate does not occur in a derivation (i.e., is not
-a \emph{dependency}) of itself, is an important subclass of general
-CHC systems for two reasons.
+In this work we focus on the subclass of CHC systems which are known
+as \emph{recursion-free}. In a recursion-free CHC system, the
+derivation of each relational predicate will contain no references to
+itself. \QZ{\DAH{I still think we need a small example here}} I.e., no
+relational predicate is \emph{dependent} on itself.
 %
-First, many modern solvers solves general CHC systems by solving 
-a series of recursion-free systems from bounded unwindings of given
-recursive CHC system and combining their solutions~\cite{bjorner13}.
+Recursion-free CHC systems are an important subclass for two reasons.
 %
-The performance of the solver for recursive systems critically depends
-on the performance of the solver for recursion-free systems that it
-uses. 
+First, recursion-free systems can be used to model safety properties
+for hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12},
+i.e., programs with control branches and procedure calls, but only
+bounded iteration and recursion.
 %
-Second, solving recursion-free systems formulates proving safety of
-hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12}, i.e.,
-programs with control branches and procedure calls, but only bounded
-iteration and recursion.
+Second (and most importantly), the task of solving a general CHC
+system can reduce to the task of solving a sequence of recursion-free
+systems.
+%
+Many \QZ{more references? Or remove many.} modern solvers generate
+such a sequence where each subsystem is a bounded unwinding of the
+input system. These solvers attempt to show that the solution to a
+given subsystem generalizes to the original system or else generate a
+larger subsystem for the next iteration~\cite{bjorner13}.
+%
+Such solvers are heavily reliant on the performance of the underlying
+recursion-free solver since they invoke this underlying solver in an
+iterative loop.
+%
+% current general techniques for solving recursion-free systems
+Typically, even recursion-free CHC systems are not solved directly.
+Instead, they are reduced to some even more specific subclass of CHC
+system. For example, some approaches generate CHC systems whose
+derivations are trees~\cite{bjorner13}.
+\QZ{In general it seems we need many more references in the intro.}
+%
+A solver can dispatch such a system by issuing interpolation queries
+to find a suitable definition for each relational predicate.
+%
+We refer to the subclass of CHC systems that can be solved by such
+techniques as \emph{directly solvable}.
 
-% current general techniques for solving recursion-free systems
-Many efficient CHC solvers, solve a recursion-free CHC system 
-by reducing it to one or a series of subclass of 
-recursion-free CHC system.
+% Time complexity
+In general, solving a recursion-free CHC system whose facts are
+restricted to the theory of integer linear arithmetic is
+co-NEXPTIME-complete~\cite{rummer13b}.
 %
-Then solvers solve these subclass of recursion-free CHC systems by
-issuing an interpolation query for each relational predicate.
+In contrast, solving a directly solvable system in the same theories
+is co-NP~~\cite{rummer13b}.
 %
-We refer to such subclass of systems as \emph{directly solvable}.
-%
-
-% state the complexity of solving general problem,and explain the 
-% why it is fast
-In general, solving a recursion-free CHC system for the theory of
-integer linear arithmetic is co-NEXPTIME-complete~\cite{rummer13b}.
-%
-In contrast, solving a directly solvalbe system for the theory of
-integer linear arithmetic is co-NP~~\cite{rummer13b}.
-%
-Therefore, the performance of a CHC solver for a recursion-free CHC system 
-typically is directly determined by the size(i.e the number of predicates) 
-of the subclass systems it reduces to.
+Therefore, the performance of a CHC solver is highly reliant on the
+order (i.e the number of predicates) of the directly solvable systems
+it produces.
 
 % current subclass
-Previous work has introduced three classes of directly solvable systems,
+Previous work has introduced three classes of directly solvable
+systems,
 % derivation trees:
-\textbf{(1)} tree structure systems~\cite{heizmann10,bjorner13,mcmillan14}, %
+\textbf{(1)} tree systems~\cite{heizmann10,bjorner13,mcmillan14}, %
 % disjunctive trees:
@@
 The first contribution of this paper is the introduction of a novel
-class of directly solvable systems that strictly
-contains the union of the classes of body-disjoint and linear systems.
+class of directly solvable systems that is a superset of both
+the body-disjoint subclass and the linear subclass.
 %
-In particular, the class consists of all recursion-free CHC systems such that, 
-for each clause $C$, the dependences of all distinct relational
-predicates in the body of $C$ are disjoint;
+In particular, the class consists of all recursion-free CHC systems
+such that, for each clause $C$, the dependences of all distinct
+relational predicates in the body of $C$ are disjoint \DAH{\QZ{Should
+we illustrate what disjoin means here?}};
 %
@@
 %
-We present a symbolic algorithm that demonstrates that CDD is directly solvable.
+We demonstrate that the class of CDD systems is directly solvable
+through a symbolic algorithm.
 %
@@
 % a new solver
-The second contribution of this paper is a novel solver for
-CHC systems, named \sys.
+The second contribution of this paper is a solver for CHC systems,
+named \sys.
 %
-Given recursion-free system $S$, \sys reduces the problem
+Given a recursion-free system $S$, \sys reduces the problem
 of solving $S$ to solving an equivalent CDD system
@@
 %
-In the worst case, the size of $S'$ may be exponential in the size of
-$S$.
+In the worst case, it is possible that the order of $S'$ may be
+exponential in the order of $S$.
 %
-In our experience, the size of $S'$ is usually close enough
-to the size of $S$ for \sys to perform significantly better
-than the best known CHC solvers.
+However, empirically we have found that the order of $S'$ is usually
+close enough to the order of $S$ that \sys often outperforms the best
+known CHC solvers.
 %
-Because CDD strictly contains the union of the classes 
-of body-disjoint and linear systems, the size of an equivalent minimum 
-CDD system is equal or less then the size of equivalent body-disjoint and
-linear system.
+Because the class of CDD systems is a superset of both body-disjoint
+and linear systems, the order of an equivalent CDD system is always
+less than or equal to that of systems in these subclasses.
 %
@@
 %
-Given a recursion system, \sys solves a series recursion-free systems 
-from bounded unwinding the original system and combines the solutions
-of recursion-free systems to generate the solution for the recursion system
-as previous work proposed~\cite{rummer13b}.
+Given a general (possibly recursive) CHC system, \sys solves a sequence of
+recursion-free systems.
 %
+Each subsystem is a bounded unwinding of the original system. \sys
+attempts to combine the solutions of these recursion-free systems to
+synthesize a soltuion of the original solution, as has been proposed
+in previous work~\cite{rummer13b}.
 
@@
 %
-The results indicate that solvers that perform better than both modern
-solvers and \sys could be designed by combining the strengths of both
-approaches (discussed in \autoref{sec:evaluation}).
+Futhermore, the results indicate that combining the strengths of \sys
+with that of other existing approaches (as discussed in
+\autoref{sec:evaluation}) is a promising direction for the future of
+CHC solving.
 
@@
 %
-\autoref{sec:overview} illustrates the operation of \sys using an
-example hierarchical program and corresponding CHC system.
+\autoref{sec:overview} illustrates the operation of \sys on a
+recursion-free CHC system.
 %

### 2018-04-10 paper/fig/example.tex
added diagram

\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node[label={0:$\{true\}$}] (L4) at (0,0) {$\cc{L}_4$};
  \coordinate[above = of L4] (L4');
  \node[label={180:$\{\cc{n} \geq 0\}$}, below left = of L4] (L6) {$\cc{L}_6$};
  \node[label={0:$\{\cc{n} < 0\}$}, below right = of L4] (L8) {$\cc{L}_8$};
  \node[label={180:$\{\cc{abs'} \geq 0\}$}, below right = of L6] (L9) {$\cc{L}_9$};
  \node[label={0:$\{\cc{res} \geq 0\}$}, below right = 1.6 and 1.2 of L9] (main) {$\cc{main}$};
  \node[label={0:$\{\cc{d} = 2*\cc{x}\}$}, above right = 1.6 and 1.2 of main] (dbl) {$\cc{dbl}$};
  \coordinate[above = of dbl] (dbl');
  \node[label={0:$\{false\}$}, below = of main, style = ultra thick] (bot) {$\bot$};
  \coordinate[above = 0.5 of main] (main');
\end{scope}

\begin{scope}[every edge/.style={draw=black, very thick}]
  \path[->, style=right] (L4') edge node {$\cc{abs} = 0$} (L4);
  \path[->, style=above left] (L4) edge node {$\cc{n} \geq 0$} (L6);
  \path[->, style=above right] (L4) edge node {$\cc{n} < 0$} (L8);
  \path[->, style=left] (L6) edge node {$\cc{abs'} = \cc{n}$} (L9);
  \path[->, style=right] (L8) edge node {$\cc{abs'} = -\cc{n}$} (L9);
  \path[-] (L9) edge[out=330, in=90] (main');
  \path[->, style=right] (dbl') edge node {$\cc{d} = 2*\cc{x}$} (dbl);
  \path[-] (dbl) edge[out=210, in=90] (main');
  \path[->, style=above right] (main') edge node {$\cc{abs'} = \cc{x} \land \cc{res} = \cc{d}$} (main);
  \path[->, style=right] (main) edge node {$\cc{res} < 0$} (bot);
\end{scope}
\end{tikzpicture}

### 2018-04-10 paper/code/multiCall.java

@@ -1,11 +1,10 @@
-int db(int m) {
-  return 2 * m; }
-int main(int n) {
-  int abs = 0;
+def dbl(int x)
+  return 2 * x
+def main(n)
+  abs = 0
   if (n >= 0)
-     abs = n;
+     abs = n
   else
-     abs = -n;
-  int res = dbl(abs);
-  return res;
-  }
+     abs = -n
+  res = dbl(abs)
+  assert(res >= 0)

### 2018-04-10 paper/appendix.tex
parts edit for appendix

@@
   % Inputs: a heap program and an error location.
-  \Input{For $\mathcal{R} \in \relpreds$, %
-    $\mathcal{S} \in \chc{\mathcal{R}}$.}
+  \Input{A recruions-free CHC system $S$.}
   % Output: inductive invariants.
-  \Output{A minimal CDD expansion of $\mathcal{S}$ and
-    correspondence to $\mathcal{S}$.} %
-  \myproc{$\expand{ \mathcal{R} }(\mathcal{S})$ %
+  \Output{A minimal CDD expansion $S'$ of $S$ and
+    correspondence from $S'$ to $S$.} %
+  \myproc{$\expand(S)$ %
     \label{line:expand-begin} }{ %
     % auxiliary procedure:
-    \myproc{$\expandaux( \mathcal{S}' )$
+    \myproc{$\expandaux(S')$
       \label{line:expand-aux-begin} }{ %
       % case split: optional sharing clause:
-      \Switch{$\sharingclause( \mathcal{S}' )$}{ %
+      \Switch{$\sharingclause(S')$}{ %
         % case: no siblings with overlapping descendants
-        \lCase{$\none$:}{ \Return{$\mathcal{ S }'$} %
+        \lCase{$\none$:}{ \Return{$S'$} %
           \label{line:expand-ret} } %
         % case: siblings with 
-        \lCase{$\mathcal{C} \in \mathcal{S}', % 
-          R \in \ctxs{ \mathcal{R} }$: }{ %
+        \lCase{$C \in S', % 
+          P \in \predof{S'}$: }{ %
           % expand the return:
           \Return{$\expandaux( %
-            \copyrel{ \mathcal{R} }( %
-            \mathcal{S}', \mathcal{C}, R))$ } %
+            \copyrel(S', C, P))$ } %
           \label{line:expand-recurse}
@@
     % base call: 
-    \Return{$( \expandaux(\mathcal{S}), \corr{ \mathcal{R} } )$ %
+    \Return{$( \expandaux(S), \corr )$ %
       \label{line:expand-base-call}}
@@
   %
-  \caption{$\expand{ \mathcal{R} }$: for $\mathcal{R} \in \relpreds$
-    given $\mathcal{S} \in \chcs{ \mathcal{R} }$, returns a minimal
-    CDD expansion of $\mathcal{S}$. }
+  \caption{$\expand$:
+    given a recursion-free CHC system $S$, returns a minimal
+    CDD expansion $S'$ of $S$ and its correspondence.}
   \label{alg:expand}
@@
 % walk through top-level procedure:
-\autoref{alg:expand}, given $\mathcal{S} \in \chcs{ \mathcal{R} }$
-(\autoref{line:expand-begin}), returns an expansion $\mathcal{S}'$ of
-$\mathcal{S}$ over relational predicates $\ctxs{ \mathcal{R} }$
-(defined below) accompanied by a correspondence from $\mathcal{S}'$ to
-$\mathcal{S}$.
+\autoref{alg:expand}, given a recruion-free CHC system $S$
+(\autoref{line:expand-begin}), returns an minimal CDD expansion $S'$ of
+$S$ with a correspondence from $S'$ to
+$S$.
 %
-$\expand{ \mathcal{R} }$ defines a procedure $\expandaux$
+$\expand$ defines a procedure $\expandaux$
 (\autoref{line:expand-aux-begin}---\autoref{line:expand-aux-end}) that
-takes a CHC system $\mathcal{S}$ and returns an expansion of
-$\mathcal{S}$.
+takes a CHC system $S$ and returns an minimal CDD expansion of 
+$S$.
 %
-$\expand{ \mathcal{R} }$ runs $\expandaux$ on $\mathcal{S}$ and
-returns the result, paired with the map $\corr{ \mathcal{R} } : \ctxs{
-  \mathcal{R} } \to \mathcal{R}$ (\autoref{line:expand-base-call};
-$\corr{\mathcal{R}}$ is defined below).
+$\expand$ runs $\expandaux$ on $S$ and
+returns the result, paired with the map $\corr: \predof{S'} \to \predof{S}$ 
+(\autoref{line:expand-base-call}).
 
 % walk through aux procedure:
-$\expandaux$, given $\mathcal{S}' \in \chcs{ \ctxs{ \mathcal{R} } }$,
-runs a procedure $\sharingclause$ on $\mathcal{S}'$, which tries to
-find $\mathcal{C} \in \mathcal{S}'$ and $R \in \ctxs{ \mathcal{R} }$
-applied in the body of $\mathcal{C}$ that is a transitive dependences of
-clause siblings in $\mathcal{S}'$.
+$\expandaux$, given a recursion-free CHC system $S'$,
+runs a procedure $\sharingclause$ on $S'$, which tries to
+find clause $C \in S'$ and predicate $P \in \bodyof{C}$
+that $P$ is in both transitive dependences of two predicates $Q_0$ and $Q_1$ are siblings 
+in $\predof{S}'$.
 %
-I.e., $\expandaux$ tries to find $\mathcal{C}$ and $R$ such that for
-some siblings $R_0, R_1 \in \ctxs{ \mathcal{R} }$, $R \in \depsof{
-  \mathcal{S}' }(R_0) \intersection \depsof{ \mathcal{S}' }(R_1)$ and
-$\relof{ \headof{ \mathcal{C} }} \in \depsof{ \mathcal{S}' }(R_0)
-\setminus \depsof{ \mathcal{S}' }(R_1)$.
-%
-In such a case, we say that $(\mathcal{C}, R)$ is a
+In such a case, we say that $(C, P)$ is a
 \emph{sibling-shared dependence}.
+
 % case: no sibling-shares
 If $\sharingclause$ determines that no sibling-shared dependence
-exists, then $\expandaux$ returns $\mathcal{S}'$
+exists, then $\expandaux$ returns $S'$
 (\autoref{line:expand-ret}).
@@
 % case: there is a sibling-shared dependence:
-Otherwise, for sibling-shared dependence $(\mathcal{C}, R)$ returned
-by $\sharingclause$, $\expandaux$ runs $\copyrel{ \mathcal{R} }$ on
-$\mathcal{S}'$, $\mathcal{C}$, and $R$, which returns an expansion of
-$\mathcal{S}'$ in which the dependents shared by $R_0$ and $R_1$ are
-those shared in $\mathcal{S}'$, without $(\mathcal{C}, R)$.
+Otherwise, for sibling-shared dependence $(C, P)$ returned
+by $\sharingclause$, $\expandaux$ runs $\copyrel $ on
+$S'$, $C$, and $P$, which returns an expansion of
+$S'$ by creates fresh copy $P'$ of $P$ and updates $\bodyof{C}$ 
+to avoid transitive dependents shared by $Q_0$ and $Q_1$.
 %
-$\expandaux$ recurses on the CHC system returned by $\copyrel{
-  \mathcal{R} }$ and returns the result
+$\expandaux$ recurses on the CHC system returned by $\copyrel $ and returns the result
 (\autoref{line:expand-recurse}).
-
-% give the representation of ctxs
-\subsubsection{Representation of $\ctxs{ \mathcal{R} }$}
-%
-$\ctxs{ \mathcal{R} }$ are relational predicates that store the
-clauses in which they occur.
-%
-\begin{defn}
-  \label{defn:ctxs}
-  $\ctxs{ \mathcal{R} }$ is the smallest space such that $\mathcal{R}
-  \subseteq \ctxs{ \mathcal{R} }$ and for $\mathcal{C} \in \clauses{
-    \mathcal{ \ctxs{R} } }$ and $A \in \appsof{ \bodyof{ \mathcal{C} }
-  }$, $(\mathcal{C}, \relof{A}) \in \ctxs{ \mathcal{R} }$.
-\end{defn}
-% give the correspondence between relational predicates:
-$\corr{ \mathcal{R} } : \ctxs{R} \to \mathcal{R}$ is such that for $R
-\in \mathcal{R}$, $\corr{ \mathcal{R} }(R) = R$ and for $\mathcal{C}
-\in \clauses{ \ctxs{R} }$ and $R \in \ctxs{R}$ such that
-$(\mathcal{C}, R) \in \ctxs{R}$, $\corr{\mathcal{R}}(\mathcal{C}, R) =
-\corr{ \mathcal{R} }(R)$.
-
-% give the function that copies instances of a relational predicate
-\subsubsection{Implementation of $\copyrel{ \mathcal{R} }$}
-%
-$\copyrel{\mathcal{R}}$, given $\mathcal{S} \in \chcs{ \mathcal{R} }$,
-$\mathcal{C} \in \mathcal{S}$, and $R \in \ctxs{ \mathcal{R} }$,
-returns the following CHC system.
-% clause: C, but the application now points to (C, A)
-Let $\mathcal{C}' \in \clauses{ \ctxs{ \mathcal{R} } }$ be
-$\mathcal{C}$, with $R$ replaced with $(\mathcal{C}, R)$ in the body
-of $\mathcal{C}$.
-%
-Let $\mathcal{S}_{ \mathcal{C} } \subseteq \clauses{ \ctxs{
-    \mathcal{R} } }$ be the clauses in $\mathcal{S}$ for which $R$ is
-the relational predicate of the clause head, and let
-$\mathcal{S}_{\mathcal{C}}'$ be each clause in $\mathcal{S}_{
-  \mathcal{C} }$, with $R$ in the clause head replaced with
-$(\mathcal{C}, R)$.
-%
-Then $\copyrel{ \mathcal{R} }$ returns $\add{ (\remove{ \mathcal{S} }{
-    \add{ \mathcal{S}_{\mathcal{C}} }{\mathcal{C} } }) }{ \mathcal{C}'
-} \union \mathcal{S}_{\mathcal{C}}'$.
 %

### 2018-04-10 paper/approach.tex
parts edit for appendix

@@
 %
-The CHC system $\mcchc$ (\autoref{sec:solve-ex}) is not body-disjoint
-or linear, but is CDD.
-%
-A proof that both the classes of body-disjoint and linear systems are
-contained in the class of CDD systems is given in \autoref{app:char}.
+The proof is given in \autoref{app:char}.
 
@@
 %
-
 $\solvecdd$, given a CDD system $\cc{S}$, it first do a topological sort on all
@@
 \begin{ex}
-  $\solvecdd$, given $\mcchc$, may generate
-  interpretations of $\mcpreds$ in any topological ordering of the
+  $\solvecdd$, given a CDD system $\mcchc$ in \autoref{sec:running-ex}, may generate
+  interpretations queries in any topological ordering of the
   dependency relation of $\mcchc$.

### 2018-04-10 paper/background.tex
parts edit for appendix

@@
 \label{sec:background}
-
-% interpolants:
-\subsection{Logical interpolation}
-\label{sec:itps}
-%
-\sys solves CHC systems, logic-programming problems in which a
-solution maps predicate symbols to first-order formulas.
-%
-All objects defined in this paper are defined over a fixed space of
-first-order logical variables $X$.
-%
-In particular, for theory $\mathcal{T}$, the space of $\mathcal{T}$
-formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.
-%
-For each formula $\varphi \in \tformulas{\mathcal{T}}$, the set of
-variables that occur in $\varphi$ (i.e., the \emph{vocabulary} of
-$\varphi$) is denoted $\vocab(\varphi)$.
-% define models, satisfaction, entailment
-For formulas $\varphi_0, \ldots, \varphi_n, \varphi \in
-\tformulas{\mathcal{T}}$, the fact that $\varphi_0, \ldots, \varphi_n$
-\emph{entail} $\varphi$ is denoted $\varphi_0, \ldots, \varphi_n
-\entails \varphi$.
-
-% Define replacement and substitution:
-For sequences of variables $Y = [ y_0, \ldots, y_n ]$ and $Z = [ z_0,
-\ldots, z_n ]$, the $\mathcal{T}$ formula constraining the equality of
-each element in $Y$ with its corresponding element in $Z$, i.e., the
-formula $\bigland_{0 \leq i \leq n} y_i = z_i$, is denoted $Y = Z$.
-%
-The repeated replacement of variables $\varphi[ z_0 / y_0 \ldots z_{n}
-/ y_{n} ]$ is denoted $\replace{\varphi}{Y}{Z}$.
-%
-For each formula $\varphi$ defined over free variables $Y$,
-$\replace{\varphi}{Y}{Z}$ is denoted alternatively as $\varphi[ Z ]$.
-%
-The number of free variables in formula $\varphi$ is denoted
-$\degreeof{\varphi}$.
-
-% introduce interpolation:
-An interpolant of mutually inconsistent $\mathcal{T}$ formulas
-$\varphi_0$ and $\varphi_1$ is a $\mathcal{T}$ formula in their common
-vocabulary that explains their inconsistency.
-%
-\begin{defn}
-  \label{defn:itps}
-  % define:
-  For $\varphi_0, \varphi_1 \in \tformulas{\mathcal{T}}$, $I \in
-  \tformulas{\mathcal{T}}$ such that
-  %
-  \textbf{(1)} $\varphi_0 \entails I$, %
-  \textbf{(2)} $I, \varphi_1 \entails \false$, and %
-  \textbf{(3)} $\vocab(I) \subseteq \vocab(\varphi_0) \intersection
-  \vocab(\varphi_1)$,
-  %
-  $I$ is an \emph{interpolant} of $\varphi_0$ and $\varphi_1$.
-\end{defn}
-%
-All spaces of formulas in the remainder of this paper will be defined
-for a fixed, arbitrary theory $\mathcal{T}$ that supports
-interpolation, such as combinations of the theories of linear
-arithmetic and the theory of uninterpreted functions with equality.
-% introduce decision procedure:
-Although determining the satisfiability of formulas in such theories
-is NP-complete in general, decision procedures~\cite{moura08} and
-interpolating theorem provers~\cite{mcmillan04} for such theories have
-been proposed that operate on such formulas efficiently.
-%
-\sys is defined as using a decision procedure for $\mathcal{T}$ named
-\issat, and an interpolating theorem prover for $\mathcal{T}$ named
-$\solveitp$.
 
@@
 %
-A body consisting of a conjunctive set of uninterpreted relational predicates applied to varialbes and a formula
-that over variables.
+A body consisting of a conjunction of a formula that over variables and a vector of uninterpreted predicates applied to varialbes.
 %
-A head can either be an uninterpreted relational predicates applied to variables or \cc{False}.
+A head can either be an uninterpreted predicates applied to variables or \cc{False}.
 %
 A Clause where the head is \cc{False} is called query.
+\QZ{make it looks nice}
 \begin{defn}
@@
 
-% define system of CHC's
-We will present \sys as a solver for recursion-free CHCs.
 %
 For a given CHC system $\cc{S}$, there is no uninterpreted predicates $\cc{P} \in \predof{S}$ 
-such that $\cc{P} \in \tdepsof{P}$, then $\cc{S}$ is a recursion-free CHCs system.
+such that $\cc{P} \in \tdepsof{P}$, then $\cc{S}$ is a \emph{recursion-free} CHCs system.
+%
 
 %
-A solution to a CHC system $S$ is an interpretation of each
-uninterpreted predicate $\cc{Q} \in \predof{S}$ that each clause holds
-after substitute all predicates by its interpretation.
+A solution to a CHC system $S$ is a map that maps each each
+uninterpreted predicate $\cc{P} \in \predof{S}$ to its interpretation which is a formula.
+%
+All clauses in $S$ holds
+after substitute all uninterpreted predicates by its interpretation.
+
+% interpolants:
+\subsection{Logical interpolation}
+\label{sec:itps}
+%
+\sys solves CHC systems, logic-programming problems in which a
+solution maps predicate symbols to first-order formulas.
+%
+All objects defined in this paper are defined over a fixed space of
+first-order logical variables $X$.
+%
+In particular, for theory $\mathcal{T}$, the space of $\mathcal{T}$
+formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.
+%
+For each formula $\varphi \in \tformulas{\mathcal{T}}$, the set of
+variables that occur in $\varphi$ (i.e., the \emph{vocabulary} of
+$\varphi$) is denoted $\vocab(\varphi)$.
+% define models, satisfaction, entailment
+For formulas $\varphi_0, \ldots, \varphi_n, \varphi \in
+\tformulas{\mathcal{T}}$, the fact that $\varphi_0, \ldots, \varphi_n$
+\emph{entail} $\varphi$ is denoted $\varphi_0, \ldots, \varphi_n
+\entails \varphi$.
+
+% Define replacement and substitution:
+For sequences of variables $Y = [ y_0, \ldots, y_n ]$ and $Z = [ z_0,
+\ldots, z_n ]$, the $\mathcal{T}$ formula constraining the equality of
+each element in $Y$ with its corresponding element in $Z$, i.e., the
+formula $\bigland_{0 \leq i \leq n} y_i = z_i$, is denoted $Y = Z$.
+%
+The repeated replacement of variables $\varphi[ z_0 / y_0 \ldots z_{n}
+/ y_{n} ]$ is denoted $\replace{\varphi}{Y}{Z}$.
+%
+For each formula $\varphi$ defined over free variables $Y$,
+$\replace{\varphi}{Y}{Z}$ is denoted alternatively as $\varphi[ Z ]$.
+%
+The number of free variables in formula $\varphi$ is denoted
+$\degreeof{\varphi}$.
+
+% introduce interpolation:
+An interpolant of mutually inconsistent $\mathcal{T}$ formulas
+$\varphi_0$ and $\varphi_1$ is a $\mathcal{T}$ formula in their common
+vocabulary that explains their inconsistency.
+%
+\begin{defn}
+  \label{defn:itps}
+  % define:
+  For $\varphi_0, \varphi_1 \in \tformulas{\mathcal{T}}$, $I \in
+  \tformulas{\mathcal{T}}$ such that
+  %
+  \textbf{(1)} $\varphi_0 \entails I$, %
+  \textbf{(2)} $I, \varphi_1 \entails \false$, and %
+  \textbf{(3)} $\vocab(I) \subseteq \vocab(\varphi_0) \intersection
+  \vocab(\varphi_1)$,
+  %
+  $I$ is an \emph{interpolant} of $\varphi_0$ and $\varphi_1$.
+\end{defn}
+%
+All spaces of formulas in the remainder of this paper will be defined
+for a fixed, arbitrary theory $\mathcal{T}$ that supports
+interpolation, such as combinations of the theories of linear
+arithmetic and the theory of uninterpreted functions with equality.
+% introduce decision procedure:
+Although determining the satisfiability of formulas in such theories
+is NP-complete in general, decision procedures~\cite{moura08} and
+interpolating theorem provers~\cite{mcmillan04} for such theories have
+been proposed that operate on such formulas efficiently.
+%
+\sys is defined as using a decision procedure for $\mathcal{T}$ named
+\issat, and an interpolating theorem prover for $\mathcal{T}$ named
+$\solveitp$.
 

### 2018-04-10 paper/fig/hierarchy.tex

\scalebox{0.75}{%
\begin{tikzpicture}
\begin{scope}
  \node (top) at (0,0) {\begin{tabular}{c}All Systems\\ \textsc{Undecidable}\end{tabular}};
    \node[below = 0.5 of top] (rec) {\begin{tabular}{c}Recursion Free\\
  \textsc{co-Nexptime Complete}\end{tabular}};
  \node[below = 0.5 of rec] (dir) {\begin{tabular}{c}Directly Solvable\\ \textsc{co-NP}\end{tabular}};
  \node[label={0:\{\textsc{Shara}\}}, draw, rounded corners=0.5cm,
      below = 0.5 of dir] (cdd) {\begin{tabular}{c}Clause-Dependent-Disjoint\\ \textsc{co-NP}\end{tabular}};
  \node[label={0:\{\textsc{Ufo}\}},
      below left = 0.5 and 0.5 of cdd] (lin) {\begin{tabular}{c}Linear\\ \textsc{co-NP}\end{tabular}};
  \node[label={180:\{\textsc{Eldarica}\cite{rummer13b}\}},
      below right = 0.5 and 0.5 of cdd] (bd) {\begin{tabular}{c}Body Disjoint\\ \textsc{co-NP}\end{tabular}};
  \node[label={180:\{\textsc{Duality}\cite{bjorner13}\}},
      below = 0.5 of bd] (tree) {\begin{tabular}{c}Tree\\ \textsc{co-NP}\end{tabular}};
\end{scope}
\begin{scope}[every edge/.style={draw=black, very thick}]
  \path[-] (tree) edge (bd);
  \path[-] (bd) edge (cdd);
  \path[-] (lin) edge (cdd);
  \path[-] (cdd) edge (dir);
  \path[-] (dir) edge (rec);
  \path[-] (rec) edge (top);
\end{scope}
\end{tikzpicture}
}

### 2018-04-10 paper/overview.tex
added diagram
pass on overview

@@
 In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
-$\mcchc$, that formulates verifying the safety of the program
-\cc{multiCall} (\autoref{fig:multicall-code}).
+$\mcchc$, that formulates verifying the safety of a small program,
+\cc{dblAbs} (\autoref{fig:multicall-code}).
 %
@@
 %
-In \autoref{sec:not-in}, we illustrate that
-$\mcchc$ is not in the directly solvable classes introduced in
-previous work.
+In \autoref{sec:not-in}, we illustrate that $\mcchc$ is not in
+directly solvable classes introduced in previous work.
 %
@@
 % introduce running example program
-\subsection{Verifying \cc{multiCall}: an example hierarchical program}
+\subsection{Verifying \cc{dblAbs}: an example hierarchical program}
 \label{sec:running-ex}
 
-\QZ{remake the graph}
 % include code of running example and some figure:
@@
     \ffigbox[.26\textwidth] %
-    { \caption{\cc{multiCall}: an example hierarchical program.} %
+    { \caption{\cc{dblAbs}: an example hierarchical program.} %
       \label{fig:multicall-code} }
-    { \input{code/multiCall.java} }
+    { \input{code/dblAbs.java} }
     %
     \ffigbox[.7\textwidth] %
-    {\caption{Dependencies of CHC system $\mcchc$, depicted as a
-        hypergraph.
+    {\caption{$\mcchc$ shown as a directed hypergraph.
         % 
-        Each relational predicate is a node (circle) and each clause
-        is a hyperedge (box). } %
-      \label{fig:ex-deps} }
-    { \includegraphics[width=\linewidth]{fig/ex-deps.pdf} }
+        Each relational predicate is depicted as a graph node
+        while each clause is represented by a hyperedge.
+        Each hyperedge is labelled by the fact in the corresponding CHC\@.
+        Each node has a valid corresponding interpretation,
+        written in braces.
+      } %
+      \label{fig:ex-graph} }
+    % { \includegraphics[width=\linewidth]{fig/ex-deps.pdf} }
+      {\input{fig/example}}
   \end{floatrow}
@@
 %
-\cc{multiCall} includes a procedure \cc{db} that, given an integer
-variable \cc{m}, returns the value of \cc{m} doubled (line 1).
-%
-\cc{main}, the entry procedure of \cc{multiCall}, given integer \cc{n}
-(line 3), first binds variable \cc{abs} to $0$ (line 4).
-%
-\cc{main} tests if \cc{n} is greater than or equal to $0$ (line
-5).
-%
-If so, \cc{main} binds the value of \cc{n} to \cc{res} (line 6).
-%
-If not, \cc{main} binds the negation value of \cc{n} to \cc{res} (line 8).
-%
-\cc{main} then calles \cc{dbl} on \cc{abs} and binds the return value to \cc{res} (line 9)
-%
-\cc{main} then returns the value stored in \cc{res} (line 10).
-
+\cc{dblAbs} is a small procedure that doubles the absolute value of
+its input and stores the result in \cc{res}.
 % state the safety problem:
-The property to be verified in \cc{multiCall} is, for each input, 
-\cc{multiCall}
-returns a value greater than or equal to $0$.
+The program also asserts that \cc{res} is greater than or equal to $0$
+before exiting.
 % talk about solving Horn Clauses:
-Verifying that \cc{multiCall} satisfies such a property can be reduced
-to solving a recursion-free CHC system over a set of uninterpreted predicates
-that each represents a control location and
-procedure name in \cc{multiCall}.
+Verifying this assertion can be reduced to solving a recursion-free
+CHC system over a set of uninterpreted predicates that each represents
+a control location and procedure name in \cc{dblAbs}.
 %
@@
 %
-\QZ{left allign}
-\begin{gather}
+\begin{align}
 \label{chc}
   % semantic constraint of dbl:
-  \cc{db}(\cc{m}, \cc{d}) \gets \cc{d} = 2 * \cc{m} \\
+\cc{dbl}(\cc{x}, \cc{d}) &\gets \cc{d} = 2 * \cc{x} \\
   % semantic constraint of line 3:
-   \cc{L4}(\cc{n}, \cc{abs}) \gets  \cc{abs} = 0 \\
+  \cc{L}_4(\cc{n}, \cc{abs}) &\gets  \cc{abs} = 0 \\
   % semantic constraint of line 5 (then branch)
-   \cc{L6}(\cc{n}, \cc{abs}) \gets \cc{L4}(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
+  \cc{L}_6(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
   % semantic constraint of line 6 (else branch)
-   \cc{L8}(\cc{n}, \cc{abs}) \gets \cc{L4}(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
+  \cc{L}_8(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
   % assign branch
-   \cc{L9}(\cc{n}, \cc{abs'}) \gets \cc{L6}(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
-   \cc{L9}(\cc{n}, \cc{abs'}) \gets \cc{L8}(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
+  \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_6(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
+   \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_8(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
   % semantic constraint of main procedure:
-    \cc{main}(\cc{n},\cc{res}) \gets \cc{L9}(\cc{n}, \cc{abs'}) \land \cc{db}(\cc{m}, \cc{d})
-    \land \cc{abs'} = \cc{m} \land \cc{res} = \cc{d} \\
+   \cc{main}(\cc{n},\cc{res}) &\gets \cc{L}_9(\cc{n}, \cc{abs'}) \land
+                                     \cc{dbl}(\cc{x}, \cc{d})
+    \land \cc{abs'} = \cc{x} \land \cc{res} = \cc{d} \\
   %
-   False \gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
-\end{gather}
+    \bot &\gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
+\end{align}
 %
 While $\mcchc$ has been presented as the result
-of a translation from hierarchical program \cc{multiCall}, \sys is
+of a translation from hierarchical program \cc{dblAbs}, \sys is
 purely a solver for CHC systems: it does not require access to the
@@
 % solving the running example
-\subsection{\cc{multiCall} as a Clause-Dependent Disjoint system}
+\subsection{\cc{dblAbs} as a Clause-Dependent Disjoint system}
 \label{sec:solve-ex}
 % state it is CDD
-The recursion-free CHC system \cc{multiCall} is a \emph{Clause-Dependent-Disjoint (CDD)}) system.
+The recursion-free CHC system $\mcchc$ is a
+\emph{Clause-Dependent-Disjoint} (CDD) system.
 % state the defn of CDD:
-\emph{Clause-Dependent-Disjoint (CDD)}) system is an recursion-free CHC system such that
-for any uninterpreted predictes in the body of the same clause, their transitive dependencies
-are disjoint(i.e,they do not have common predicates).
+A CDD system is a recursion-free CHC system such that
+each clause is dependence disjoint.
+A clause is dependence disjoint if no two predicates in the body share
+any transitive dependencies (i.e, they have no common predicates) and
+there are no two predicates that are the same.
 %
@@
 % give a example
-In particular, for clause $(7)$, uninterpreted predicates \cc{L6} and \cc{db} shows in
-the body of same clause.
+In particular, clause $(7)$ uses both $\cc{L}_9$ and $\cc{dbl}$ in its
+body.
 %
-The transitive dependency of \cc{L6} is \{\cc{L6}\,\cc{L8}\,\cc{L4}\}, 
-and the transitive dependency of \cc{db} is $\varnothing$.
+The transitive dependency of $\cc{L}_9$ is the set
+$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\}$, while the transitive dependency of
+\cc{dbl} is the empty set.
 %
-Their transitive dependencies are disjoint, because \{\cc{L6}\,\cc{L8}\,\cc{L4}\} $\cap \varnothing = \varnothing$.
+Therefore, their transitive dependencies are disjoint:
+$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\} \cap \varnothing = \varnothing$.
+If there were a conflict in the transitive dependencies, then $\mcchc$
+would not be a CDD system.
 %
@@
 %
-And the other clauses in \cc{multiCall} only has one uninterpreted predicates in the body, so
-\cc{multiCall} is a CDD system.
+All other clauses in $\mcchc$ have at most one uninterpreted predicate
+in the body, so they are trivially disjoint dependent.
 %
-The fact that $\mcchc$ is CDD corresponds to the fact that there no statements executed more than
-once within one execution path.
+Since every rule in $\mcchc$ is disjoint dependent, $\mcchc$ is a CDD
+system.
+%
+The fact that $\mcchc$ is CDD corresponds to the fact that there no
+statements executed more than once within one execution path.
+\DAH{elaborate}
 %
@@
 % restate defn of CDD systems:
-\sys solves $\mcchc$ directly by issuing a binary 
-interpolation queriey for each uninterpreted predicates in $\mcchc$ in a topological order.
+\sys solves $\mcchc$ directly by issuing a binary interpolation query
+for each uninterpreted predicate in topological order.
 %
-For each uninterpreted predicate $P$ in $\mathcal{S}$, the solution of
-$P$ is the interpolant of two formulas: %
-\textbf{(1)} a \emph{pre}-formula, constructed as 
-the solution of each uninterpreted predicate in dependence of $P$, constraints
-in corresponding clauses and boolean indicator variable for 
-each uninterpreted predicates in $\mcchc$%
-\textbf{(2)} a \emph{post}-formula, constructed as  
-all constraints that occur after $P$ and boolean indicator variable for 
-each uninterpreted predicates in $\mcchc$.
+Each uninterpreted predicate $P$ has a valid interpretation which can
+be computed by finding the interpolant of two formulas:
+\textbf{(1)} the \emph{pre}-formula, constructed from the solutions of
+each uninterpreted predicate $P$ directly depends on, constraints in
+clauses where $P$ is the head, and boolean indicator variables for
+each uninterpreted predicate in $\mcchc$%
+\textbf{(2)} the \emph{post}-formula, constructed from
+all constraints that occur after $P$ and boolean indicator
+variables for each uninterpreted predicate in $\mcchc$.
 %
 
-% walk through the example
-E.g., when \sys solving $\mcchc$, \sys issues interpolation queries of 
-each uninterpreted predicates in the order \cc{L4}, \cc{L6}, \cc{L8}, 
-\cc{L9}, \cc{db} and \cc{main}.
+We will walk through the interpolation query for $\cc{L}_9$ by hand.
+By the time \sys reaches $\cc{L}_9$ it will already have solutions for
+$\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$. Possible interpretations of these
+predicates are shown in \autoref{fig:ex-graph}.
 %
-In particular, when \sys issues interpolation query for uninterpreted predicates \cc{L9},
-\sys has generated solutions for \cc{L4}, \cc{L6}, \cc{L8} which shows in\QZ{refer to graph}.
-%
-
-The solution of \cc{L9} is an interpolant of \textbf{(1)} 
-a \emph{pre}-formula, constructed as solutions generated for \cc{L4}, \cc{L6}
+\DAH{WORK FROM HERE}
+A solution for \cc{L9} can be constructed by interpolating \textbf{(1)}
+the \emph{pre}-formula, constructed as solutions generated for \cc{L4}, \cc{L6}
 the constraints in clauses (5) and (6) and boolean variables%
-\textbf{(2)} a \emph{post}-formula constructed from constraints in
-$\mcchc$ that all constrains after \cc{L9}, constrains for \cc{db}
+\textbf{(2)} the \emph{post}-formula constructed from constraints in
+$\mcchc$ that all constraints after \cc{L9}, constraints for \cc{dbl}
 and boolean variables.
@@
 \begin{gather}
-((b_{L8}\land abs'= -n) \lor (b_{L6} \land abs=n))\land (\neg b_{L6} \lor n \ge 0) \land (\neg b_{L8} \lor n<0)
+  ((b_{L_8}\land \cc{abs'}= -\cc{n})
+    \lor (b_{L_6} \land \cc{abs}=\cc{n}))
+  \land (\neg b_{L_6} \lor \cc{n} \ge 0)
+    \land (\neg b_{L_8} \lor \cc{n}<0)
 \end{gather}
@@
 \begin{gather}
-(b_{main} \land res' < 0 ) \land (\neg b_{main} \lor (b_{db} \land abs'= m \land res =d )) 
-\land (\neg b_{db} \lor d = 2*m)
+  (b_{main} \land \cc{res'} < 0 )
+    \land (\neg b_{main} \lor (b_{dbl} \land \cc{abs'}= \cc{x} \land \cc{res}=\cc{d} ))
+  \land (\neg b_{dbl} \lor \cc{d} = 2*\cc{x})
 \end{gather}
-, $b_{main}$ and $b_{db}$ are boolean variables corresponding to predicates \cc{main} and \cc{db}.
+, $b_{main}$ and $b_{dbl}$ are boolean variables corresponding to
+predicates \cc{main} and \cc{dbl}.
 %
@@
 
-\subsection{\cc{multiCall} is not in other system}
+\subsection{\cc{dblAbs} is not in other system}
 \label{sec:not-in}
@@
 $\mcchc$ is not a linear system because the body of clause (7) has two
-predicates \cc{L9} and \cc{db}.
+predicates \cc{L9} and \cc{dbl}.
 %
@@
 %
-E.g., a recursion-free system that models \cc{multiCall} 
+E.g., a recursion-free system that models \cc{dblAbs} 
 call \cc{dbl} twice on \cc{abs} in sequence would not

### 2018-04-11 paper/overview.tex
more work on overview

@@
 % talk about solving Horn Clauses:
-Verifying this assertion can be reduced to solving a recursion-free
-CHC system over a set of uninterpreted predicates that each represents
-a control location and procedure name in \cc{dblAbs}.
+Verifying this assertion reduces to solving a recursion-free
+CHC system over a set of uninterpreted predicates that represent
+the control locations in \cc{dblAbs}.
 %
@@
 % solving the running example
-\subsection{\cc{dblAbs} as a Clause-Dependent Disjoint system}
+\subsection{$\mcchc$ as a Clause-Dependent Disjoint System}
 \label{sec:solve-ex}
@@
 A CDD system is a recursion-free CHC system such that
-each clause is dependence disjoint.
-A clause is dependence disjoint if no two predicates in the body share
-any transitive dependencies (i.e, they have no common predicates) and
-there are no two predicates that are the same.
+the dependencies between predicates in the system obey certain rules.
 %
-The formal definition of CDD and its key properties are given in
-\autoref{sec:CDD-defn}.
-
+One predicate is directly dependent on another if the second appears
+in the body of a clause where the first is the head.
+%
+Based on this, we say that a clause is dependence disjoint if
+\textbf{(1)} no two predicates in the body share any transitive
+dependencies on other predicates and \textbf{(2)} no two predicates in
+the body are the same.
 % give a example
-In particular, clause $(7)$ uses both $\cc{L}_9$ and $\cc{dbl}$ in its
-body.
+As an example, clause $(7)$ is dependence disjoint. Clause $(7)$ uses
+both $\cc{L}_9$ and $\cc{dbl}$ in its body.
 %
@@
 $\{\cc{L}_4,\cc{L}_6,\cc{L}_8\} \cap \varnothing = \varnothing$.
-If there were a conflict in the transitive dependencies, then $\mcchc$
-would not be a CDD system.
-%
-The formal definition of transitive dependency is given in\QZ{reference}.
 %
@@
 %
-Since every rule in $\mcchc$ is disjoint dependent, $\mcchc$ is a CDD
-system.
+When each clause in a CHC system is dependence disjoint, it is a CDD
+system. Therefore $\mcchc$ is a CDD system.
 %
@@
 %
+The formal definition of CDD and its key properties are given in
+\autoref{sec:CDD-defn}.
+The formal definition of transitive dependency is given in\QZ{reference}.
+%
 
 % restate defn of CDD systems:
-\sys solves $\mcchc$ directly by issuing a binary interpolation query
-for each uninterpreted predicate in topological order.
+\sys solves CDD systems directly by issuing a binary interpolation
+query for each uninterpreted predicate in topological order.
 %
 Each uninterpreted predicate $P$ has a valid interpretation which can
-be computed by finding the interpolant of two formulas:
+be computed by interpolating
 \textbf{(1)} the \emph{pre}-formula, constructed from the solutions of
-each uninterpreted predicate $P$ directly depends on, constraints in
-clauses where $P$ is the head, and boolean indicator variables for
-each uninterpreted predicate in $\mcchc$%
+each uninterpreted predicate $P$ directly depends on and constraints
+in clauses where $P$ is the head and
 \textbf{(2)} the \emph{post}-formula, constructed from
-all constraints that occur after $P$ and boolean indicator
-variables for each uninterpreted predicate in $\mcchc$.
+all clauses where the head transitively depends on $P$.
 %
 
-We will walk through the interpolation query for $\cc{L}_9$ by hand.
-By the time \sys reaches $\cc{L}_9$ it will already have solutions for
-$\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$. Possible interpretations of these
-predicates are shown in \autoref{fig:ex-graph}.
+We will walk through the construction of an interpolation query for
+$\cc{L}_9$ by hand.  By the time \sys reaches $\cc{L}_9$ it will
+already have solutions for $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$.
+Possible interpretations of these predicates are shown in
+\autoref{fig:ex-graph}.
 %
-\DAH{WORK FROM HERE}
 A solution for \cc{L9} can be constructed by interpolating \textbf{(1)}
@@
 
-\subsection{\cc{dblAbs} is not in other system}
+\subsection{$\mcchc$ is not in other system}
 \label{sec:not-in}

### 2018-04-11 paper/abstract.tex
pass on first half of paper

@@
   Recursion-free Constrained Horn Clauses (CHCs) are logic-programming
-  problems that formulate verifying safety of programs with bounded
+  problems that can model safety properties of programs with bounded
   iteration and recursion.
   %
-  They both formulate bounded verification problems and are generated
-  by solvers that attempt to verify safety of unbounded iterative or
-  recursive systems.
-  %
-  Efficient solvers of recursion-free systems reduce the problem to
-  solving a series of systems that can each be solved tractably.
+  In addition, many recursive CHC system solvers reduce inputs to a
+  series of recursion-free CHC systems that can each be solved
+  efficiently.
 
   % contribution:
-  In this paper, we define a novel tractable class of recursion-free
-  systems, named \emph{Clause-Dependence Disjoint (CDD)} systems, that
-  strictly generalizes tractable classes defined in previous work.
+  In this paper, we define a novel class of recursion-free systems,
+  named \emph{Clause-Dependence Disjoint} (CDD), that
+  generalizes classes defined in previous work.
   %
-  We implemented a novel solver for recursion-free CHCs, named \sys,
-  that reduces the problem of solving a given recursion-free system to
-  solving an equivalent CDD system.
+  The advantage of this class is that many systems are significantly
+  smaller than equivalent systems in other classes which allows them
+  to be solved more efficiently.
   %
-  Our evaluation of \sys on standard benchmarks indicates that, in
-  many practical cases, it performs significantly better than a
-  state-of-the-art implementation.
+  Based on the class of CDD systems, we implemented a CHC solver named
+  \sys.  \sys~solves arbitrary CHC systems by reducing the input to a
+  series of CDD systems.
+  %
+  Our evaluation indicates that \sys~outperforms state-of-the-art
+  implementations in many practical cases.
 \end{abstract}

### 2018-04-11 paper/introduction.tex
pass on first half of paper

@@
 % context, problem defn:
-Many critical problems in program verification, such as the
-verification of safety properties of recursive programs and
-concurrent programs, can be reduced to solving systems of
-\emph{Constrained Horn Clauses} (CHCs), a class of logic-programming
+Many critical problems in program verification can be reduced to
+solving systems of Constrained Horn Clauses (CHCs), a class of
+logic-programming
 problems~\cite{bjorner13,flanagan03,rummer13a,rummer13b}.
-
-% general introduce chc system
-A CHC consists of a body of applications of uninterpreted relational
-predicates, a constraint in a first-order theory, and a head
-application of a relational-predicate.
 %
-A CHC system consists of a set of CHCs and a query relational
-predicate.
+A CHC is a logical implication where the antecedent is a conjunction
+of uninterpreted relational predicates with a fact in a background
+theory and the consequent is an uninterpreted relational predicate. In
+the context of a CHC we call the antecedent the body and the
+consequent the head.  A CHC system is a set of CHCs.
 %
-A solution of a system is an interpretation of each relational
-predicate as a formula such that each clause body with predicate
-symbols substituted with their interpretation entails the interpreted
-clause head.
-\QZ{This seems really out of place -- DAH}
+The goal of a CHC solving problem is to find suitable interpretations
+for each relational predicate such that each CHC is logically
+consistent in isolation.
 
@@
 derivation of each relational predicate will contain no references to
-itself. \QZ{\DAH{I still think we need a small example here}} I.e., no
-relational predicate is \emph{dependent} on itself.
+itself. \QZ{\DAH{I still think we need a small example here}}
 %
@@
 First, recursion-free systems can be used to model safety properties
-for hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12},
-i.e., programs with control branches and procedure calls, but only
-bounded iteration and recursion.
+for hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12}
+(programs with only bounded iteration and recursion).
 %
-Second (and most importantly), the task of solving a general CHC
+Second and most importantly, the task of solving a general CHC
 system can reduce to the task of solving a sequence of recursion-free
-systems.
-%
+systems \QZ{reference}.
+
 Many \QZ{more references? Or remove many.} modern solvers generate
-such a sequence where each subsystem is a bounded unwinding of the
-input system. These solvers attempt to show that the solution to a
-given subsystem generalizes to the original system or else generate a
+a sequence such that each subsystem is a bounded unwinding of the
+input system. These solvers either attempt to show that the solution to a
+given subsystem generalizes to the original system or generate a
 larger subsystem for the next iteration~\cite{bjorner13}.
 %
-Such solvers are heavily reliant on the performance of the underlying
-recursion-free solver since they invoke this underlying solver in an
-iterative loop.
+Since such solvers inspect recursion-free systems in an iterative
+loop, they are heavily reliant on the performance of the underlying
+recursion-free solver.
 %
@@
 Typically, even recursion-free CHC systems are not solved directly.
-Instead, they are reduced to some even more specific subclass of CHC
-system. For example, some approaches generate CHC systems whose
-derivations are trees~\cite{bjorner13}.
-\QZ{In general it seems we need many more references in the intro.}
+Instead, they are reduced to a more specific subclass of
+recursion-free CHC system.
 %
-A solver can dispatch such a system by issuing interpolation queries
-to find a suitable definition for each relational predicate.
+These classes include those of
+\emph{body disjoint} (or \emph{tree structure})
+systems~\cite{heizmann10,bjorner13,mcmillan14,rummer13a,rummer13b} and
+of \emph{linear} systems~\cite{albarghouthi12a}.
 %
-We refer to the subclass of CHC systems that can be solved by such
-techniques as \emph{directly solvable}.
+We will discuss these classes in \autoref{sec:overview} and
+\autoref{sec:related-work}.
+%
+Such classes have proved practical because a solver can find
+suitable definitions for relational predicates by issuing
+\emph{interpolation queries}.
+%
+When a class of CHC systems can be solved by interpolation queries, we
+say the class is \emph{directly solvable}.
 
@@
 In general, solving a recursion-free CHC system whose facts are
-restricted to the theory of integer linear arithmetic is
+restricted to the theory of linear integer arithmetic is
 co-NEXPTIME-complete~\cite{rummer13b}.
 %
-In contrast, solving a directly solvable system in the same theories
-is co-NP~~\cite{rummer13b}.
+In contrast, solving a directly solvable system with the same theories
+is in co-NP~\cite{rummer13b}.
 %
 Therefore, the performance of a CHC solver is highly reliant on the
-order (i.e the number of predicates) of the directly solvable systems
-it produces.
-
-% current subclass
-Previous work has introduced three classes of directly solvable
-systems,
-% derivation trees:
-\textbf{(1)} tree systems~\cite{heizmann10,bjorner13,mcmillan14}, %
-% disjunctive trees:
-\textbf{(2)} body-disjoint systems, which represent multiple
-derivation trees with a single disjunctive tree
-system~\cite{rummer13a,rummer13b}, and
-% linear systems:
-\textbf{(3)} linear CHC systems, which can compactly represent
-multiple linear derivations with a single system structured as a
-Directed Acyclic Graph (DAG)~\cite{albarghouthi12a}.
-%
-The class of body-disjoint systems strictly contains the class of
-derivation trees, and is independent of the class of linear systems
-(i.e., the classes overlap, and neither contains the
-other)~\cite{rummer13a,rummer13b}.
+size of the directly solvable systems it produces.
 
@@
 The first contribution of this paper is the introduction of a novel
-class of directly solvable systems that is a superset of both
-the body-disjoint subclass and the linear subclass.
+class of directly solvable systems that we refer to as
+\emph{Clause-Dependence Disjoint} (CDD).
 %
-In particular, the class consists of all recursion-free CHC systems
-such that, for each clause $C$, the dependences of all distinct
-relational predicates in the body of $C$ are disjoint \DAH{\QZ{Should
-we illustrate what disjoin means here?}};
+The class consists of all recursion-free CHC systems such that each
+clause in the system has at most one dependency on the same relational
+predicate.
 %
-we thus refer to such systems as \emph{Clause-Dependent Disjoint
-(CDD)} systems.
-%
-We demonstrate that the class of CDD systems is directly solvable
-through a symbolic algorithm.
-%
+The key characteristic of this class is that CDD systems are often
+smaller than equivalent systems in other directly solvable classes,
+and therefore solving recursion free systems by rewriting them in CDD
+form is often less computationally expensive.
 
@@
 %
-In the worst case, it is possible that the order of $S'$ may be
-exponential in the order of $S$.
+In the worst case, it is possible that the size of $S'$ may be
+exponential in the size of $S$.
 %
-However, empirically we have found that the order of $S'$ is usually
-close enough to the order of $S$ that \sys often outperforms the best
+However, empirically we have found that the size of $S'$ is usually
+close enough to the size of $S$ that \sys often outperforms the best
 known CHC solvers.
 %
-Because the class of CDD systems is a superset of both body-disjoint
-and linear systems, the order of an equivalent CDD system is always
-less than or equal to that of systems in these subclasses.
+The procedure implemented in \sys is a generalization of existing
+techniques that synthesize compact verification conditions for
+hierarchical programs~\cite{flanagan01,lal-qadeer15}.
 %
-The procedure implemented in \sys to construct $S'$ is a
-generalization of existing procedures for generating compact
-verification conditions of hierarchical
-programs~\cite{flanagan01,lal-qadeer15}.
-%
-Given a general (possibly recursive) CHC system, \sys solves a sequence of
-recursion-free systems.
+Given a general (possibly recursive) CHC system, \sys solves a
+sequence of recursion-free systems.
 %
@@
 attempts to combine the solutions of these recursion-free systems to
-synthesize a soltuion of the original solution, as has been proposed
+synthesize a soltuion to the original problem, as has been proposed
 in previous work~\cite{rummer13b}.
@@
 % experience:
-\QZ{need to change if need}
 We implemented \sys within the \duality CHC solver~\cite{bjorner13},
@@
 %
-The results indicate that in a strong majority of practical cases,
+The results indicate that in a strong majority of cases,
 \sys performs better than modern solvers.

### 2018-04-11 paper/overview.tex
pass on first half of paper

@@
 In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
-$\mcchc$, that formulates verifying the safety of a small program,
+$\mcchc$, that models the safety of a small program,
 \cc{dblAbs} (\autoref{fig:multicall-code}).
 %
-In \autoref{sec:solve-ex}, we show that $\mcchc$ is a
-\emph{Clause-Dependent Disjoint (CDD)} system and how \sys can solve it by
-encoding it into binary interpolants.
+In \autoref{sec:solve-ex}, we show that $\mcchc$ is a CDD system and
+how \sys can solve it by encoding it into binary interpolants.
 %
@@
     \ffigbox[.7\textwidth] %
-    {\caption{$\mcchc$ shown as a directed hypergraph.
+    {\caption{$\mcchc$ as a directed hypergraph.
         % 
@@
 While $\mcchc$ has been presented as the result
-of a translation from hierarchical program \cc{dblAbs}, \sys is
+of a translation from \cc{dblAbs}, \sys is
 purely a solver for CHC systems: it does not require access to the
@@
 % solving the running example
-\subsection{$\mcchc$ as a Clause-Dependent Disjoint System}
+\subsection{$\mcchc$ as a Clause-Dependence Disjoint System}
 \label{sec:solve-ex}
@@
 The recursion-free CHC system $\mcchc$ is a
-\emph{Clause-Dependent-Disjoint} (CDD) system.
+\emph{Clause-Dependence-Disjoint} (CDD) system.
 % state the defn of CDD:
-A CDD system is a recursion-free CHC system such that
-the dependencies between predicates in the system obey certain rules.
+A CHC system can be classified as CDD when the clauses obey
+rules based on dependencies between the predicates.
 %
@@
 The transitive dependency of $\cc{L}_9$ is the set
-$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\}$, while the transitive dependency of
+$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\}$ while the transitive dependency of
 \cc{dbl} is the empty set.
@@
 %
-The fact that $\mcchc$ is CDD corresponds to the fact that there no
-statements executed more than once within one execution path.
-\DAH{elaborate}
 %
@@
 be computed by interpolating
-\textbf{(1)} the \emph{pre}-formula, constructed from the solutions of
-each uninterpreted predicate $P$ directly depends on and constraints
-in clauses where $P$ is the head and
+\textbf{(1)} the \emph{pre}-formula, constructed from clauses where
+$P$ is the head and
 \textbf{(2)} the \emph{post}-formula, constructed from
@@
 %
-A solution for \cc{L9} can be constructed by interpolating \textbf{(1)}
-the \emph{pre}-formula, constructed as solutions generated for \cc{L4}, \cc{L6}
-the constraints in clauses (5) and (6) and boolean variables%
-\textbf{(2)} the \emph{post}-formula constructed from constraints in
-$\mcchc$ that all constraints after \cc{L9}, constraints for \cc{dbl}
-and boolean variables.
+A solution for $\cc{L}_9$ can be constructed by interpolating
+the appropriate pre and post formulas.
 %
-In particular, the \emph{pre}-formula for \cc{L9} is 
+The pre-formula is constructed from the bodies of clauses where
+$\cc{L}_9$ is the head. Each relational
+predicate, $P$,
+is replaced by a corresponding boolean indicator variable, $\cc{b}_P$, and each
+boolean indicator variable implies the solution for its predicate.
+%
+In particular, the pre-formula for $\cc{L}_9$ is constructed from
+clauses (5) and (6):
 \begin{gather}
-  ((b_{L_8}\land \cc{abs'}= -\cc{n})
-    \lor (b_{L_6} \land \cc{abs}=\cc{n}))
-  \land (\neg b_{L_6} \lor \cc{n} \ge 0)
-    \land (\neg b_{L_8} \lor \cc{n}<0)
+  ((\cc{b}_{\cc{L}_6} \land \cc{abs'} = \cc{n})
+  \lor
+  (\cc{b}_{\cc{L}_8} \land \cc{abs'} = -\cc{n}))
+  \land
+  (\cc{b}_{\cc{L}_6} \Rightarrow \cc{n} \ge 0)
+  \land
+  (\cc{b}_{\cc{L}_8} \Rightarrow \cc{n} < 0)
 \end{gather}
 %
-, $b_{L8}$ and $b_{L6}$ are boolean variables corresponding to predicates \cc{L8} and \cc{L6}.
+The post-formula is constructed from clauses that transitively depend
+on $\cc{L}_9$. Again, we replace relational predicates by
+corresponding boolean indicators. However, we omit the boolean indicator for
+$\cc{L}_9$. The post-formula is composed from clauses (1), (7), and
+(8):
+\begin{gather}
+  (\cc{b}_{\cc{main}} \land \cc{res'} < 0)
+  \land
+  (\cc{b}_{\cc{main}} \Rightarrow
+    (\cc{b}_{\cc{dbl}}
+    \land \cc{abs'}= \cc{x}
+    \land \cc{res}=\cc{d} ))
+  \land (\cc{b}_{\cc{dbl}} \Rightarrow \cc{d} = 2*\cc{x})
+\end{gather}
 %
-The \emph{post}-formula for \cc{L9} is
-\begin{gather}
-  (b_{main} \land \cc{res'} < 0 )
-    \land (\neg b_{main} \lor (b_{dbl} \land \cc{abs'}= \cc{x} \land \cc{res}=\cc{d} ))
-  \land (\neg b_{dbl} \lor \cc{d} = 2*\cc{x})
-\end{gather}
-, $b_{main}$ and $b_{dbl}$ are boolean variables corresponding to
-predicates \cc{main} and \cc{dbl}.
-%
-A detailed and formal description for the procedure of solving a CDD system is given in
-\autoref{sec:solve-cdd}.
+Interpolating the pre and post formulas yields a valid intepretation
+of $\cc{L}_9$: $\cc{abs'} \geq 0$.
+The procedure for solving a CDD system is described in formal detail
+in \autoref{sec:solve-cdd}.
 
-\subsection{$\mcchc$ is not in other system}
+\subsection{$\mcchc$ is not in other recursion-free classes}
 \label{sec:not-in}
 %
-There are three classes of directly solvable systems has been introduced
-in previous work(as described in \autoref{sec:intro}).
-%
-However, $\mcchc$ does not belong to any of these three classes.
+In this section, we show that $\mcchc$ cannot be placed in other
+known classes of recursion-free CHC systems. Specifically, we will
+discuss body-disjoint systems and linear systems.
 %
@@
 % tree solver
-In particular, previous work has introduced tree structure systems\cite{bjorner13,heizmann10},
-which each uninterpreted predicate in the system is the head of at most one clause.
+Tree structure
+systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
+class of recursion-free CHC system where each uninterpreted predicate
+appears in the body of at most one clause (and appears in such a
+clause exactly once).
 %
-Such systems do not include $\mcchc$, because predicate \cc{L9} is the head 
-of clause (5) and (6).
-%
-CHC solvers based on tree structure system has to reduce solving $\mcchc$ to
-solving two tree strucutre systems that contains by $\mcchc$, that one corresponds
-to the tree branch and the other corresponds to the false branch.
-% not that bad 
-There are works has been proposed which in some cases enumerating 
-and solving all derivation trees 
-of a given recursion-free
-system~\cite{mcmillan14}.
-%
-However, such solvers may enumerate all derivations tree of a given system
-in the worst case.
-
-% Body-Disjoint:
-Previous work has introduced \emph{body-disjoint}
-systems~\cite{rummer13a,rummer13b}, 
-which each uninterpreted predicate shows in the body of at most one
-clause, and shows in the body of such a clause at most once;
-%
-However, such systems cannot model a program with multiple
-control paths that share a common subpath, typically modeled as a CHC
-system with a uninterpreted predicate that occurs in the body of multiple clauses.
+Such systems cannot model a program with multiple control paths that
+share a common subpath, typically modeled as a CHC system with a
+uninterpreted predicate that occurs in the body of multiple clauses.
 %
 $\mcchc$ is not a body-disjoint system because the uninterpreted
-predicates \cc{L4} shows in the body of both clauses (3) and (4).
+predicates $\cc{L}_4$ shows in the body of both clause (3) and clause
+(4).
 %
-Such a solver, given $\mcchc$, generates a body-disjoint system
-consisting of two copies of the sub-systems with head relational
-predicates \cc{L4}.
+In order to find an interpretation for $\mcchc$, a solver which uses
+tree-structure systems would have to create a fresh copy of
+$\cc{L}_4$, as well copies for all relational predicates that
+$\cc{L}_4$ depends on.
 
@@
 Previous work also introduced linear system~\cite{albarghouthi12a},
-which each clause's body has at most one uninterpreted predicate.
+where each clause's body has at most one uninterpreted predicate.
 %
@@
 %
-CHC solvers that solve linear systems, $\mcchc$, effectively inline constraints on
-relational predicates that occur in non-linear clauses~\cite{albarghouthi12b}.
+CHC solvers that solve linear systems effectively inline the
+constraints for relational predicates that occur in non-linear
+clauses~\cite{albarghouthi12b}.
 %
-Such approaches can, in general, generate an inlined system that is
-exponentially larger than a given system.
+In the case of $\mcchc$, inlining the constraints of $\cc{dbl}$ is
+efficient, but in general such approaches can generate systems that
+are exponentially larger than the input.
 
 %% talk about expansion:
-The class of CDD systems strictly contains these three classes, 
-but does not contain all
-recursion-free systems.
+\begin{figure}
+  \input{fig/hierarchy}
+\end{figure}
+
+The class CDD of CDD systems is a superclass of the classes of
+body-disjoint and linear systems.
+%
+% The fact that $\mcchc$ is CDD corresponds to the fact that there no
+% statements executed more than once within one execution path.
+% \DAH{elaborate}
 %

### 2018-04-11 paper/background.tex
updates

@@
 % definition of CHC
-A Constrained Horn Clause contains a body and a head.
+As stated in \autoref{sec:intro} a Constrained Horn Clause is a
+logical implication where the antecedent is called the body and the
+consequent is called the head.
 %
-A body consisting of a conjunction of a formula that over variables and a vector of uninterpreted predicates applied to varialbes.
+The body is a conjunction of a logical formula, called the fact,
+and a vector of uninterpreted predicates. The fact is an arbitrary
+formula in some background logic, such as linear integer arithmetic.
 %
-A head can either be an uninterpreted predicates applied to variables or \cc{False}.
+The uninterpreted predicates are applied to variables which may or may
+not appear in the fact.
+%
+A head can be either an uninterpreted predicate applied to variables or \cc{False}.
 %
 A Clause where the head is \cc{False} is called query.
-\QZ{make it looks nice}
 \begin{defn}
-  A Constrained Horn Clause is definied as follows: \\
-  $\cc{chc} \Coloneqq \cc{head} \gets \cc{body}$ \\
-  $\cc{head} \Coloneqq \cc{False} \mid \cc{pred}$  \\
-  $\cc{body} \Coloneqq \varphi \wedge \cc{preds}$ \\
-  $\cc{preds} \Coloneqq \cc{True} \mid \cc{pred} \wedge \cc{preds}$ \\
-  $\cc{pred} \Coloneqq$ an uninterpreted predicate applied to variables \\
-  $\varphi \Coloneqq$ a formula that all variables are vocabulary \\
+  CHCs can be defined structurally:
+\begin{align*}
+  \cc{chc} &\Coloneqq \cc{head} \gets \cc{body} \\
+  \cc{head} &\Coloneqq \cc{False} \\&\mid \cc{pred}  \\
+  \cc{body} &\Coloneqq \varphi \wedge \cc{preds} \\
+  \cc{preds} &\Coloneqq \cc{True} \\&\mid \cc{pred} \wedge \cc{preds} \\
+  \cc{pred} &\Coloneqq \textsf{an uninterpreted predicate applied to variables} \\
+  \varphi &\Coloneqq \textsf{a formula that all variables are vocabulary} \\
+\end{align*}
 \end{defn}
@@
 % define a body function, head function, and recursion-free
-For a given constrained horn clause $C$, $\bodyof{C}$ denotes
-the vector of uninterpreted predicates in the body of $C$, $\consof{C}$
-denotes the formula in the body.
+For a given CHC $C$, $\bodyof{C}$ denotes the vector of uninterpreted
+predicates in the body of $C$ and $\consof{C}$ denotes the fact in the
+body.
 %
-It is possible for a uninterpreted predicate shows multiple times in the
-body of $C$.
+It is possible for an uninterpreted predicate to apper multiple times
+in the body of $C$.
 %
-For a given constrained horn clause $C$ that is not a query,
-$\headof{c}$ denotes the uninterpreted predicates of the head of $C$.
+If $C$ is not a query, then $\headof{C}$ denotes the uninterpreted
+predicate in the head.
 %
@@
 %define a CHC system
-A Constrained Horn Clauses system is a set of constrained horn clauses 
-that it has one and only one query.
+A CHC system is a set of CHCs where exactly one clause is a query
 %
-For a given Constrained Horn Clauses system $S$,$\predof{S}$ denotes 
-the set of all uninterpreted predicates in this CHC system.
+For a given CHC system $S$, $\predof{S}$ denotes the set of all
+uninterpreted predicates.
 

### 2018-04-11 paper/fig/hierarchy.tex
updates

@@
-\scalebox{0.75}{%
+% \scalebox{0.75}{%
+% }
+
 \begin{tikzpicture}
-\begin{scope}
-  \node (top) at (0,0) {\begin{tabular}{c}All Systems\\ \textsc{Undecidable}\end{tabular}};
-    \node[below = 0.5 of top] (rec) {\begin{tabular}{c}Recursion Free\\
-  \textsc{co-Nexptime Complete}\end{tabular}};
-  \node[below = 0.5 of rec] (dir) {\begin{tabular}{c}Directly Solvable\\ \textsc{co-NP}\end{tabular}};
-  \node[label={0:\{\textsc{Shara}\}}, draw, rounded corners=0.5cm,
-      below = 0.5 of dir] (cdd) {\begin{tabular}{c}Clause-Dependent-Disjoint\\ \textsc{co-NP}\end{tabular}};
-  \node[label={0:\{\textsc{Ufo}\}},
-      below left = 0.5 and 0.5 of cdd] (lin) {\begin{tabular}{c}Linear\\ \textsc{co-NP}\end{tabular}};
-  \node[label={180:\{\textsc{Eldarica}\cite{rummer13b}\}},
-      below right = 0.5 and 0.5 of cdd] (bd) {\begin{tabular}{c}Body Disjoint\\ \textsc{co-NP}\end{tabular}};
-  \node[label={180:\{\textsc{Duality}\cite{bjorner13}\}},
-      below = 0.5 of bd] (tree) {\begin{tabular}{c}Tree\\ \textsc{co-NP}\end{tabular}};
-\end{scope}
-\begin{scope}[every edge/.style={draw=black, very thick}]
-  \path[-] (tree) edge (bd);
-  \path[-] (bd) edge (cdd);
-  \path[-] (lin) edge (cdd);
-  \path[-] (cdd) edge (dir);
-  \path[-] (dir) edge (rec);
-  \path[-] (rec) edge (top);
+\tikzstyle{every node}=[draw=black, thin, ellipse]
+  \begin{scope}[fill opacity=0.1,text opacity=1]
+
+  \draw node[minimum width=6.2cm,minimum height=6cm,
+    label={[anchor=north, inner sep=5pt]north:
+      \begin{tabular}{c}Recursion-Free\\\textsc{co-NEXPTIME Complete}\end{tabular}} ]
+    at (0,0)
+    {};
+
+  \draw node[fill=blue, minimum width=5.5cm,minimum height=4.5cm,
+    label={[anchor=south, inner sep=0pt]south:
+      \begin{tabular}{c}Directly Solvable\\\textsc{co-NP}\end{tabular}} ]
+    at (0,-6mm)
+    {};
+
+  \draw node[fill=yellow, minimum width=5cm,minimum height=3cm,
+    label={[anchor=north, inner sep=5pt]north:CDD}]
+    at (0,-1mm)
+    {};
+
+  \draw node[fill=red, minimum width=2.7cm,minimum height=2cm]
+    at (10mm,-2mm)
+    {\begin{tabular}{c}Body\\Disjoint\end{tabular}};
+
+  \draw node[fill=green, minimum width=2.7cm,minimum height=2cm]
+    at (-10mm,-2mm)
+    {Linear};
+
+    % \begin{tabular}{c}Directly Solvable\\\textsc{co-NP}\end{tabular}} ]
   \end{scope}
 \end{tikzpicture}
-}

### 2018-04-11 paper/appendix.tex

@@
 % CDD systems:
-$\expand{\mathcal{R}}$ always returns a CDD expansion of its input
+$\expand$ always returns a CDD expansion of its input
 (see \autoref{app:corr}, \autoref{lem:expand-corr}) that is minimal.
 % discussion: there are variations:
-$\expand{ \mathcal{R} }$ is certainly not unique as an algorithm for
+$\expand$ is certainly not unique as an algorithm for
 generating a minimal CDD expansion.
 %
-In particular, feasible variations of $\expand{ \mathcal{R} }$ can be
+In particular, feasible variations of $\expand$ can be
 generated from different implementations of $\sharingclause$, each of
@@
   % contains linear systems:
-  For goal \textbf{(1)}, let $\mathcal{S}$ be an arbitrary linear
+  For goal \textbf{(1)}, let $S$ be an arbitrary linear
   system.
   %
-  $\mathcal{S}$ is CDD if for each clause $\mathcal{C}$ in
-  $\mathcal{S}$, for each pair of distinct relational predicates in
-  the body of $\mathcal{C}$, the transitive dependence are disjoint
+  $S$ is CDD if for each clause $C$ in
+  $S$, each pair of distinct predicates in
+  the body of $C$, the transitive dependence are disjoint, and
+  no predicates shows more than once in the body of $C$.
   (\autoref{defn:cdds}).
   %
-  Let $\mathcal{C}$ be an arbitrary clause in $\mathcal{S}$.
+  Let $C$ be an arbitrary clause in $S$.
   %
-  $\mathcal{C}$ has no pairs of distinct clauses, by the fact that
-  $\mathcal{S}$ is linear.
+  The body of $C$ has no pair of distinct predicates or a predicates shows more than once, by the fact that
+  $S$ is linear.
   %
-  Therefore, $\mathcal{S}$ is CDD.
+  Therefore, $S$ is CDD.
 
   % contains body-disjoint systems:
-  For goal \textbf{(2)}, let $\mathcal{S}$ be an arbitrary
+  For goal \textbf{(2)}, let $S$ be an arbitrary
   body-disjoint system.
   %
-  The dependence relation of $\mathcal{S}$ is a tree $T$, by the
+  The dependence relation of $S$ is a tree $T$, by the
   definition of a body-disjoint system.
   %
-  Let $\mathcal{C}$ be an arbitrary clause in $\mathcal{S}$, with
+  Let $C$ be an arbitrary clause in $S$, with
   distinct relational predicates $R_0$ and $R_1$ in its body.
@@
   %
-  Thus, $\mathcal{S}$ is CDD, by \autoref{defn:cdds}.
+  Thus, $S$ is CDD, by \autoref{defn:cdds}.
 
@@
 by \sys, namely \textsc{Collapse} (\autoref{lem:expansion-sound} and
-\autoref{lem:expansion-complete}), $\expand{\mathcal{R}}$
-(\autoref{lem:expand-corr}), and $\solvecdd{ \mathcal{R} }$
+\autoref{lem:expansion-complete}), $\expand$
+(\autoref{lem:expand-corr}), and $\solvecdd$
 (\autoref{lem:vc}, \autoref{lem:cdd-soln-sound}, and
@@
 % lem: result connecting expansions to constructing solutions
-For CHC systems $\mathcal{S}$ and $\mathcal{S}'$, if $\mathcal{S}'$ is
-an expansion of $\mathcal{S}$, then the result of collapsing a
-solution of $\mathcal{S}'$ is a solution of $\mathcal{S}$.
+For recursion-free CHC systems $S$ and $S'$, if $S'$ is
+an expansion of $S$, then the result of collapsing a
+solution of $S'$ is a solution of $S$.
 %
@@
   \label{lem:expansion-sound}
-  For $\mathcal{R}, \mathcal{R}' \in \relpreds$, $\mathcal{S} \in
-  \chcs{ \mathcal{R} }$, $\mathcal{S}' \in \chcs{ \mathcal{R}' }$,
-  $\sigma' \in \interps{ \mathcal{R}' }$, and $\eta : \mathcal{R}' \to
-  \mathcal{R}$ a correspondence from $\mathcal{S}'$ to $\mathcal{S}$,
-  $\collapse{\eta}{\sigma'}$ is a solution of $\mathcal{S}$.
+  For recursion-free CHC system $S'$ and $S$,
+  $\sigma'$ is a solution of $S'$, and $\eta : \predof{S}' \to
+  \predof{S}$ a correspondence from $\predof{S}'$ to $\predof{S}$,
+  $\collapse{\eta}{\sigma'}$ is a solution of $S$.
 \end{lem}
@@
   %
-  For each $R' \in \mathcal{R}'$ such that $\eta(R') = R$, there is
-  some $\mathcal{B}' \in \bodies{ \mathcal{R}' }$ such that
-  $((\mathcal{B}', \varphi), R'(\vars{ \mathcal{S} }(R))) \in
-  \mathcal{S}'$, by the fact that $\mathcal{S}'$ is an expansion of
-  $\mathcal{S}$ and the definition of an expansion.
+  For each predicate $P' \in \predof{S}'$ such that $\eta(P') = P$, there is
+  some clause $C' \in S'$ such that $P' \in \bodyof{C'}$ by the fact that $S'$ is an expansion of
+  $S$ and the definition of an expansion.
   %
-  $\sigma'(\mathcal{B}'), \varphi \entails \sigma'(\mathcal{R}')$ by
-  the fact that $\sigma'$ is a solution of $\mathcal{S}'$.
+  Let preicate $Q' \in \predof{S}'$ such that $Q'$ is the head of clause $C'$.
   % 
-  $\collapse{ \eta }{ \sigma' }(\mathcal{B}), \varphi \entails
-  \sigma'(\mathcal{R}')$, by the definition of
-  $\collapse{\eta}{\sigma'}$.
+  $\sigma'[P'] \land \consof{C'} \entails \sigma'[Q']$ by
+  the fact that $\sigma'$ is a solution of $S'$.
   %
@@
   %
-  \[ \collapse{ \eta }{ \sigma' }( \mathcal{B} ), \varphi \entails %
-  \left( \bigland_{ \substack{R' \in \mathcal{R}' \\ \eta(R') = R} }
-  \sigma(R') \right) = %
-  \collapse{\eta}{\sigma'}(\mathcal{R}') 
+  \[ \collapse{ \eta }{ \sigma' }[P] \land \consof{C'} \entails %
+  \left( \bigland_{ \substack{Q' \in \predof{S}' \\ \eta(Q') = Q} }
+  \sigma(Q') = %
+  \collapse{\eta}{\sigma'}[Q'] \right)
   \]
   % 
-  Therefore, $\collapse{\eta}{\sigma'}$ is a solution of clause
-  $((\mathcal{B}, \varphi), R(\varsof{ \mathcal{S} }(R)))$.
+  Therefore, $\collapse{\eta}{\sigma'}$ is a solution of $P$.
   %
   Therefore, $\collapse{\eta}{\sigma'}$ is a solution of
-  $\mathcal{S}$.
+  $S$.
 \end{proof}
@@
 % completeness of expansions:
-For CHC solvable system $\mathcal{S}$, each expansion of $\mathcal{S}$
+For a recursion-free CHC system $S$ is solvalbe, each expansion of $S$
 is solvable.
@@
   \label{lem:expansion-complete}
-  For $\mathcal{R}, \mathcal{R}' \in \relpreds$, $\mathcal{S} \in
-  \chcs{ \mathcal{R} }$ such that $\mathcal{S}$ is solvable, and
-  $\mathcal{S}' \in \chcs{ \mathcal{R}' }$ such that $\mathcal{S'}$ is
-  an expansion of $\mathcal{S}$, $\mathcal{S}'$ is solvable.
+  For a recursion-free CHC system $S$ is solvable, and
+  $S'$ is
+  an expansion of $S$, $S'$ is solvable.
 \end{lem}
@@
 \begin{proof}
-  Let $\sigma : \mathcal{R} \to \formulas$ be a solution of
-  $\mathcal{S}$, and let $\eta : \mathcal{R}' \to \mathcal{R}$ be a
-  correspondence from $\mathcal{R}'$ to $\mathcal{R}$.
+  Let $\sigma : P \to \formulas$ be a solution of
+  $S$, and let $\eta : \predof{S}' \to \predof{S}$ be a
+  correspondence from $S'$ to $S$.
   %
-  Let $\sigma' : \mathcal{R}' \to \formulas$ be such that for each $R'
-  \in \mathcal{R}'$, $\sigma'(R') = \sigma(\eta(R'))$.
+  Let $\sigma' : \predof{S}' \to \formulas$ be such that for each $P'
+  \in \predof{S}'$, $\sigma'(P') = \sigma(\eta(P'))$.
   %
-  Then $\sigma$ is a solution of $\mathcal{S}'$.
+  Then $\sigma'$ is a solution of $S'$.
 \end{proof}
@@
 % CDD systems:
-$\expand{\mathcal{R}}$ always returns a CDD expansion of its input.
+$\expand$ always returns a CDD expansion of its input.
 %
@@
   \label{lem:expand-corr}
-  For $\mathcal{R} \in \relpreds$ and $\mathcal{S} \in \chcs{
-    \mathcal{R} }$, $\mathcal{S} \in \chcs{ \ctxs{ \mathcal{R} } }$,
-  and $\eta : \ctxs{ \mathcal{R} } \to \mathcal{R}$ such that
-  $(\mathcal{S}', \eta) = \expand{\mathcal{R}}(\mathcal{S})$, $\eta$
-  is a correspondence from $\mathcal{S}'$ to $\mathcal{S}$.
+  For a recursion-free CHC system $S$ and $S'$, and $\eta : \predof{S'} \to \predof{S}$ such that
+  $(S', \eta) = \expand(S)$, $\eta$
+  is a correspondence from $S'$ to $S$, $S'$ is an CDD system and an expansion
+  of $S$.
 \end{lem}
@@
 \begin{proof}
-  Proof by induction on the evaluation of $\expand{ \mathcal{R} }$ on
-  an arbitrary $\mathcal{S} \in \chcs{ \mathcal{R} }$.
+  Proof by induction on the evaluation of $\expand$ on
+  an arbitrary recursion-free CHC system $S$.
   % inductive fact:
   The inductive fact is that in each step of evaluation \expandaux,
-  $\corr{ \mathcal{R} }$ is a correspondence from argument
-  $\mathcal{S}'$ to $\mathcal{S}$.
+  $\corr$ is a correspondence from argument
+  $S'$ to $S$.
   % base case:
-  For the base case, \expandaux is called initially on $\mathcal{S}$,
+  For the base case, \expandaux is called initially on $S$,
   by \autoref{alg:expand}.
   % 
-  $\corr{ \mathcal{R} }$ is a correspondence from $\mathcal{S}$ to
-  itself, by the definition of $\corr{ \mathcal{R} }$
+  $\corr$ is a correspondence from $S$ to
+  itself, by the definition of $\corr$
   (\autoref{app:cons-cdd}).
@@
   % inductive case:
-  For the inductive case, for $\mathcal{C} \in \mathcal{S}$ and $R \in
-  \ctxs{ \mathcal{R} }$, $\expandaux$ calls itself on $\copyrel{
-    \mathcal{R} }(\mathcal{S}, \mathcal{C}, R)$.
+  For the inductive case, for $C \in S$ and $P \in \predof{S}$, $\expandaux$ calls itself on 
+  $\copyrel(S, C, P)$.
   %
-  For each $\mathcal{S}' \in \chcs{ \ctxs{ \mathcal{R} } }$ such that
-  $\corr{ \mathcal{R} }$ is a correspondence from $\mathcal{S}'$ to
-  $\mathcal{S}$, $\corr{ \mathcal{R} }$ is a correspondence from
-  $\copyrel{ \mathcal{R} }( \mathcal{S}', \mathcal{C}, R)$ to
-  $\mathcal{S}$, by definition of $\copyrel{ \mathcal{R} }$
+  For each recursion-free CHC system $S'$ that generated by $\copyrel(S, C, P)$,
+  $\corr$ is a correspondence from $S'$ to
+  $S$ by definition of $\copyrel$
   (\autoref{app:cons-cdd}).
   %
-  By this fact and the inductive hypothesis, $\corr{ \mathcal{R}}$ is
-  a correspondence from $\copyrel{ \mathcal{R} }( \mathcal{S}',
-  \mathcal{C}, R)$ to $\mathcal{S}$.
+  By this fact and the inductive hypothesis, $\corr$ is
+  a correspondence from $\copyrel(S',
+  C, P)$ to $S$.
 
   % final claim: expand returns an expansion:
-  $\expandaux{ \mathcal{R} }$ returns its parameter at some step, by
+  $\expandaux$ returns its parameter at some step, by
   \autoref{alg:expand}.
   %
-  Therefore, $\expandaux$ returns an expansion of $\mathcal{S}$.
+  Therefore, $\expandaux$ returns an expansion of $S$.
 
   % prove that returned system is CDD:
-  For $\mathcal{S}' \in \chcs{ \ctxs{ \mathcal{R} } }$ and $\eta :
-  \ctxs{\mathcal{R}} \to \mathcal{R}$, if $(\mathcal{S}', \eta) =
-  \expand{ \mathcal{R} }( \mathcal{S} )$, then $\sharingclause(
-  \mathcal{S}') = \none$, by the definition of $\expand{ \mathcal{R}
-  }$.
+  For a given recursion-free CHC system $S'$, if $(S', \eta) =
+  \expand(S')$, 
+  then $\sharingclause(S') = \none$, by the definition of $\expand$.
   %
-  If $\sharingclause(\mathcal{S}') = \none$, then $\mathcal{S}'$ is
+  If $\sharingclause(S') = \none$, then $S'$ is
   CDD, by the definition of $\sharingclause$ and CDD systems
@@
   %
-  Therefore, $\mathcal{S}'$ is CDD.
+  Therefore, $S'$ is CDD.
 \end{proof}
 %
-Furthermore, $\expand{ \mathcal{R} }$ returns a \emph{minimal} CDD
+Furthermore, $\expand$ returns a \emph{minimal} CDD
 expansion of its input.
@@
 % lem: VC's characterize system sat:
-For each CHC system $\mathcal{S}$, $\mathcal{S}$ has a solution if and
-only if the verification condition of $\mathcal{S}$ is unsatisfiable.
+For each recursion-free CHC system $S$, $S$ has a solution if and
+only if all interpolants queries return interpolants.
 %
@@
   \label{lem:vc}
-  For $\mathcal{R} \in \relpreds$ and $\mathcal{S} \in \chcs{
-    \mathcal{R} }$ that is CDD and solvable, $\vc{ \mathcal{R} }(
-  \mathcal{S} )$ is unsatisfiable.
+  Given a recursion-free CHC system $S$ that is CDD and solvable, for all predicates $P \in \predof{S}$,
+   $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns a interpolatn $I$.
 \end{lem}
@@
   %
-  Assume that $\mathcal{S}$ has solution $\sigma$ and $m$ is a model
-  of $\vc{ \mathcal{R} }( \mathcal{S})$.
+  Assume that $S$ has solution $\sigma$ and there are
+  some predicates $P \in \predof{S}$ such that $\solveitp(\prectr(\cc{P},\sigma'),\postctr(\cc{P},\sigma'))$
+  returns $SAT$, which means there is a model $m$ for the conjuntion of $\prectr(\cc{P},\sigma')$
+  and $\postctr(\cc{P},\sigma')$.
   %
-  There are some relational predicates $\mathcal{R}' \subseteq
-  \mathcal{R}$ containing $\queryof{ \mathcal{S} }$ such that for each
-  $R' \in \mathcal{R}'$, $m(\used{R})$ holds, by the definition of
-  $\vc{ \mathcal{R} }$ (\autoref{sec:solve-cdd}).
-  %
-  For each $R'$, the interpretation of $m$ over $\varsof{ \mathcal{R}'
-  }$ must satisfy $\sigma(R')$, by well-founded induction on $\deps_S$
-  restricted to $\mathcal{R}'$.
-  %
-  But $\sigma(\queryof{ \mathcal{S} }) = \false$, by the definition of
-  a solution of a CHC system (\autoref{defn:chc-soln}).
+  But $\postctr(\cc{P},\sigma) = \false$, by the definition of
+  a solution of a CHC system.
   %
@@
   %
-  Therefore, $\vc{ \mathcal{R} }( \mathcal{S} )$ is unsatisfiable.
+  Therefore, $\solveitp(\prectr(\cc{P},\sigma'),\postctr(\cc{P},\sigma'))$ always returns interpolant.
 \end{proof}
@@
   \label{lem:solve-aux}
-  For $\mathcal{R} \in \relpreds$ and $\mathcal{S} \in \chcs{
-    \mathcal{R} }$ such that $\vc{ \mathcal{R} }(\mathcal{S})$ is
-  unsatisfiable, $\solvecdd{ \mathcal{R} }(\mathcal{S})$ is a solution
-  of $\mathcal{S}$.
+  Given a recursion-free CHC system $S$ that is CDD, if for all predciates $P \in \predof{S}$, 
+  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns a interpolatn $I$, 
+  then $\solvecdd(S)$ is a solution of $S$.
 \end{lem}
@@
   % overview of structure:
-  Proof by induction on the evaluation of $\solveaux$ on
-  interpretation $\sigma$.
+  Proof by induction on the $\solvecdd(S)$ calles $\solveitp$ on all predicates $P \in \predof{S}$
+  in a topolgical order.
   %
-  The inductive fact is that at each step of evaluation, $\sigma$ is a
-  partial solution of $\mathcal{S}$ such that $\sigma(\mathcal{S})$
+  The inductive fact is that at each step of calling $\solveitp$, $\sigma$ is a
+  partial solution of $S$.
   % base case:
-  For the base case, \solveaux is initially called on the empty
-  interpretation.
+  For the base case, $\predof{S} = \varnothing$,
   %
-  Therefore, $\vc{\mathcal{R}}(\sigma(\mathcal{S})) = \vc{
-    \mathcal{R}}(\mathcal{S})$, which is unsatisfiable.
+  Therefore, $\sigma$ is a solution of $S$.
 
   % inductive case:
-  For the inductive case, in each step, \solveaux calls itself on an
-  interpretation $\sigma'$, which is its argument $\sigma$, extended
-  to map relational predicate $R$ to an interpolant of $\prectr{
-    \mathcal{R} }( \mathcal{S}, \sigma, R)$ and $\postctr{ \mathcal{R}
-  }( \mathcal{S}, \sigma, R)$.
+  For the inductive case, in each step, $\solvecdd$ calls $\solveitp$ on
+  a predicate $P \in \predof{S}$ with partial solution $\sigma$, which
+  contains all solutions of predciates $P \in \depsof{S}$.
   %
-  $\sigma'$ is a partial solution of $\mathcal{S}$ by a combination of
-  the inductive hypothesis that $\sigma$ is a partial solution, the
-  fact that any extension of $\sigma$ that maps $R$ to a formula
-  implied by $\prectr{ \mathcal{R} }( \mathcal{S}, \sigma, R)$ is a
-  partial solution, and the definition of an interpolant
-  (\autoref{defn:itps}).
+  Based on the definition of an interpolant(\autoref{defn:itps}), $\sigma$
+  is a partial solution of $S$, then $\prectr(\cc{P},\sigma')$ and $\postctr(\cc{P},\sigma))$
+  are inconsistent.
   %
-  $\vc{ \mathcal{R} }(\sigma'(\mathcal{S}))$ is unsatisfiable by a
-  combination of the inductive hypothesis that $\vc{ \mathcal{R}
-  }(\sigma(\mathcal{ S }))$ is unsatisfiable, that for any extension
-  $\sigma''$ of $\sigma$ that maps $R$ to a formula inconsistent with
-  $\postctr{ \mathcal{R} }( \mathcal{S}, \sigma, R)$, $\vc{
-    \mathcal{R} }(\sigma''(\mathcal{S}))$ is unsatisfiable, and the
-  definition of an interpolant (\autoref{defn:itps}).
+  The interpolant $I$ returns by $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ entails
+  by all clause that head is $P$ and $\bodyof{P}$ is substituted by partial solution $\sigma$.
+  %
+  $I$ also is inconsistment with all constrains shows later that support the query clause.
+  %
+  Then $\solvecdd$ updates $\sigma$ by binding $P$ to $I$, which makes the new $\sigma$ also a partial
+  solution of $S$.
 
-  % prove goal:
-  $\mathcal{R}'$ (defined at \autoref{alg:solve-cdd},
-  \autoref{line:cons-nodeps}) is empty only if $\mathcal{R} \setminus
-  \domainof{ \sigma }$ is empty, by induction on evaluation of
-  \solveaux.
   %
-  Therefore, \solveaux only returns its parameter $\sigma$ if $\sigma$
-  is in fact a complete interpretation.
-  %
-  This fact, combined with the inductive fact, imply that \solveaux
-  returns a solution of $\mathcal{S}$.
+  Base case combined with the inductive case, imply that \solvecdd
+  returns a solution of $S$.
 \end{proof}
@@
 % SolveCDD: establish soundness
-For $\mathcal{R} \in \relpreds$, $\solvecdd{ \mathcal{R} }$ is a
-correct CHC solver for CDD systems over $\mathcal{R}$.
+$\solvecdd$ is a
+correct CHC solver for CDD systems $S$.
 %
@@
   \label{lem:cdd-soln-sound}
-  For $\mathcal{R} \in \relpreds$, %
-  $\mathcal{S} \in \chcs{ \mathcal{R} }$ such that $\mathcal{S}$ is
-  CDD, and %
-  $\sigma \in \interps{ \mathcal{R} }$ such that %
-  $\sigma = \solvecdd{ \mathcal{R} }(\mathcal{S})$, %
+  For a given CDD system $S$, and $\sigma = \solvecdd(S)$, %
   $\sigma$ is a solution of $\mathcal{S}$.
@@
   %
-  $\vc{ \mathcal{R} }(\mathcal{S})$ is unsatisfiable, by the fact that
-  $\sigma \in \interps{ \mathcal{R} }$ \autoref{alg:solve-cdd},
-  \autoref{line:check-sat}.
+  $\solvecdd{S}$ returns $\sigma$ implies that for all predicates 
+  $P \in \predof{S}$, $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns a interpolatn $I$, 
+  by the definition of $\solvecdd$ \autoref{alg:solve-cdd}. 
   %
-  Therefore, $\solvecdd{ \mathcal{R} }(\mathcal{S})$ is the result of
-  calling \solveaux, by \autoref{alg:solve-cdd},
-  \autoref{line:check-sat}.
-  %
-  Therefore, \autoref{lem:solve-aux} and the fact that $\vc{
-    \mathcal{R} }(\mathcal{S})$ is unsatisfiable imply that \solveaux
-  returns a complete solution of $\mathcal{S}$.
+  Therefore, \autoref{lem:solve-aux} implies that \solvecdd
+  returns a complete solution of $S$.
 \end{proof}
@@
   \label{lem:cdd-soln-complete}
-  For $\mathcal{R} \in \relpreds$ and %
-  $\mathcal{S} \in \chcs{ \mathcal{R} }$ such that $\mathcal{S}$ is
-  solvable and CDD, %
-  there is some $\sigma \in \interps{ \mathcal{R} }$ such that %
-  $\sigma = \solvecdd{ \mathcal{R} }(\mathcal{S})$.
+  For a CDD system $S$ such that $S$ is
+  solvable, %
+  there is some $\sigma$ such that %
+  $\sigma = \solvecdd(S)$.
 \end{lem}
@@
   % final claim:
-  $\vc{ \mathcal{R} }(\mathcal{S})$ is not satisfiable, by
+  for all predicates $P \in \predof{S}$, $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns a interpolatn $I$, by
   \autoref{lem:vc} and the fact that $\mathcal{S}$ is solvable.
   %
-  Therefore, $\solvecdd{ \mathcal{R} }$ is the result of running
-  \solveaux, by \autoref{alg:solve-cdd}.
-  %
   Therefore, by \autoref{lem:solve-aux} and the fact that
-  $\mathcal{S}$ is solvable, $\solvecdd{ \mathcal{R} }$ returns a
-  solution of $\mathcal{S}$.
+  $S$ is solvable, $\solvecdd(S)$ returns a
+  solution of $S$.
 \end{proof}
@@
 \begin{proof}
-  % if returns a map, then S is a solution
-  Assume that for $\sigma \in \interps{ \mathcal{R} }$, $\sigma =
-  \shara{ \mathcal{R} }( \mathcal{S} ) \in \interps{ \mathcal{R} }$.
+  Given recurion-free CHC systems $S$ and $S'$, and $\eta :\predof{S'} \to 
+  \predof{S}$ such that $(S', \eta) = \expand(S)$, 
+  then $S'$ is minimal CDD expansion of 
+  $S$, and $\eta$ is a correspondence from $S'$ to $S$ by \autoref{lem:expand-corr}.
   %
-  Let $\mathcal{S'} \in \chcs{ \ctxs{ \mathcal{R} } }$ and $\eta :
-  \ctxs{ \mathcal{R} } \to \mathcal{R}$ be such that $(\mathcal{S}',
-  \eta) = \expand{ \mathcal{R} }( \mathcal{S} )$;
+  There is some $\sigma'$ such that
+  $\sigma' = \solvecdd(S' )$, by
+  the definition of $\shara$.
   %
-  $\eta$ is a correspondence from $\mathcal{S'}$ to $\mathcal{S}$, by
-  \autoref{lem:expand-corr}.
-  %
-  There is some $\sigma' \in \interps{ \mathcal{R}' }$ such that
-  $\sigma' = \solvecdd{ \ctxs{ \mathcal{R} } }( \mathcal{S}' )$, by
-  the definition of $\shara{ \mathcal{R} }$.
-  %
-  $\sigma'$ is a solution of $\mathcal{S}'$, by
+  $\sigma'$ is a solution of $S'$, by
   \autoref{lem:cdd-soln-sound}.
   %
-  $\collapse{ \eta }{ \sigma' }$ is a solution of $\mathcal{S}$, by
+  $\collapse{ \eta }{ \sigma' }$ is a solution of $S$, by
   \autoref{lem:expansion-sound}.
@@
   % if S is solvable, then shara returns a solution
-  Assume that $\mathcal{S}$ is solvable.
+  Assume that a recursion-free $S$ is solvable.
   %
-  Let $\mathcal{S'} \in \chcs{ \ctxs{ \mathcal{R} } }$ and $\eta :
-  \ctxs{ \mathcal{R} } \to \mathcal{R}$ be such that $(\mathcal{S}',
-  \eta) = \expand{ \mathcal{R} }( \mathcal{S} )$.
+  Let a recursion-free CHC system $S'$ and $\eta :
+  \predof{S'} \to \predof{S}$ be such that $(S',
+  \eta) = \expand(S)$.
   %
-  $\mathcal{S}'$ is a CDD expansion of $\mathcal{S}$, by
+  $S'$ is a CDD expansion of $S$, by
   \autoref{lem:expand-corr}.
   %
-  $\mathcal{S}'$ is solvable, by \autoref{lem:expansion-complete}.
+  $S'$ is solvable, by \autoref{lem:expansion-complete}.
   %
-  $\solvecdd{ \ctxs{ \mathcal{R} } }( \mathcal{S'}) \in \interps{
-    \mathcal{R'} } $, by \autoref{lem:cdd-soln-complete}.
+  $\solvecdd(S')$ returns a solution $\sigma$ of $S'$ by \autoref{lem:cdd-soln-complete}.
   %
-  $\shara{\mathcal{R}} \in \interps{ \mathcal{R} }$, by
+  $\shara$ returns a solution of $S$ by calling $\collapse{ \eta }{ \sigma' }$, based on
   \autoref{alg:shara}.

### 2018-04-11 paper/introduction.tex

@@
 A CHC is a logical implication where the antecedent is a conjunction
-of uninterpreted relational predicates with a fact in a background
-theory and the consequent is an uninterpreted relational predicate. In
-the context of a CHC we call the antecedent the body and the
+of uninterpreted predicates with a fact in a background
+theory and the consequent is an uninterpreted predicate. 
+%
+In the context of a CHC we call the antecedent the body and the
 consequent the head.  A CHC system is a set of CHCs. 
@@
 The goal of a CHC solving problem is to find suitable interpretations
-for each relational predicate such that each CHC is logically
-consistent in isolation.
+for each predicate such that each CHC is logically
+consistent in isolation.\QZ{re-write this based on other papers}
 
@@
 as \emph{recursion-free}. In a recursion-free CHC system, the
-derivation of each relational predicate will contain no references to
+derivation of each predicate will not contain
 itself. \QZ{\DAH{I still think we need a small example here}}
@@
 (programs with only bounded iteration and recursion).
+
 %
-Second and most importantly, the task of solving a general CHC
-system can reduce to the task of solving a sequence of recursion-free
+Second and most importantly, A well-known approach for solving a general CHC
+system can reduce to solving a sequence of recursion-free
 systems \QZ{reference}.
+%
+Many solvers attempt to synthesize the solution of the original system 
+by using the solutions of constructed recruion-free system~\cite{bjorner13}.
+%
+Because of this, the performance of solving general CHC system is heavily reliant on the 
+performance of the solving recursion-free CHC system.
+%
 
-Many \QZ{more references? Or remove many.} modern solvers generate
-a sequence such that each subsystem is a bounded unwinding of the
-input system. These solvers either attempt to show that the solution to a
-given subsystem generalizes to the original system or generate a
-larger subsystem for the next iteration~\cite{bjorner13}.
+% current general techniques for solving recursion-free systems
+Typically, recursion-free CHC systems are not solved directly.
 %
-Since such solvers inspect recursion-free systems in an iterative
-loop, they are heavily reliant on the performance of the underlying
-recursion-free solver.
-%
-% current general techniques for solving recursion-free systems
-Typically, even recursion-free CHC systems are not solved directly.
 Instead, they are reduced to a more specific subclass of
@@
 These classes include those of
-\emph{body disjoint} (or \emph{tree structure})
+\emph{body disjoint} (or \emph{derivation tree})
 systems~\cite{heizmann10,bjorner13,mcmillan14,rummer13a,rummer13b} and
@@
 %
-Such classes have proved practical because a solver can find
-suitable definitions for relational predicates by issuing
-\emph{interpolation queries}.
+Such classes can be solved by issuing
+\emph{interpolation queries} for uninterpreted predicate and 
+each constrains shows only one time in the interpolation query.
 %
-When a class of CHC systems can be solved by interpolation queries, we
-say the class is \emph{directly solvable}.
+We
+refer such class as \emph{directly solvable}.
 
 % Time complexity
-In general, solving a recursion-free CHC system whose facts are
-restricted to the theory of linear integer arithmetic is
+In general, solving a recursion-free CHC system for 
+propositional logic and theory of linear integer arithmetic is
 co-NEXPTIME-complete~\cite{rummer13b}.
 %
-In contrast, solving a directly solvable system with the same theories
+In contrast, solving a directly solvable system with the same logic and theories
 is in co-NP~\cite{rummer13b}.
 %
-Therefore, the performance of a CHC solver is highly reliant on the
-size of the directly solvable systems it produces.
+Therefore, the performance of solving a recurion-free system is highly reliant on the
+size of the directly solvable systems it is reduced to.
 
@@
 clause in the system has at most one dependency on the same relational
-predicate.
+predicate.\QZ{repharse this}
 %

### 2018-04-11 paper/related-work.tex

@@
 Such an approach corresponds to attempting to solve a recursive CHC
-system $\mathcal{S}$ by selecting derivation trees of $\mathcal{S}$
+system $S$ by selecting derivation trees of $S$
 and solving each tree.
@@
 Previous work has proposed solvers for recursive systems that, given a
-system $\mathcal{S}$, attempt to solve $\mathcal{S}$ by generating and
-solving a series of recursion-free unwindings of $\mathcal{S}$.
+system $S$, attempt to solve $S$ by generating and
+solving a series of recursion-free unwindings of $S$.
+
 %
-In particular, \duality attempts to solve each unwinding
-$\mathcal{S}'$ by selecting and solving derivation-trees of
-$\mathcal{S}'$~\cite{bjorner13}.
+In particular, \eldarica attempts to solve each unwinding
+$S'$ by reducing to and solving body-disjoint systems~\cite{rummer13a,rummer13b}.
+%
+
+%
+\duality attempts to avoid solving all 
+derivation-trees (i.e body-disjoint system) by using lazy annotation~\cite{bjorner13}.
 %
@@
 %
-\eldarica attempts to solve each recursion-free CHC system by lazily
-copying its subsystems to form a body-disjoint
-over-approximation~\cite{rummer13a,rummer13b}.
 
-% DAG interpolation:
 \whale attempts to verify a sequential recursive program by generating
@@
 %
-To solve a particular recursion-free system $\mathcal{S}$, \whale
-generates a linear inlining $\mathcal{S}'$ of $\mathcal{S}$ and solves
+To solve a particular recursion-free system $S$, \whale
+generates a linear inlining $S'$ of $S$ and solves
 it using a procedure \vinta~\cite{albarghouthi12a}.
 %
-In general, $\mathcal{S}'$ may have size exponential in the size of
-$\mathcal{S}$.
+In general, $S'$ may have size exponential in the size of
+$S$.
 
@@
 recursion-free CHC systems in that it reduces the problem of solving a
-given recursion-free CHC system $\mathcal{S}$ to solving a CHC system
+given recursion-free CHC system $S$ to solving a CHC system
 in a directly-solvable class.
 %
+
 \sys is distinct from all of the approaches given above in that it
@@
 themselves be solved directly.
+
+%
+For solving genreal CHC system, \sys uses the same strategy that proposed
+by above approaches, that solving a series recursion-free CHC system that unwinding 
+from the original CHC system, and try to synthesize the solution for the original 
+CHC system by using solutions of recursion-free CHC systems.
 
@@
 \sys contains a procedure that, given a recursion-free Horn Clause
-system $\mathcal{S}$, attempts to construct a compact CDD system
-$\mathcal{S}'$ such that each solution of $\mathcal{S}'$ defines a
-solution of $\mathcal{S}$.
+system $S$, attempts to construct a compact CDD system
+$S'$ such that each solution of $S'$ defines a
+solution of $S$.
 %

### 2018-04-11 paper/background.tex
changes to background

@@
 For a given CHC $C$, $\bodyof{C}$ denotes the vector of uninterpreted
-predicates in the body of $C$ and $\consof{C}$ denotes the fact in the
+predicates in the body and $\consof{C}$ denotes the fact in the
 body.
 %
-It is possible for an uninterpreted predicate to apper multiple times
+It is possible for an uninterpreted predicate to appear multiple times
 in the body of $C$.
@@
 %define a CHC system
-A CHC system is a set of CHCs where exactly one clause is a query
+A CHC system is a set of CHCs where exactly one clause is a query.
 %
@@
 \begin{defn}
-  Given a CHC system $\cc{S}$ and two uninterpreted predicates $\cc{P}$ and
-  $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$ such that $\cc{P} = \headof{C}$
-  and $\cc{Q} \in \bodyof{C}$, then $\cc{Q}$ is a dependency predicate of $\cc{P}$.
+  Given a CHC system $\cc{S}$ and two uninterpreted predicates
+  $\cc{P}$ and $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$
+  such that $\cc{P} = \headof{C}$ and $\cc{Q} \in \bodyof{C}$, then
+  $\cc{Q}$ is a predicate dependency of $\cc{P}$.
 \end{defn}
@@
 \begin{ex}
-  Because uninterpreted predicates \cc{L4} is in the body of clause (4) and uninterpreted predicates \cc{L8}
-  is the head of clause (4), \cc{L4} is a dependency predicate of \cc{L8}.\autoref{chc}
+  In $\mcchc$, because $\cc{L}_4$ is in the body of clause (4) and
+  $\cc{L}_8$ is the head of clause (4), $\cc{L}_4$ is a predicate
+  dependency of $\cc{L}_8$.
 \end{ex}
 %
-Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$, $\depsof{P}$
-denotes the set of all dependency predicates of $\cc{P}$ in $\cc{S}$.
+Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
+$\depsof{P}$ denotes the set of all predicate dependencies of $\cc{P}$
+in $\cc{S}$.
 %
@@
 \begin{defn}
-  Given a CHC system $\cc{S}$ and three uninterpreted predicates $\cc{P},
-  \cc{Q}$ and $\cc{R} \in \predof{S}$, if $\cc{Q} \in \depsof{P}$ then \cc{Q} is a transitive dependency
-  predicate of \cc{P}. 
+  Given a CHC system $\cc{S}$ and three uninterpreted predicates
+  $\cc{P}, \cc{Q}$ and $\cc{R} \in \predof{S}$, if $\cc{Q} \in
+  \depsof{P}$ then \cc{Q} is a transitive predicate dependency of
+  \cc{P}.
   %
-  if \cc{Q} is a transitive dependency predicate of \cc{P}, and \cc{R} is a transitive dependency predicate
-  of \cc{Q},then $\cc{R}$ is a transitive dependency predicate of $\cc{P}$.
+  If \cc{Q} is a transitive predicate dependency of \cc{P} and \cc{R}
+  is a transitive predicate dependency of \cc{Q}, then $\cc{R}$ is a
+  transitive predicate dependency of $\cc{P}$.
 \end{defn}
@@
 \begin{ex}
-  Because uninterpreted predicates \cc{L4} is a dependency predicate of \cc{L8}, 
-  \cc{L4} is a transitive dependency predicate of \cc{L8}.
+  In $\mcchc$, because $\cc{L}_4$ is a predicate dependency of
+  $\cc{L}_8$, $\cc{L}_4$ is a transitive predicate dependency of
+  $\cc{L}_8$.
   %
-  And because \cc{L8} is an transitive dependency predicate of \cc{L9}, 
-  \cc{L4} is a transitive dependency predicate of \cc{L9}.\autoref{chc}
+  And because $\cc{L}_8$ is an transitive predicate dependency of
+  $\cc{L}_9$, $\cc{L}_4$ is a transitive predicate dependency of
+  $\cc{L}_9$.
 \end{ex}
 %
-Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$, $\tdepsof{P}$
-denotes the set of all transitive dependency predicates of $\cc{P}$ in $\cc{S}$.
+Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
+$\tdepsof{P}$ denotes the set of all transitive predicate dependencies
+of $\cc{P}$ in $\cc{S}$.
 %
@@
 \begin{defn}
-  Given a CHC system $\cc{S}$ and two uninterpreted predicates $\cc{P}$ and
-  $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$ such that $\cc{P} \in \bodyof{C}$
-  and $\cc{Q} \in \bodyof{C}$, then $\cc{Q}$ and $\cc{P}$ are siblings of each other.
+  Given a CHC system $\cc{S}$ and two uninterpreted predicates
+  $\cc{P}$ and $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$
+  such that $\cc{P} \in \bodyof{C}$ and $\cc{Q} \in \bodyof{C}$, then
+  $\cc{Q}$ and $\cc{P}$ are siblings.
 \end{defn}
@@
 \begin{ex}
-  Because uninterpreted predicates \cc{L9} and \cc{db} both shows in the body of
-  clause (6), \cc{L9} and \cc{db} are siblings of each other.
+  Because uninterpreted predicates $\cc{L}_9$ and $\cc{db}$ both
+  appear in the body of clause (6), $\cc{L}_9$ and $\cc{db}$ are
+  siblings.
   %
@@
 %
-Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$, $\siblingof{P}$
-denotes the set of all siblings predicates of $\cc{P}$ in $\cc{S}$.
+Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
+$\siblingof{P}$ denotes the set of all predicate siblings of $\cc{P}$
+in $\cc{S}$.
+%
+For a given CHC system $\cc{S}$, if there are no uninterpreted
+predicate $\cc{P} \in \predof{S}$ such that $\cc{P} \in \tdepsof{P}$,
+then $\cc{S}$ is a \emph{recursion-free} CHCs system.
 %
 
+A solution to a CHC system $S$ is a map from each uninterpreted
+predicate $\cc{P} \in \predof{S}$ to its corresponding interpretation
+which is a formula.
 %
-For a given CHC system $\cc{S}$, there is no uninterpreted predicates $\cc{P} \in \predof{S}$ 
-such that $\cc{P} \in \tdepsof{P}$, then $\cc{S}$ is a \emph{recursion-free} CHCs system.
-%
-
-%
-A solution to a CHC system $S$ is a map that maps each each
-uninterpreted predicate $\cc{P} \in \predof{S}$ to its interpretation which is a formula.
-%
-All clauses in $S$ holds
-after substitute all uninterpreted predicates by its interpretation.
+For a solution to be valid, each clause in $S$ must be valid after
+substituting each uninterpreted predicate by its interpretation.
 

### 2018-04-11 paper/overview.tex

@@
 % state the defn of CDD:
-A CHC system can be classified as CDD when the clauses obey
-rules based on dependencies between the predicates.
+A CHC system can be classified as CDD when all clauses are satisified
+the following rules:
 %
-One predicate is directly dependent on another if the second appears
-in the body of a clause where the first is the head.
+\textbf{(1)} no two predicates in the body of the same clause share any transitive
+dependencies on other predicates(i.e. dependence disjoint) 
+and \textbf{(2)} no clause has more than one same predicate in its body.
+% give a example
+As an example, clause $(7)$ is dependence disjoint. 
 %
-Based on this, we say that a clause is dependence disjoint if
-\textbf{(1)} no two predicates in the body share any transitive
-dependencies on other predicates and \textbf{(2)} no two predicates in
-the body are the same.
-% give a example
-As an example, clause $(7)$ is dependence disjoint. Clause $(7)$ uses
-both $\cc{L}_9$ and $\cc{dbl}$ in its body.
+Two predicates $\cc{L}_9$ and $\cc{dbl}$ are in its body.
 %
@@
 %
-When each clause in a CHC system is dependence disjoint, it is a CDD
-system. Therefore $\mcchc$ is a CDD system.
+And there is no clauses has more than one same predicate in its body.
+%
+Therefore $\mcchc$ is a CDD system.
 %
@@
 \autoref{sec:CDD-defn}.
-The formal definition of transitive dependency is given in\QZ{reference}.
+The formal definition of transitive dependency is given in
+\QZ{reference}.
 %
@@
 %
-Each uninterpreted predicate $P$ has a valid interpretation which can
+Each interpretation of uninterpreted predicate $P$ can
 be computed by interpolating
@@
 
-We will walk through the construction of an interpolation query for
-$\cc{L}_9$ by hand.  By the time \sys reaches $\cc{L}_9$ it will
-already have solutions for $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$.
+For example, when \sys try to solve $\cc{L}_9$, it has partial solution that 
+map $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$ to their interpretation.
+%
 Possible interpretations of these predicates are shown in
@@
 %
-A solution for $\cc{L}_9$ can be constructed by interpolating
+
+%
+A interpretation of $\cc{L}_9$ can be constructed by interpolating
 the appropriate pre and post formulas.
@@
 The pre-formula is constructed from the bodies of clauses where
-$\cc{L}_9$ is the head. Each relational
-predicate, $P$,
-is replaced by a corresponding boolean indicator variable, $\cc{b}_P$, and each
-boolean indicator variable implies the solution for its predicate.
+$\cc{L}_9$ is the head. 
+%
+Each relational predicate, $P$,
+is replaced by a corresponding boolean indicator variable, $\cc{b}_P$.
+%
+Each
+boolean indicator variable implies the solution for its predicate which is encoded
+as the disjunction of the negation of the boolean indicator variable and the solution.
 %
@@
   \land
-  (\cc{b}_{\cc{L}_6} \Rightarrow \cc{n} \ge 0)
+  (\neg \cc{b}_{\cc{L}_6} \lor \cc{n} \ge 0)
   \land
-  (\cc{b}_{\cc{L}_8} \Rightarrow \cc{n} < 0)
+  (\neg \cc{b}_{\cc{L}_8} \lor \cc{n} < 0)
 \end{gather}
@@
   \land
-  (\cc{b}_{\cc{main}} \Rightarrow
+  (\neg \cc{b}_{\cc{main}} \lor
     (\cc{b}_{\cc{dbl}}
@@
     \land \cc{res}=\cc{d} ))
-  \land (\cc{b}_{\cc{dbl}} \Rightarrow \cc{d} = 2*\cc{x})
+  \land (\neg \cc{b}_{\cc{dbl}} \lor \cc{d} = 2*\cc{x})
 \end{gather}
 %
-Interpolating the pre and post formulas yields a valid intepretation
+Interpolating the pre and post formulas yields a intepretation
 of $\cc{L}_9$: $\cc{abs'} \geq 0$.
+%
 The procedure for solving a CDD system is described in formal detail
@@
 %
-In this section, we show that $\mcchc$ cannot be placed in other
-known classes of recursion-free CHC systems. Specifically, we will
+In this section, we show that $\mcchc$ is not in other
+known classes of recursion-free CHC systems. 
+%
+Specifically, we will
 discuss body-disjoint systems and linear systems.
@@
 % tree solver
-Tree structure
+Body-disjoint systems(i.e derivation tree)
 systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
 class of recursion-free CHC system where each uninterpreted predicate
-appears in the body of at most one clause (and appears in such a
-clause exactly once).
+appears in the body of at most one clause and appears in such a
+clause exactly once.
 %
 Such systems cannot model a program with multiple control paths that
-share a common subpath, typically modeled as a CHC system with a
+share a common pre-subpath, typically modeled as a CHC system with a
 uninterpreted predicate that occurs in the body of multiple clauses.
@@
 
-The class CDD of CDD systems is a superclass of the classes of
+The class CDD of CDD systems is a superclass of union of the classes of
 body-disjoint and linear systems.

### 2018-04-11 paper/overview.tex
revised overview figures

@@
   \begin{floatrow}[2]
-    \ffigbox[.26\textwidth] %
-    { \caption{\cc{dblAbs}: an example hierarchical program.} %
+    \ffigbox[.3\textwidth] {%
+      \caption{\cc{dblAbs}: an example hierarchical program.} %
       \label{fig:multicall-code} }
-    { \input{code/dblAbs.java} }
+    {\lstinputlisting[
+      basicstyle=\small,
+      xleftmargin=\parindent,
+      numbers=left,
+      morekeywords={def, return, if, else, assert}]{code/dblAbs.java}
+        % \input{code/dblAbs.java}}
+    }
     %
-    \ffigbox[.7\textwidth] %
-    {\caption{$\mcchc$ as a directed hypergraph.
-        % 
-        Each relational predicate is depicted as a graph node
-        while each clause is represented by a hyperedge.
-        Each hyperedge is labelled by the fact in the corresponding CHC\@.
-        Each node has a valid corresponding interpretation,
-        written in braces.
-      } %
-      \label{fig:ex-graph} }
-    % { \includegraphics[width=\linewidth]{fig/ex-deps.pdf} }
-      {\input{fig/example}}
-  \end{floatrow}
-\end{figure}
-% walk through code of running example:
-%
-\cc{dblAbs} is a small procedure that doubles the absolute value of
-its input and stores the result in \cc{res}.
-% state the safety problem:
-The program also asserts that \cc{res} is greater than or equal to $0$
-before exiting.
-% talk about solving Horn Clauses:
-Verifying this assertion reduces to solving a recursion-free
-CHC system over a set of uninterpreted predicates that represent
-the control locations in \cc{dblAbs}.
-%
-In particular, one such system $\mcchc$ is
-%
+    \ffigbox[.66\textwidth]{%
+      \caption{A CHC system that models the safety condition of $\cc{dblAbs}$,
+    named $\mcchc$.}}{%
+      \label{fig:chc}
       \begin{align}
-\label{chc}
         % semantic constraint of dbl:
@@
       \end{align}
+    }
+  \end{floatrow}
+\end{figure}
+% walk through code of running example:
+%
+\cc{dblAbs} is a small procedure that doubles the absolute value of
+its input and stores the result in \cc{res}.
+% state the safety problem:
+The program also asserts that \cc{res} is greater than or equal to $0$
+before exiting.
+% talk about solving Horn Clauses:
+Verifying this assertion reduces to solving a recursion-free
+CHC system over a set of uninterpreted predicates that represent
+the control locations in \cc{dblAbs}.
+%
+In particular, one such system $\mcchc$ is shown in \autoref{fig:chc}.
+%
 %
@@
 %% talk about expansion:
-\begin{figure}
-  \input{fig/hierarchy}
+\begin{figure}[t]
+  \centering
+  \begin{floatrow}[2]
+    \ffigbox[.45\textwidth] {%
+      {\caption{$\mcchc$ as a directed hypergraph. Each relational
+          predicate is depicted as a graph node while each clause is
+          represented by a hyperedge. Each hyperedge is labelled by the
+          fact in the corresponding CHC\@. Each node has a valid
+          corresponding interpretation, written in braces.
+      }\label{fig:ex-graph}}}
+      {\input{fig/example}}
+    \ffigbox[.45\textwidth]
+      {\caption{The hierarchy of classes of recursion-free CHC
+          systems. Body Disjoint and Linear systems are subsumed by
+          CDD systems. Solving Directly Solvable CHC systems is in
+          co-NP while solving general, recursion-free systems is
+          co-NEXPTIME Complete.
+      }\label{fig:hierarchy}}
+      {\input{fig/hierarchy}}
+  \end{floatrow}
 \end{figure}

### 2018-04-11 paper/overview.tex
commit change

--- 2018-04-11 paper/overview.tex
@@
 %
-In order to find an interpretation for $\mcchc$, a solver which uses
-tree-structure systems would have to create a fresh copy of
-$\cc{L}_4$, as well copies for all relational predicates that
+In order to solve for $\mcchc$, a solver which uses
+body-disjoint systems would have to create a fresh copy of
+$\cc{L}_4$, as well copies for all clauses and relational predicates that
 $\cc{L}_4$ depends on.

### 2018-04-11 paper/background.tex

@@
 % definition of CHC
-As stated in \autoref{sec:intro} a Constrained Horn Clause is a
-logical implication where the antecedent is called the body and the
-consequent is called the head.
+A Constrained Horn Clause is a logical implication where the
+antecedent is called the body and the consequent is called the head.
 %
@@
 %
-A Clause where the head is \cc{False} is called query.
-\begin{defn}
-  CHCs can be defined structurally:
+A clause where the head is \cc{False} is called a query. A CHC can be
+defined structurally:
 \begin{align*}
@@
 \end{align*}
-\end{defn}
 %
@@
 Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
-$\siblingof{P}$ denotes the set of all predicate siblings of $\cc{P}$
+$\siblingof{P}$ denotes the set of all siblings of $\cc{P}$
 in $\cc{S}$.
@@
 %
-\sys solves CHC systems, logic-programming problems in which a
-solution maps predicate symbols to first-order formulas.
-%
 All objects defined in this paper are defined over a fixed space of
@@
 %
-In particular, for theory $\mathcal{T}$, the space of $\mathcal{T}$
+In particular, for a theory $\mathcal{T}$, the space of $\mathcal{T}$
 formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.
@@
 $\varphi$) is denoted $\vocab(\varphi)$.
+%
 % define models, satisfaction, entailment
@@
   % define:
-  For $\varphi_0, \varphi_1 \in \tformulas{\mathcal{T}}$, $I \in
-  \tformulas{\mathcal{T}}$ such that
+  For $\varphi_0, \varphi_1, I \in \tformulas{\mathcal{T}}$, if
   %
   \textbf{(1)} $\varphi_0 \entails I$, %
-  \textbf{(2)} $I, \varphi_1 \entails \false$, and %
+  \textbf{(2)} $I \land \varphi_1 \entails \false$, and %
   \textbf{(3)} $\vocab(I) \subseteq \vocab(\varphi_0) \intersection
@@
   %
-  $I$ is an \emph{interpolant} of $\varphi_0$ and $\varphi_1$.
+  then $I$ is an \emph{interpolant} of $\varphi_0$ and $\varphi_1$.
 \end{defn}
 %
-All spaces of formulas in the remainder of this paper will be defined
-for a fixed, arbitrary theory $\mathcal{T}$ that supports
+For the remainder of this paper, all spaces of formulas will be
+defined for a fixed, arbitrary theory $\mathcal{T}$ that supports
 interpolation, such as combinations of the theories of linear
@@
 %
-\sys is defined as using a decision procedure for $\mathcal{T}$ named
-\issat, and an interpolating theorem prover for $\mathcal{T}$ named
-$\solveitp$.
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "p"
-%%% End:
+We define \sys in terms of an abstract decision procedure for
+$\mathcal{T}$ named \issat, and an abstract interpolating theorem
+prover for $\mathcal{T}$ named $\solveitp$.

### 2018-04-11 paper/introduction.tex

@@
 (programs with only bounded iteration and recursion).
-
 %
-Second and most importantly, A well-known approach for solving a general CHC
-system can reduce to solving a sequence of recursion-free
-systems \QZ{reference}.
+Second and most importantly, a well-known approach for solving a
+general CHC system can reduces the input problem to solving a sequence
+of recursion-free systems \QZ{reference}.
 %
-Many solvers attempt to synthesize the solution of the original system 
-by using the solutions of constructed recruion-free system~\cite{bjorner13}.
+Such approaches attempt to synthesize a solution for the   original
+system from the solutions of recursion-free systems~\cite{bjorner13}.
 %
-Because of this, the performance of solving general CHC system is heavily reliant on the 
-performance of the solving recursion-free CHC system.
+Because of this, the performance of solving general CHC system relies
+heavily on the performance of solving recursion-free CHC systems.
 %
@@
 Such classes can be solved by issuing
-\emph{interpolation queries} for uninterpreted predicate and 
-each constrains shows only one time in the interpolation query.
+\emph{interpolation queries} to find suitable definitions for the
+uninterpreted predicates.
 %
-We
-refer such class as \emph{directly solvable}.
+We refer such classes as \emph{directly solvable}.
 
@@
 In general, solving a recursion-free CHC system for
-propositional logic and theory of linear integer arithmetic is
+propositional logic and the theory of linear integer arithmetic is
 co-NEXPTIME-complete~\cite{rummer13b}.
 %
-In contrast, solving a directly solvable system with the same logic and theories
-is in co-NP~\cite{rummer13b}.
+In contrast, solving a directly solvable system with the same logic
+and theories is in co-NP~\cite{rummer13b}.
 %
-Therefore, the performance of solving a recurion-free system is highly reliant on the
-size of the directly solvable systems it is reduced to.
+Therefore, the performance of solving a recursion-free system is
+highly reliant on the size of the directly solvable system it reduces
+to.
 
@@
 attempts to combine the solutions of these recursion-free systems to
-synthesize a soltuion to the original problem, as has been proposed
+synthesize a solution to the original problem, as has been proposed
 in previous work~\cite{rummer13b}.
@@
 \autoref{sec:related-work} compares \sys to related work.
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "p"
-%%% End:

### 2018-04-12 paper/fig/example.tex
intro and overview

@@
+\scalebox{0.75}{%
 \begin{tikzpicture}
@@
 \begin{scope}[every edge/.style={draw=black, very thick}]
-  \path[->, style=right] (L4') edge node {$\cc{abs} = 0$} (L4);
-  \path[->, style=above left] (L4) edge node {$\cc{n} \geq 0$} (L6);
-  \path[->, style=above right] (L4) edge node {$\cc{n} < 0$} (L8);
-  \path[->, style=left] (L6) edge node {$\cc{abs'} = \cc{n}$} (L9);
-  \path[->, style=right] (L8) edge node {$\cc{abs'} = -\cc{n}$} (L9);
+  \path[->, style=right] (L4') edge node {$\cc{abs} = 0$ \textbf{(2)}} (L4);
+  \path[->, style=above left] (L4) edge node {$\cc{n} \geq 0$ \textbf{(3)}} (L6);
+  \path[->, style=above right] (L4) edge node {$\cc{n} < 0$ \textbf{(4)}} (L8);
+  \path[->, style=left] (L6) edge node {$\cc{abs'} = \cc{n}$ \textbf{(5)}} (L9);
+  \path[->, style=right] (L8) edge node {$\cc{abs'} = -\cc{n}$ \textbf{(6)}} (L9);
   \path[-] (L9) edge[out=330, in=90] (main');
-  \path[->, style=right] (dbl') edge node {$\cc{d} = 2*\cc{x}$} (dbl);
+  \path[->, style=right] (dbl') edge node {$\cc{d} = 2*\cc{x}$ \textbf{(1)}} (dbl);
   \path[-] (dbl) edge[out=210, in=90] (main');
-  \path[->, style=above right] (main') edge node {$\cc{abs'} = \cc{x} \land \cc{res} = \cc{d}$} (main);
-  \path[->, style=right] (main) edge node {$\cc{res} < 0$} (bot);
+  \path[->, style=above right] (main') edge node {$\cc{abs'} = \cc{x} \land \cc{res} = \cc{d}$ \textbf{(7)}} (main);
+  \path[->, style=right] (main) edge node {$\cc{res} < 0$ \textbf{(8)}} (bot);
 \end{scope}
 \end{tikzpicture}
+}

### 2018-04-12 paper/introduction.tex
intro and overview

@@
 Second and most importantly, a well-known approach for solving a
-general CHC system can reduces the input problem to solving a sequence
-of recursion-free systems \QZ{reference}.
+general CHC system reduces the input problem to solving a sequence
+of recursion-free systems.
 %
@@
 %
-Because of this, the performance of solving general CHC system relies
+The performance of such solvers relies
 heavily on the performance of solving recursion-free CHC systems.
@@
 %
-Instead, they are reduced to a more specific subclass of
+Instead, they are rewritten to a more specific subclass of
 recursion-free CHC system.
@@
 %
-We refer such classes as \emph{directly solvable}.
+We refer to such classes as \emph{directly solvable}.
 
@@
 Therefore, the performance of solving a recursion-free system is
-highly reliant on the size of the directly solvable system it reduces
-to.
+highly reliant on the size of the synthesized directly solvable
+system.
 
@@
 clause in the system has at most one dependency on the same relational
-predicate.\QZ{repharse this}
+predicate.
 %
@@
 smaller than equivalent systems in other directly solvable classes,
-and therefore solving recursion free systems by rewriting them in CDD
+and therefore solving recursion free systems by rewriting them to CDD
 form is often less computationally expensive.
@@
 %
-The results indicate that in a strong majority of cases,
-\sys performs better than modern solvers.
+The results indicate that \sys outperforms modern solvers in a strong
+majority of cases.
 %

### 2018-04-12 paper/overview.tex
intro and overview

--- 2018-04-11 paper/overview.tex
@@
 In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
-$\mcchc$, that models the safety of a small program,
+$\mcchc$ (\autoref{fig:chc}), that models the safety of a small program,
 \cc{dblAbs} (\autoref{fig:multicall-code}).
@@
 %
-In particular, one such system $\mcchc$ is shown in \autoref{fig:chc}.
+In particular, one such system, $\mcchc$, is shown in
+\autoref{fig:chc}.
 %
@@
 % state the defn of CDD:
-A CHC system can be classified as CDD when all clauses are satisified
-the following rules:
+A CHC system can be classified as CDD when each clause satisfies the
+following rules:
 %
-\textbf{(1)} no two predicates in the body of the same clause share any transitive
-dependencies on other predicates(i.e. dependence disjoint) 
-and \textbf{(2)} no clause has more than one same predicate in its body.
+\textbf{(1)} no two predicates in the body of the same clause share
+any transitive dependencies on other predicates and
+\textbf{(2)} no clause has more than one occurrence of a given
+predicate in the body.
 % give a example
@@
 %
-And there is no clauses has more than one same predicate in its body.
-%
 Therefore $\mcchc$ is a CDD system.
@@
 %
-Each interpretation of uninterpreted predicate $P$ can
-be computed by interpolating
+Each interpretation of a predicate $P$ can be computed by
+interpolating
 \textbf{(1)} the \emph{pre}-formula, constructed from clauses where
-$P$ is the head and
+$P$ is the head, and
 \textbf{(2)} the \emph{post}-formula, constructed from
@@
 
-For example, when \sys try to solve $\cc{L}_9$, it has partial solution that 
-map $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$ to their interpretation.
+For example, consider $\cc{L}_9$. By the time \sys attempts to
+synthesize an interpretation for $\cc{L}_9$ it will have solutions for
+$\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$.
 %
@@
 \autoref{fig:ex-graph}.
-%
-
-%
-A interpretation of $\cc{L}_9$ can be constructed by interpolating
-the appropriate pre and post formulas.
 %
@@
 %
-Each
-boolean indicator variable implies the solution for its predicate which is encoded
-as the disjunction of the negation of the boolean indicator variable and the solution.
+Each boolean indicator variable implies the solution for its
+predicate,
+encoded as the disjunction of the negation of the boolean
+indicator variable and the solution.
 %
@@
 \begin{gather}
-  (\cc{b}_{\cc{main}} \land \cc{res'} < 0)
+  (\neg \cc{b}_{\cc{dbl}} \lor \cc{d} = 2*\cc{x})
   \land
@@
     \land \cc{res}=\cc{d} ))
-  \land (\neg \cc{b}_{\cc{dbl}} \lor \cc{d} = 2*\cc{x})
+  \land
+  (\cc{b}_{\cc{main}} \land \cc{res'} < 0)
 \end{gather}
 %
-Interpolating the pre and post formulas yields a intepretation
+Interpolating the pre and post formulas yields an intepretation
 of $\cc{L}_9$: $\cc{abs'} \geq 0$.
@@
 % tree solver
-Body-disjoint systems(i.e derivation tree)
+Body-disjoint (or derivation tree)
 systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
-class of recursion-free CHC system where each uninterpreted predicate
-appears in the body of at most one clause and appears in such a
-clause exactly once.
+a class of recursion-free CHC system where each uninterpreted
+predicate appears in the body of at most one clause and appears in
+such a clause exactly once.
 %
@@
 %
-$\mcchc$ is not a body-disjoint system because the uninterpreted
-predicates $\cc{L}_4$ shows in the body of both clause (3) and clause
-(4).
+$\mcchc$ is not a body-disjoint system because $\cc{L}_4$ shows in the
+body of both clause (3) and clause (4).
 %
@@
 %linear systems:
-Previous work also introduced linear system~\cite{albarghouthi12a},
-where each clause's body has at most one uninterpreted predicate.
+Previous work has also introduced the class of linear
+systems~\cite{albarghouthi12a}, where the body of each has at most
+one uninterpreted predicate.
 %
@@
 %
-CHC solvers that solve linear systems effectively inline the
+CHC solvers that use linear systems effectively inline the
 constraints for relational predicates that occur in non-linear
@@
 efficient, but in general such approaches can generate systems that
-are exponentially larger than the input.
+are exponentially larger than the input. For example, if the procedure
+$\cc{dbl}$ were called more than once in $\cc{dblsAbs}$ then multiple
+copies of the body of the procedure would be inlined. And if the body
+of this procedure were large, the inlining would quickly become
+prohibitively expensive.
 
@@
 
-The class CDD of CDD systems is a superclass of union of the classes of
-body-disjoint and linear systems.
+The hierarchy of discussed classes of recursion-free systems is
+depicted in \autoref{fig:hierarchy}.
 %
-% The fact that $\mcchc$ is CDD corresponds to the fact that there no
-% statements executed more than once within one execution path.
-% \DAH{elaborate}
+As shown, the class of CDD systems is a superclass of both the class
+of body-disjoint systems and the class of linear systems. So any
+recursion-free system that is efficiently expressible in body-disjoint
+or linear form is also efficiently expressible in CDD form.
 %
-E.g., a recursion-free system that models \cc{dblAbs} 
-call \cc{dbl} twice on \cc{abs} in sequence would not
-be CDD.
+In addition, some systems which are expensive to express in
+body-disjoint or linear form are efficiently expressible in CDD form.
 %
-\sys, given an arbitrary recursion-free CHC system $S$,
-solves it by generating an equivalent CDD system $S'$ and
-solving $S'$ directly.
+\sys~takes advantage of this fact when solving input systems.
+Given an arbitrary recursion-free CHC system $S$, \sys~generates an
+equivalent CDD system $S'$ and solves $S'$ directly.
 %
@@
 %
-In practice, a suitable equivalent system can be generated using
-heuristics analogous to those used to generate compact verification
-conditions of hierarchical programs~\cite{flanagan01,lal-qadeer15}.
+However, \sys~generates equivalent systems via heuristics analogous to
+those used to generate compact verification conditions of hierarchical
+programs~\cite{flanagan01,lal-qadeer15}.
+%
+In practice these heuristics often yield CDD systems which are
+small with respect to the input system.
 %
@@
 system is given in \autoref{app:cons-cdd}.
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "p"
-%%% End:

### 2018-04-12 paper/introduction.tex
replaced venn diagram with tree

@@
 These classes include those of
-\emph{body disjoint} (or \emph{derivation tree})
+\emph{body-disjoint} (or \emph{derivation tree})
 systems~\cite{heizmann10,bjorner13,mcmillan14,rummer13a,rummer13b} and

### 2018-04-12 paper/background.tex

@@
 %
-A head can be either an uninterpreted predicate applied to variables or \cc{False}.
+A head can be either an uninterpreted predicate applied to variables or $False$.
 %
-A clause where the head is \cc{False} is called a query. A CHC can be
+A clause where the head is $False$ is called a query. A CHC can be
 defined structurally:
 \begin{align*}
-\cc{chc} &\Coloneqq \cc{head} \gets \cc{body} \\
-\cc{head} &\Coloneqq \cc{False} \\&\mid \cc{pred}  \\
-\cc{body} &\Coloneqq \varphi \wedge \cc{preds} \\
-\cc{preds} &\Coloneqq \cc{True} \\&\mid \cc{pred} \wedge \cc{preds} \\
-\cc{pred} &\Coloneqq \textsf{an uninterpreted predicate applied to variables} \\
-\varphi &\Coloneqq \textsf{a formula that all variables are vocabulary} \\
+\cc{chc} \Coloneqq&~\cc{head} \gets \cc{body} \\
+\cc{head} \Coloneqq&~False \\
+  \mid&~\cc{pred} \\
+\cc{body} \Coloneqq&~\varphi \wedge \cc{preds} \\
+\cc{preds} \Coloneqq&~True \\
+  \mid&~\cc{pred} \wedge \cc{preds} \\
+\cc{pred} \Coloneqq&~\textsf{\emph{an uninterpreted predicate applied to variables}} \\
+\varphi \Coloneqq&~\textsf{\emph{a formula}} \\
 \end{align*}
@@
 %
-It is possible for an uninterpreted predicate to appear multiple times
-in the body of $C$.
-%
 If $C$ is not a query, then $\headof{C}$ denotes the uninterpreted
@@
 %
-
 %define a CHC system
@@
 
+To explain the structure of a CDD system, we need terminology that relates
+predicates in a CHC system including the terms \emph{predicate dependency},
+\emph{transitive predicate dependency}, and \emph{sibling}.
 % define dependency predicates
@@
 %
-
 %define transitive dependency predicates
@@
 %
-
 %define siblings
@@
   Because uninterpreted predicates $\cc{L}_9$ and $\cc{db}$ both
-  appear in the body of clause (6), $\cc{L}_9$ and $\cc{db}$ are
+  appear in the body of clause (7), $\cc{L}_9$ and $\cc{db}$ are
   siblings.
@@
 %
-All objects defined in this paper are defined over a fixed space of
-first-order logical variables $X$.
+All logical objects in this paper are defined over a fixed space of first-order
+variables, $X$.
 %
-In particular, for a theory $\mathcal{T}$, the space of $\mathcal{T}$
+For a theory $\mathcal{T}$, the space of $\mathcal{T}$
 formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.

### 2018-04-12 paper/fig/hierarchy.tex

--- 2018-04-10 paper/fig/hierarchy.tex
@@
-\scalebox{0.75}{%
 \begin{tikzpicture}
 \begin{scope}
-  \node (top) at (0,0) {\begin{tabular}{c}All Systems\\ \textsc{Undecidable}\end{tabular}};
-    \node[below = 0.5 of top] (rec) {\begin{tabular}{c}Recursion Free\\
-  \textsc{co-Nexptime Complete}\end{tabular}};
-  \node[below = 0.5 of rec] (dir) {\begin{tabular}{c}Directly Solvable\\ \textsc{co-NP}\end{tabular}};
-  \node[label={0:\{\textsc{Shara}\}}, draw, rounded corners=0.5cm,
-      below = 0.5 of dir] (cdd) {\begin{tabular}{c}Clause-Dependent-Disjoint\\ \textsc{co-NP}\end{tabular}};
-  \node[label={0:\{\textsc{Ufo}\}},
-      below left = 0.5 and 0.5 of cdd] (lin) {\begin{tabular}{c}Linear\\ \textsc{co-NP}\end{tabular}};
-  \node[label={180:\{\textsc{Eldarica}\cite{rummer13b}\}},
-      below right = 0.5 and 0.5 of cdd] (bd) {\begin{tabular}{c}Body Disjoint\\ \textsc{co-NP}\end{tabular}};
-  \node[label={180:\{\textsc{Duality}\cite{bjorner13}\}},
-      below = 0.5 of bd] (tree) {\begin{tabular}{c}Tree\\ \textsc{co-NP}\end{tabular}};
+  \node at (0, 0) (rec) {%
+    \begin{tabular}{c}
+      Recursion Free\\\textsc{co-Nexptime Complete}
+    \end{tabular}
+  };
+
+  \node[below = 0.5 of rec] (dir) {%
+    \begin{tabular}{c}
+      Directly Solvable\\\textsc{co-NP}
+    \end{tabular}
+  };
+
+  \node[draw, rounded corners=0.5cm, below = 0.5 of dir] (cdd) {%
+    \begin{tabular}{c}
+      Clause-Dependence Disjoint\\\textsc{co-NP}
+    \end{tabular}
+  };
+
+  \node[below left = 0.5 and -1.1 of cdd] (lin) {%
+    \begin{tabular}{c}
+      Linear\\\textsc{co-NP}
+    \end{tabular}
+  };
+
+  \node[below right = 0.5 and -1.1 of cdd] (bd) {%
+    \begin{tabular}{c}
+      Body-Disjoint\\\textsc{co-NP}
+    \end{tabular}
+  };
+
 \end{scope}
 \begin{scope}[every edge/.style={draw=black, very thick}]
-  \path[-] (tree) edge (bd);
   \path[-] (bd) edge (cdd);
@@
   \path[-] (dir) edge (rec);
-  \path[-] (rec) edge (top);
 \end{scope}
 \end{tikzpicture}
-}

### 2018-04-12 paper/overview.tex
added motivation for definitions

@@
         %
-        \bot &\gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
+        False &\gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
       \end{align}

### 2018-04-12 paper/approach.tex
pass through 4.2.1

@@
 The key contribution of our work is the introduction of the class of
-Clause-Dependence Disjoint (CDD) systems.
-%
-Solving an CHC system in CDD is coNP time.
-%
-CDD strictly contains
-the union of classes of recursion-free systems that can be solved
-in coNP time that have been introduced in previous work.
+Clause-Dependence Disjoint (CDD) systems. As discussed in
+\autoref{sec:overview}, CDD systems \textbf{(1)} are a superclass of other
+practical classes of directly solvable CHC systems and \textbf{(2)} can be
+solved in co-NP.
 %
@@
   %
-  if for all pair of uninterpreted predicates $P, Q \in \predof{S}$ that are siblings(\autoref{sec:chcs}), the 
-  transitive dependences of $P$ and $Q$ are disjoint $\tdepsof{P} \cap \tdepsof{Q} = \emptyset$, and there is 
-  no uninterpreted predicate shows more than once in a body of clause, then $\cc{S}$ is
-  \emph{Clause-Dependence-Disjoint (CDD)}.
+  if for all sibling pairs, $P, Q \in \predof{S}$,
+  %
+  the transitive dependences of $P$ and $Q$ are disjoint ($\tdepsof{P} \cap
+  \tdepsof{Q} = \emptyset$)
+  %
+  and no predicate is a sibling of itself,
+  %
+  then $\cc{S}$ is \emph{Clause-Dependence Disjoint (CDD)}.
 \end{defn}
 %
-CDD systems model hierarchical programs with branching and
-procedure calls that in each execution path that each statments be
-exected at most once.
+CDD systems model hierarchical programs with branches and procedure calls such
+that each execution path invokes each statement at most once.
 %
 \begin{ex}
-  The CHC system $\mcchc$ (given in \autoref{sec:solve-ex}) is a CDD
-  system.
-  %
-  The only clauses in $\mcchc$ that its body has multiple uninterpreted
-  predicates are clause (7) that contains $\cc{L9}$ and $\cc{db}$.
-  %
-  The transitive dependences of $\cc{L9}$ and $\cc{db}$ are \{\cc{L6}\,\cc{L8}\,\cc{L4}\} and $\varnothing$, which are disjoint.
+  The CHC system $\mcchc$ is a CDD system. An argument is given in
+  \autoref{sec:solve-ex}.
 \end{ex}
 
-% talk about body disjoint
-The class of CDD systems contains tree structure system, and body-disjoint and linear
+% talk about previous system
+The class of CDD systems contains body-disjoint and linear
 systems introduced in previous work.
-%
-Tree structure system is contained by body-disjoint system. 
 %
@@
 %
-The proof is given in \autoref{app:char}.
+Proof given in \autoref{app:char}.
 
@@
        \lCase{\cc{SAT}:}{\Return{$\nosoln$}}
-       \lCase{$\cc{I}$:}{$\sigma$ [$\cc{P} \to \cc{I}$]}
+       \lCase{$\cc{I}$:}{$\sigma$ [$\cc{P}$] $\assign$ $\cc{I}$}
      }
@@
 % introduce solveCDD
-the procedure $\solvecdd$ (\autoref{alg:solve-cdd}) is a solver for CDD systems.
+\autoref{alg:solve-cdd} presents $\solvecdd$, a procedure designed to solve
+CDD systems.
 %
-$\solvecdd$, given a CDD system $\cc{S}$, it first do a topological sort on all
-uninterpreted predicates in $\cc{S}$ based on their dependency relations.
+Given a CDD system $\cc{S}$, $\solvecdd$ topologically sorts the uninterpreted
+predicates in $\cc{S}$ based on their dependency relations.
 %
-In the order of topological sort, for each uninterpreted predicates $\cc{P}$ 
- $\solvecdd$ calles $\solveitp$ to compute the binary interpolation of 
- encoded pre and post formulas with the current solution.
+Then, the algorithm calculates interpretations for each predicate in this order
+by invoking $\solveitp$.
 %
- If the conjunction of pre and post formulas is satisifiable, then $\solvecdd$
- returns $\nosoln$ to indicate this CDD system $\cc{S}$ is not solvable.
+$\solveitp$ computes a binary interpolant of the pre and post formulas of the
+given predicate, where these formulas are based on the current, partial
+solution.
 %
- Otherwise, $\solvecdd$ binds the uninterpreted predicate $\cc{P}$ to the
- interpolation result $\cc{I}$ as part of the soltuion.
+The pre and post formulas are computed respectively by $\prectr$ and
+$\postctr$, which we define below.
 %
- The encoding procedure $\prectr$ and $\postctr$ are defined below.
+It is possible that the pre and post formulas may be mutually satisifiable, in
+which case $\solveitp$ returns \cc{SAT}. In this case, $\solvecdd$ returns
+$\nosoln$ to indicate that $\cc{S}$ is not solvable.
 %
- $\solvecdd$ returns the soltuion $\sigma$.
+Otherwise, $\solvecdd$ modifies the partial solution by setting the
+interpretation of $\cc{P}$ to $\cc{I}$.
+%
+Once all predicates have been interpolated, $\solvecdd$ returns the complete
+solution, $\sigma$.
 
@@
 \begin{ex}
-  $\solvecdd$, given a CDD system $\mcchc$ in \autoref{sec:running-ex}, may generate
-  interpretations queries in any topological ordering of the
-  dependency relation of $\mcchc$.
+  Given the CDD system $\mcchc$, $\solvecdd$ may generate interpolation
+  queries in any topological ordering of the dependency relations.
   %
-  One such ordering is \cc{L4}, \cc{L6}, \cc{L8}, 
-   \cc{L9}, \cc{db} and \cc{main}.
+  One such ordering is $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$, $\cc{L}_9$, \cc{db},
+  \cc{main}.
 \end{ex}
 
-\subsubsection{Constructing constrains of uninterpreted 
-predicate}
+\subsubsection{Constructing constraints for predicates}
 %
-For a CDD system $S$, given a uninterpreted predicate $\cc{P} \in \predof{S}$, and a partial 
-solution $\sigma$ that maps some uninterpreted predicate to its solution formula, 
-,the formula $\ctrof{P,\sigma}$ is a compact representation of constrains of $\cc{P}$ based on parital
-solution $\sigma$, defined as follows.
-% for each relational predicate, define constraint:
-If partial solution $\sigma$ does not contains $P$, 
-the constraint of $P$ under $\sigma$ that denoted as $\ctrof{P,\sigma}$ 
-is constructed as: for all clauses $C \in S$ in which $P$ is the head of $\cc{C}$
-\textbf{(1)} each predicate $Q \in \bodyof{C}$ is used and 
-\textbf{(2)} the constrained of the clause $\consof{C}$ is encoded.
+For a CDD system $S$, a predicate $\cc{P} \in \predof{S}$, and a partial
+solution $\sigma$ that maps predicates to their solutions, the formula
+$\ctrof{P,\sigma}$ is a compact representation of the constraints of $\cc{P}$.
 %
-I.e., $\ctrof{P,\sigma}$ is
+If $\sigma$ does not contain $P$, then the constraint of $P$, denoted
+$\ctrof{P,\sigma}$, is constructed from the clauses where $P$.  When $\sigma$
+does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$:
+%
 \[
-\biglor_{ (\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
+  \ctrof{P,\sigma}=
+  \begin{dcases}
+    \biglor_{(\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
     \left( \consof{C_i} \land %
-\bigland_{ \cc{Q} \in \bodyof{C_i}} %
-b_Q \right)
+      \bigland_{ \cc{Q} \in \bodyof{C_i}} \cc{b}_Q
+    \right),
+  &\text{if } P \notin \sigma\\
+  \sigma[P], &\text{if } P \in \sigma
+  \end{dcases}
 \]
 %
-If partial solution $\sigma$ contains $P$, then $\ctrof{P,\sigma}$ is the solution formula $\sigma[P]$.
-
 % introduce the isused predicates:
-For each $\cc{P} \in \predof{S}$, there is a corresponding boolean variable
-$b_P$.
+Each $\cc{P} \in \predof{S}$ has a corresponding boolean variable $b_P$.
 %
-The counterexample characterization of $P$ constrains that
-if $P$ is used, which means $b_{P}$ is \cc{True}, then its constraint must hold.
+The counterexample characterization of $P$ constrains that if $P$ is used,
+(meaning $\cc{b}_{P} = True$) then its constraint must hold:
 %
-I.e., $\vc{P,\sigma}$ is $\neg b_{P} \lor \ctrof{P,\sigma}$.
+$\vc{P,\sigma} = \neg \cc{b}_{P} \lor \ctrof{P,\sigma}$.
 
@@
   \label{ex:ctr}
-  In the CHC system $\mcchc$ (given in \autoref{sec:solve-ex}), 
-  when $\mcsolve$ solves the uninterpreted predicate \cc{L9}, it 
-  has a partial solution $\sigma$ that maps \cc{L6} to $n \ge 0$.
-  %
-  For predicate \cc{L6}, the the $\ctrof{L6,\sigma}$ is $n \ge 0$.
-  %
-  For the 
-  uninterpreted \cc{L9}, the $\ctrof{L9,\sigma}$ is $(abs' = n \land b_{L6}) \lor (abs' = n \land b_{L8})$.
-  %
-  The $\vc{L9,\sigma}$ is $\neg b_{L9} \lor \ctrof{L9,\sigma}$.
+  When $\mcsolve$ solves the predicate $\cc{L}_9$ in $\mcchc$, it generates a
+  constraint based on clauses (5) and (6):
+  $\ctrof{\cc{L}_9,\sigma} =
+    (\cc{abs'} = \cc{n} \land \cc{b}_{\cc{L}_6})
+    \lor
+    (\cc{abs'} = -\cc{n} \land \cc{b}_{\cc{L}_8})$.
+  The counterexample characterization for $\cc{L}_9$ is based on its boolean
+  indicator and this constraint:
+  $\vc{\cc{L}_9,\sigma} = \neg \cc{b}_{\cc{L}_9} \lor \ctrof{\cc{L}_9,\sigma}$.
 \end{ex}

### 2018-04-12 paper/approach.tex
rewrote most of approach

@@
 The key contribution of our work is the introduction of the class of
-Clause-Dependence Disjoint (CDD) systems. As discussed in
-\autoref{sec:overview}, CDD systems \textbf{(1)} are a superclass of other
-practical classes of directly solvable CHC systems and \textbf{(2)} can be
-solved in co-NP.
+Clause-Dependence Disjoint (CDD) systems. Critically, CDD systems can be solved
+in co-NP time by interpolation of the uninterpreted relational predicates.
 %
@@
 % talk about previous system
-The class of CDD systems contains body-disjoint and linear
-systems introduced in previous work.
+As discussed in \autoref{sec:overview}, CDD is a superclass of the
+class of body-disjoint systems and the class of linear systems.
 %
-For recursion-free system $S$, if for each uninterpreted predicate
-$Q \in \predof{S}$, there is at most one caluse $C$ that $Q \in \bodyof{C}$,
-and there is no uninterpreted predicate shows more than once in a body of clause, then
-$S$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
+For a given recursion-free system $S$, if each uninterpreted predicate $Q \in
+\predof{S}$, appears in at most one clause $C$ ($Q \in \bodyof{C}$) and no
+predicate appears more than once in the body of a single clause,
+then $S$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
 %
-If the body of each clause in $\cc{S}$ contains at most
-relational predicate, then $\cc{S}$ is
-\emph{linear}~\cite{albarghouthi12a}.
+If the body of each clause in $\cc{S}$ contains at most one relational
+predicate, then $\cc{S}$ is \emph{linear}~\cite{albarghouthi12a}.
 
@@
     $\cc{Preds} \assign \topSort(\predof{S})$ \\
-    \For {$\cc{P} \in \cc{Preds}$}
-    {$\cc{interpolant} \assign \solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$\\
+    \For {$P \in \cc{Preds}$}
+    {$\cc{interpolant} \assign \solveitp(\prectr(P,\sigma),\postctr(P,\sigma))$\\
      \Switch{\cc{interpolant}}{
        \lCase{\cc{SAT}:}{\Return{$\nosoln$}}
-       \lCase{$\cc{I}$:}{$\sigma$ [$\cc{P}$] $\assign$ $\cc{I}$}
+       \lCase{$\cc{I}$:}{$\sigma$ [$P$] $\assign$ $\cc{I}$}
      }
@@
 The pre and post formulas are computed respectively by $\prectr$ and
-$\postctr$, which we define below.
+$\postctr$, which we define in \autoref{sec:cons-pre} and \autoref{sec:cons-post}.
 %
@@
 Otherwise, $\solvecdd$ modifies the partial solution by setting the
-interpretation of $\cc{P}$ to $\cc{I}$.
+interpretation of $P$ to $\cc{I}$.
 %
@@
   %
-  One such ordering is $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$, $\cc{L}_9$, \cc{db},
+  One such ordering is $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$, $\cc{L}_9$, \cc{dbl},
   \cc{main}.
@@
 %
-For a CDD system $S$, a predicate $\cc{P} \in \predof{S}$, and a partial
+For a CDD system $S$, a predicate $P \in \predof{S}$, and a partial
 solution $\sigma$ that maps predicates to their solutions, the formula
-$\ctrof{P,\sigma}$ is a compact representation of the constraints of $\cc{P}$.
+$\ctrof{P,\sigma}$ is a compact representation of the constraints of $P$.
 %
-If $\sigma$ does not contain $P$, then the constraint of $P$, denoted
-$\ctrof{P,\sigma}$, is constructed from the clauses where $P$.  When $\sigma$
-does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$:
+If $\sigma$ does not contain $P$, then the constraint of $P$ is constructed
+from the clauses where $P$.  When $\sigma$ does contain $P$, $\ctrof{P,\sigma}$
+is a lookup from $\sigma$:
 %
@@
   \begin{dcases}
-    \biglor_{(\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
+    \biglor_{(\cc{C}_i \in \cc{S}) \land (\headof{C_i} = P)}
     \left( \consof{C_i} \land %
@@
 % introduce the isused predicates:
-Each $\cc{P} \in \predof{S}$ has a corresponding boolean variable $b_P$.
+Each $P \in \predof{S}$ has a corresponding boolean variable $\cc{b}_P$.
 %
@@
 (meaning $\cc{b}_{P} = True$) then its constraint must hold:
-%
-$\vc{P,\sigma} = \neg \cc{b}_{P} \lor \ctrof{P,\sigma}$.
+\[
+  \vc{P,\sigma} = \neg \cc{b}_{P} \lor \ctrof{P,\sigma}
+\]
 
@@
   \label{ex:ctr}
-  When $\mcsolve$ solves the predicate $\cc{L}_9$ in $\mcchc$, it generates a
+  When $\solvecdd$ solves the predicate $\cc{L}_9$ in $\mcchc$, it generates a
   constraint based on clauses (5) and (6):
-  $\ctrof{\cc{L}_9,\sigma} =
+  $$\ctrof{\cc{L}_9,\sigma} =
     (\cc{abs'} = \cc{n} \land \cc{b}_{\cc{L}_6})
     \lor
-    (\cc{abs'} = -\cc{n} \land \cc{b}_{\cc{L}_8})$.
+    (\cc{abs'} = -\cc{n} \land \cc{b}_{\cc{L}_8})$$
   The counterexample characterization for $\cc{L}_9$ is based on its boolean
   indicator and this constraint:
-  $\vc{\cc{L}_9,\sigma} = \neg \cc{b}_{\cc{L}_9} \lor \ctrof{\cc{L}_9,\sigma}$.
+  $$\vc{\cc{L}_9,\sigma} = \neg \cc{b}_{\cc{L}_9} \lor \ctrof{\cc{L}_9,\sigma}$$
 \end{ex}
@@
 % pre-constraint:
-\subsubsection{Constructing the pre-constraint of a relational predicate}
+\subsubsection{Constructing pre-formulas for predicates}
 \label{sec:cons-pre}
 %
-For a CDD system $\cc{S}$, given a uninterpreted predicate $\cc{P} \in \predof{\cc{S}}$
-, because $\solvecdd$ solve each uninterpreted predicate in a toplogical order, 
-it has a partial solution $\sigma : \cc{Q} \to \formula$ for all uninterpreted predicates in
-$\depsof{P}$.
+$\prectr(P, \sigma)$ denotes the pre-formula for an arbitrary predicate $P$
+with respect to the partial solution map, $\sigma$.
 %
-the pre-constraint $\prectr(\cc{P},\sigma)$ is a formula that the conjunction of the constrains
-of $\cc{P}$ and solutions of all dependencies of $P$ with their boolean varialbe.
+Due to the topological ordering, when $\solvecdd$ attempts to solve $P$, the
+interpretations for all dependencies of $P$ will be stored in
+$\sigma$.
 %
-In particular, $\prectr(\cc{P},\sigma)$ is:
-\[\ctrof{P,\sigma} \land
+The pre-formula is built from these interpretations together with boolean
+indicators of the dependencies and the constraint of $P$:
+%
+\[
+\prectr(P,\sigma) =
+  \ctrof{P,\sigma} \land
     \left(
@@
   \label{ex:pre-ctr}
-  In the CHC system $\mcchc$ (given in \autoref{sec:solve-ex}), 
-  when $\mcsolve$ solves the uninterpreted predicate \cc{L9}, because it solves
-  in topological order, partial solution $\sigma$ contains \cc{L6} and \cc{L8}.
+  When $\solvecdd$ solves the uninterpreted predicate $\cc{L}_9$ in $\mcchc$,
+  $\sigma$ contains $\cc{L}_6$ and $\cc{L}_8$.
   %
-  $\sigma$ maps
-  \cc{L6} to $n \ge 0$ and \cc{L8} to $n < 0$.
+  $\sigma$ maps $\cc{L}_6$ to $n \ge 0$ and $\cc{L}_8$ to $n < 0$.
   %
-  The pre-formula for \cc{L9} under $\sigma$ is $\ctrof{L9} \land (\neg b_{L6} \lor n \ge 0) \land (\neg b_{L8} \lor n < 0)$.
+  The pre-formula for $\cc{L}_9$ under $\sigma$ is therefore:
+  $$\ctrof{\cc{L}_9}
+  \land (\neg \cc{b}_{\cc{L}_6} \lor n \ge 0) \land (\neg \cc{b}_{\cc{L}_8}
+  \lor n < 0)$$
   %
-  The forumula of $\ctrof{L9,\sigma}$ is given in~\autoref{ex:ctr}.
+  The formula $\ctrof{\cc{L}_9,\sigma}$ is given in~\autoref{ex:ctr}.
 \end{ex}
@@
   predicate}
-\label{sec:cons-ctx}
+\label{sec:cons-post}
+$\postctr(P, \sigma)$ denotes the post-formula for an arbitrary predicate $P$
+with respect to the partial solution map, $\sigma$. A valid post-constraint is
+mutually inconsistent with the solution of $P$.
 %
-For $\sigma$ as a partial solution of
-$S$ and $P \in \predof{S}$, the post-constraint $\postctr(P, \sigma)$ is mutually
-inconsistent with the solution of $P$ in each solution of
-$S$ that extends $\sigma$.
+The post formula is constructed based on the predicates which depend on $P$.
+%
 % dependents of R:
-In particular, let $D_0$ be the reflexive transitive dependents of $P$
-in $S$,
+Let $D_0$ be the reflexive transitive dependents of $P$ in $S$,
 % clause siblings of dependents:
-let $D_1$ be the siblings in $S$ of
-$(D_0 \union {P})$,
+let $D_1$ be the siblings in $S$ of $(D_0 \union {P})$,
 %
@@
 % define instantiation of C
-The post constraint for $P$ under $\sigma$ is the conjunction of
-counterexample characterization of all predicates $Q \in D$ and the query clause.
+The post constraint for $P$ under $\sigma$ is the conjunction of counterexample
+characterization of all predicates $Q \in D$ and the query clause.
 %
-I.e., $\postctr(R, \sigma) =\cc{query} \land (\bigland_{\cc{Q} \in D} \vc{Q,\sigma})$.
+\[
+\postctr(R, \sigma) =\cc{query} \land (\bigland_{\cc{Q} \in D} \vc{Q,\sigma})
+\]
 % example post constraint:
@@
   \label{ex:ctx-ctr}
-  Consider the case in which $\mcsolve$, given $\mcchc$, generates an
-  interpretation for \cc{L9}, introduced in \autoref{ex:pre-ctr}.
+  When $\solvecdd$ solves $\cc{L}_9$ in $\mcchc$, it must consider
+  the dependents of $\cc{L}_9$.
   %
-  The reflexive transitive dependents $D_0$ of \cc{L9} is \{\cc{main}\}.
+  The transitive dependents $D_0$ of $\cc{L}_9$ is \{\cc{main}\}.
   %
-  The siblings set $D_1$ is \{\cc{db}\}.
+  The siblings set $D_1$ is \{\cc{dbl}\}.
   %
@@
   %
-  Then the $D$ set is \{\cc{main},\cc{db}\}
+  Therefore, the $D$ set is \{\cc{main},\cc{dbl}\}.
   %
-  The \cc{query} is $b_{main} \land res'<0$.
+  The \cc{query} is $\cc{b}_{main} \land res'<0$.
   %
-  The post constraint for $L9$ under $\sigma$ is $\cc{query} \land \vc{main,\sigma} \land \vc{db,\sigma}$.
-  %
-  The whole formula is given in\QZ{add reference}. 
+  The post constraint for $\cc{L}_9$ under $\sigma$ is
+  $$\cc{query} \land \vc{\cc{main},\sigma} \land \vc{\cc{dbl},\sigma}$$
 \end{ex}
@@
 % give the general procedure:
-$\shara$, given an recursion-free CHC system $S$,
-constructs a CDD system $S'$ equivalent to $S$.
+$\shara$, given a recursion-free CHC system $S$,
+constructs an equivalent CDD system $S'$.
 %
-\sys then directly solves $S'$ and from its solution,
-constructs a solution of $S$.
+\sys then directly solves $S'$ and, from this solution, constructs a solution
+for $S$.
 % define an expansion of a CHC system:
-For two given recursion-free CHC system $S$ and $S'$, if
+For two given recursion-free CHC systems $S$ and $S'$, if
 there is a homomorphism from $\predof{S'}$ to $\predof{S}$ that
@@
   % preserves clause structure:
-  \textbf{(2)} for all clause $C' \in S'$, the new clause $C$
-  by substituting all predicates $P'$ by $f(P')$ is in $S$.
+  \textbf{(2)} for each clause $C' \in S'$, the clause $C$ (constructed
+  by substituting all predicates $P'$ by $f(P')$) is in $S$.
   %
@@
 %
-If there is a correspondence from $S'$ to $S$,
-$S'$ is an \emph{expansion} of $S$, denoted
-$S \expandsto S'$.
+If there is a correspondence from $S'$ to $S$, then $S'$ is an \emph{expansion}
+of $S$, denoted $S \expandsto S'$.
 %
@@
 % walk through the sys algorithm:
-$\shara$ (\autoref{alg:shara}), given an recursion-free CHC system $S$ (\autoref{line:shara-begin}), 
+\sys (\autoref{alg:shara}), given a recursion-free CHC system $S$
+(\autoref{line:shara-begin}), 
 returns a solution
@@
 %
-$\shara$ first runs a procedure $\expand{}$ on $S$ to obtain a CDD expansion $S'$ (\autoref{defn:min-cdd-expansion})
-of $S$.
+\sys first runs a procedure $\expand{}$ on $S$ to obtain a CDD expansion
+$S'$ of $S$.
 %
-$\shara$ then runs a procedure $\solvecdd$ on $S'$, which either returns the value
-$\none$ to denote that $S'$ has no solution or a solution
-$\sigma'$ (\autoref{line:shara-case}).
+\sys then invokes $\solvecdd$ on $S'$.
 %
-If $\solvecdd$ returns that $S'$ has
-no solution, then \sys returns that $S$ has no solution
-(\autoref{line:shara-ret-none}).
+When \solvecdd~returns that $S'$ has
+no solution, \sys~propogates $\none$ (\autoref{line:shara-ret-none}).
 %
-Otherwise, if $\solvecdd$ returns a solution
-$\sigma'$, then $\sys$ returns, as a solution to $S$,
-$\collapse{ \eta }{ \sigma' }$, defined below
-(\autoref{line:shara-ret-collapse}).
+Otherwise, \sys constructs a solution from the CDD solution, $\sigma'$, by
+invoking $\collapse{ \eta }{ \sigma' }$ (\autoref{line:shara-ret-collapse}).
 

### 2018-04-12 paper/approach.tex
one pass for approach

@@
 \autoref{sec:core-solver} describes how \sys solves a given
-recursion-free system by solving an equivalent CDD system.
+recursion-free system by solving an CDD system.
 %
@@
 The key contribution of our work is the introduction of the class of
-Clause-Dependence Disjoint (CDD) systems. Critically, CDD systems can be solved
-in co-NP time by interpolation of the uninterpreted relational predicates.
+Clause-Dependence Disjoint (CDD) systems.
+%
+CDD systems can be solved
+in co-NP time and contains the union of all previous introduced classes
+that can be solved in co-NP time.
 %
@@
   %
-  and no predicate is a sibling of itself,
+  and no predicate shows more than once in the body of a single clause,
   %
@@
 For a given recursion-free system $S$, if each uninterpreted predicate $Q \in
-\predof{S}$, appears in at most one clause $C$ ($Q \in \bodyof{C}$) and no
+\predof{S}$, appears in at most one in the body of a clause and, no
 predicate appears more than once in the body of a single clause,
@@
 %
-Proof given in \autoref{app:char}.
+Proof is given in \autoref{app:char}.
 
@@
 %
-Otherwise, $\solvecdd$ modifies the partial solution by setting the
+Otherwise, $\solvecdd$ updates the partial solution by setting the
 interpretation of $P$ to $\cc{I}$.
@@
 \end{ex}
+
+% solve-cdd in coNP
+\begin{thm}
+  \label{thm:solve-cp}
+  Given a CDD system $S$, $\solvecdd$ either returns the solution of $S$ or $\none$
+  in coNP time.
+\end{thm}
 
@@
 % dependents of R:
-Let $D_0$ be the reflexive transitive dependents of $P$ in $S$,
+Let $D_0$ be the \emph{reflexive transitive dependents} of $P$ in $S$,
 % clause siblings of dependents:
@@
   %
-  The transitive dependents $D_0$ of $\cc{L}_9$ is \{\cc{main}\}.
+  The \emph{reflexive transitive dependents} $D_0$ of $\cc{L}_9$ is \{\cc{main}\}.
   %
@@
   %
-  The \cc{query} is $\cc{b}_{main} \land res'<0$.
+  The \cc{query} is $\cc{b}_{main} \land res<0$.
   %
@@
 $\shara$, given a recursion-free CHC system $S$,
-constructs an equivalent CDD system $S'$.
+constructs an CDD system $S'$.
 %
@@
 %
+
+%
+\begin{defn}
+  If $S'$ is CDD, $S \expandsto S'$, and there is no CDD system $S''$ such that $S \expandsto S' \expandsto S''$ and
+  $S'' \neq S'$, then $S'$ is a minimal CDD expansion of $S$.
+\end{defn}
 
@@
 %
-\sys first runs a procedure $\expand{}$ on $S$ to obtain a CDD expansion
+\sys first runs a procedure $\expand$ on $S$ to obtain a CDD expansion
 $S'$ of $S$.
@@
 %
-$\expand{ \mathcal{R} }$ is given in \autoref{app:cons-cdd}.
+$\expand$ is given in \autoref{app:cons-cdd}.
 
@@
 \end{thm}
+Proof is given in \autoref{app:corr}.
 

### 2018-04-12 paper/background.tex
one pass for approach

@@
 %
-For a theory $\mathcal{T}$, the space of $\mathcal{T}$
-formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.
+For a theory $T$, the space of $T$
+formulas over $X$ is denoted $\tformulas{T}$.
 %
-For each formula $\varphi \in \tformulas{\mathcal{T}}$, the set of
+For each formula $\varphi \in \tformulas{T}$, the set of
 variables that occur in $\varphi$ (i.e., the \emph{vocabulary} of
@@
 For formulas $\varphi_0, \ldots, \varphi_n, \varphi \in
-\tformulas{\mathcal{T}}$, the fact that $\varphi_0, \ldots, \varphi_n$
+\tformulas{Ts}$, the fact that $\varphi_0, \ldots, \varphi_n$
 \emph{entail} $\varphi$ is denoted $\varphi_0, \ldots, \varphi_n
@@
 
-% Define replacement and substitution:
-For sequences of variables $Y = [ y_0, \ldots, y_n ]$ and $Z = [ z_0,
-\ldots, z_n ]$, the $\mathcal{T}$ formula constraining the equality of
-each element in $Y$ with its corresponding element in $Z$, i.e., the
-formula $\bigland_{0 \leq i \leq n} y_i = z_i$, is denoted $Y = Z$.
-%
-The repeated replacement of variables $\varphi[ z_0 / y_0 \ldots z_{n}
-/ y_{n} ]$ is denoted $\replace{\varphi}{Y}{Z}$.
-%
-For each formula $\varphi$ defined over free variables $Y$,
-$\replace{\varphi}{Y}{Z}$ is denoted alternatively as $\varphi[ Z ]$.
-%
-The number of free variables in formula $\varphi$ is denoted
-$\degreeof{\varphi}$.
-
 % introduce interpolation:
-An interpolant of mutually inconsistent $\mathcal{T}$ formulas
-$\varphi_0$ and $\varphi_1$ is a $\mathcal{T}$ formula in their common
-vocabulary that explains their inconsistency.
+An interpolant of a pair of mutually inconsistent formulas
+$\varphi_0$ and $\varphi_1$ in $\tformulas{T}$ is a formula $I$ in $\tformulas{T}$ over 
+their common vocabulary that explains their inconsistency.
 %
@@
 For the remainder of this paper, all spaces of formulas will be
-defined for a fixed, arbitrary theory $\mathcal{T}$ that supports
-interpolation, such as combinations of the theories of linear
-arithmetic and the theory of uninterpreted functions with equality.
+defined for a fixed, arbitrary theory $T$ that supports
+interpolation, such as the theory of linear
+arithmetic.
 % introduce decision procedure:
@@
 %
-We define \sys in terms of an abstract decision procedure for
-$\mathcal{T}$ named \issat, and an abstract interpolating theorem
-prover for $\mathcal{T}$ named $\solveitp$.
+
+%
+We define \sys in terms of an abstract interpolating theorem
+prover for $T$ named $\solveitp$.
+%
+Given two formulas $\varphi_0$ and $\varphi_1$, if $\varphi_0$ and $\varphi_1$ are mutually 
+inconsistent, $\solveitp$ returns the interpolant of $\varphi_0$ and $\varphi_1$.
+%
+Otherwise, $\solveitp$ returns $\none$.

### 2018-04-13 paper/abstract.tex
pass over all non-appendix sections

@@
   %
-  In addition, many recursive CHC system solvers reduce inputs to a
-  series of recursion-free CHC systems that can each be solved
-  efficiently.
+  In addition, many CHC solvers which handle recursive systems reduce
+  their inputs to a series of recursion-free CHC systems that can each
+  be solved efficiently.
 
@@
   %
-  The advantage of this class is that many systems are significantly
-  smaller than equivalent systems in other classes which allows them
-  to be solved more efficiently.
+  The advantage of this class is that many CDD systems are smaller
+  than systems which express the same constraints but are part of a
+  different class.
+  %
+  This advantage in size allows CDD systems to be solved more
+  efficiently than their counterparts in other classes.
   %

### 2018-04-13 paper/approach.tex
pass over all non-appendix sections

@@
 %
-Proofs of all theorems stated in this section are contained in appendix
+Proofs of all theorems stated in this section are in appendix
 \autoref{app:char} and \autoref{app:corr}.
@@
 solution $\sigma$ that maps predicates to their solutions, the formula
-$\ctrof{P,\sigma}$ is a compact representation of the constraints of $P$.
+$\ctrof{P,\sigma}$ is a compact representation of the constraints of
+$P$.
 %
-If $\sigma$ does not contain $P$, then the constraint of $P$ is constructed
-from the clauses where $P$.  When $\sigma$ does contain $P$, $\ctrof{P,\sigma}$
-is a lookup from $\sigma$:
+If $\sigma$ does not contain $P$, then the constraint of $P$ is
+constructed from the clauses where $P$ is the head.  When $\sigma$
+does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$:
 %
@@
   When $\solvecdd$ solves the uninterpreted predicate $\cc{L}_9$ in $\mcchc$,
-  $\sigma$ contains $\cc{L}_6$ and $\cc{L}_8$.
-  %
   $\sigma$ maps $\cc{L}_6$ to $n \ge 0$ and $\cc{L}_8$ to $n < 0$.
@@
 
-\subsubsection{Constructing the post-constraint of a relational
-  predicate}
+\subsubsection{Constructing post-formulas for predicates}
 \label{sec:cons-post}
@@
 of $S$, denoted $S \expandsto S'$.
-%
 

### 2018-04-13 paper/background.tex
pass over all non-appendix sections

@@
 %
-For a given CHC system $\cc{S}$, if there are no uninterpreted
+For a given CHC system $\cc{S}$, if there is no uninterpreted
 predicate $\cc{P} \in \predof{S}$ such that $\cc{P} \in \tdepsof{P}$,
@@
 
-A solution to a CHC system $S$ is a map from each uninterpreted
-predicate $\cc{P} \in \predof{S}$ to its corresponding interpretation
-which is a formula.
+A solution to a CHC system $S$ is a map from each predicate $\cc{P}
+\in \predof{S}$ to its corresponding interpretation which is a
+formula.
 %
 For a solution to be valid, each clause in $S$ must be valid after
-substituting each uninterpreted predicate by its interpretation.
+substituting each predicate by its interpretation.
 

### 2018-04-13 paper/evaluation.tex
pass over all non-appendix sections

@@
 % describe systems:
-We performed an empirical evaluation of \sys to determine in what
-cases \sys performs better or worse than existing solvers for
-recursion-free CHC systems.
+We performed an empirical evaluation of \sys to determine how it
+compares to existing solvers for recursion-free CHC systems.
 %
@@
 were used to evaluate \duality in previous work~\cite{mcmillan14}.
+
+We also attempted to compare \sys to the \eldarica CHC solver, but
+\eldarica could not parse the CHC systems generated by \seahorn.
 %
-We plan to make \sys publicly available.
+We compared \sys and \eldarica on an alternative set of benchmarks
+generated by the UFO model checker~\cite{albarghouthi12c}, and found
+that \sys outperformed \eldarica by at least an order of magnitude on
+an overwhelming number of cases.
+%
+As a result, we focus our discussion on a comparison of \sys and
+\duality.
 
@@
 % summarize the result
-Out of 4,309 benchmarks, \sys solved or refuted 2,408, timed out on
-762, and for 1,139 benchmarks, generated a constraint that caused
-\zthree's interpolating theorem prover to fail, which caused \sys to
-neither be able to solve or refute the CHC system.
+Out of 4,309 benchmarks, \sys solved or refuted 2,408 while \duality
+solved or refuted 2,321. \sys timed out on 762 benchmarks and \duality
+timed out on 1,145.
 %
-\duality solved or refuted 2,321 benchmarks, timed out on 1,145, and
-failed due to an error thrown by \zthree's interpolating theorem
-prover on 843.
+On the remaining benchmarks, some constraint caused \zthree's
+interpolating theorem prover to fail, meaning the result was neither a
+solve nor a refutation. \sys reached this failure on 1,139 benchmarks
+while \duality failed on 843.
 %
@@
 interpolation queries.
-
-%
-We also attempted to compare \sys to the \eldarica CHC solver, but
-it could not parse the CHC systems generated by \seahorn.
-%
-We compared \sys and \eldarica on an alternative set of benchmarks
-generated by the UFO model checker~\cite{albarghouthi12c}, and found
-that \sys outperformed \eldarica by at least an order of magnitude on
-an overwhelming number of cases.
-%
-As a result, we focus our discussion on a comparison of \sys and
-\duality.
 
@@
 % introduce figure, summarize data not included:
-The results of our evaluation are contained in
+The results of our evaluation are in
 \autoref{fig:complete-data}.
 %
-Of the 4,040 benchmarks on which both solvers took a negligible amount
+Of the 4,040 benchmarks on which both solvers took a short amount
 of time---less than five seconds---\sys solved the benchmarks in an
@@
 %
-\autoref{fig:complete-data} contains data on only the benchmarks that
-were not solved in a negligible amount of time.
-% give key numbers on non-negligible cases:
-Out of the 269 non-negligible benchmarks, \sys solved 185 in less time
+\autoref{fig:complete-data} contains data for benchmarks which took
+longer than five second for both systems to solve.
+%
+Out of these 269 benchmarks, \sys solved 185 in less time
 than \duality, and solved 159 in less than half the time of \duality.
@@
 %
-Because the majority of files have size ranging between 1,000 and
-100,000 lines, \autoref{fig:size} is restricted to this range to
-clarify the presentation.
+The majority of files have between 1,000 and 100,000 lines, so
+\autoref{fig:size} is restricted to this range.
 %

### 2018-04-13 paper/introduction.tex
pass over all non-appendix sections

@@
 for each predicate such that each CHC is logically
-consistent in isolation.\QZ{re-write this based on other papers}
+consistent in isolation.
 
@@
 derivation of each predicate will not contain
-itself. \QZ{\DAH{I still think we need a small example here}}
+itself.
 %
@@
 %
-Instead, they are rewritten to a more specific subclass of
+Instead, they are reduced to a more specific subclass of
 recursion-free CHC system.
@@
 %
-We refer to such classes as \emph{directly solvable}.
 
@@
 %
-In contrast, solving a directly solvable system with the same logic
-and theories is in co-NP~\cite{rummer13b}.
+In contrast, solving a linear system or body-disjoint system with the
+same logic and theories is in co-NP~\cite{rummer13b}.
 %
-Therefore, the performance of solving a recursion-free system is
-highly reliant on the size of the synthesized directly solvable
-system.
+We refer to such classes that are solvable in co-NP time as
+\emph{directly solvable}.
+%
+Because solving an arbitrary recursion-free system is harder than
+solving a directly solvable system, solvers which reduce to directly
+solvable systems are highly reliant on the size of the reductions.
 
@@
 %
-The key characteristic of this class is that CDD systems are often
-smaller than equivalent systems in other directly solvable classes,
-and therefore solving recursion free systems by rewriting them to CDD
-form is often less computationally expensive.
+The key characteristic of this class is that when an arbitrary
+recursion-free system is reduced to a CDD system and to a system from
+a different directly solvable class, the CDD system is frequently the
+smaller of the two.
+%
+Therefore, solving recursion free systems by reducing them to CDD form
+is often less computationally expensive than reducing them to a
+system in a different class.
 
@@
 %
-Given a recursion-free system $S$, \sys reduces the problem
-of solving $S$ to solving an equivalent CDD system
-$S'$.
+Given a recursion-free system $S$, \sys reduces the problem of solving
+$S$ to solving a CDD system $S'$.
 %
@@
 However, empirically we have found that the size of $S'$ is usually
-close enough to the size of $S$ that \sys often outperforms the best
+close enough to the size of $S$ that \sys frequently outperforms the best
 known CHC solvers.

### 2018-04-13 paper/overview.tex
pass over all non-appendix sections

@@
 body-disjoint systems would have to create a fresh copy of
-$\cc{L}_4$, as well copies for all clauses and relational predicates that
-$\cc{L}_4$ depends on.
+$\cc{L}_4$. 
+%
+Worse, if $\cc{L}_4$ had dependencies, each would also need to be copied.
 
@@
 Previous work has also introduced the class of linear
-systems~\cite{albarghouthi12a}, where the body of each has at most
-one uninterpreted predicate.
+systems~\cite{albarghouthi12a}, where the body of each claouse has at
+most one uninterpreted predicate.
 %
@@
 are exponentially larger than the input. For example, if the procedure
-$\cc{dbl}$ were called more than once in $\cc{dblsAbs}$ then multiple
+$\cc{dbl}$ were called more than once in $\cc{dblAbs}$ then multiple
 copies of the body of the procedure would be inlined. And if the body

### 2018-04-13 paper/related-work.tex
pass over all non-appendix sections

@@
 solving a series of recursion-free unwindings of $S$.
-
 %
@@
 %
-
-%
-\duality attempts to avoid solving all 
-derivation-trees (i.e body-disjoint system) by using lazy annotation~\cite{bjorner13}.
+\duality attempts to avoid solving all derivation-trees (i.e
+body-disjoint system) by using lazy annotation~\cite{bjorner13}.
 %
@@
 
-\whale attempts to verify a sequential recursive program by generating
-and solving hierarchical programs (i.e., programs that may contain
-conditional branches and procedure calls, but do not contain loops or
-recursion), which correspond to recursion-free CHC
+\whale attempts to verify sequential recursive programs by generating
+and solving hierarchical programs, which correspond to recursion-free CHC
 systems~\cite{albarghouthi12b}.
 %
-To solve a particular recursion-free system $S$, \whale
-generates a linear inlining $S'$ of $S$ and solves
-it using a procedure \vinta~\cite{albarghouthi12a}.
+To solve a particular recursion-free system, \whale solves a linear
+inlining of the input using a procedure named
+\vinta~\cite{albarghouthi12a}.
 %
-In general, $S'$ may have size exponential in the size of
-$S$.
+In general, the linear inlining may be exponentially larger than the
+input.
 
-% compare sys to all of the previous work:
-\sys is similar to all of the approaches given above for solving 
-recursion-free CHC systems in that it reduces the problem of solving a
-given recursion-free CHC system $S$ to solving a CHC system
-in a directly-solvable class.
+\sys is similar to the recursion free CHC approaches given above in
+that it reduces the problem to solving a CHC system in a
+directly-solvable class.
 %
+\sys is distinct in that it reduces to solving Clause-Dependent
+Disjoint (CDD) systems.
+%
+As discussed in \autoref{sec:overview}, the class of CDD systems if a
+superset of classes used by the approaches above. And CDD systems can
+also be solved directly.
 
-\sys is distinct from all of the approaches given above in that it
-reduces solving a recursion-free CHC system to solving a
-Clause-Dependent Disjoint (CDD) system.
+\sys solves general CHC systems using the same strategy as proposed by
+the above approaches.
 %
-CDD systems strictly contain the union of all classes of
-directly-solvable CHC systems used by the above approaches, and can
-themselves be solved directly.
+Specifically, it solves a series of recursion-free unwindings of the
+original system, and tries to synthesize a general solution from the
+recursion-free solutions.
 
-%
-For solving genreal CHC system, \sys uses the same strategy that proposed
-by above approaches, that solving a series recursion-free CHC system that unwinding 
-from the original CHC system, and try to synthesize the solution for the original 
-CHC system by using solutions of recursion-free CHC systems.
-
-% bullshit CADE papers:
-Previous work has given solvers for non-linear Horn clauses over
+Previous work describes solvers for non-linear Horn clauses over
 particular theories.
@@
 %
-Other work gives a solver for the class of \emph{timed pushdown
+Other work describes a solver for the class of \emph{timed pushdown
 systems}, a subclass of CHC systems over the theory of linear real
@@
 %
-Unlike both approaches, \sys can solve systems over any theory that
+Unlike these approaches, \sys can solve systems over any theory that
 supports interpolation.
@@
 % DAG inlining:
-DAG inlining, given a hierarchical program $P$, attempts to generate a
-compact verification condition for $P$~\cite{lal-qadeer15}.
+DAG inlining attempts to generate compact verification conditions for
+hierarchical programs~\cite{lal-qadeer15}.
 %
-\sys contains a procedure that, given a recursion-free Horn Clause
-system $S$, attempts to construct a compact CDD system
-$S'$ such that each solution of $S'$ defines a
-solution of $S$.
+\sys attempts to solve recursion-free CHC systems by attempting to
+reduce them to compact CDD systems.
 %
-Because hierarchical programs and recursion-free Horn Clauses
-correspond closely to each other, algorithms that operate on
-hierarchical programs directly correspond to algorithms that operate
-on recursion-free Horn Clauses.
+Because hierarchical programs and recursion-free CHC systems are
+closely related, algorithms that operate on hierarchical programs
+correspond to algorithms that operate on recursion-free Horn Clauses.
 %
-However, an algorithm for constructing a verification condition of
-hierarchical programs cannot apparently be directly used to synthesize
-a solution of a recursion-free CHC system.
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "p"
-%%% End:
+However, it is not apparent that such algorithms can be used directly
+to synthesize solutions for recursion-free CHC systems.

### 2018-04-13 paper/appendix.tex
one pass for approach
add the rough proof

@@
 \end{proof}
+
+\section{Proof of the upper bound of $\solvecdd$ is coNP}
+\label{app:solve-cp}
+The following is a proof of \autoref{thm:solve-cp}.
+%
+In this section, we give a proof that the upper bound of $\solvecdd$ is coNP.
+%
+It is easy to see that $\prectr$ and $\postctr$ construct the formula as the linear size
+of the whole CHC system, and with satisfiability problem in NP for linear arithmetic.
+%
+$\solvecdd$ at most issues linear number of interpolation queries in terms of number of predicate.
+%
+Therefore, the upper bound of $\solvecdd$ is coNP.
 

### 2018-04-13 paper/overview.tex
add the rough proof

@@
 \sys~takes advantage of this fact when solving input systems.
-Given an arbitrary recursion-free CHC system $S$, \sys~generates an
-equivalent CDD system $S'$ and solves $S'$ directly.
+Given an arbitrary recursion-free CHC system $S$, \sys~reduce $S$ to an
+CDD system $S'$ and solves $S'$ directly.
 %
@@
 %
-However, \sys~generates equivalent systems via heuristics analogous to
+However, \sys~generates CDD systems via heuristics analogous to
 those used to generate compact verification conditions of hierarchical

### 2018-04-13 paper/appendix.tex
pass over all except proof of correctness

--- 2018-04-11 paper/appendix.tex
@@
 \end{algorithm}
+%
 % walk through top-level procedure:
-\autoref{alg:expand}, given a recruion-free CHC system $S$
-(\autoref{line:expand-begin}), returns an minimal CDD expansion $S'$ of
-$S$ with a correspondence from $S'$ to
-$S$.
+Given a recursion-free CHC system $S$, \autoref{alg:expand} returns a
+minimal CDD expansion of $S$ (\autoref{defn:min-expansion}).
 %
@@
 (\autoref{line:expand-aux-begin}---\autoref{line:expand-aux-end}) that
-takes a CHC system $S$ and returns an minimal CDD expansion of 
-$S$.
+takes a CHC system $S$ and returns a minimal CDD expansion of $S$.
 %
@@
 runs a procedure $\sharingclause$ on $S'$, which tries to
-find clause $C \in S'$ and predicate $P \in \bodyof{C}$
-that $P$ is in both transitive dependences of two predicates $Q_0$ and $Q_1$ are siblings 
-in $\predof{S}'$.
+find a clause $C \in S'$ and a predicate $P \in \bodyof{C}$
+such that $P$ is in the transitive dependencies of two sibling
+predicates.
 %
 In such a case, we say that $(C, P)$ is a
-\emph{sibling-shared dependence}.
+\emph{sibling-shared dependency}.
 
 % case: no sibling-shares
-If $\sharingclause$ determines that no sibling-shared dependence
+If $\sharingclause$ determines that no sibling-shared dependency
 exists, then $\expandaux$ returns $S'$
@@
 
-% case: there is a sibling-shared dependence:
-Otherwise, for sibling-shared dependence $(C, P)$ returned
-by $\sharingclause$, $\expandaux$ runs $\copyrel $ on
-$S'$, $C$, and $P$, which returns an expansion of
-$S'$ by creates fresh copy $P'$ of $P$ and updates $\bodyof{C}$ 
-to avoid transitive dependents shared by $Q_0$ and $Q_1$.
+% case: there is a sibling-shared dependency:
+Otherwise, $\sharingclause$ must have located a sibling-shared
+dependency $(C, P)$. In this case, $\expandaux$ runs $\copyrel $ on
+$S'$, $C$, and $P$, which returns an expansion of $S'$ by creating a
+fresh copy of $P$ and updating $\bodyof{C}$ to avoid the shared
+dependency.
 %
-$\expandaux$ recurses on the CHC system returned by $\copyrel $ and returns the result
+$\expandaux$ recurses on this expansion and returns the result
 (\autoref{line:expand-recurse}).
@@
   % proof structure:
-  To prove that the union of the classes of linear systems and
-  body-disjoint systems are strictly contained by the class of CDD
-  systems, we prove %
-  \textbf{(1)} the class of linear systems are contained by the class
-  of CDD systems, %
-  \textbf{(2)} the class of body-disjoint systems are contained by the
-  class of CDD systems, and %
-  \textbf{(3)} there is some CDD system that is not linear or
+  To prove that CDD is a strict superset of the union of the class
+  of linear systems and the class of body-disjoint systems, we prove
+  %
+  \textbf{(1)} CDD contains the class of linear systems,
+  \textbf{(2)} CDD contains the class of body-disjoint systems, and
+  \textbf{(3)} there is some CDD system that is neither linear nor
   body-disjoint.
@@
   $S$, each pair of distinct predicates in
-  the body of $C$, the transitive dependence are disjoint, and
-  no predicates shows more than once in the body of $C$.
+  the body of $C$, the transitive dependencies are disjoint and
+  no predicates appears more than once in the body of $C$.
   (\autoref{defn:cdds}).
@@
   %
-  The body of $C$ has no pair of distinct predicates or a predicates shows more than once, by the fact that
-  $S$ is linear.
+  Since $C$ is a linear clause, it has at most one relational
+  predicate in its body. Therefore, the transitive dependencies are
+  trivially disjoint and there can be no repeated predicate.
   %
@@
   %
-  All dependences of $R_0$ and $R_1$ are in subtrees of $T$, which
+  All dependencies of $R_0$ and $R_1$ are in subtrees of $T$, which
   are disjoint by the definition of a tree.
@@
   % distinguishing example:
-  For goal \textbf{(3)}, the system $\mcchc$ is CDD, but is not linear
-  or body-disjoint.
+  For goal \textbf{(3)}, the system $\mcchc$ is CDD, but is neither linear
+  nor body-disjoint.
+\end{proof}
+
+\section{Proof $\solvecdd$ is in co-NP}
+\label{app:solve-cp}
+The following is a proof of \autoref{thm:solve-cp}.
+%
+Namely, the upper bound of $\solvecdd$ is in co-NP.
+%
+\begin{proof}
+  $\prectr$ and $\postctr$ construct formulas linear in the size of the
+  CHC system. The satisfiability problem for the constructed formulas
+  are in NP for linear arithmetic.
+  %
+  $\solvecdd$ at issues (at worst) a linear number of interpolation
+  queries in terms of number of predicate.
+  %
+  Therefore, the upper bound of $\solvecdd$ is co-NP.
 \end{proof}

### 2018-04-13 paper/approach.tex
add the rough proof
pass over all except proof of correctness

@@
   %
-  the transitive dependences of $P$ and $Q$ are disjoint ($\tdepsof{P} \cap
+  the transitive dependencies of $P$ and $Q$ are disjoint ($\tdepsof{P} \cap
   \tdepsof{Q} = \emptyset$)
@@
 \label{thm:cdd-contains}
-  The class of CDD systems strictly contains the union of the classes
-  of body-disjoint and linear systems.
+  The class of CDD systems is a strict superset of the union of the
+  class of body-disjoint systems and the class of linear systems.
 \end{thm}
@@
   \myproc{$\solvecdd(\cc{S})$ %
-    \label{line:solve-begin}}{ %
+    \label{line:solve-begin}}{%
     $\sigma \assign \emptyset$ \\
     $\cc{Preds} \assign \topSort(\predof{S})$ \\
+    \label{line:top-sort}}{%
     \For {$P \in \cc{Preds}$}
-    {$\cc{interpolant} \assign \solveitp(\prectr(P,\sigma),\postctr(P,\sigma))$\\
-     \Switch{\cc{interpolant}}{
-       \lCase{\cc{SAT}:}{\Return{$\nosoln$}}
-       \lCase{$\cc{I}$:}{$\sigma$ [$P$] $\assign$ $\cc{I}$}
+    {$\cc{interpolant} \assign \solveitp(\prectr(P,\sigma),\postctr(P,\sigma))$
+      \label{line:interp}\\
+     \Switch{\cc{interpolant}}{%
+       \lCase{\cc{SAT}:}{\Return{$\nosoln$}}\label{line:interp-sat}
+       \lCase{$\cc{I}$:}{$\sigma$ [$P$] $\assign$ $\cc{I}$}\label{line:interp-valid}
      }
     }
-    \Return{$\sigma$}
+    \Return{$\sigma$}\label{line:solve-done}
     }
@@
 Given a CDD system $\cc{S}$, $\solvecdd$ topologically sorts the uninterpreted
-predicates in $\cc{S}$ based on their dependency relations.
+predicates in $\cc{S}$ based on their dependency relations
+(\autoref{line:top-sort}).
 %
 Then, the algorithm calculates interpretations for each predicate in this order
-by invoking $\solveitp$.
+by invoking $\solveitp$ (\autoref{line:interp}).
 %
@@
 %
-It is possible that the pre and post formulas may be mutually satisifiable, in
-which case $\solveitp$ returns \cc{SAT}. In this case, $\solvecdd$ returns
+It is possible that the pre and post formulas may be mutually
+satisifiable, in which case $\solveitp$ returns \cc{SAT}
+(\autoref{line:interp-sat}). In this case, $\solvecdd$ returns
 $\nosoln$ to indicate that $\cc{S}$ is not solvable.
@@
 Otherwise, $\solvecdd$ updates the partial solution by setting the
-interpretation of $P$ to $\cc{I}$.
+interpretation of $P$ to $\cc{I}$ (\autoref{line:interp-valid}).
 %
 Once all predicates have been interpolated, $\solvecdd$ returns the complete
-solution, $\sigma$.
+solution, $\sigma$ (\autoref{line:solve-done}).
 
@@
   Given a CDD system $S$, $\solvecdd$ either returns the solution of $S$ or $\none$
-  in coNP time.
+  in co-NP time.
 \end{thm}
+Proof is given in \autoref{app:solve-cp}.
 
+%
 \subsubsection{Constructing constraints for predicates}
@@
 %
-The counterexample characterization of $P$ constrains that if $P$ is used,
-(meaning $\cc{b}_{P} = True$) then its constraint must hold:
+The counterexample characterization of $P$ constrains that if $P$ is used
+(meaning $\cc{b}_{P} = True$), then its constraint must hold:
 \[
@@
   \label{ex:ctr}
-  When $\solvecdd$ solves the predicate $\cc{L}_9$ in $\mcchc$, it generates a
+  When $\solvecdd$ solves predicate $\cc{L}_9$ in $\mcchc$, it generates a
   constraint based on clauses (5) and (6):
@@
   \label{ex:pre-ctr}
-  When $\solvecdd$ solves the uninterpreted predicate $\cc{L}_9$ in $\mcchc$,
+  When $\solvecdd$ solves predicate $\cc{L}_9$ in $\mcchc$,
   $\sigma$ maps $\cc{L}_6$ to $n \ge 0$ and $\cc{L}_8$ to $n < 0$.
@@
 $\postctr(P, \sigma)$ denotes the post-formula for an arbitrary predicate $P$
-with respect to the partial solution map, $\sigma$. A valid post-constraint is
-mutually inconsistent with the solution of $P$.
+with respect to the partial solution map, $\sigma$. A valid
+post-formula is mutually inconsistent with the solution of $P$.
 %
-The post formula is constructed based on the predicates which depend on $P$.
+and is constructed based on the predicates which depend on $P$.
 %
@@
 %
-let $D_2$ be all transitive dependences of $D_1$, and
+let $D_2$ be all transitive dependencies of $D_1$, and
 % collect all predicates that we define a constraint over:
@@
 % define instantiation of C
-The post constraint for $P$ under $\sigma$ is the conjunction of counterexample
-characterization of all predicates $Q \in D$ and the query clause.
+The post-formula for $P$ under $\sigma$ is the conjunction of counterexample
+characterization of all predicates $Q \in D$ and the query clause:
 %
@@
   %
-  The transitive dependences of $D_1$ is $\varnothing$. 
+  The transitive dependencies of $D_1$ is $\varnothing$. 
   %
@@
   %
-  The post constraint for $\cc{L}_9$ under $\sigma$ is
+  The post-formula for $\cc{L}_9$ under $\sigma$ is:
   $$\cc{query} \land \vc{\cc{main},\sigma} \land \vc{\cc{dbl},\sigma}$$
@@
 \begin{defn}
+  \label{defn:min-expansion}
   If $S'$ is CDD, $S \expandsto S'$, and there is no CDD system $S''$ such that $S \expandsto S' \expandsto S''$ and

### 2018-04-13 paper/overview.tex
pass over all except proof of correctness

@@
 The recursion-free CHC system $\mcchc$ is a
-\emph{Clause-Dependence-Disjoint} (CDD) system.
+\emph{Clause-Dependence Disjoint} (CDD) system.
 % state the defn of CDD:
@@
 The formal definition of transitive dependency is given in
-\QZ{reference}.
+\autoref{sec:chcs}.
 %
@@
 %
-In order to solve for $\mcchc$, a solver which uses
+In order to solve for $\mcchc$, a solver that uses
 body-disjoint systems would have to create a fresh copy of
@@
 Previous work has also introduced the class of linear
-systems~\cite{albarghouthi12a}, where the body of each claouse has at
+systems~\cite{albarghouthi12a}, where the body of each clause has at
 most one uninterpreted predicate.
@@
 copies of the body of the procedure would be inlined. And if the body
-of this procedure were large, the inlining would quickly become
+of this procedure were large, the inlining could become
 prohibitively expensive.
@@
 \sys~takes advantage of this fact when solving input systems.
-Given an arbitrary recursion-free CHC system $S$, \sys~reduce $S$ to an
-CDD system $S'$ and solves $S'$ directly.
+Given an arbitrary recursion-free CHC system $S$, \sys~reduces $S$ to
+a CDD system $S'$ and solves $S'$ directly.
 %

### 2018-04-13 paper/related-work.tex
pass over all except proof of correctness

@@
 %
-As discussed in \autoref{sec:overview}, the class of CDD systems if a
-superset of classes used by the approaches above. And CDD systems can
+As discussed in \autoref{sec:overview}, the class of CDD systems is a
+superset of classes used by the approaches above. CDD systems can
 also be solved directly.

### 2018-04-13 paper/approach.tex
added less formal explanations to intro and approach

@@
 The key contribution of our work is the introduction of the class of
-Clause-Dependence Disjoint (CDD) systems.
-%
-CDD systems can be solved
-in co-NP time and contains the union of all previous introduced classes
-that can be solved in co-NP time.
+Clause-Dependence Disjoint (CDD) CHC systems:
 %
@@
 %
+In order to construct efficiently sized pre and post formulas for
+relational predicates, we need a method of compactly expressing the
+constraints on a given predicate.
+%
 For a CDD system $S$, a predicate $P \in \predof{S}$, and a partial
@@
 constructed from the clauses where $P$ is the head.  When $\sigma$
-does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$:
+does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$.
+%
+Each $P \in \predof{S}$ has a corresponding boolean variable
+$\cc{b}_P$:
 %
@@
 % introduce the isused predicates:
-Each $P \in \predof{S}$ has a corresponding boolean variable $\cc{b}_P$.
+The counterexample characterization of $P$ is a small extension of the
+compact constraint of $P$.
 %
-The counterexample characterization of $P$ constrains that if $P$ is used
-(meaning $\cc{b}_{P} = True$), then its constraint must hold:
+It states that if $P$ is used
+(meaning $\cc{b}_{P} = True$), then the constraint of $P$ must hold:
 \[
@@
   The counterexample characterization for $\cc{L}_9$ is based on its boolean
-  indicator and this constraint:
+  indicator and its constraint:
   $$\vc{\cc{L}_9,\sigma} = \neg \cc{b}_{\cc{L}_9} \lor \ctrof{\cc{L}_9,\sigma}$$
@@
 %
-Otherwise, \sys constructs a solution from the CDD solution, $\sigma'$, by
-invoking $\collapse{ \eta }{ \sigma' }$ (\autoref{line:shara-ret-collapse}).
-
-% define collapse:
-For $\sigma'$ and %
-correspondence $\eta$ from $P' \in \predof{S'}$ to $P \in \predof{S}$,
-$\collapse{\eta}{\sigma'}$ maps solution $\sigma'$ to
-$\sigma$[$P \to \bigland_{ \eta(P') = P}
-\sigma'(P')$].
+Otherwise, \sys constructs a solution from the CDD solution,
+$\sigma'$, by invoking $\collapse{ \eta }{ \sigma' }$
+(\autoref{line:shara-ret-collapse}).
+%
+$\textsc{Collapse}$ is designed to convert the solution for the CDD system
+back to a sensible solution for the original problem. It does this by
+taking the conjunction of all interpretations of predicates which
+correspond to one predicate in the original problem.
+%
+That is, given a CDD solution $\sigma'$ and a correspondence $\eta$ from
+$P' \in \predof{S'}$ to $P \in \predof{S}$,
+$\collapse{\eta}{\sigma'}$ generates an entry in $\sigma$ for each
+predicate in the original system:
+$\sigma[P] \assign \bigland_{ \eta(P') = P} \sigma'(P')$.
 %

### 2018-04-13 paper/introduction.tex
added less formal explanations to intro and approach

@@
 %
-A CHC is a logical implication where the antecedent is a conjunction
-of uninterpreted predicates with a fact in a background
-theory and the consequent is an uninterpreted predicate.
+A CHC is a logical implication with the following form:
+$$
+  R_1(\vec{v_1}) \leftarrow R_2(\vec{v_2}) \land R_3(\vec{v_3}) \land
+  ... \land \varphi(\vec{v_0}, \vec{v_1}, \vec{v_2}, \vec{v_3},...)
+$$
+Here, the left side of the implication, called the head, contains an
+uninterpreted relational predicate applied to a vector of variables.
 %
-In the context of a CHC we call the antecedent the body and the
-consequent the head.  A CHC system is a set of CHCs. 
+The right side has any number of such predicates conjoined together
+with a \emph{fact} ($\varphi$). The fact is a logical formula in a background
+theory and may use variables named by the predicates.
 %
@@
 as \emph{recursion-free}. In a recursion-free CHC system, the
-derivation of each predicate will not contain
-itself.
+derivation of each predicate will not contain itself.
+%
+Less formally, a recursion-free CHC system is one where following
+implication arrows through the system will never reach the same clause
+twice.
 %

### 2018-04-13 paper/appendix.tex
pass over most of correctness proof

@@
 %
-In this section, we give a proof that \sys is a correct solver for
+In this section, we prove that \sys is correct when applied to
 recursion-free CHC systems.
@@
 %
-We combine the lemmas to prove correctness of \sys
-(\autoref{thm:corr}).
+We combine the lemmas to prove \sys is correct (\autoref{thm:corr}).
 
 % lem: result connecting expansions to constructing solutions
-For recursion-free CHC systems $S$ and $S'$, if $S'$ is
+For two recursion-free CHC systems $S$ and $S'$, if $S'$ is
 an expansion of $S$, then the result of collapsing a
@@
   \label{lem:expansion-sound}
-  For recursion-free CHC system $S'$ and $S$,
-  $\sigma'$ is a solution of $S'$, and $\eta : \predof{S}' \to
-  \predof{S}$ a correspondence from $\predof{S}'$ to $\predof{S}$,
-  $\collapse{\eta}{\sigma'}$ is a solution of $S$.
+  For two recursion-free CHC system $S'$ and $S$ such that $\sigma'$
+  is a solution of $S'$ and $\eta$ is a correspondence from
+  $S'$ to $S$, $\collapse{\eta}{\sigma'}$ is a
+  solution of $S$.
 \end{lem}
@@
   % introduce 
-  Let $\mathcal{B} \in \bodies{\mathcal{R}}$, $\varphi \in \formulas$,
-  and $R \in \mathcal{R}$ be such that $((\mathcal{B}, \varphi),
-  R(\vars{ \mathcal{S} }(R))) \in \mathcal{S}$.
   %
-  For each predicate $P' \in \predof{S}'$ such that $\eta(P') = P$, there is
-  some clause $C' \in S'$ such that $P' \in \bodyof{C'}$ by the fact that $S'$ is an expansion of
-  $S$ and the definition of an expansion.
+  For each predicate $P' \in \predof{S}'$ such that $\eta(P') = P$,
+  there must exist some clause $C' \in S'$ such that $P' \in \bodyof{C'}$
+  because $S'$ is an expansion of $S$.
   %
-  Let preicate $Q' \in \predof{S}'$ such that $Q'$ is the head of clause $C'$.
+  Let predicate $Q' \in \predof{S}'$ be the head of $C'$.
   % 
@@
   % 
-  Therefore, $\collapse{\eta}{\sigma'}$ is a solution of $P$.
+  Therefore, $\collapse{\eta}{\sigma'}$ has a solution for $P$.
   %
-  Therefore, $\collapse{\eta}{\sigma'}$ is a solution of
-  $S$.
+  Since $\collapse{\eta}{\sigma'}$ has a solution for each predicate
+  in $S$, $\collapse{\eta}{\sigma'}$ is a solution of $S$.
 \end{proof}
 
+Every expansion of a solvable recursion-free CHC system is also
+solvable.
 % completeness of expansions:
-For a recursion-free CHC system $S$ is solvalbe, each expansion of $S$
-is solvable.
 \begin{lem}
   \label{lem:expansion-complete}
-  For a recursion-free CHC system $S$ is solvable, and
-  $S'$ is
-  an expansion of $S$, $S'$ is solvable.
+  If a recursion-free CHC system $S$ is solvable and
+  $S'$ is an expansion of $S$, then $S'$ is solvable.
 \end{lem}
@@
 \begin{proof}
-  Let $\sigma : P \to \formulas$ be a solution of
-  $S$, and let $\eta : \predof{S}' \to \predof{S}$ be a
+  Let $\sigma$ be a solution of $S$, and let $\eta$ be a
   correspondence from $S'$ to $S$.
   %
-  Let $\sigma' : \predof{S}' \to \formulas$ be such that for each $P'
-  \in \predof{S}'$, $\sigma'(P') = \sigma(\eta(P'))$.
+  Let $\sigma'$ be such that for each $P' \in \predof{S}'$,
+  $\sigma'(P') = \sigma(\eta(P'))$.
   %
@@
   \label{lem:expand-corr}
-  For a recursion-free CHC system $S$ and $S'$, and $\eta : \predof{S'} \to \predof{S}$ such that
-  $(S', \eta) = \expand(S)$, $\eta$
-  is a correspondence from $S'$ to $S$, $S'$ is an CDD system and an expansion
-  of $S$.
+  For two recursion-free CHC systems $S$ and $S'$ and a correspondence
+  from $S'$ to $S$, $\eta$, such that $(S', \eta) = \expand(S)$,
+  $S'$ is an CDD system and an expansion of $S$.
 \end{lem}
@@
 \begin{proof}
-  Proof by induction on the evaluation of $\expand$ on
+  By induction over the evaluation of $\expand$ on
   an arbitrary recursion-free CHC system $S$.
   % inductive fact:
-  The inductive fact is that in each step of evaluation \expandaux,
+  The inductive fact is that for each evaluation step, \expandaux,
   $\corr$ is a correspondence from argument
@@
   % base case:
-  For the base case, \expandaux is called initially on $S$,
+  In the base case, \expandaux is called initially on $S$,
   by \autoref{alg:expand}.
   %
-  $\corr$ is a correspondence from $S$ to
-  itself, by the definition of $\corr$
+  $\corr$ is a correspondence from $S$ to itself, by the definition of
+  $\corr$
   (\autoref{app:cons-cdd}).
@@
   % inductive case:
-  For the inductive case, for $C \in S$ and $P \in \predof{S}$, $\expandaux$ calls itself on 
-  $\copyrel(S, C, P)$.
+  In the inductive case,
+  $\expandaux$ constructs an argument
+  $\copyrel(S, C, P)$
+  where $C$ is a clause and $P$ is a predicate in $S$.
+  $\expandaux$ recusively invokes itself with this argument.
   %
-  For each recursion-free CHC system $S'$ that generated by $\copyrel(S, C, P)$,
+  For each recursion-free CHC system $S'$ generated by $\copyrel(S, C, P)$,
   $\corr$ is a correspondence from $S'$ to
@@
 % lem: VC's characterize system sat:
-For each recursion-free CHC system $S$, $S$ has a solution if and
-only if all interpolants queries return interpolants.
+For each recursion-free CHC system $S$, $S$ has a solution if and only
+if all interpolants queries return interpolants.
 %
@@
   \label{lem:vc}
-  Given a recursion-free CHC system $S$ that is CDD and solvable, for all predicates $P \in \predof{S}$,
-   $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns a interpolatn $I$.
+  Given a recursion-free CHC system $S$ that is CDD and solvable, for
+  all predicates $P \in \predof{S}$,
+  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns
+  an interpolant $I$.
 \end{lem}
@@
   %
-  Assume that $S$ has solution $\sigma$ and there are
-  some predicates $P \in \predof{S}$ such that $\solveitp(\prectr(\cc{P},\sigma'),\postctr(\cc{P},\sigma'))$
-  returns $SAT$, which means there is a model $m$ for the conjuntion of $\prectr(\cc{P},\sigma')$
-  and $\postctr(\cc{P},\sigma')$.
+  Assume that $S$ has a solution $\sigma$ and there are some
+  predicates $P \in \predof{S}$ such that
+  $\solveitp(\prectr(\cc{P},\sigma'),\postctr(\cc{P},\sigma'))$
+  returns $SAT$. This means there must be a model $m$ for the
+  conjunction of $\prectr(\cc{P},\sigma')$ and
+  $\postctr(\cc{P},\sigma')$.
   %
@@
   %
-  Therefore, $\solveitp(\prectr(\cc{P},\sigma'),\postctr(\cc{P},\sigma'))$ always returns interpolant.
+  Therefore,
+  $\solveitp(\prectr(\cc{P},\sigma'),\postctr(\cc{P},\sigma'))$ always
+  returns interpolant.
 \end{proof}
@@
   \label{lem:solve-aux}
-  Given a recursion-free CHC system $S$ that is CDD, if for all predciates $P \in \predof{S}$, 
-  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns a interpolatn $I$, 
-  then $\solvecdd(S)$ is a solution of $S$.
+  Given a recursion-free CHC system $S$ that is CDD, if for all
+  predicates $P \in \predof{S}$,
+  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns
+  an interpolant $I$, then $\solvecdd(S)$ is a solution of $S$.
 \end{lem}
@@
   % overview of structure:
-  Proof by induction on the $\solvecdd(S)$ calles $\solveitp$ on all predicates $P \in \predof{S}$
-  in a topolgical order.
+  By induction on the $\solvecdd(S)$ calls to $\solveitp$ over all
+  predicates $P \in \predof{S}$
+  in topological order.
   %
-  The inductive fact is that at each step of calling $\solveitp$, $\sigma$ is a
-  partial solution of $S$.
+  The inductive fact is that after each call to $\solveitp$, $\sigma$
+  is a partial solution of $S$.
   % base case:
-  For the base case, $\predof{S} = \varnothing$,
+  In the base case, $\predof{S} = \varnothing$,
   %
@@
   % inductive case:
-  For the inductive case, in each step, $\solvecdd$ calls $\solveitp$ on
-  a predicate $P \in \predof{S}$ with partial solution $\sigma$, which
-  contains all solutions of predciates $P \in \depsof{S}$.
+  In the inductive case, $\solvecdd$ calls $\solveitp$ on a predicate
+  $P \in \predof{S}$ with partial solution $\sigma$.  Due to the
+  topological ordering, $\sigma$ contains interpretations for each
+  predicate $P \in \depsof{S}$.
   %
-  Based on the definition of an interpolant(\autoref{defn:itps}), $\sigma$
-  is a partial solution of $S$, then $\prectr(\cc{P},\sigma')$ and $\postctr(\cc{P},\sigma))$
-  are inconsistent.
+  Based on the definition of an interpolant(\autoref{defn:itps}),
+  $\prectr(\cc{P},\sigma)$ and $\postctr(\cc{P},\sigma))$ are
+  inconsistent.
   %
-  The interpolant $I$ returns by $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ entails
-  by all clause that head is $P$ and $\bodyof{P}$ is substituted by partial solution $\sigma$.
+  The interpolant $I$ returned by
+  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ is
+  entailed by each clause $C$ where $P$ is the head where the
+  predicates in $\bodyof{C}$ are substituted by their interpretations
+  in$\sigma$.
   %
-  $I$ also is inconsistment with all constrains shows later that support the query clause.
+  $I$ is also inconsistment with all constraints that appear after $P$
+  that support the query clause.
   %
-  Then $\solvecdd$ updates $\sigma$ by binding $P$ to $I$, which makes the new $\sigma$ also a partial
-  solution of $S$.
-
-  %
-  Base case combined with the inductive case, imply that \solvecdd
-  returns a solution of $S$.
+  Therefore, when $\solvecdd$ updates $\sigma$ by binding $P$ to $I$
+  the result is a partial solution of $S$.
 \end{proof}
@@
 % SolveCDD: establish soundness
-$\solvecdd$ is a
-correct CHC solver for CDD systems $S$.
+$\solvecdd$ is correct for CDD systems.
 %
@@
   For a given CDD system $S$, and $\sigma = \solvecdd(S)$, %
-  $\sigma$ is a solution of $\mathcal{S}$.
+  $\sigma$ is a solution of $S$.
 \end{lem}
@@
   %
-  $\solvecdd{S}$ returns $\sigma$ implies that for all predicates 
-  $P \in \predof{S}$, $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns a interpolatn $I$, 
-  by the definition of $\solvecdd$ \autoref{alg:solve-cdd}. 
+  The fact that $\solvecdd(S)$ returns $\sigma$ implies that for each
+  predicate $P \in \predof{S}$,
+  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns
+  a valid interpolant (\autoref{alg:solve-cdd}).
   %
-  Therefore, \autoref{lem:solve-aux} implies that \solvecdd
-  returns a complete solution of $S$.
+  Therefore, \autoref{lem:solve-aux} implies that \solvecdd~returns a
+  complete solution of $S$.
 \end{proof}

### 2018-04-13 paper/evaluation.tex
pass over all except proof of correctness
pass over most of correctness proof

@@
 % introduce figure, summarize data not included:
-The results of our evaluation are in
+The results of our evaluation are shown in
 \autoref{fig:complete-data}.
@@
 of both \duality and \sys, perhaps by lazily unwinding a given
-recursion free system into a series of CDD systems instead of
+recursion-free system into a series of CDD systems instead of
 derivation trees, could yield further improvements.

### 2018-04-13 paper/introduction.tex
pass over most of correctness proof

@@
 %
-Therefore, solving recursion free systems by reducing them to CDD form
+Therefore, solving recursion-free systems by reducing them to CDD form
 is often less computationally expensive than reducing them to a

### 2018-04-13 paper/overview.tex
finished pass on appendix

@@
 %
-Interpolating the pre and post formulas yields an intepretation
+Interpolating the pre and post formulas yields an interpretation
 of $\cc{L}_9$: $\cc{abs'} \geq 0$.
@@
           predicate is depicted as a graph node while each clause is
-          represented by a hyperedge. Each hyperedge is labelled by the
+          represented by a hyperedge. Each hyperedge is labeled by the
           fact in the corresponding CHC\@. Each node has a valid

### 2018-04-13 paper/overview.tex
fixed minus sign in program

@@
       \label{fig:multicall-code} }
-    {\lstinputlisting[
+    {
+      \lstinputlisting[
         basicstyle=\small,
@@
         numbers=left,
-      morekeywords={def, return, if, else, assert}]{code/dblAbs.java}
-        % \input{code/dblAbs.java}}
+        escapeinside={(*@}{@*)},
+        morekeywords={def, return, if, else, assert}]
+        {code/dblAbs.java}
     }

### 2018-04-13 paper/overview.tex
aligned numbers in program correctly

@@
         basicstyle=\small,
-        xleftmargin=\parindent,
+        % xleftmargin=\parindent,
+        xleftmargin=2em,framexleftmargin=1.5em,
         numbers=left,

### 2018-04-13 paper/evaluation.tex
a pass for the paper

@@
 To generate CHC systems, we ran the \seahorn~\cite{gurfinkel15}
-verification framework with its default settings, set to timeout at 90
+verification framework with its default settings which not inlining procedure call and each 
+loop-free fragment is a clause, set to timeout at 90
 seconds.

### 2018-04-13 paper/introduction.tex
a pass for the paper

@@
 %
-The class consists of all recursion-free CHC systems such that each
-clause in the system has at most one dependency on the same relational
-predicate.
+The formal definition of CDD is given at ~\autoref{defn:cdds}.
+%
+CDD is a strict superset of the union of all previous introduced directly
+solvable systems.
 %

### 2018-04-13 paper/overview.tex
a pass for the paper

--- 2018-04-13 paper/overview.tex
@@
 In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
-$\mcchc$ (\autoref{fig:chc}), that models the safety of a small program,
-\cc{dblAbs} (\autoref{fig:multicall-code}).
+$\mcchc$ (\autoref{fig:chc}), that models the safety of program \cc{dblAbs} (\autoref{fig:multicall-code}).
 %
@@
 %
-\cc{dblAbs} is a small procedure that doubles the absolute value of
+\cc{dblAbs} is a procedure that doubles the absolute value of
 its input and stores the result in \cc{res}.
@@
   \land
-  (\cc{b}_{\cc{main}} \land \cc{res'} < 0)
+  (\cc{b}_{\cc{main}} \land \cc{res} < 0)
 \end{gather}
 %
-Interpolating the pre and post formulas yields an interpretation
+Interpolating the pre and post formulas yields an intepretation
 of $\cc{L}_9$: $\cc{abs'} \geq 0$.
@@
           predicate is depicted as a graph node while each clause is
-          represented by a hyperedge. Each hyperedge is labeled by the
+          represented by a hyperedge. Each hyperedge is labelled by the
           fact in the corresponding CHC\@. Each node has a valid

### 2018-04-13 paper/appendix.tex

@@
   all predicates $P \in \predof{S}$,
-  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns
+  $\solveitp$ returns
   an interpolant $I$.
@@
   predicates $P \in \predof{S}$ such that
-  $\solveitp(\prectr(\cc{P},\sigma'),\postctr(\cc{P},\sigma'))$
+  $\solveitp$
   returns $SAT$. This means there must be a model $m$ for the
@@
   Therefore,
-  $\solveitp(\prectr(\cc{P},\sigma'),\postctr(\cc{P},\sigma'))$ always
+  $\solveitp$ always
   returns interpolant.
@@
   predicates $P \in \predof{S}$,
-  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns
+  $\solveitp$ returns
   an interpolant $I$, then $\solvecdd(S)$ is a solution of $S$.
@@
   %
-  The interpolant $I$ returned by
-  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ is
-  entailed by each clause $C$ where $P$ is the head where the
+  The interpolant of these two formulas returned by $\solveitp$, $I$,
+  is entailed by each clause $C$ where $P$ is the head where the
   predicates in $\bodyof{C}$ are substituted by their interpretations
-  in$\sigma$.
+  in $\sigma$.
   %
-  $I$ is also inconsistment with all constraints that appear after $P$
+  $I$ is also inconsistent with all constraints that appear after $P$
   that support the query clause.
@@
 % SolveCDD: establish soundness
-$\solvecdd$ is correct for CDD systems.
+The output of $\solvecdd$ is correct for a given input CDD system.
 %
@@
   predicate $P \in \predof{S}$,
-  $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns
+  $\solveitp$ returns
   a valid interpolant (\autoref{alg:solve-cdd}).
@@
   % final claim:
-  for all predicates $P \in \predof{S}$, $\solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$ returns a interpolatn $I$, by
+  for all predicates $P \in \predof{S}$, $\solveitp$ returns a
+  interpolant $I$, by
   \autoref{lem:vc} and the fact that $\mathcal{S}$ is solvable.
@@
 % proof of main theorem:
-The following is a proof of correctness of \sys
+The output of \sys is correct for a given input CDD system.
 (\autoref{sec:approach}, \autoref{thm:corr}).
@@
 \begin{proof}
-  Given recurion-free CHC systems $S$ and $S'$, and $\eta :\predof{S'} \to 
-  \predof{S}$ such that $(S', \eta) = \expand(S)$, 
-  then $S'$ is minimal CDD expansion of 
-  $S$, and $\eta$ is a correspondence from $S'$ to $S$ by \autoref{lem:expand-corr}.
+  Given two recursion-free CHC systems $S$ and $S'$ and a
+  $\eta$ such that $(S', \eta) = \expand(S)$,
+  $S'$ is minimal CDD expansion of $S$ and $\eta$ is a correspondence
+  from $S'$ to $S$ (\autoref{lem:expand-corr}).
   %
-  There is some $\sigma'$ such that
-  $\sigma' = \solvecdd(S' )$, by
-  the definition of $\shara$.
+  Assume that $S$ is solvable. Then so is $S'$, by
+  \autoref{lem:expansion-complete}.
   %
-  $\sigma'$ is a solution of $S'$, by
-  \autoref{lem:cdd-soln-sound}.
+  Therefore, there exists some $\sigma'$ such that $\sigma' =
+  \solvecdd(S')$, by the definition of $\shara$.
+  %
+  $\sigma'$ is a solution of $S'$, by \autoref{lem:cdd-soln-sound}.
   %
@@
   \autoref{lem:expansion-sound}.
-
-  % if S is solvable, then shara returns a solution
-  Assume that a recursion-free $S$ is solvable.
   %
-  Let a recursion-free CHC system $S'$ and $\eta :
-  \predof{S'} \to \predof{S}$ be such that $(S',
-  \eta) = \expand(S)$.
-  %
-  $S'$ is a CDD expansion of $S$, by
-  \autoref{lem:expand-corr}.
-  %
-  $S'$ is solvable, by \autoref{lem:expansion-complete}.
-  %
-  $\solvecdd(S')$ returns a solution $\sigma$ of $S'$ by \autoref{lem:cdd-soln-complete}.
-  %
-  $\shara$ returns a solution of $S$ by calling $\collapse{ \eta }{ \sigma' }$, based on
-  \autoref{alg:shara}.
+  Therefore, $\shara$ returns a valid solution of $S$.
 \end{proof}
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "p"
-%%% End:

### 2018-04-13 paper/approach.tex
finished pass on appendix
a pass for the paper

@@
 % talk about previous system
-As discussed in \autoref{sec:overview}, CDD is a superclass of the
+As discussed in \autoref{sec:overview}, CDD is a superclass of the union of
 class of body-disjoint systems and the class of linear systems.
@@
 When \solvecdd~returns that $S'$ has
-no solution, \sys~propogates $\none$ (\autoref{line:shara-ret-none}).
+no solution, \sys~propagates $\none$ (\autoref{line:shara-ret-none}).
 %

### 2018-04-13 paper/overview.tex

--- 2018-04-13 paper/overview.tex
@@
   \land
-  (\cc{b}_{\cc{main}} \land \cc{res'} < 0)
+  (\cc{b}_{\cc{main}} \land \cc{res} < 0)
 \end{gather}

### 2018-04-13 paper/evaluation.tex
pass

--- 2018-04-13 paper/evaluation.tex
@@
 To generate CHC systems, we ran the \seahorn~\cite{gurfinkel15}
-verification framework with its default settings, set to timeout at 90
+verification framework with its default settings (procedures are not
+inlined and each loop-free fragment is a clause), set to timeout at 90
 seconds.

### 2018-04-13 paper/overview.tex
pass

@@
 In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
-$\mcchc$ (\autoref{fig:chc}), that models the safety of program \cc{dblAbs} (\autoref{fig:multicall-code}).
+$\mcchc$ (\autoref{fig:chc}), that models the safety of the program
+\cc{dblAbs} (\autoref{fig:multicall-code}).
 %

### 2018-04-13 paper/related-work.tex
pass over most of correctness proof
pass

@@
 
-\sys is similar to the recursion free CHC approaches given above in
+\sys is similar to the recursion-free CHC approaches given above in
 that it reduces the problem to solving a CHC system in a
@@
 However, it is not apparent that such algorithms can be used directly
-to synthesize solutions for recursion-free CHC systems.
+to synthesize solutions.

### 2018-04-13 paper/appendix.tex
pass
more changes

@@
   % Inputs: a heap program and an error location.
-  \Input{A recruions-free CHC system $S$.}
+  \Input{A recursion-free CHC system $S$.}
   % Output: inductive invariants.
   \Output{A minimal CDD expansion $S'$ of $S$ and
-    correspondence from $S'$ to $S$.} %
+    a correspondence from $S'$ to $S$.} %
   \myproc{$\expand(S)$ %
@@
   $S$ is CDD if for each clause $C$ in
-  $S$, each pair of distinct predicates in
-  the body of $C$, the transitive dependencies are disjoint and
-  no predicates appears more than once in the body of $C$.
+  $S$ \textbf{(1)} each pair of distinct predicates in
+  the body of $C$ has disjoint transitive dependencies and
+  \textbf{(2)} no predicate appears more than once in the body of $C$.
   (\autoref{defn:cdds}).
@@
   %
-  $\solvecdd$ at issues (at worst) a linear number of interpolation
+  $\solvecdd$ issues (at worst) a linear number of interpolation
   queries in terms of number of predicate.
@@
   %
-  For each predicate $P' \in \predof{S}'$ such that $\eta(P') = P$,
+  For each predicate $P' \in \predof{S'}$ such that $\eta(P') = P$,
   there must exist some clause $C' \in S'$ such that $P' \in \bodyof{C'}$
@@
   %
-  Let predicate $Q' \in \predof{S}'$ be the head of $C'$.
+  Let predicate $Q' \in \predof{S'}$ be the head of $C'$.
   % 
@@
   \[ \collapse{ \eta }{ \sigma' }[P] \land \consof{C'} \entails %
-  \left( \bigland_{ \substack{Q' \in \predof{S}' \\ \eta(Q') = Q} }
+  \left( \bigland_{ \substack{Q' \in \predof{S'} \\ \eta(Q') = Q} }
   \sigma(Q') = %
@@
   %
-  Let $\sigma'$ be such that for each $P' \in \predof{S}'$,
+  Let $\sigma'$ be such that for each $P' \in \predof{S'}$,
   $\sigma'(P') = \sigma(\eta(P'))$.
@@
   % inductive fact:
-  The inductive fact is that for each evaluation step, \expandaux,
+  The inductive fact is that for each evaluation step
   $\corr$ is a correspondence from argument
@@
   %
-  Based on the definition of an interpolant(\autoref{defn:itps}),
-  $\prectr(\cc{P},\sigma)$ and $\postctr(\cc{P},\sigma))$ are
+  Based on the definition of an interpolant (\autoref{defn:itps}),
+  $\prectr(\cc{P},\sigma)$ and $\postctr(\cc{P},\sigma)$ are
   inconsistent.
@@
   % final claim:
-  for all predicates $P \in \predof{S}$, $\solveitp$ returns a
+  For all predicates $P \in \predof{S}$, $\solveitp$ returns a
   interpolant $I$, by
-  \autoref{lem:vc} and the fact that $\mathcal{S}$ is solvable.
+  \autoref{lem:vc} and the fact that $S$ is solvable.
   %

### 2018-04-13 paper/approach.tex
pass
more changes

@@
 %
-Proofs of all theorems stated in this section are in appendix
-\autoref{app:char} and \autoref{app:corr}.
+Proofs of all theorems stated in this section are in the appendix.
 
@@
 As discussed in \autoref{sec:overview}, CDD is a superclass of the union of
-class of body-disjoint systems and the class of linear systems.
+the class of body-disjoint systems and the class of linear systems.
 %
@@
 $\shara$, given a recursion-free CHC system $S$,
-constructs an CDD system $S'$.
+constructs a CDD system $S'$.
 %
@@
 preserves the relationship between the clauses of $S'$ in
-the clauses of $S$, then $S'$ is an expansion of
+the clauses of $S$, then $S'$ is an \emph{expansion} of
 $S$ (all definitions in this section will be over fixed
@@
   % Inputs: a heap program and an error location.
-  \Input{For a recursion-free CHC system $S$.}
+  \Input{A recursion-free CHC system $S$.}
   % Output: inductive invariants.

### 2018-04-13 paper/evaluation.tex
more changes

@@
 We modified the algorithm used by \duality to generate recursion-free
-unwindings of a given recursive system so that in each iteration, it
+unwindings of a given recursive system so that, in each iteration, it
 generates an unwinding with a maximal set of relational predicates and

### 2018-04-13 paper/overview.tex
more changes

@@
 %
-In order to solve for $\mcchc$, a solver that uses
+In order to handle $\mcchc$, a solver that uses
 body-disjoint systems would have to create a fresh copy of
@@
 %
-Worse, if $\cc{L}_4$ had dependencies, each would also need to be copied.
+Worse, if $\cc{L}_4$ had dependencies, then each dependency would also
+need to be copied.
 
@@
 $\mcchc$ is not a linear system because the body of clause (7) has two
-predicates \cc{L9} and \cc{dbl}.
+predicates, \cc{L9} and \cc{dbl}.
 %

### 2018-04-13 paper/appendix.tex
an CDD

@@
   from $S'$ to $S$, $\eta$, such that $(S', \eta) = \expand(S)$,
-  $S'$ is an CDD system and an expansion of $S$.
+  $S'$ is a CDD system and an expansion of $S$.
 \end{lem}

### 2018-04-14 paper/abstract.tex
worked in cvoss comments

@@
   %
-  In addition, many CHC solvers which handle recursive systems reduce
-  their inputs to a series of recursion-free CHC systems that can each
-  be solved efficiently.
+  In addition, many CHC solvers reduce recursive systems to a series
+  of recursion-free CHC systems that can each be solved efficiently.
 
@@
   %
-  Based on the class of CDD systems, we implemented a CHC solver named
+  We implemented a CHC solver named
   \sys.  \sys~solves arbitrary CHC systems by reducing the input to a

### 2018-04-14 paper/appendix.tex
worked in cvoss comments

@@
 generated from different implementations of $\sharingclause$, each of
-which chooses clause-relations pairs to return based on different
+which chooses clause-relation pairs to return based on different
 heuristics.
@@
   $S$ is CDD if for each clause $C$ in
-  $S$ \textbf{(1)} each pair of distinct predicates in
+  $S$ \textbf{(1)} and each pair of distinct predicates in
   the body of $C$ has disjoint transitive dependencies and
@@
   Since $C$ is a linear clause, it has at most one relational
-  predicate in its body. Therefore, the transitive dependencies are
-  trivially disjoint and there can be no repeated predicate.
+  predicate in its body. And since the system is recursion-free, the
+  transitive dependencies are trivially disjoint and there can be no
+  repeated predicate.
   %
@@
 %
-Namely, the upper bound of $\solvecdd$ is in co-NP.
+Namely, $\solvecdd$ is in co-NP.
 %
@@
 For each recursion-free CHC system $S$, $S$ has a solution if and only
-if all interpolants queries return interpolants.
+if all interpolation queries return interpolants.
 %

### 2018-04-14 paper/approach.tex
worked in cvoss comments

@@
 % talk about previous system
-As discussed in \autoref{sec:overview}, CDD is a superclass of the union of
+As discussed in \autoref{sec:overview}, CDD is a superset of the union of
 the class of body-disjoint systems and the class of linear systems.
@@
 For a given recursion-free system $S$, if each uninterpreted predicate $Q \in
-\predof{S}$, appears in at most one in the body of a clause and, no
+\predof{S}$ appears in the body of at most one clause and no
 predicate appears more than once in the body of a single clause,
@@
   \label{thm:solve-cp}
-  Given a CDD system $S$, $\solvecdd$ either returns the solution of $S$ or $\none$
-  in co-NP time.
+  Given a CDD system $S$ over the theory of linear integer arithmetic,
+  $\solvecdd$ either returns the solution of $S$ or $\none$ in co-NP
+  time.
 \end{thm}
 Proof is given in \autoref{app:solve-cp}.
+
+In order to solve a CDD system, we construct efficiently sized pre and
+post formulas for each relational predicate and interpolate over these
+formulas. These pre and post formulas are built from \textbf{(1)} the
+\emph{constraint} of a given predicate, which explains under what
+conditions the predicate holds, and \textbf{(2)} the
+\emph{counterexample characterization}, which explains what condition
+must be true if the predicate holds.
 
@@
 In order to construct efficiently sized pre and post formulas for
-relational predicates, we need a method of compactly expressing the
+relational predicates, we use a method for compactly expressing the
 constraints on a given predicate.
@@
 % introduce the isused predicates:
-The counterexample characterization of $P$ is a small extension of the
-compact constraint of $P$.
+The \emph{counterexample characterization} of $P$ is a small extension
+of the compact constraint of $P$.
 %
@@
 % dependents of R:
-Let $D_0$ be the \emph{reflexive transitive dependents} of $P$ in $S$,
+Let $D_0$ be the \emph{transitive dependents} of $P$ in $S$ (i.e the
+predicates that have $P$ as a transitive dependency),
 % clause siblings of dependents:
@@
   %
-  The \emph{reflexive transitive dependents} $D_0$ of $\cc{L}_9$ is \{\cc{main}\}.
+  The \emph{transitive dependents} $D_0$ of $\cc{L}_9$ is \{\cc{main}\}.
   %
@@
   %
-  The transitive dependencies of $D_1$ is $\varnothing$. 
+  The set of transitive dependencies of $D_1$ is $\varnothing$. 
   %
-  Therefore, the $D$ set is \{\cc{main},\cc{dbl}\}.
+  Therefore, $D$ is \{\cc{main},\cc{dbl}\}.
   %
@@
 % give the general procedure:
-$\shara$, given a recursion-free CHC system $S$,
-constructs a CDD system $S'$.
+Given a recursion-free CHC system $S$, \shara~constructs a CDD system
+$S'$.
 %
@@
   \label{defn:expansion}
-  Let $f : \predof{S'} \to \predof{S}$ be such that
+  Let $\eta : \predof{S'} \to \predof{S}$ be such that
   % preserves arity:
   \textbf{(1)} for all $P' \in \predof{S'}$, $P'$ has the
-  same vector of parameters as $f(P')$;
+  same parameters as $\eta(P')$;
   % preserves clause structure:
-  \textbf{(2)} for each clause $C' \in S'$, the clause $C$ (constructed
-  by substituting all predicates $P'$ by $f(P')$) is in $S$.
+  \textbf{(2)} for each clause $C' \in S'$, the clause $C$, constructed
+  by substituting all predicates $P'$ by $\eta(P')$, is in $S$; and
+  \textbf{(3)} each predicate $P$ in $S$ has at least one predicate
+  $P'$ in $S'$ such that $\eta(P') = P$.
   %
-  Then $f$ is a \emph{correspondence} from $S'$ to $S$.
+  %
+  Then $\eta$ is a \emph{correspondence} from $S'$ to $S$.
 \end{defn}
@@
 \sys first runs a procedure $\expand$ on $S$ to obtain a CDD expansion
-$S'$ of $S$.
+$S'$ of $S$ ($\expand$ is given in \autoref{app:cons-cdd}).
 %
@@
 $\textsc{Collapse}$ is designed to convert the solution for the CDD system
-back to a sensible solution for the original problem. It does this by
+back to a solution for the original problem. It does this by
 taking the conjunction of all interpretations of predicates which
-correspond to one predicate in the original problem.
+correspond to the same predicate in the original problem.
 %
@@
 $\sigma[P] \assign \bigland_{ \eta(P') = P} \sigma'(P')$.
-%
-$\expand$ is given in \autoref{app:cons-cdd}.
 
@@
   \label{thm:corr}
-  $S$ is solvable if and only if $\shara$ returns solution $\sigma$.
+  $S$ is solvable if and only if $\shara$ returns a solution $\sigma$.
 \end{thm}

### 2018-04-14 paper/background.tex
pass over all except proof of correctness
more changes
worked in cvoss comments

@@
 For a given CHC system $S$, $\predof{S}$ denotes the set of all
-uninterpreted predicates.
+uninterpreted predicates and $\cc{query}$ denotes the body of the
+query clause.
 
@@
   such that $\cc{P} = \headof{C}$ and $\cc{Q} \in \bodyof{C}$, then
-  $\cc{Q}$ is a predicate dependency of $\cc{P}$.
+  $\cc{Q}$ is a \emph{predicate dependency} of $\cc{P}$.
 \end{defn}
@@
   Given a CHC system $\cc{S}$ and three uninterpreted predicates
-  $\cc{P}, \cc{Q}$ and $\cc{R} \in \predof{S}$, if $\cc{Q} \in
-  \depsof{P}$ then \cc{Q} is a transitive predicate dependency of
+  $\cc{P}, \cc{Q}$, and $\cc{R} \in \predof{S}$, if $\cc{Q} \in
+  \depsof{P}$ then \cc{Q} is a \emph{transitive predicate dependency} of
   \cc{P}.
@@
   is a transitive predicate dependency of \cc{Q}, then $\cc{R}$ is a
-  transitive predicate dependency of $\cc{P}$.
+  \emph{transitive predicate dependency} of $\cc{P}$.
 \end{defn}
@@
 \begin{ex}
-  Because uninterpreted predicates $\cc{L}_9$ and $\cc{db}$ both
-  appear in the body of clause (7), $\cc{L}_9$ and $\cc{db}$ are
+  Because uninterpreted predicates $\cc{L}_9$ and $\cc{dbl}$ both
+  appear in the body of clause (7), $\cc{L}_9$ and $\cc{dbl}$ are
   siblings.
@@
 %
-Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
-$\siblingof{P}$ denotes the set of all siblings of $\cc{P}$
-in $\cc{S}$.
 %
@@
 %
-For each formula $\varphi \in \tformulas{T}$, the set of
+For each formula $\varphi \in \tformulas{T}$, the set of free
 variables that occur in $\varphi$ (i.e., the \emph{vocabulary} of
@@
 For formulas $\varphi_0, \ldots, \varphi_n, \varphi \in
-\tformulas{Ts}$, the fact that $\varphi_0, \ldots, \varphi_n$
+\tformulas{T}$, the fact that $\varphi_0, \ldots, \varphi_n$
 \emph{entail} $\varphi$ is denoted $\varphi_0, \ldots, \varphi_n
@@
   % define:
-  For $\varphi_0, \varphi_1, I \in \tformulas{\mathcal{T}}$, if
+  For $\varphi_0, \varphi_1, I \in \tformulas{T}$, if
   %

### 2018-04-14 paper/introduction.tex
pass
added intro sentence
worked in cvoss comments

@@
 %
-The goal of a CHC solving problem is to find suitable interpretations
+A CHC system is a set of CHCs.
+%
+The goal of the CHC solving problem is to find suitable interpretations
 for each predicate such that each CHC is logically
@@
 In this work we focus on the subclass of CHC systems which are known
-as \emph{recursion-free}. In a recursion-free CHC system, the
-derivation of each predicate will not contain itself.
+as \emph{recursion-free}. In a recursion-free CHC system, no
+derivation of a predicate will invoke that predicate.
 %
@@
 % current general techniques for solving recursion-free systems
-Typically, recursion-free CHC systems are not solved directly.
+Typically, even recursion-free CHC systems are not solved directly.
 %
@@
 %
-CDD is a strict superset of the union of all previous introduced directly
-solvable systems.
+CDD is a strict superset of the union of previously introduced classes
+of directly solvable systems.
 %
@@
 However, empirically we have found that the size of $S'$ is usually
-close enough to the size of $S$ that \sys frequently outperforms the best
-known CHC solvers.
+close enough to the size of $S$ that \sys frequently outperforms
+\duality, one of the best known CHC solvers.
 %
@@
 %
-The results indicate that \sys outperforms modern solvers in a strong
-majority of cases.
+The results indicate that \sys outperforms modern solvers many cases.
 %

### 2018-04-14 paper/overview.tex
worked in cvoss comments

@@
 Such systems cannot model a program with multiple control paths that
-share a common pre-subpath, typically modeled as a CHC system with a
+share a common prefix, typically modeled as a CHC system with an
 uninterpreted predicate that occurs in the body of multiple clauses.
 %
-$\mcchc$ is not a body-disjoint system because $\cc{L}_4$ shows in the
+$\mcchc$ is not a body-disjoint system because $\cc{L}_4$ appears in the
 body of both clause (3) and clause (4).
@@
 In order to handle $\mcchc$, a solver that uses
-body-disjoint systems would have to create a fresh copy of
+body-disjoint systems would have to duplicate
 $\cc{L}_4$. 
@@
 Worse, if $\cc{L}_4$ had dependencies, then each dependency would also
-need to be copied.
+need to be duplicated.
 
@@
 %
-As shown, the class of CDD systems is a superclass of both the class
+As shown, the class of CDD systems is a superset of both the class
 of body-disjoint systems and the class of linear systems. So any

### 2018-04-14 paper/related-work.tex
worked in cvoss comments

@@
 % solving linear systems:
-\impact attempts to verify a given sequential procedure, which
-corresponds to solving a recursive linear CHC
+\impact attempts to verify a given sequential procedure by iteratively
+selecting paths and synthesizing invariants for each path. This
+approach corresponds to solving a recursive linear CHC
 system~\cite{mcmillan06}.
-%
-\impact attempts to verify a given procedure by iteratively selecting
-paths and synthesizing invariants for each path.
-%
-Such an approach corresponds to iteratively and selecting and solving
-derivations of a corresponding linear CHC system.
 
@@
 \duality attempts to avoid solving all derivation-trees (i.e
-body-disjoint system) by using lazy annotation~\cite{bjorner13}.
+body-disjoint systems) by using lazy annotation~\cite{bjorner13}.
 %
-Other optimizations selects derivation trees to solve using symbolic
+Other optimizations select derivation trees to solve using
 symbolic analogs of Prolog evaluation with
@@
 %
-\sys attempts to solve recursion-free CHC systems by attempting to
-reduce them to compact CDD systems.
+\sys attempts to solve recursion-free CHC systems by reducing them to
+compact CDD systems.
 %
@@
 %
-However, it is not apparent that such algorithms can be used directly
+However, it is not apparent whether such algorithms can be used directly
 to synthesize solutions.

### 2018-04-15 paper/background.tex
final pass

@@
 %
-The body is a conjunction of a logical formula, called the fact,
-and a vector of uninterpreted predicates. The fact is an arbitrary
+The body is a conjunction of a logical formula, called the constraint,
+and a vector of uninterpreted predicates. The constraint is an arbitrary
 formula in some background logic, such as linear integer arithmetic.
@@
 The uninterpreted predicates are applied to variables which may or may
-not appear in the fact.
+not appear in the constraint.
 %
@@
 For a given CHC $C$, $\bodyof{C}$ denotes the vector of uninterpreted
-predicates in the body and $\consof{C}$ denotes the fact in the
+predicates in the body and $\consof{C}$ denotes the constraint in the
 body.

### 2018-04-25 src/Formula/Z3.hs
cleaned up cut

@@ -5,15 +5,12 @@ module Formula.Z3 where
 import           Control.Lens
 import           Control.Monad.State
 import           Control.Monad.Except
-
-import           Data.List (partition)
+import           Data.List (partition, nub)
 import           Data.Maybe
 import qualified Data.Map as M
 import           Data.Map (Map)
-
 import           Formula (Expr((:@)), Chc, Type((:=>)), Var(..))
 import qualified Formula as F
-
 import           Z3.Monad hiding (local)
 
 data Env = Env { _envVars :: Map Var AST

### 2018-04-25 src/Shara/GrammarCut.hs

@@ -1,4 +1,4 @@
-module Shara.Cut where
+module Shara.GrammarCut where
 
 import           Control.Lens        hiding (below)
 import           Control.Monad.State
@@ -8,17 +8,77 @@ import qualified Data.Map            as M
 import           Data.Ratio
 import           Data.Set            (Set)
 import qualified Data.Set            as S
-import           Shara.Graph
+import           Shara.Grammar       hiding (partition)
 import qualified Shara.MinCut        as MC
+import qualified Shara.Reg           as R
 
-cut :: Graph -> (Set Symbol, Set Symbol, Set Symbol)
+type Symbol = Int
+
+data HyperGraph = HyperGraph
+  { graphForward  :: Map Symbol [Symbol]
+  , graphBackward :: Map Symbol [[Symbol]]
+  } deriving (Show, Read, Eq, Ord)
+
+successors :: Symbol -> HyperGraph -> [Symbol]
+successors s = M.findWithDefault [] s . graphForward
+
+predecessors :: Symbol -> HyperGraph -> [[Symbol]]
+predecessors s = M.findWithDefault [] s . graphBackward
+
+symbols :: HyperGraph -> Set Symbol
+symbols g = M.keysSet (graphForward g) `S.union` M.keysSet (graphBackward g)
+
+initials :: HyperGraph -> Set Symbol
+initials g = S.filter (\s -> all null (predecessors s g)) (symbols g)
+
+terminals :: HyperGraph -> Set Symbol
+terminals g = symbols g `S.difference` M.keysSet (graphForward g)
+
+clear :: Set Symbol -> HyperGraph -> HyperGraph
+clear toClear (HyperGraph for back) = HyperGraph for' back'
+  where
+    for' =
+      M.filterWithKey (\k _ -> k `notElem` toClear) $
+      fmap (filter (`notElem` toClear)) for
+    back' =
+      M.filterWithKey (\k _ -> k `notElem` toClear) $
+      fmap (map (filter (`notElem` toClear))) back
+
+grammarToGraph :: Grammar a -> HyperGraph
+grammarToGraph (SGrammar _ rs) = HyperGraph forw back
+  where
+    back = fmap ruleSymbols rs
+    forw = foldr addForw M.empty (M.toList back)
+    addForw ::
+         (Symbol, [[Symbol]]) -> Map Symbol [Symbol] -> Map Symbol [Symbol]
+    addForw (snk, srcs) m =
[…]

### 2018-04-25 src/Shara/Graph.hs

@@ -1,67 +1,0 @@
-module Shara.Graph where
-
-import           Data.Map      (Map)
-import qualified Data.Map      as M
-import           Data.Set      (Set)
-import qualified Data.Set      as S
-import           Shara.Grammar
-import qualified Shara.Reg     as R
-
-type Symbol = Int
-
-data Graph = Graph
-  { graphForward  :: Map Symbol [Symbol]
-  , graphBackward :: Map Symbol [[Symbol]]
-  } deriving (Show, Read, Eq, Ord)
-
-successors :: Symbol -> Graph -> [Symbol]
-successors s = M.findWithDefault [] s . graphForward
-
-predecessors :: Symbol -> Graph -> [[Symbol]]
-predecessors s = M.findWithDefault [] s . graphBackward
-
-symbols :: Graph -> Set Symbol
-symbols g = M.keysSet (graphForward g) `S.union` M.keysSet (graphBackward g)
-
-initials :: Graph -> Set Symbol
-initials g = S.filter (\s -> all null (predecessors s g)) (symbols g)
-
-terminals :: Graph -> Set Symbol
-terminals g = symbols g `S.difference` M.keysSet (graphForward g)
-
-clear :: Set Symbol -> Graph -> Graph
-clear toClear (Graph for back) = Graph for' back'
-  where
-    for' =
-      M.filterWithKey (\k _ -> k `notElem` toClear) $
-      fmap (filter (`notElem` toClear)) for
-    back' =
-      M.filterWithKey (\k _ -> k `notElem` toClear) $
-      fmap (map (filter (`notElem` toClear))) back
-
-grammarToGraph :: Grammar a -> Graph
-grammarToGraph (SGrammar _ rs) = Graph forw back
-  where
-    back = fmap ruleSymbols rs
-    forw = foldr addForw M.empty (M.toList back)
-    addForw ::
-         (Symbol, [[Symbol]]) -> Map Symbol [Symbol] -> Map Symbol [Symbol]
-    addForw (snk, srcs) m =
-      foldr (\src -> M.insertWith (++) src [snk]) m (concat srcs)
-    ruleSymbols :: Rule a -> [[Symbol]]
-    ruleSymbols =
-      \case
-        R.Null -> []
-        R.Eps -> []
-        R.Neg a -> ruleSymbols a
-        R.Alt a b -> ruleSymbols a ++ ruleSymbols b
-        R.Seq a b ->
-          let as = ruleSymbols a
-              bs = ruleSymbols b
-          in if null as
[…]

### 2018-04-25 src/Shara/Interpolate.hs

@@ -9,9 +9,8 @@ import           Data.Maybe                (fromMaybe)
 import qualified Data.Set                  as S
 import           Formula
 import qualified Formula.Z3                as Z3
-import           Shara.Cut
 import           Shara.Grammar
-import           Shara.Graph
+import           Shara.GrammarCut
 import qualified Shara.Reg                 as R
 
 import           Data.Text.Prettyprint.Doc

### 2018-04-25 src/Shara/MinCut.hs

@@ -201,31 +201,3 @@ dfsFrom s = do
 
 search :: Ord k => k -> Map k (Set a) -> Set a
 search = M.findWithDefault S.empty
-
-testG :: Graph
-testG =
-  empty
-  & addEdge 0 1
-  & addEdge 0 2
-  & addEdge 1 2
-  & addEdge 1 3
-  & addEdge 1 4
-  & addEdge 2 4
-  & addEdge 3 5
-  & addEdge 4 3
-  & addEdge 4 5
-  & addInitial 0
-  & addTerminal 5
-
-testCap :: Map (Symbol, Symbol) Rational
-testCap =
-  mempty
-  & M.insert (0, 1) 10
-  & M.insert (0, 2) 10
-  & M.insert (1, 2) 2
-  & M.insert (1, 3) 4
-  & M.insert (1, 4) 8
-  & M.insert (2, 4) 9
-  & M.insert (3, 5) 10
-  & M.insert (4, 3) 6
-  & M.insert (4, 5) 10

### 2018-04-27 solver.tex
can solve paper example

\documentclass{article}

\usepackage{relsize}
\usepackage{amsmath}

\newcommand{\cc}[1]{\mbox{\smaller[0.5]\texttt{#1}}}

\newcommand{\DUALITY}{\textsc{Duality}}
\newcommand{\SHARA}{\textsc{Shara}}
\newcommand{\RHORT}{\textsc{Rhort}}

\begin{document}
When \DUALITY~attempts to interpolate a DAG, it is forced to unwind the DAG to a
tree.
%
When this DAG happens to have been constructed from complex proof rules (such
as in the case of Relational, Higher-Order Refinement Type Inference) this tree
will probably be far too large for \DUALITY~to solve in a reasonable amount of
time.
%
There are three sources of inefficiency which we can fix with a more
appropriate solver for this type of problem.

\begin{enumerate}
\item The solver need not unwind the DAG to a tree, but rather only a disjoin
dependency DAG (\SHARA). The importance of this optimization is greatly
exaggerated by proof systems that include a context, since these contexts often
end up being used across various branches of the same proof.
%
This means that it is very likely that \SHARA~could reduce the number of
required interpolants exponentially compared to \DUALITY~in most cases.
%
\item The Solver need not calculate every interpolant. The reason for this is
quite simple: We \emph{can't} use certain interpolants to help us construct a
proof. So there is no reason to calculate an interpolant there in the first
place. The biggest source of this is, again, contexts. In \RHORT, every Lambda
and Match expression spawns new context vertices. However, none of the facts at
these interpolants can be used to help generalize a proof.
%
\item The interpolants need not be solved in any particular order. The approach
taken in \SHARA~was to solve the interpolants in topological order. However,
this is not essential, merely convenient. A more appropriate strategy would be
to try to maximally reduce the number of dependencies in the DAG at each step.

\end{enumerate}

I propose a solver which takes advantage of these three facts. First, I propose
a modification to the solver's type. \DUALITY~can be seen as having the following type:

\begin{align*}
\DUALITY : \mathsf{CHC} \rightarrow \mathsf{Query} \rightarrow (\mathsf{Model} + \mathsf{Proof})
\end{align*}

\DUALITY~takes a (potentially recursive) CHC system and a query, and finds a
counterexample model or a valid proof. \SHARA~should instead have the following type:

\begin{align*}
\SHARA &: \mathsf{ProofStructure} \\
  &\rightarrow \mathsf{NonrecCHC} \\
  &\rightarrow \mathsf{Query} \\
  &\rightarrow (\mathsf{Model} + \mathsf{PartialProof} + \mathsf{Proof})
\end{align*}

In addition to asking the CHC system to be nonrecursive, this new signature
asks the user to provide a structure for the current proof. This structure
contains the information needed to construct entailments whose validity would
generalize a hierarchical proof.
%
\SHARA~can use this information to 1) reduce the number of interpolant
calculations and to 2) parallelize computing entailments of the interpolants.
%
\SHARA~has three possible terminating outcomes.
\begin{enumerate}
\item A counterexample model was found.
\item A proof of the hierarchical program was found, but it did not generalize.
In this case, the partial proof will indicate which interpolants generalized
and which did not so that the user can intelligently restrict construction of
their next hierarchical approximation.
\item A proof of the hierarchical program was found, and it generalized. This
would indicate the user is now done with their proof entirely.
\end{enumerate}

The operation of \SHARA~is to be a parallel, recursive algorithm which at each
step attempts to reduce the number of dependencies in the DAG. At any point, if
there are no \emph{relevant} interpolants left to calculate, that branch of the
computation can be killed. An interpolant is relevant if it appears somewhere
in the proof structure.

\SHARA~works in the following way: First, it constructs a disjoint dependency
DAG based on the input DAG. It then performs the following recursive algorithm:

\begin{enumerate}
\item \SHARA~calculates a \emph{minimum cut} of the current subgraph. The
minimum cut is calculated by assigning each vertex in the subgraph a weight
inversely related to the number of dependencies which cross that vertex. Two
vertices are dependent on one another if they exist on the same hyperpath in
the DAG. This cut partitions the subgraph into three pieces: The vertices which
must be interpolated \emph{now} to perform the cut, and the two new subgraphs
on either side of this cut.
\item \SHARA~interpolates the vertices along the cut before recursively
dispatching the two subgraphs \emph{in parallel}.
\item \SHARA~combines the resulting interpolant solutions from the cut and the
two parallel calls.
\item If the current invocation has the solution for all instances of a
relevant vertex from the original DAG, it combines those solutions to construct
an interpolant for the original DAG.
\item If the current invocation has the solution to both relevant vertices
involved in an entailment in the proof structure, it can query the entailment
before returning.
\end{enumerate}

In addition to this, the following optimizations can be made:
\begin{enumerate}
\item The proof structure can contain distinct, optional ways to complete the
proof tree at each level. If \SHARA~finds that one of these options is valid,
the remaining options can be removed, potentially reducing the number of
relevant interpolants.
\item If the current subgraph contains only a small number of relevant
interpolants (below some threshold), it can calculate them all sequentially
rather than continuing to cut. This could reduce the number of interpolants to
calculate since each cut might contain irrelevant interpolants.
\end{enumerate}

In short, the purpose of this algorithm is to try to gain two advantages over
\DUALITY: 1) To aggresively reduce the number of required interpolant
calculations 2) To parallelize SMT calls (both interpolation and and
entailment) as much as possible.
\end{document}

### 2018-04-27 app/Main.hs

@@ -4,8 +4,8 @@ import           Control.Monad.State
 import           Data.Map                  (Map)
 import qualified Data.Map                  as M
 import           Data.Text.Prettyprint.Doc
-import           Formula
-import           Prelude                   hiding (seq)
+import           Formula                   hiding (res)
+import           Prelude                   hiding (abs, seq)
 import           Shara.CDD
 import           Shara.Grammar
 import           Shara.Interpolate
@@ -30,6 +30,64 @@ test1 =
 test1M :: Map NT [Var]
 test1M = M.fromList [(0, [Var "x" Int])]
 
+test2 :: SGrammar [Var] Expr
+test2 =
+  SGrammar
+    (SNonterm ([n, res], m) `R.seq` STerm [expr|res < 0|])
+    (M.fromList
+       [ ( m
+         , mconcat
+             [ SNonterm ([n, abs'], l9)
+             , SNonterm ([x, d], dbl)
+             , STerm [expr|abs' = x && res = d|]
+             ])
+       , ( l9
+         , R.alt
+             (SNonterm ([n, abs], l8) `R.seq` STerm [expr|abs' = (0 - n)|])
+             (SNonterm ([n, abs], l6) `R.seq` STerm [expr|abs' = n|]))
+       , (l8, SNonterm ([n, abs], l4) `R.seq` STerm [expr|n < 0|])
+       , (l6, SNonterm ([n, abs], l4) `R.seq` STerm [expr|n >= 0|])
+       , (l4, STerm [expr|abs = 0|])
+       , (dbl, STerm [expr|d = 2 * x|])
+       ])
+  where
+
+
+test2M :: Map NT [Var]
+test2M =
+  M.fromList
+    [ (m, [n, res])
+    , (l9, [n, abs'])
+    , (l8, [n, abs])
+    , (l6, [n, abs])
+    , (l4, [n, abs])
+    , (dbl, [x, d])
+    ]
+
+m = 0
+
+dbl = 1
+
+l9 = 2
+
+l8 = 3
+
+l6 = 4
+
+l4 = 5
+
+n = Var "n" Int
+
+x = Var "x" Int
+
+d = Var "d" Int
+
+abs = Var "abs" Int
+
+abs' = Var "abs'" Int
+
+res = Var "res" Int
+
 main :: IO ()
 main = do
   sk <-
@@ -45,13 +103,6 @@ main = do
                     else LicketySplit SequentialInterpolation
[…]

### 2018-04-27 src/Shara/CDD.hs

@@ -16,6 +16,8 @@ import           Shara.Grammar
 import           Shara.Inf
 import qualified Shara.Reg           as R
 
+import Debug.Trace
+
 type Branch = Int
 type Unrolling = Int
 
@@ -90,8 +92,6 @@ cdd g = currentBranch .= 0 >> cddRule (start g & allFresh 0)
         R.Seq a b ->
           Infinite <$>
           (seq <$> (getInfinite <$> cddRule a) <*> (getInfinite <$> cddRule b))
-        R.Neg a ->
-          Infinite . neg . getInfinite <$> cddRule a
         STerm t -> pure (iflat $ Term (Fin t))
         R.Alt a b -> do
           -- We need to invent 2 new branch points, one for each branch
@@ -104,15 +104,13 @@ cdd g = currentBranch .= 0 >> cddRule (start g & allFresh 0)
           -- Also, manage the dependencies
           elem nt <$> use visited >>= \case
             -- We have alrady seen this nonterminal, so postpone expanding it.
-            True -> do
+            True ->
               pure (
                 iflat $ Term (Inf (do
                 visited .= S.empty
                 cddRule (SNonterm (vs, nt)))))
             False -> do
               -- We have not yet seen this nonterminal, so mark it.
-              visited %= S.insert nt
-
               cb <- use currentBranch
               -- Create a fresh nonterminal if needed, or use one we do not
               -- already depend on.
@@ -128,7 +126,9 @@ cdd g = currentBranch .= 0 >> cddRule (start g & allFresh 0)
                   nt' <- freshNT nt
                   addDep cb nt'
                   vocab <- substMap vs nt nt'
+                  visited %= S.insert nt
                   g' <- getInfinite <$> cddRule (ruleFor nt g & allFresh nt')
+                  visited %= S.delete nt
                   pure $ Infinite $ (flat (Term (Fin vocab)) `seq` abstract () nt' g')
 
 mkBranch :: MonadState CDDState m => Branch -> m ()

### 2018-04-27 src/Shara/Grammar.hs

@@ -2,6 +2,7 @@
 
 module Shara.Grammar where
 
+import           Control.Arrow
 import           Control.Applicative
 import           Control.Lens
 import           Control.Monad.Reader
@@ -52,9 +53,6 @@ alt :: SGrammar s a -> SGrammar s a -> SGrammar s a
 alt (SGrammar s rs) (SGrammar s' rs') =
   SGrammar (s `R.alt` s') (M.unionWith R.alt rs rs')
 
-neg :: SGrammar s a -> SGrammar s a
-neg (SGrammar s rs) = SGrammar (R.Neg s) rs
-
 abstract :: s -> NT -> SGrammar s a -> SGrammar s a
 abstract sc nt (SGrammar s rs) =
   SGrammar (SNonterm (sc, nt)) (M.insertWith R.alt nt s rs)
@@ -87,19 +85,18 @@ after nt g =
         R.Seq x y -> go x >> go y
         _ -> pure ()
 
--- | This only works for commutative semirings.
-reverseRules :: Grammar a -> Map NT (Rule a)
-reverseRules g = runReader (go $ start g) R.Eps
+reverseRules :: Grammar a -> Map NT [(Maybe NT, Rule a)]
+reverseRules g = runReader (go $ start g) (Nothing, R.Eps)
   where
     go =
       \case
         R.Seq a b -> do
-          M.union <$> local (R.seq b) (go a) <*> local (R.seq a) (go b)
+          M.unionWith (++) <$> local (second (R.seq b)) (go a) <*> local (second (R.seq a)) (go b)
         R.Alt a b -> M.union <$> go a <*> go b
         Nonterm nt -> do
-          r <- ask
-          M.union <$> local (const (R.Neg $ Nonterm nt)) (go (ruleFor nt g)) <*>
-            pure (M.singleton nt r)
+          ctx <- ask
+          M.unionWith (++) <$> local (const (Just nt, R.Eps)) (go (ruleFor nt g)) <*>
+            pure (M.singleton nt [ctx])
         _ -> pure mempty
 
 grammarFromMap :: NT -> Map NT (SRule s a) -> SGrammar s a
@@ -144,8 +141,6 @@ partition nts g = (bef, aft)
              SGrammar (R.alt st st') (M.unionWith R.alt rs rs')) <$>
           go x <*>
           go y
-        R.Neg x -> go x >>= \case
-                     SGrammar st rs -> pure $ SGrammar (R.Neg st) rs
[…]

### 2018-04-27 src/Shara/GrammarCut.hs

@@ -35,14 +35,14 @@ terminals :: HyperGraph -> Set Symbol
 terminals g = symbols g `S.difference` M.keysSet (graphForward g)
 
 clear :: Set Symbol -> HyperGraph -> HyperGraph
-clear toClear (HyperGraph for back) = HyperGraph for' back'
+clear toKeep (HyperGraph for back) = HyperGraph for' back'
   where
     for' =
-      M.filterWithKey (\k _ -> k `notElem` toClear) $
-      fmap (filter (`notElem` toClear)) for
+      M.filterWithKey (\k _ -> k `elem` toKeep) $
+      fmap (filter (`elem` toKeep)) for
     back' =
-      M.filterWithKey (\k _ -> k `notElem` toClear) $
-      fmap (map (filter (`notElem` toClear))) back
+      M.filterWithKey (\k _ -> k `elem` toKeep) $
+      fmap (map (filter (`elem` toKeep))) back
 
 grammarToGraph :: Grammar a -> HyperGraph
 grammarToGraph (SGrammar _ rs) = HyperGraph forw back
@@ -58,7 +58,6 @@ grammarToGraph (SGrammar _ rs) = HyperGraph forw back
       \case
         R.Null -> []
         R.Eps -> []
-        R.Neg a -> ruleSymbols a
         R.Alt a b -> ruleSymbols a ++ ruleSymbols b
         R.Seq a b ->
           let as = ruleSymbols a

### 2018-04-27 src/Shara/Interpolate.hs

@@ -1,19 +1,17 @@
 module Shara.Interpolate where
 
-import           Control.Concurrent.Async  (concurrently)
+import           Control.Concurrent.Async (concurrently)
 import           Control.Monad.Except
 import           Control.Monad.State
-import           Data.Map                  (Map)
-import qualified Data.Map                  as M
-import           Data.Maybe                (fromMaybe)
-import qualified Data.Set                  as S
+import           Data.Map                 (Map)
+import qualified Data.Map                 as M
+import           Data.Set (Set)
+import qualified Data.Set                 as S
 import           Formula
-import qualified Formula.Z3                as Z3
+import qualified Formula.Z3               as Z3
 import           Shara.Grammar
 import           Shara.GrammarCut
-import qualified Shara.Reg                 as R
-
-import           Data.Text.Prettyprint.Doc
+import qualified Shara.Reg                as R
 
 data InterpolationStrategy
   = SequentialInterpolation
@@ -27,7 +25,9 @@ licketySplit ::
   => InterpolationStrategy
   -> Grammar Expr
   -> m (Either Model (Map NT Expr))
-licketySplit strategy g = liftIO (loop strategy g (reverseRules g) M.empty)
+licketySplit strategy g =
+  let ss = nonterminals g in
+      liftIO (loop strategy ss g (reverseRules g) M.empty)
 
 -- | The core loop for interpolating a DAG. The loop requires two versions of
 -- the DAG, one which is a full representation and another which is a
@@ -38,25 +38,25 @@ licketySplit strategy g = liftIO (loop strategy g (reverseRules g) M.empty)
 -- restrictions which can be interpolated in parallel.
 loop ::
      InterpolationStrategy
+  -> Set NT
   -> Grammar Expr
-  -> Map NT (Rule Expr)
+  -> Map NT [(Maybe NT, Rule Expr)]
   -> Map NT Expr
   -> IO (Either Model (Map NT Expr))
-loop strategy g@(SGrammar _ rs) rev m =
-  if null (M.keysSet rs S.\\ M.keysSet m)
+loop strategy targets g rev m = do
+  if null (targets S.\\ M.keysSet m)
     then pure (Right m)
     else do
[…]

### 2018-04-27 src/Shara/Reg.hs

@@ -16,7 +16,6 @@ data Reg a
         (Reg a) -- Sequence two expressions one after the other.
   | Alt (Reg a)
         (Reg a) -- Alternation: Choose one or the other.
-  | Neg (Reg a)
   | Eps -- The expression that matches the empty string.
   | Null -- The expression that matches nothing.
   | Symbol a -- The expression that matches exactly the given symbol.
@@ -47,7 +46,6 @@ instance Functor Reg where
     \case
       Seq a b -> seq (fmap f a) (fmap f b)
       Alt a b -> alt (fmap f a) (fmap f b)
-      Neg a -> Neg (fmap f a)
       Null -> Null
       Eps -> Eps
       Symbol a -> Symbol (f a)
@@ -60,7 +58,6 @@ instance Monad Reg where
         \case
           Seq a b -> seq (join' a) (join' b)
           Alt a b -> alt (join' a) (join' b)
-          Neg a -> Neg (join' a)
           Eps -> Eps
           Null -> Null
           Symbol a -> a
@@ -86,7 +83,6 @@ instance Arbitrary a => Arbitrary (Reg a) where
     oneof
       [ seq <$> arbitrary <*> arbitrary
       , alt <$> arbitrary <*> arbitrary
-      , Neg <$> arbitrary
       , pure Eps
       , pure Null
       , Symbol <$> arbitrary

### 2018-04-27 src/Shara/Shara.hs

@@ -2,19 +2,17 @@ module Shara.Shara where
 
 import           Control.Lens
 import           Control.Monad.State
-import           Data.List.Split           (splitOn)
-import           Data.Map                  (Map)
-import qualified Data.Map                  as M
-import           Data.Set                  (Set)
-import qualified Data.Set                  as S
+import           Data.List.Split     (splitOn)
+import           Data.Map            (Map)
+import qualified Data.Map            as M
+import           Data.Set            (Set)
+import qualified Data.Set            as S
 import           Formula
-import qualified Formula.Z3                as Z3
+import qualified Formula.Z3          as Z3
 import           Shara.CDD
 import           Shara.Grammar
 import           Shara.Interpolate
-import qualified Shara.Reg                 as R
-
-import           Data.Text.Prettyprint.Doc
+import qualified Shara.Reg           as R
 
 data SolveKind
   = Topological
@@ -29,13 +27,12 @@ shara ::
   -> m (Either Model (Map NT Expr))
 shara sk vocab sg = evalStateT (go =<< cdd sg) (emptyCDDState vocab)
   where
-    go g =
+    go g = do
       solveDirect sk (finitePrefix g) >>= \case
         Left m -> pure (Left m)
         Right m -> do
           cs <- use clones
           let sol = collapse cs m
-          -- liftIO $ print (fmap pretty sol)
           inductive vocab sol sg >>= \case
             Nothing -> pure (Right sol)
             Just inds -> go =<< unrollCDD g {- inds -}
@@ -53,9 +50,7 @@ inductive ::
   -> Map NT Expr
   -> SGrammar [Var] Expr
   -> m (Maybe (Set NT))
-inductive vars sols g
-  -- liftIO $ print sols
- = do
+inductive vars sols g = do
   s <- (S.fromList . concat) <$> M.traverseWithKey indRule (rules g)
   pure $
     if null (nonterminals g S.\\ s)

### 2018-04-27 src/Shara/GrammarCut.hs
working version

@@ -38,7 +38,7 @@ clear :: Set Symbol -> HyperGraph -> HyperGraph
 clear toKeep (HyperGraph for back) = HyperGraph for' back'
   where
     for' =
-      M.filterWithKey (\k _ -> k `elem` toKeep) $
+      M.filterWithKey (\k v -> k `elem` toKeep && not (null v)) $
       fmap (filter (`elem` toKeep)) for
     back' =
       M.filterWithKey (\k _ -> k `elem` toKeep) $
@@ -70,15 +70,16 @@ grammarToGraph (SGrammar _ rs) = HyperGraph forw back
         Nonterm nt -> [[nt]]
         Term _ -> []
 
-cut :: HyperGraph -> (Set Symbol, Set Symbol, Set Symbol)
-cut g =
-  let (mcg, caps) = mkCutGraph g
+cut :: Set Symbol -> HyperGraph -> (Set Symbol, Set Symbol, Set Symbol)
+cut uncuttable g =
+  let (mcg, caps) = mkCutGraph uncuttable g
   in MC.mincut caps mcg
 
 -- | Construct a simple, non-hyper graph and a map of costs from an input
 -- hypergraph.
-mkCutGraph :: HyperGraph -> (MC.Graph, Map (Symbol, Symbol) Rational)
-mkCutGraph g
+mkCutGraph ::
+     Set Symbol -> HyperGraph -> (MC.Graph, Map (Symbol, Symbol) Rational)
+mkCutGraph uncuttable g
   -- Find each pairing where each pairing represents a hyperedge. The list of
   -- symbols are the sources of the edge and the distinguished symbol is the
   -- sink.
@@ -115,51 +116,50 @@ mkCutGraph g
       -> Symbol
       -> (MC.Graph, Map (Symbol, Symbol) Rational)
       -> (MC.Graph, Map (Symbol, Symbol) Rational)
-    singlePair (snk, below) aboves src (mcg, cap)
-      -- Find the distinguished source's above count and separate it from the others.
-     =
-      let ([(_, srcUp)], aboves') = partition ((== src) . fst) aboves
-          -- The vertices above the source are dependent on the vertices below the
-          -- sink as well as all vertices above all other sources.
-          allDown = below + sum (map snd aboves')
-          -- The number of dependencies along this partial edge is the product of
-          -- the dependencies above the edge versus those below.
-          numDeps = allDown * srcUp
-          -- The cost of the edge is the inverse of the number of dependencies (or
[…]

### 2018-04-27 src/Shara/Interpolate.hs

@@ -5,6 +5,7 @@ import           Control.Monad.Except
 import           Control.Monad.State
 import           Data.Map                 (Map)
 import qualified Data.Map                 as M
+import           Data.Semigroup
 import           Data.Set (Set)
 import qualified Data.Set                 as S
 import           Formula
@@ -47,16 +48,20 @@ loop strategy targets g rev m = do
   if null (targets S.\\ M.keysSet m)
     then pure (Right m)
     else do
-      let gr = clear (targets S.\\ M.keysSet m) $ grammarToGraph g
-      -- let gr = grammarToGraph g
-      let (now, half1, half2) = cut gr
-      runStateT (runExceptT $ mapM_ (\nt -> interpolateNT' g rev nt) now) m >>=
+      let gr = clear targets $ grammarToGraph g
+      let (now', half1', half2') = cut (M.keysSet m) gr
+      let (now, half1, half2) = if null (now' S.\\ M.keysSet m)
+          then (targets S.\\ M.keysSet m, S.empty, S.empty)
+          else (now', half1', half2')
+      runStateT (runExceptT $ mapM_ (interpolateNT' g rev) (now S.\\ M.keysSet m)) m >>=
       -- We only proceed when none of the solved symbols failed.
        \case
         (Left m', _) -> pure (Left m')
         (_, m') -> do
           (m1, m2) <-
-            evaluator (loop strategy half1 g rev m') (loop strategy half2 g rev m')
+            evaluator
+              (loop strategy (now <> half1) g rev m')
+              (loop strategy (now <> half2) g rev m')
         -- A simple union of the two maps suffices since there should be no
         -- possibility that the two subiterations computed interpolants for the
         -- same vertices.

### 2018-04-27 src/Shara/MinCut.hs

@@ -173,10 +173,11 @@ dfsFrom s = do
     -- Otherwise, we can try to proceed both forward and backward from all
     -- edges connected to this vertex.
     | otherwise -> do
-      g <- view theGraph
-      succPaths <- concatMapM for $ S.toList (search s (view forward g))
-      predPaths <- concatMapM back $ S.toList (search s (view backward g))
-      pure (succPaths ++ predPaths)
+      local (visited %~ S.insert s) $ do
+        g <- view theGraph
+        succPaths <- concatMapM for $ S.toList (search s (view forward g))
+        predPaths <- concatMapM back $ S.toList (search s (view backward g))
+        pure (succPaths ++ predPaths)
   where
     -- We can proceed forward when there is capacity left in the edge.
     for :: Symbol -> Reader ExploreCtxt [Set Edge]
@@ -186,7 +187,7 @@ dfsFrom s = do
       -- We recurse, making sure to add the current edge to the eventual traversal and
       -- noting that we've already seen the current vertex.
       if weight < cap
-      then map (S.insert (s, t, Forward)) <$> local (visited %~ S.insert s) (dfsFrom t)
+      then map (S.insert (s, t, Forward)) <$> dfsFrom t
       else pure []
 
     -- We can proceed backward when there is weight in the edge.
@@ -196,7 +197,7 @@ dfsFrom s = do
       -- Note that we lookup the weight using the forward direction of the edge.
       let w = M.findWithDefault 0 (t, s) ws
       if w > 0
-      then map (S.insert (s, t, Backward)) <$> local (visited %~ S.insert s) (dfsFrom t)
+      then map (S.insert (s, t, Backward)) <$> dfsFrom t
       else pure []
 
 search :: Ord k => k -> Map k (Set a) -> Set a

### 2018-04-27 src/Shara/Shara.hs

@@ -27,7 +27,7 @@ shara ::
   -> m (Either Model (Map NT Expr))
 shara sk vocab sg = evalStateT (go =<< cdd sg) (emptyCDDState vocab)
   where
-    go g = do
+    go g =
       solveDirect sk (finitePrefix g) >>= \case
         Left m -> pure (Left m)
         Right m -> do

### 2018-04-27 paper/sat-2017-rebuttals/rebuttal.txt

We thank the reviewers for their detailed comments. Responses are
included below, where each summarized comment is labeled with the
reviews that contained it. Each of the points will be addressed in a
future version of the paper.

--------------------------------------------------------------------

Q2 (Review 1): What is the meaning of "Query_S that is the dependency
of no relation predicate", page 8?

A: An alternative phrasing is that Query_S is not the dependency of
any relational predicate.

---------------------------------------------------------------------

Q3 (Review 1): Does Defnition 6 imply that the dependency graph of
each body is a forest?

A: Not necessarily; the following system is CDD:

(1) R1 => R0
(2) R3 => R1
(3) R4 => R1
(4) R5 => R3
(5) R5 => R4

However, the dependency graph of the body of clause (1) is a DAG.

--------------------------------------------------------------------

Q4 (Review 1): What is Use[R] in "Constructing the Counterexample
Characterization of a CDD"? In this section, why is T used as a
metavariable for an arbitrary CHC system?

A: Use[R] is a boolean predicate to indicate if relational predicate R
occurs in a hypothetical counterexample. The constraints on it are
described in detail in paragraph "Constructing the Counterexample
Characterization of a CDD", page 10.

We have used a fresh variable T so that the counterexample
characterization can be defined once, in general, for an arbitrary
system and instantiated for multple systems, each introduced in
"Constructing the post constraint of a relational predicate."

The variable S is fixed prior to this section as an arbitrary CHC
system to be solved.

--------------------------------------------------------------------

Q5 (Review 1): Can the pre-constraint and post-constraint be
characterized precisely?

A: After giving the intuition behind each constraint, we have given
the formula itself at the end of the paragraph.

---------------------------------------------------------------------

Q1 (Review 1): Was the Expand procedure introduced in previous work?

A: We have included the definition of Expand in Appendix A.

-------------------------------------------------------------------------

Q5 (Review 1): Why can SHARA not parse benchmarks generated by SeaHorn?

A: SHARA correctly parses benchmarks generated by SeaHorn, and these
benchmarks are the subject of our evaluation. The Eldarica CHC solver
[20], a solver to which we attempted to compare SHARA, could not parse
such systems.

---------------------------------------------------------------------------

Q6 (Review 2): Why did you evaluate SHARA compared only to Duality?
Did you try to compare SHARA to other tools?

A: We primarily evaluated SHARA compared to Duality because it is the
most reliable competitive solver that we could find. We also attempted
to compare SHARA to Eldarica on benchmarks generated by SeaHorn;
Eldarica failed to parse such benchmarks. Previous results strongly
indicate that Duality can consistently and significantly outperform
SeaHorn [18]. These were confirmed by our own informal experiments,
which we can report in a future version of this paper.

---------------------------------------------------------------------

Q7 (Review 2): The paper is missing a conclusion and discussion of
potential optimizations as future work.

A: The current paper speculates one potential optimization in Sec. 5,
final paragraph. We will include a more extensive discussion in a
future version of the paper.

---------------------------------------------------------------------

Q8 (Review 3): The paper is in a specialized area of verification that
is unlikely to be interesting to the SAT audience.

A: While we have only discussed one relatively simple example of a
safety verification problem formulated as a CHC system, previous work
has established that CHC systems can formulate an extremely expressive
set of safety-verification problems, including safety of sequential
recursive programs, safety of concurrent programs, and
information-flow security of programs.

CHC solving, which in general is defined as an instance of Constraint
Programming over SMT theories independent of any verification problem,
has also been shown to be useful in modeling a variety of systems,
automata, and concurrent processes [Jaffar], and in knowledge
representation within AI [Brachman].

While the expressive power of CHC systems has been well-studied,
effective techniques for solving them, particularly those based on
interpolation over SMT theories, are relatively new. We believe that
just as techniques communicated between the SAT/SMT and verification
communities for solving SAT and SMT formulas have significantly
strengthened both fields, both communities will benefit from the
mutual study of this branch of Constraint Programming.

---------------------------------------------------------------------

References:

Brachman, Ronald J., Hector J. Levesque, and Raymond Reiter. Knowledge
representation. MIT press, 1992.

Jaffar, Joxan. "A CLP approach to modelling systems." Lecture Notes in
Computer Science (2004).

### 2018-04-27 paper/evaluation.tex

--- 2018-04-13 paper/evaluation.tex
@@
 We performed an empirical evaluation of \sys to determine how it
-compares to existing solvers for recursion-free CHC systems.
+compares to existing CHC solvers.
 %
@@
 We modified the algorithm used by \duality to generate recursion-free
-unwindings of a given recursive system so that in each iteration, it
-generates an unwinding with a maximal set of relational predicates and
-clauses.
+unwindings of a given recursive system so that, in each iteration, it
+generates an unwinding which is converted to CDD form.
 %
-In the following context, ``\sys'' refers to \duality modified to
-solve general CHC systems.
+In the following context, ``\sys'' refers to this modified version of
+\duality.
 
 % describe benchmarks:
-We evaluated \sys and \duality on 4,309 CHC systems generated from
-programs in the SV-COMP 2015~\cite{svcomp15} verification benchmark
-suite.
+We evaluated \sys and an unmodified version of \duality on 4,309 CHC
+systems generated from programs in the SV-COMP 2015~\cite{svcomp15}
+verification benchmark suite.
 %
@@
 
-We also attempted to compare \sys to the \eldarica CHC solver, but
+We also compared \sys to the \eldarica CHC solver, but
 \eldarica could not parse the CHC systems generated by \seahorn.
@@
 %
-The data indicates that that performance improvement of \sys compared
+The data indicates that the performance improvement of \sys compared
 to \duality is consistent across systems of all sizes available.
@@
 algorithm yield better results.
+%
+Specifically, the unmodified version of \duality~uses a technique
+called lazy annotation to avoid enumerating all derivation trees.
+\sys~does not use lazy annotation.
 %
@@
 Our results indicate that a third approach that combines the strengths
-of both \duality and \sys, perhaps by lazily unwinding a given
-recursion-free system into a series of CDD systems instead of
-derivation trees, could yield further improvements.
+of both \duality and \sys, perhaps by \emph{lazily} unwinding a given
+system into a series of CDD systems, could yield further improvements.
 

### 2018-04-27 paper/fig/hierarchy2.tex

% \scalebox{0.75}{%
% }

\begin{tikzpicture}
\tikzstyle{every node}=[draw=black, thin, ellipse]
  \begin{scope}[fill opacity=0.1,text opacity=1]

  \draw node[minimum width=6.2cm,minimum height=6cm,
    label={[anchor=north, inner sep=5pt]north:
      \begin{tabular}{c}Recursion-Free\\\textsc{co-NEXPTIME Complete}\end{tabular}} ]
    at (0,0)
    {};

  \draw node[minimum width=5.5cm,minimum height=4.5cm,
    label={[anchor=south, inner sep=0pt]south:
      \begin{tabular}{c}Directly Solvable\\\textsc{co-NP}\end{tabular}} ]
    at (0,-6mm)
    {};

  \draw node[very thick, minimum width=5cm,minimum height=3cm,
    label={[anchor=north, inner sep=5pt]north:CDD}]
    at (0,-1mm)
    {};

  \draw node[minimum width=2.7cm,minimum height=2cm]
    at (10mm,-2mm)
    {\begin{tabular}{c}Body\\Disjoint\end{tabular}};

  \draw node[minimum width=2.7cm,minimum height=2cm]
    at (-10mm,-2mm)
    {Linear};

    % \begin{tabular}{c}Directly Solvable\\\textsc{co-NP}\end{tabular}} ]
  \end{scope}
\end{tikzpicture}

### 2018-04-27 paper/introduction.tex

@@
 The right side has any number of such predicates conjoined together
-with a \emph{fact} ($\varphi$). The fact is a logical formula in a background
+with a \emph{constraint} ($\varphi$). The constraint is a logical formula in a background
 theory and may use variables named by the predicates.

### 2018-04-27 src/Shara/CDD.hs
refactoring

@@ -188,19 +188,16 @@ addDep b nt = do
   mapM_ (connect b) bs
   ntBranches . at nt <>= Just (S.singleton b)
   branchDeps . at b <>= Just (S.singleton nt)
-
--- | Connect two branches by adding them as neighbors.
-connect :: MonadState CDDState m => Branch -> Branch -> m ()
-connect a b = do
-  branchSibs . at a <>= Just (S.singleton b)
-  branchSibs . at b <>= Just (S.singleton a)
+  where
+    connect a b = do
+      branchSibs . at a <>= Just (S.singleton b)
+      branchSibs . at b <>= Just (S.singleton a)
 
 -- | Construct a variable whose name reflects the nonterminal it belongs to.
-freshVar :: NT -> Var -> Var
-freshVar nt (Var n t) = Var (n ++ "#" ++ show nt) t
-
 allFresh :: Data a => NT -> a -> a
-allFresh nt x = x & vars %~ freshVar nt
+allFresh nt x = x & vars %~ fresh
+  where
+    fresh (Var n t) = Var (n ++ "#" ++ show nt) t
 
 findSet :: Ord a => a -> Map a (Set b) -> Set b
 findSet = M.findWithDefault S.empty

### 2018-04-27 src/Shara/Grammar.hs

@@ -66,25 +66,6 @@ flat r = SGrammar r M.empty
 ruleFor :: NT -> SGrammar s a -> SRule s a
 ruleFor nt = M.findWithDefault R.Null nt . rules
 
-after :: NT -> SGrammar s a -> SGrammar s a
-after nt g =
-  let s = ruleFor nt g
-  in SGrammar s (fst $ execState (go s) (M.empty, S.empty))
-  where
-    go =
-      \case
-        SNonterm (_, nt') ->
-          elem nt' <$> use _2 >>= \case
-            True -> pure ()
-            False ->
-              let r = ruleFor nt' g
-              in do _1 %= M.insert nt' r
-                    _2 %= S.insert nt'
-                    go r
-        R.Alt x y -> go x >> go y
-        R.Seq x y -> go x >> go y
-        _ -> pure ()
-
 reverseRules :: Grammar a -> Map NT [(Maybe NT, Rule a)]
 reverseRules g = runReader (go $ start g) (Nothing, R.Eps)
   where
@@ -102,54 +83,6 @@ reverseRules g = runReader (go $ start g) (Nothing, R.Eps)
 grammarFromMap :: NT -> Map NT (SRule s a) -> SGrammar s a
 grammarFromMap st m =
   SGrammar (M.findWithDefault R.Null st m) (M.delete st m)
-
-before :: NT -> SGrammar s a -> SGrammar s a
-before nt g =
-  let (s, (rs, _)) = runState (go (start g)) (M.empty, S.empty)
-  in SGrammar s rs
-  where
-    go =
-      \case
-        SNonterm (vs, nt') -> do
-          when (nt /= nt') $
-            elem nt <$> use _2 >>= \case
-              True -> pure ()
-              False -> do
-                r' <- go (ruleFor nt' g)
-                _1 %= M.insert nt' r'
-                _2 %= S.insert nt'
-          pure (SNonterm (vs, nt'))
-        R.Alt x y -> R.alt <$> go x <*> go y
-        R.Seq x y -> R.seq <$> go x <*> go y
-        x -> pure x
-
-partition :: [NT] -> Grammar a -> (Grammar a, Grammar a)
-partition nts g = (bef, aft)
-  where
-    bef = evalState (go (start g)) (S.fromList nts)
-    aft =
-      evalState
-        (asum <$> mapM (\nt -> fmap (abstract () nt) $ go $ ruleFor nt g) nts)
-        (nonterminals bef)
-    go =
-      \case
[…]

### 2018-04-27 src/Shara/GrammarCut.hs

@@ -8,7 +8,7 @@ import qualified Data.Map            as M
 import           Data.Ratio
 import           Data.Set            (Set)
 import qualified Data.Set            as S
-import           Shara.Grammar       hiding (partition)
+import           Shara.Grammar
 import qualified Shara.MinCut        as MC
 import qualified Shara.Reg           as R
 
@@ -34,18 +34,8 @@ initials g = S.filter (\s -> all null (predecessors s g)) (symbols g)
 terminals :: HyperGraph -> Set Symbol
 terminals g = symbols g `S.difference` M.keysSet (graphForward g)
 
-clear :: Set Symbol -> HyperGraph -> HyperGraph
-clear toKeep (HyperGraph for back) = HyperGraph for' back'
-  where
-    for' =
-      M.filterWithKey (\k v -> k `elem` toKeep && not (null v)) $
-      fmap (filter (`elem` toKeep)) for
-    back' =
-      M.filterWithKey (\k _ -> k `elem` toKeep) $
-      fmap (map (filter (`elem` toKeep))) back
-
-grammarToGraph :: Grammar a -> HyperGraph
-grammarToGraph (SGrammar _ rs) = HyperGraph forw back
+grammarToGraph :: Set Symbol -> Grammar a -> HyperGraph
+grammarToGraph toKeep (SGrammar _ rs) = clear toKeep $ HyperGraph forw back
   where
     back = fmap ruleSymbols rs
     forw = foldr addForw M.empty (M.toList back)
@@ -69,6 +59,15 @@ grammarToGraph (SGrammar _ rs) = HyperGraph forw back
                       else [a' ++ b' | a' <- as, b' <- bs]
         Nonterm nt -> [[nt]]
         Term _ -> []
+    clear :: Set Symbol -> HyperGraph -> HyperGraph
+    clear toKeep (HyperGraph for back) = HyperGraph for' back'
+      where
+        for' =
+          M.filterWithKey (\k v -> k `elem` toKeep && not (null v)) $
+          fmap (filter (`elem` toKeep)) for
+        back' =
+          M.filterWithKey (\k _ -> k `elem` toKeep) $
+          fmap (map (filter (`elem` toKeep))) back
 
 cut :: Set Symbol -> HyperGraph -> (Set Symbol, Set Symbol, Set Symbol)
 cut uncuttable g =

### 2018-04-27 src/Shara/Inf.hs

@@ -3,7 +3,6 @@
 module Shara.Inf where
 
 import           Control.Monad
-import Shara.Semiring
 
 newtype Infinite m f a = Infinite
   { getInfinite :: f (Either a (m (Infinite m f a)))

### 2018-04-27 src/Shara/Interpolate.hs

@@ -48,7 +48,7 @@ loop strategy targets g rev m = do
   if null (targets S.\\ M.keysSet m)
     then pure (Right m)
     else do
-      let gr = clear targets $ grammarToGraph g
+      let gr = grammarToGraph targets g
       let (now', half1', half2') = cut (M.keysSet m) gr
       let (now, half1, half2) = if null (now' S.\\ M.keysSet m)
           then (targets S.\\ M.keysSet m, S.empty, S.empty)
@@ -117,9 +117,6 @@ interpolateNT solns g rg target =
 mkForm :: NT -> Map NT Expr -> Grammar Expr -> Expr
 mkForm st m g = evalState (ruleExpr m g (ruleFor st g)) S.empty
 
-mark :: NT -> Expr
-mark nt = V $ Var ("__b" ++ show nt) Bool
-
 mkRevForm ::
      NT -> Map NT Expr -> Grammar Expr -> Map NT [(Maybe NT, Rule Expr)] -> Expr
 mkRevForm st m g rg =
@@ -164,3 +161,6 @@ ruleExpr m g r = do
               Nothing -> go (ruleFor nt g)
               Just phi' -> pure (phi', [])
           pure (mark nt, (mkImpl (mark nt) phi) : is)
+
+mark :: NT -> Expr
+mark nt = V $ Var ("__b" ++ show nt) Bool

### 2018-04-27 app/Main.hs
started working on tree interpolation for subproblems

@@ -100,9 +100,9 @@ main = do
            if a == "topo"
              then Topological
              else if a == "par"
-                    then LicketySplit ConcurrentInterpolation
-                    else LicketySplit SequentialInterpolation
-         _ -> LicketySplit SequentialInterpolation)
+                    then LicketySplit (LicketySplitOptions True True)
+                    else LicketySplit (LicketySplitOptions False True)
+         _ -> LicketySplit (LicketySplitOptions False True))
   print sk
   shara sk test1M test1 >>= \case
     Left m -> print (fmap pretty m)

### 2018-04-27 src/Shara/Interpolate.hs

@@ -1,5 +1,7 @@
+{-# LANGUAGE TemplateHaskell #-}
 module Shara.Interpolate where
 
+import           Control.Lens hiding (Context)
 import           Control.Concurrent.Async (concurrently)
 import           Control.Monad.Except
 import           Control.Monad.State
@@ -14,21 +16,26 @@ import           Formula
 import qualified Formula.Z3               as Z3
 import           Shara.GrammarCut
 
-data InterpolationStrategy
-  = SequentialInterpolation
-  | ConcurrentInterpolation
-  deriving (Show, Read, Eq, Ord)
+data LicketySplitOptions = LicketySplitOptions
+  -- When lickety split divides the problem in two, should it continue on the
+  -- two subproblems concurrently?
+  { _concurrentInterpolation :: Bool
+  -- If lickety split finds that the current subproblem is treelike, should it
+  -- invoke the tree interpolator?
+  , _useTreeInterpolator :: Bool
+  } deriving (Show, Read, Eq, Ord)
+makeLenses ''LicketySplitOptions
 
 -- | `licketySplit` computes interpolants of a DAG. It attempts to do this
 -- maximally in parallel by cutting the DAG consideration at each iteration.
 licketySplit ::
      MonadIO m
-  => InterpolationStrategy
+  => LicketySplitOptions
   -> Grammar Expr
   -> m (Either Model (IntMap Expr))
-licketySplit strategy g =
+licketySplit opts g =
   let ss = nonterminals g
-  in liftIO (loop strategy ss g (contextualize g) M.empty)
+  in liftIO (loop opts ss g (contextualize g) M.empty)
 
 -- | The core loop for interpolating a DAG. The loop requires two versions of
 -- the DAG, one which is a full representation and another which is a
@@ -38,16 +45,19 @@ licketySplit strategy g =
 -- current iteration. The two subgraphs on either side of the cut are the new
 -- restrictions which can be interpolated in parallel.
 loop ::
-     InterpolationStrategy
+     LicketySplitOptions
   -> IntSet
   -> Grammar Expr
   -> Context Expr
   -> IntMap Expr
   -> IO (Either Model (IntMap Expr))
-loop strategy targets g rev m =
-  if S.null (targets S.\\ M.keysSet m)
-    then pure (Right m)
-    else do
+loop opts targets g rev m = do
+  if treelike (targets S.\\ M.keysSet m) rev
+  then treeSolve targets g rev m
[…]

### 2018-04-27 src/Shara/Shara.hs

@@ -17,7 +17,7 @@ import           Shara.Interpolate
 
 data SolveKind
   = Topological
-  | LicketySplit InterpolationStrategy
+  | LicketySplit LicketySplitOptions
   deriving (Show, Read, Eq, Ord)
 
 shara ::

### 2018-04-29 app/Main.hs
outsourced languages

@@ -1,59 +1,60 @@
 {-# LANGUAGE QuasiQuotes #-}
 
 import           Control.Monad.State
-import           Data.Map                  (Map)
-import qualified Data.Map                  as M
+import           Data.IntMap                 (IntMap)
+import qualified Data.IntMap                 as M
+import           Data.Language.Grammar
+import qualified Data.Language.Reg           as R
+import qualified Data.Language.ScopedGrammar as SG
 import           Data.Text.Prettyprint.Doc
-import           Formula                   hiding (res)
-import           Prelude                   hiding (abs, seq)
+import           Formula                     hiding (res)
+import           Prelude                     hiding (abs, seq)
 import           Shara.CDD
-import           Shara.Grammar
 import           Shara.Interpolate
-import qualified Shara.Reg                 as R
 import           Shara.Shara
-import           System.Environment        (getArgs)
+import           System.Environment          (getArgs)
 
-test1 :: SGrammar [Var] Expr
+test1 :: SG.Grammar [Var] Expr
 test1 =
-  SGrammar
-    (STerm [expr|x = 41|] `R.seq` SNonterm ([x], 0))
+  SG.Grammar
+    (SG.Term [expr|x = 41|] `R.seq` SG.Nonterm [x] 0)
     (M.fromList
        [ ( 0
          , R.alt
-             (STerm [expr|x = 0|])
-             (STerm [expr|x = x' + 2|] `R.seq` SNonterm ([x'], 0)))
+             (SG.Term [expr|x = 0|])
+             (SG.Term [expr|x = x' + 2|] `R.seq` SG.Nonterm [x'] 0))
        ])
   where
     x = Var "x" Int
     x' = Var "x'" Int
 
-test1M :: Map NT [Var]
+test1M :: IntMap [Var]
 test1M = M.fromList [(0, [Var "x" Int])]
 
-test2 :: SGrammar [Var] Expr
+test2 :: SG.Grammar [Var] Expr
 test2 =
-  SGrammar
-    (SNonterm ([n, res], m) `R.seq` STerm [expr|res < 0|])
+  SG.Grammar
+    (SG.Nonterm [n, res] m `R.seq` SG.Term [expr|res < 0|])
     (M.fromList
        [ ( m
          , mconcat
[…]

### 2018-04-29 src/Formula/Parser.hs

@@ -108,7 +108,7 @@ addop = (op "+" >> return (app2 $ Add T.Int))
     <|> (op "-" >> return (app2 $ Sub T.Int))
 
 integer :: (Read i, Integral i) => CharParser st i
-integer = lexeme $ do { ds <- many1 digit ; return (read ds) }
+integer = lexeme $ read <$> many1 digit
 
 parseChc :: CharParser st [Chc]
 parseChc = many parseChc'

### 2018-04-29 src/Formula/Vocab.hs

@@ -27,12 +27,12 @@ instance MonadReader s m => MonadReader s (VocabT m) where
 instance Monad m => MonadVocab (VocabT m) where
   fresh n' = do
     let n = baseName n'
-    Vocab (M.lookup n <$> get >>= \case
+    Vocab (gets (M.lookup n) >>= \case
       Nothing -> modify (M.insert n 0)
       Just i  -> modify (M.insert n (i+1)))
     fetch n
 
-  fetch n = Vocab (M.lookup (baseName n) <$> get >>= \case
+  fetch n = Vocab (gets (M.lookup (baseName n)) >>= \case
     Nothing -> pure (baseName n)
     Just i -> pure (baseName n ++ "#" ++ show i))
 

### 2018-04-29 src/Formula/Z3.hs

@@ -5,13 +5,13 @@ module Formula.Z3 where
 import           Control.Lens
 import           Control.Monad.State
 import           Control.Monad.Except
-import           Data.List (partition, nub)
+import           Data.List (partition)
 import           Data.Maybe
 import qualified Data.Map as M
 import           Data.Map (Map)
 import           Formula (Expr((:@)), Chc, Type((:=>)), Var(..))
 import qualified Formula as F
-import           Z3.Monad hiding (local)
+import           Z3.Monad
 
 data Env = Env { _envVars :: Map Var AST
                , _envFuns :: Map Var FuncDecl
@@ -225,9 +225,7 @@ formFromApp n args range
   | n == "true"  = return $ F.LBool True
   | n == "false" = return $ F.LBool False
   -- The 'app' is just a variable
-  | null args = do
-    typ <- sortToType range
-    return $ F.V $ F.Var n typ
+  | null args = F.V . F.Var n <$> sortToType range
   | n == "ite" || n == "if" = do
     c <- astToExpr (head args)
     e1 <- astToExpr (args !! 1)

### 2018-04-29 src/Shara/CDD.hs

@@ -1,60 +1,62 @@
 {-# LANGUAGE ScopedTypeVariables #-}
-{-# LANGUAGE TemplateHaskell #-}
+{-# LANGUAGE TemplateHaskell     #-}
 
 module Shara.CDD where
 
 import           Control.Lens
 import           Control.Monad.State
-import           Data.Data           (Data)
-import           Data.Map            (Map)
-import qualified Data.Map            as M
-import           Data.Set            (Set)
-import qualified Data.Set            as S
+import           Data.Data                   (Data)
+import           Data.IntMap                 (IntMap)
+import qualified Data.IntMap                 as M
+import           Data.IntSet                 (IntSet)
+import qualified Data.IntSet                 as S
+import           Data.Language.Grammar
+import           Data.Language.Inf
+import qualified Data.Language.Reg           as R
+import qualified Data.Language.ScopedGrammar as SG
 import           Formula
-import           Prelude             hiding (seq)
-import           Shara.Grammar
-import           Shara.Inf
-import qualified Shara.Reg           as R
-
-import Debug.Trace
+import           Prelude                     hiding (null, seq)
 
 type Branch = Int
+
 type Unrolling = Int
 
 data CDDState = CDDState
   -- Which nonterminals have been seen in the current unrolling?
-  { _visited :: Set NT
+  { _visited          :: IntSet
   -- Tracks the highest branch point thus far.
-  , _branchCount :: Branch
+  , _branchCount      :: Branch
   -- What is the identity of the current branch point?
-  , _currentBranch :: Branch
+  , _currentBranch    :: Branch
   -- Which nonterminals is each branch point directly dependent on?
-  , _branchDeps :: Map Branch (Set NT)
+  , _branchDeps       :: IntMap IntSet
   -- Which branch points are siblings of a given branch point?
-  , _branchSibs :: Map Branch (Set Branch)
+  , _branchSibs       :: IntMap IntSet
   -- Which branch points are parents of a given branch point?
-  , _branchPars :: Map Branch (Set Branch)
+  , _branchPars       :: IntMap IntSet
   -- Which branches are local to the given nonterminal?
-  , _ntBranches :: Map NT (Set Branch)
[…]

### 2018-04-29 src/Shara/Grammar.hs

@@ -1,150 +1,0 @@
-{-# LANGUAGE PatternSynonyms #-}
-
-module Shara.Grammar where
-
-import           Control.Arrow
-import           Control.Applicative
-import           Control.Lens
-import           Control.Monad.Reader
-import           Control.Monad.State
-import           Data.Data            (Data)
-import           Data.Foldable        (asum)
-import           Data.Map             (Map)
-import qualified Data.Map             as M
-import           Data.Semigroup
-import           Data.Set             (Set)
-import qualified Data.Set             as S
-import           Prelude              hiding (seq)
-import           Shara.Inf            (Infinite)
-import           Shara.Reg            (Reg)
-import qualified Shara.Reg            as R
-
-type NT = Int
-
-type SRule s a = Reg (Either (s, NT) a)
-
-type Rule a = Reg (Either ((), NT) a)
-
-pattern SNonterm :: a -> Reg (Either a b)
-pattern SNonterm x = R.Symbol (Left x)
-pattern STerm :: b -> Reg (Either a b)
-pattern STerm x = R.Symbol (Right x)
-pattern Nonterm :: a -> Reg (Either ((), a) b)
-pattern Nonterm x = R.Symbol (Left ((), x))
-pattern Term :: b -> Reg (Either ((), a) b)
-pattern Term x = R.Symbol (Right x)
-
-data SGrammar s a = SGrammar
-  { start :: SRule s a
-  , rules :: Map NT (SRule s a)
-  } deriving (Show, Read, Eq, Ord, Data, Functor, Foldable, Traversable)
-
-type Grammar a = SGrammar () a
-
-type ISGrammar m s a = Infinite m (SGrammar s) a
-
-type IGrammar m a = Infinite m (SGrammar ()) a
-
-seq :: SGrammar s a -> SGrammar s a -> SGrammar s a
-seq (SGrammar s rs) (SGrammar s' rs') =
-  SGrammar (s `R.seq` s') (M.unionWith R.alt rs rs')
-
-alt :: SGrammar s a -> SGrammar s a -> SGrammar s a
-alt (SGrammar s rs) (SGrammar s' rs') =
-  SGrammar (s `R.alt` s') (M.unionWith R.alt rs rs')
-
-abstract :: s -> NT -> SGrammar s a -> SGrammar s a
-abstract sc nt (SGrammar s rs) =
[…]

### 2018-04-29 src/Shara/GrammarCut.hs

@@ -1,48 +1,49 @@
 module Shara.GrammarCut where
 
-import           Control.Lens        hiding (below)
+import           Control.Lens
 import           Control.Monad.State
-import           Data.List           (partition)
-import           Data.Map            (Map)
-import qualified Data.Map            as M
+import           Data.IntMap           (IntMap)
+import qualified Data.IntMap           as IM
+import           Data.IntSet           (IntSet)
+import qualified Data.IntSet           as S
+import           Data.Language.Grammar hiding (null)
+import qualified Data.Language.Reg     as R
+import           Data.List             (partition)
+import           Data.Map              (Map)
+import qualified Data.Map              as M
 import           Data.Ratio
-import           Data.Set            (Set)
-import qualified Data.Set            as S
-import           Shara.Grammar
-import qualified Shara.MinCut        as MC
-import qualified Shara.Reg           as R
+import qualified Shara.MinCut          as MC
 
 type Symbol = Int
 
 data HyperGraph = HyperGraph
-  { graphForward  :: Map Symbol [Symbol]
-  , graphBackward :: Map Symbol [[Symbol]]
+  { graphForward  :: IntMap [Symbol]
+  , graphBackward :: IntMap [[Symbol]]
   } deriving (Show, Read, Eq, Ord)
 
 successors :: Symbol -> HyperGraph -> [Symbol]
-successors s = M.findWithDefault [] s . graphForward
+successors s = IM.findWithDefault [] s . graphForward
 
 predecessors :: Symbol -> HyperGraph -> [[Symbol]]
-predecessors s = M.findWithDefault [] s . graphBackward
+predecessors s = IM.findWithDefault [] s . graphBackward
 
-symbols :: HyperGraph -> Set Symbol
-symbols g = M.keysSet (graphForward g) `S.union` M.keysSet (graphBackward g)
+symbols :: HyperGraph -> IntSet
+symbols g = IM.keysSet (graphForward g) `S.union` IM.keysSet (graphBackward g)
 
-initials :: HyperGraph -> Set Symbol
+initials :: HyperGraph -> IntSet
 initials g = S.filter (\s -> all null (predecessors s g)) (symbols g)
 
-terminals :: HyperGraph -> Set Symbol
[…]

### 2018-04-29 src/Shara/Inf.hs

@@ -1,43 +1,0 @@
-{-# LANGUAGE PatternSynonyms #-}
-
-module Shara.Inf where
-
-import           Control.Monad
-
-newtype Infinite m f a = Infinite
-  { getInfinite :: f (Either a (m (Infinite m f a)))
-  }
-
-pattern Fin :: a -> Either a b
-pattern Fin a = Left a
-
-pattern Inf :: b -> Either a b
-pattern Inf a = Right a
-
--- | Collect the current finite prefix of the structure by replacing infinite
--- elements by the default structure.
-finite :: Monad f => f a -> Infinite m f a -> f a
-finite def =
-  join .
-  fmap
-    (\case
-       Fin x -> pure x
-       Inf _ -> def) .
-  getInfinite
-
-unroll ::
-     (Applicative m, Monad f, Traversable f)
-  => Infinite m f a
-  -> m (Infinite m f a)
-unroll =
-  fmap (Infinite . join) .
-  traverse
-    (\case
-       Fin x -> pure (pure (Fin x))
-       Inf ac -> getInfinite <$> ac) .
-  getInfinite
-
-type IList m a = Infinite m [] a
-
-singleton :: (Functor m, Applicative f) => m a -> m (Infinite m f a)
-singleton x = Infinite . pure . Fin <$> x

### 2018-04-29 src/Shara/Interpolate.hs

@@ -3,16 +3,16 @@ module Shara.Interpolate where
 import           Control.Concurrent.Async (concurrently)
 import           Control.Monad.Except
 import           Control.Monad.State
-import           Data.Map                 (Map)
-import qualified Data.Map                 as M
+import           Data.IntMap              (IntMap)
+import qualified Data.IntMap              as M
+import           Data.IntSet              (IntSet)
+import qualified Data.IntSet              as S
+import           Data.Language.Grammar
+import qualified Data.Language.Reg        as R
 import           Data.Semigroup
-import           Data.Set (Set)
-import qualified Data.Set                 as S
 import           Formula
 import qualified Formula.Z3               as Z3
-import           Shara.Grammar
 import           Shara.GrammarCut
-import qualified Shara.Reg                as R
 
 data InterpolationStrategy
   = SequentialInterpolation
@@ -25,10 +25,10 @@ licketySplit ::
      MonadIO m
   => InterpolationStrategy
   -> Grammar Expr
-  -> m (Either Model (Map NT Expr))
+  -> m (Either Model (IntMap Expr))
 licketySplit strategy g =
-  let ss = nonterminals g in
-      liftIO (loop strategy ss g (reverseRules g) M.empty)
+  let ss = nonterminals g
+  in liftIO (loop strategy ss g (contextualize g) M.empty)
 
 -- | The core loop for interpolating a DAG. The loop requires two versions of
 -- the DAG, one which is a full representation and another which is a
@@ -39,21 +39,25 @@ licketySplit strategy g =
 -- restrictions which can be interpolated in parallel.
 loop ::
      InterpolationStrategy
-  -> Set NT
+  -> IntSet
   -> Grammar Expr
-  -> Map NT [(Maybe NT, Rule Expr)]
-  -> Map NT Expr
-  -> IO (Either Model (Map NT Expr))
-loop strategy targets g rev m = do
-  if null (targets S.\\ M.keysSet m)
+  -> Context Expr
+  -> IntMap Expr
+  -> IO (Either Model (IntMap Expr))
+loop strategy targets g rev m =
+  if S.null (targets S.\\ M.keysSet m)
     then pure (Right m)
[…]

### 2018-04-29 src/Shara/MinCut.hs

@@ -1,106 +1,120 @@
 {-# LANGUAGE TemplateHaskell #-}
+
 module Shara.MinCut where
 
 import           Control.Lens
-import           Control.Monad.State
+import           Control.Monad.Extra  (concatMapM)
 import           Control.Monad.Reader
-import           Control.Monad.Extra (concatMapM)
-import           Data.Map (Map)
-import qualified Data.Map as M
-import           Data.Set (Set)
-import qualified Data.Set as S
+import           Control.Monad.State
+import           Data.IntMap          (IntMap)
+import qualified Data.IntMap          as IM
+import           Data.IntSet          (IntSet)
+import qualified Data.IntSet          as IS
+import           Data.Map             (Map)
+import qualified Data.Map             as M
+import           Data.Set             (Set)
+import qualified Data.Set             as S
 
 type Symbol = Int
 
 data Graph = Graph
-  { _forward   :: Map Symbol (Set Symbol)
-  , _backward  :: Map Symbol (Set Symbol)
-  , _initials  :: Set Symbol
-  , _terminals :: Set Symbol
+  { _forward   :: IntMap IntSet
+  , _backward  :: IntMap IntSet
+  , _initials  :: IntSet
+  , _terminals :: IntSet
   } deriving (Show, Read, Eq, Ord)
+
 makeLenses ''Graph
 
 addEdge :: Symbol -> Symbol -> Graph -> Graph
 addEdge s t g =
-  g & forward  %~ M.insertWith S.union s (S.singleton t)
-    & backward %~ M.insertWith S.union t (S.singleton s)
+  g & forward %~ IM.insertWith IS.union s (IS.singleton t) & backward %~
+  IM.insertWith IS.union t (IS.singleton s)
 
 deleteEdge :: Symbol -> Symbol -> Graph -> Graph
 deleteEdge s t g =
-  g & forward %~ M.adjust (S.delete t) s
-    & backward %~ M.adjust (S.delete s) t
+  g & forward %~ IM.adjust (IS.delete t) s & backward %~
+  IM.adjust (IS.delete s) t
 
 addInitial :: Symbol -> Graph -> Graph
-addInitial s = initials %~ S.insert s
+addInitial s = initials %~ IS.insert s
 
 addTerminal :: Symbol -> Graph -> Graph
-addTerminal s = terminals %~ S.insert s
+addTerminal s = terminals %~ IS.insert s
 
 empty :: Graph
[…]

### 2018-04-29 src/Shara/Reg.hs

@@ -1,89 +1,0 @@
-module Shara.Reg where
-
-import           Control.Applicative
-import           Data.Data           (Data)
-import           Data.Semigroup
-import           Prelude             hiding (seq)
-import           Shara.Inf
-import           Shara.Semiring
-import           Test.QuickCheck
-
--- | A finite (no Kleene star) regular expression that supports scoping.
--- Scoping supports operations such as a change of vocabulary in the expression
--- terms.
-data Reg a
-  = Seq (Reg a)
-        (Reg a) -- Sequence two expressions one after the other.
-  | Alt (Reg a)
-        (Reg a) -- Alternation: Choose one or the other.
-  | Eps -- The expression that matches the empty string.
-  | Null -- The expression that matches nothing.
-  | Symbol a -- The expression that matches exactly the given symbol.
-  deriving (Show, Read, Eq, Ord, Data, Foldable, Traversable)
-
-type IReg m a = Infinite m Reg a
-
--- | Associative sequencing where Null is the a 0 and Eps is a unit.
-seq :: Reg a -> Reg a -> Reg a
-seq Null _ = Null
-seq _ Null = Null
-seq Eps x = x
-seq x Eps = x
-seq (Seq x y) z = Seq x (seq y z)
-seq x y = Seq x y
-
--- | Associative alternation where Null is a unit and Eps is not duplicated.
-alt :: Reg a -> Reg a -> Reg a
-alt Eps Eps = Eps
-alt Eps x = alt x Eps
-alt Null x = x
-alt x Null = x
-alt (Alt x y) z = Alt x (alt y z)
-alt x y = Alt x y
-
-instance Functor Reg where
-  fmap f =
-    \case
-      Seq a b -> seq (fmap f a) (fmap f b)
-      Alt a b -> alt (fmap f a) (fmap f b)
-      Null -> Null
-      Eps -> Eps
-      Symbol a -> Symbol (f a)
-
--- | Finite regular expressions form a monad where each expression is inlined.
-instance Monad Reg where
-  x >>= f = join' (fmap f x)
-    where
-      join' =
-        \case
-          Seq a b -> seq (join' a) (join' b)
-          Alt a b -> alt (join' a) (join' b)
-          Eps -> Eps
-          Null -> Null
-          Symbol a -> a
-
--- | Since they form a Monad, regular expressions form an applicative in a
--- straightforward manner.
[…]

### 2018-04-29 src/Shara/Semiring.hs

@@ -1,13 +1,0 @@
-module Shara.Semiring where
-
-class Semiring a where
-  (<.>) :: a -> a -> a
-  (<+>) :: a -> a -> a
-  zero :: a
-  one :: a
-
-instance (Semiring a, Semiring b) => Semiring (a, b) where
-  (a, b) <.> (a', b') = (a <.> a', b <.> b')
-  (a, b) <+> (a', b') = (a <+> a', b <+> b')
-  zero = (zero, zero)
-  one = (one, one)

### 2018-04-29 src/Shara/Shara.hs

@@ -2,17 +2,18 @@ module Shara.Shara where
 
 import           Control.Lens
 import           Control.Monad.State
-import           Data.List.Split     (splitOn)
-import           Data.Map            (Map)
-import qualified Data.Map            as M
-import           Data.Set            (Set)
-import qualified Data.Set            as S
+import           Data.IntMap                 (IntMap)
+import qualified Data.IntMap                 as M
+import           Data.IntSet                 (IntSet)
+import qualified Data.IntSet                 as S
+import           Data.Language.Grammar
+import qualified Data.Language.Reg           as R
+import qualified Data.Language.ScopedGrammar as SG
+import           Data.List.Split             (splitOn)
 import           Formula
-import qualified Formula.Z3          as Z3
+import qualified Formula.Z3                  as Z3
 import           Shara.CDD
-import           Shara.Grammar
 import           Shara.Interpolate
-import qualified Shara.Reg           as R
 
 data SolveKind
   = Topological
@@ -22,9 +23,9 @@ data SolveKind
 shara ::
      MonadIO m
   => SolveKind
-  -> Map NT [Var]
-  -> SGrammar [Var] Expr
-  -> m (Either Model (Map NT Expr))
+  -> IntMap [Var]
+  -> SG.Grammar [Var] Expr
+  -> m (Either Model (IntMap Expr))
 shara sk vocab sg = evalStateT (go =<< cdd sg) (emptyCDDState vocab)
   where
     go g =
@@ -35,10 +36,11 @@ shara sk vocab sg = evalStateT (go =<< cdd sg) (emptyCDDState vocab)
           let sol = collapse cs m
           inductive vocab sol sg >>= \case
             Nothing -> pure (Right sol)
-            Just inds -> go =<< unrollCDD g {- inds -}
+            Just _ -> go =<< unrollCDD g {- inds -}
+             {- inds -}
 
 solveDirect ::
-     MonadIO m => SolveKind -> Grammar Expr -> m (Either Model (Map NT Expr))
+     MonadIO m => SolveKind -> Grammar Expr -> m (Either Model (IntMap Expr))
 solveDirect =
   \case
     Topological -> topologicalInterpolation
@@ -46,18 +48,18 @@ solveDirect =
 
 inductive ::
      MonadIO m
[…]

### 2018-04-30 src/Formula/Expr.hs
outsourced formula to a library

@@ -1,287 +1,0 @@
-{-# LANGUAGE TemplateHaskell #-}
-module Formula.Expr where
-
-import           Control.Lens
-import           Control.Monad.State
-
-import           Data.Set (Set)
-import           Data.Map (Map)
-import qualified Data.Map as M
-import           Data.Data (Data)
-import           Data.Data.Lens (uniplate)
-import           Data.List (sort)
-import           Data.Text.Prettyprint.Doc
-
-import           Formula.Type (Type((:=>)))
-import qualified Formula.Type as T
-import           Formula.Var
-
-data Expr
-  = Expr :@ Expr
-  | V Var
-
-  | If Type
-
-  | Not
-  | Impl
-  | Iff
-  | And
-  | Or
-
-  | Add Type
-  | Mul Type
-  | Sub Type
-  | Div Type
-  | Mod Type
-
-  | Distinct Type
-  | Eql Type
-  | Nql Type
-  | Lt  Type
-  | Le  Type
-  | Gt  Type
-  | Ge  Type
-
-  | LUnit
-  | LBool Bool
-  | LInt Integer
-  deriving (Show, Read, Eq, Ord, Data)
-
-infixl 9 :@
-
-makePrisms ''Expr
-
-instance Plated Expr where plate = uniplate
-
-instance Monoid Expr where
-  mappend = mkAnd
-  mempty = LBool True
-
-exprType :: Expr -> Type
-exprType = \case
-  V v         -> v ^. varType
-  o :@ _      -> case exprType o of
-                   _ :=> t -> t
-                   _ -> error "bad function application type"
-
-  If t        -> T.Bool :=> t :=> t :=> t
-
-  Not         -> T.Bool :=> T.Bool
-  Impl        -> T.Bool :=> T.Bool
-  Iff         -> T.Bool :=> T.Bool
-  And         -> T.Bool :=> T.Bool
-  Or          -> T.Bool :=> T.Bool
-
-  Add t       -> t :=> t :=> t
-  Mul t       -> t :=> t :=> t
-  Sub t       -> t :=> t :=> t
-  Div t       -> t :=> t :=> t
-  Mod t       -> t :=> t :=> t
-
-  Distinct t  -> T.List t :=> T.Bool
-  Eql t       -> t :=> t :=> T.Bool
-  Nql t       -> t :=> t :=> T.Bool
[…]

### 2018-04-30 src/Formula/Parser.hs

@@ -1,249 +1,0 @@
-{-# LANGUAGE TemplateHaskell #-}
-module Formula.Parser where
-
-import           Control.Lens hiding (op)
-
-import           Data.Data (Data)
-import           Data.Generics (extQ)
-
-import           Text.Parsec
-import           Text.ParserCombinators.Parsec.Char
-import qualified Text.Parsec.Token as T
-import           Text.Parsec.Language (emptyDef)
-
-import qualified Language.Haskell.TH as TH
-import           Language.Haskell.TH.Quote
-
-import           Formula.Type (Type)
-import qualified Formula.Type as T
-import           Formula.Expr
-import           Formula.Var
-import           Formula.Chc
-
-lexeme :: Stream s m Char => ParsecT s u m b -> ParsecT s u m b
-lexeme p = do { x <- p; spaces; return x  }
-
-parseExpr :: CharParser st Expr
-parseExpr = ex1
-
-ex1 :: CharParser st Expr
-ex1 = ex2 `chainl1` impop
-
-ex2 :: CharParser st Expr
-ex2 = ex3 `chainl1` (op "||" >> return (\x y -> manyOr [x, y]))
-
-ex3 :: CharParser st Expr
-ex3 = ex4 `chainl1` (op "&&" >> return (\x y -> manyAnd [x, y]))
-
-ex4 :: CharParser st Expr
-ex4 = ex6 `chainl1` compareop
-
-ex6 :: CharParser st Expr
-ex6 = ex7 `chainl1` addop
-
-ex7 :: CharParser st Expr
-ex7 = exInf `chainl1` mulop
-
-exInf :: CharParser st Expr
-exInf = atom
-
-atom :: CharParser st Expr
-atom = try app <|> nonapp
-  where
-    nonapp = parens parseExpr <|> (V <$> try var) <|> bool <|> (LInt <$> integer)
-    app = (res "not"      >> (:@) Not                    <$> nonapp)
-      <|> (res "if"       >> appMany (If T.Int)          <$> sequence [nonapp, nonapp, nonapp])
-      <|> (res "distinct" >> appMany (Distinct T.Int)    <$> many1 nonapp)
-      <|> (res "and"      >> manyAnd                     <$> many1 nonapp)
-      <|> (res "or"       >> manyOr                      <$> many1 nonapp)
-      <|> (res "add"      >> appMany (Add T.Int)         <$> many1 nonapp)
[…]

### 2018-04-30 src/Formula/Var.hs

@@ -1,47 +1,0 @@
-{-# LANGUAGE TemplateHaskell #-}
-module Formula.Var where
-
-import           Control.Lens
-
-import           Data.Data (Data)
-import           Data.Data.Lens (biplate)
-import           Data.Map (Map)
-import qualified Data.Map as M
-import           Data.Set (Set)
-import qualified Data.Set as S
-import           Data.Text.Prettyprint.Doc
-
-import           Formula.Type
-
-data Var = Var
-  { _varName :: String
-  , _varType :: Type
-  } deriving (Show, Read, Eq, Ord, Data)
-makeLenses ''Var
-
-instance Pretty Var
-  where pretty (Var n _) = pretty n
-
--- | A traversal which targets all of the variables in a given expression.
-vars :: Data a => Traversal' a Var
-vars = biplate
-
--- | Perform substitution over the variables in the expression. If a given
--- variable does not appear in the mapping, it is left untouched.
-subst :: Data a => Map Var Var -> a -> a
-subst m = over vars (\v -> M.findWithDefault v v m)
-
--- | The set of all variables in the expression.
-varSet :: Data a => a -> Set Var
-varSet x = S.fromList (x ^.. vars)
-
--- | Replace bound variables in the structure by those in the list.
-instantiate :: Data a => [Var] -> a -> a
-instantiate vs =
-  let ts = map (view varType) vs
-      bs = zipWith bound [0..] ts
-      m = M.fromList (zip bs vs)
-  in subst m
-
-bound :: Integer -> Type -> Var
-bound i = Var ("!" ++ show i)

### 2018-04-30 src/Formula/Vocab.hs

@@ -1,56 +1,0 @@
-{-# LANGUAGE MultiParamTypeClasses, UndecidableInstances #-}
-module Formula.Vocab where
-
-import           Control.Lens
-import           Control.Monad.State
-import           Control.Monad.Reader
-
-import qualified Data.Map as M
-import           Data.Map (Map)
-import           Data.List.Split
-
-class Monad m => MonadVocab m where
-  fresh :: String -> m String
-  fetch :: String -> m String
-
-newtype VocabT m a = Vocab { getVocab :: StateT (Map String Int) m a }
-  deriving (Functor, Applicative, Monad, MonadIO, MonadTrans)
-
-instance MonadState s m => MonadState s (VocabT m) where
-  state = lift . state
-
-instance MonadReader s m => MonadReader s (VocabT m) where
-  ask = lift ask
-  local = mapVocabT . local
-    where mapVocabT f = Vocab . mapStateT f . getVocab
-
-instance Monad m => MonadVocab (VocabT m) where
-  fresh n' = do
-    let n = baseName n'
-    Vocab (gets (M.lookup n) >>= \case
-      Nothing -> modify (M.insert n 0)
-      Just i  -> modify (M.insert n (i+1)))
-    fetch n
-
-  fetch n = Vocab (gets (M.lookup (baseName n)) >>= \case
-    Nothing -> pure (baseName n)
-    Just i -> pure (baseName n ++ "#" ++ show i))
-
-instance MonadVocab m => MonadVocab (StateT s m) where
-  fresh = lift . fresh
-  fetch = lift . fetch
-
-instance MonadVocab m => MonadVocab (ReaderT s m) where
-  fresh = lift . fresh
-  fetch = lift . fetch
-
-type Vocab a = VocabT Identity a
-
-runVocab :: Vocab a -> a
-runVocab (Vocab a) = evalState a M.empty
-
-runVocabT :: Monad m => VocabT m a -> m a
-runVocabT (Vocab a) = evalStateT a M.empty
-
-baseName :: String -> String
-baseName = head . splitOn "#"

### 2018-05-18 app/Main.hs
finished tree interpolator

@@ -104,6 +104,6 @@ main = do
                     else LicketySplit (LicketySplitOptions False True)
          _ -> LicketySplit (LicketySplitOptions False True))
   print sk
-  shara sk test2M test2 >>= \case
+  shara sk test1M test1 >>= \case
     Left m -> print (fmap pretty m)
     Right m -> print (fmap pretty m)

### 2018-05-18 src/Shara/Interpolate.hs

@@ -56,22 +56,17 @@ loop ::
   -> IO (Either Model (IntMap Expr))
 loop opts targets g rev m =
   if
-    | S.null (targets S.\\ M.keysSet m) -> pure (Right m)
-    | treelike (targets S.\\ M.keysSet m) rev -> do
-      putStrLn "treelike"
-      -- mapM print (M.toList $ getContext rev)
-      print targets
+    | S.null (targets S.\\ M.keysSet m) ->
+      pure (Right m)
+    | treelike (targets S.\\ M.keysSet m) rev ->
       treeSolve targets g rev m
     | otherwise -> do
-      putStrLn "not treelike"
-      print targets
       let gr = grammarToGraph targets g
       let (now', half1', half2') = cut (M.keysSet m) gr
       let (now, half1, half2) =
             if S.null (now' S.\\ M.keysSet m)
               then (targets S.\\ M.keysSet m, S.empty, S.empty)
               else (now', half1', half2')
-      print now
       runStateT
         (runExceptT $
          mapM_ (interpolateNT' g rev) (S.toList $ now S.\\ M.keysSet m))

### 2018-05-22 app/Main.hs
added new interpolation method

@@ -3,11 +3,13 @@
 import           Control.Monad.State
 import           Data.IntMap                 (IntMap)
 import qualified Data.IntMap                 as M
+import           Data.Language.Inf
 import qualified Data.Language.Reg           as R
 import qualified Data.Language.ScopedGrammar as SG
 import           Data.Text.Prettyprint.Doc
 import           Formula                     hiding (res)
 import           Prelude                     hiding (abs)
+import           Shara.CDD
 import           Shara.Interpolate
 import           Shara.Shara
 import           System.Environment          (getArgs)
@@ -15,7 +17,7 @@ import           System.Environment          (getArgs)
 test1 :: SG.Grammar [Var] Expr
 test1 =
   SG.Grammar
-    (SG.Term [expr|x = 41|] `R.seq` SG.Nonterm [x] 0)
+    (SG.Term [expr|x = 5|] `R.seq` SG.Nonterm [x] 0)
     (M.fromList
        [ ( 0
          , R.alt
@@ -104,6 +106,6 @@ main = do
                     else LicketySplit (LicketySplitOptions False True)
          _ -> LicketySplit (LicketySplitOptions False True))
   print sk
-  shara sk test1M test1 >>= \case
+  shara sk test2M test2 >>= \case
     Left m -> print (fmap pretty m)
     Right m -> print (fmap pretty m)

### 2018-05-22 src/Shara/Interpolate.hs

@@ -17,8 +17,6 @@ import           Formula
 import qualified Formula.Z3               as Z3
 import           Shara.GrammarCut
 
-import Data.Text.Prettyprint.Doc
-
 data LicketySplitOptions = LicketySplitOptions
   -- When lickety split divides the problem in two, should it continue on the
   -- two subproblems concurrently?

### 2018-05-22 src/Shara/Plan.hs

@@ -1,0 +1,115 @@
+module Shara.Plan where
+
+import           Control.Lens hiding (Context)
+import           Control.Monad.Except
+import           Control.Monad.State
+import           Data.IntMap               (IntMap)
+import qualified Data.IntMap               as M
+import           Data.IntSet               (IntSet)
+import qualified Data.IntSet               as S
+import           Data.Language.Grammar
+import qualified Data.Language.Reg         as R
+import           Data.List.Split
+import           Data.Tree                 as Tree
+import           Formula
+import qualified Formula.Z3                as Z3
+import           Shara.Interpolate
+
+data Task
+  = Join NT
+  | TreeLike (Tree NT)
+  deriving (Show, Read, Eq)
+
+interp :: Grammar Expr -> IO (Either Model (IntMap Expr))
+interp g = do
+  (a, s) <- runStateT (runExceptT (execute g (contextualize g) (plan g))) M.empty
+  case a of
+    Left m -> pure (Left m)
+    Right _ -> pure (Right s)
+
+plan :: Grammar a -> [Task]
+plan g = tree (start g) (-1)
+  where
+    js = joinPoints g
+    tree r nt =
+      let (t, ijs) = runState (go r nt) S.empty
+          rest = concatMap (\n -> tree (ruleFor n g) n) (S.toList ijs)
+      in map Join (S.toList ijs) ++ [TreeLike t] ++ rest
+    go r nt = do
+      let (localJoins, others) =
+            S.partition (`S.member` js) (ruleNonterminals r)
+      modify (S.union localJoins)
+      Tree.Node nt <$> mapM (\n -> go (ruleFor n g) n) (S.toList others)
+
+execute ::
+     Grammar Expr
+  -> Context Expr
+  -> [Task]
+  -> ExceptT Model (StateT (IntMap Expr) IO) ()
+execute g rev = \case
+  [] -> pure ()
+  (Join nt:rest) -> interpolateNT' g rev nt >> execute g rev rest
+  (TreeLike t:rest) -> interpolateTree g rev t >> execute g rev rest
+
+interpolateTree ::
+     Grammar Expr
+  -> Context Expr
+  -> Tree NT
+  -> ExceptT Model (StateT (IntMap Expr) IO) ()
+interpolateTree g rev ntTree = do
+  let root = Tree.rootLabel ntTree
+  m <- get
+  let context = case root of
[…]

### 2018-05-22 src/Shara/Shara.hs

@@ -14,6 +14,7 @@ import           Formula
 import qualified Formula.Z3                  as Z3
 import           Shara.CDD
 import           Shara.Interpolate
+import           Shara.Plan
 
 data SolveKind
   = Topological
@@ -54,7 +55,8 @@ sharaStep ::
   -> IGrammar m' Expr
   -> m (Either Model (IntMap Expr))
 sharaStep sk g =
-  solveDirect sk (finitePrefix g) >>= \case
+  liftIO (interp (finitePrefix g)) >>= \case
+  -- solveDirect sk (finitePrefix g) >>= \case
     Left m -> pure (Left m)
     Right m -> do
       cs <- use clones

### 2018-05-22 app/Main.hs
removed lots of cruft

@@ -92,20 +92,6 @@ abs' = Var "abs'" Int
 res = Var "res" Int
 
 main :: IO ()
-main = do
-  sk <-
-    getArgs >>=
-    (\as ->
-       pure $
-       case as of
-         [a] ->
-           if a == "topo"
-             then Topological
-             else if a == "par"
-                    then LicketySplit (LicketySplitOptions True True)
-                    else LicketySplit (LicketySplitOptions False True)
-         _ -> LicketySplit (LicketySplitOptions False True))
-  print sk
-  shara sk test2M test2 >>= \case
-    Left m -> print (fmap pretty m)
-    Right m -> print (fmap pretty m)
+main = shara test2M test2 >>= \case
+  Left m -> print (fmap pretty m)
+  Right m -> print (fmap pretty m)

### 2018-05-22 src/Shara/Chc.hs

@@ -21,14 +21,14 @@ import           Shara.Shara
 
 chcStrategy ::
      Monad m
-  => (SolveKind -> IntMap [Var] -> SG.Grammar [Var] Expr -> m (Either a (IntMap Expr)))
+  => (IntMap [Var] -> SG.Grammar [Var] Expr -> m (Either a (IntMap Expr)))
   -> [Chc]
   -> m (Either a (Map Var Expr))
 chcStrategy solver hcs =
   let (relToNt, ntToRel) = mapRels hcs
       (vs, mapBack) = varMapping relToNt hcs
       g = grammar relToNt vs hcs
-   in solver (LicketySplit $ LicketySplitOptions False True) vs g >>= \case
+   in solver vs g >>= \case
        Left m -> pure (Left m)
        Right m -> pure (Right (transcribe mapBack ntToRel m))
 

### 2018-05-22 src/Shara/GrammarCut.hs

@@ -1,166 +1,0 @@
-module Shara.GrammarCut where
-
-import           Control.Lens          hiding (below)
-import           Control.Monad.State
-import           Data.IntMap           (IntMap)
-import qualified Data.IntMap           as IM
-import           Data.IntSet           (IntSet)
-import qualified Data.IntSet           as S
-import           Data.Language.Grammar hiding (null)
-import qualified Data.Language.Reg     as R
-import           Data.List             (partition)
-import           Data.Map              (Map)
-import qualified Data.Map              as M
-import           Data.Ratio
-import qualified Shara.MinCut          as MC
-
-type Symbol = Int
-
-data HyperGraph = HyperGraph
-  { graphForward  :: IntMap [Symbol]
-  , graphBackward :: IntMap [[Symbol]]
-  } deriving (Show, Read, Eq, Ord)
-
-successors :: Symbol -> HyperGraph -> [Symbol]
-successors s = IM.findWithDefault [] s . graphForward
-
-predecessors :: Symbol -> HyperGraph -> [[Symbol]]
-predecessors s = IM.findWithDefault [] s . graphBackward
-
-symbols :: HyperGraph -> IntSet
-symbols g = IM.keysSet (graphForward g) `S.union` IM.keysSet (graphBackward g)
-
-initials :: HyperGraph -> IntSet
-initials g = S.filter (\s -> all null (predecessors s g)) (symbols g)
-
-terminals :: HyperGraph -> IntSet
-terminals g = symbols g `S.difference` IM.keysSet (graphForward g)
-
-grammarToGraph :: IntSet -> Grammar a -> HyperGraph
-grammarToGraph toKeep (Grammar _ rs) = clear $ HyperGraph forw' back'
-  where
-    back' = fmap ruleSymbols rs
-    forw' = foldr addForw IM.empty (IM.toList back')
-    addForw :: (Symbol, [[Symbol]]) -> IntMap [Symbol] -> IntMap [Symbol]
-    addForw (snk, srcs) m =
-      foldr (\src -> IM.insertWith (++) src [snk]) m (concat srcs)
-    ruleSymbols :: Rule a -> [[Symbol]]
-    ruleSymbols =
-      \case
-        R.Null -> []
-        R.Eps -> []
-        R.Alt a b -> ruleSymbols a ++ ruleSymbols b
-        R.Seq a b ->
-          let as = ruleSymbols a
-              bs = ruleSymbols b
-          in if null as
-               then bs
[…]

### 2018-05-22 src/Shara/Interpolate.hs

@@ -5,6 +5,7 @@ import           Control.Lens hiding (Context)
 import           Control.Concurrent.Async (concurrently)
 import           Control.Monad.Except
 import           Control.Monad.State
+import           Control.Monad.Reader
 import           Data.IntMap              (IntMap)
 import qualified Data.IntMap              as M
 import           Data.IntSet              (IntSet)
@@ -13,186 +14,121 @@ import           Data.Language.Grammar
 import qualified Data.Language.Reg        as R
 import           Data.Semigroup
 import           Data.Tree as Tree
+import           Data.List.Split (splitOn)
 import           Formula
 import qualified Formula.Z3               as Z3
-import           Shara.GrammarCut
 
-data LicketySplitOptions = LicketySplitOptions
-  -- When lickety split divides the problem in two, should it continue on the
-  -- two subproblems concurrently?
-  { _concurrentInterpolation :: Bool
-  -- If lickety split finds that the current subproblem is treelike, should it
-  -- invoke the tree interpolator?
-  , _useTreeInterpolator :: Bool
-  } deriving (Show, Read, Eq, Ord)
-makeLenses ''LicketySplitOptions
+type Interpolate a =
+  ReaderT (Grammar Expr, Context Expr) (ExceptT Model (StateT (IntMap Expr) IO)) a
 
--- | `licketySplit` computes interpolants of a DAG. It attempts to do this
--- maximally in parallel by cutting the DAG consideration at each iteration.
-licketySplit ::
-     MonadIO m
-  => LicketySplitOptions
-  -> Grammar Expr
-  -> m (Either Model (IntMap Expr))
-licketySplit opts g =
-  let ss = nonterminals g
-  in liftIO (loop opts ss g (contextualize g) M.empty)
+interpolate :: Grammar Expr -> IO (Either Model (IntMap Expr))
+interpolate g = do
+  (a, s) <-
+    runStateT (
+      runExceptT (
+        runReaderT execute (g, contextualize g)
+   )) M.empty
+  case a of
+    Left m -> pure (Left m)
+    Right _ -> pure (Right s)
 
--- | The core loop for interpolating a DAG. The loop requires two versions of
--- the DAG, one which is a full representation and another which is a
[…]

### 2018-05-22 src/Shara/MinCut.hs

@@ -1,210 +1,0 @@
-{-# LANGUAGE TemplateHaskell #-}
-
-module Shara.MinCut where
-
-import           Control.Lens
-import           Control.Monad.Extra  (concatMapM)
-import           Control.Monad.Reader
-import           Control.Monad.State
-import           Data.IntMap          (IntMap)
-import qualified Data.IntMap          as IM
-import           Data.IntSet          (IntSet)
-import qualified Data.IntSet          as IS
-import           Data.Map             (Map)
-import qualified Data.Map             as M
-import           Data.Set             (Set)
-import qualified Data.Set             as S
-
-type Symbol = Int
-
-data Graph = Graph
-  { _forward   :: IntMap IntSet
-  , _backward  :: IntMap IntSet
-  , _initials  :: IntSet
-  , _terminals :: IntSet
-  } deriving (Show, Read, Eq, Ord)
-
-makeLenses ''Graph
-
-addEdge :: Symbol -> Symbol -> Graph -> Graph
-addEdge s t g =
-  g & forward %~ IM.insertWith IS.union s (IS.singleton t) & backward %~
-  IM.insertWith IS.union t (IS.singleton s)
-
-deleteEdge :: Symbol -> Symbol -> Graph -> Graph
-deleteEdge s t g =
-  g & forward %~ IM.adjust (IS.delete t) s & backward %~
-  IM.adjust (IS.delete s) t
-
-addInitial :: Symbol -> Graph -> Graph
-addInitial s = initials %~ IS.insert s
-
-addTerminal :: Symbol -> Graph -> Graph
-addTerminal s = terminals %~ IS.insert s
-
-empty :: Graph
-empty = Graph IM.empty IM.empty IS.empty IS.empty
-
-data CutState = CutState
-  { _capacities :: Map (Symbol, Symbol) Rational
-  , _weights    :: Map (Symbol, Symbol) Rational
-  } deriving (Show, Read, Eq, Ord)
-
-makeLenses ''CutState
-
-type Cut = State CutState
-
-data Direction
-  = Forward
-  | Backward
-  deriving (Show, Read, Eq, Ord)
-
-data ExploreCtxt = ExploreCtxt
-  { _direction :: Direction
-  , _visited   :: IntSet
-  , _cutState  :: CutState
-  , _theGraph  :: Graph
-  } deriving (Show, Read, Eq, Ord)
-
[…]

### 2018-05-22 src/Shara/Plan.hs

@@ -1,115 +1,0 @@
-module Shara.Plan where
-
-import           Control.Lens hiding (Context)
-import           Control.Monad.Except
-import           Control.Monad.State
-import           Data.IntMap               (IntMap)
-import qualified Data.IntMap               as M
-import           Data.IntSet               (IntSet)
-import qualified Data.IntSet               as S
-import           Data.Language.Grammar
-import qualified Data.Language.Reg         as R
-import           Data.List.Split
-import           Data.Tree                 as Tree
-import           Formula
-import qualified Formula.Z3                as Z3
-import           Shara.Interpolate
-
-data Task
-  = Join NT
-  | TreeLike (Tree NT)
-  deriving (Show, Read, Eq)
-
-interp :: Grammar Expr -> IO (Either Model (IntMap Expr))
-interp g = do
-  (a, s) <- runStateT (runExceptT (execute g (contextualize g) (plan g))) M.empty
-  case a of
-    Left m -> pure (Left m)
-    Right _ -> pure (Right s)
-
-plan :: Grammar a -> [Task]
-plan g = tree (start g) (-1)
-  where
-    js = joinPoints g
-    tree r nt =
-      let (t, ijs) = runState (go r nt) S.empty
-          rest = concatMap (\n -> tree (ruleFor n g) n) (S.toList ijs)
-      in map Join (S.toList ijs) ++ [TreeLike t] ++ rest
-    go r nt = do
-      let (localJoins, others) =
-            S.partition (`S.member` js) (ruleNonterminals r)
-      modify (S.union localJoins)
-      Tree.Node nt <$> mapM (\n -> go (ruleFor n g) n) (S.toList others)
-
-execute ::
-     Grammar Expr
-  -> Context Expr
-  -> [Task]
-  -> ExceptT Model (StateT (IntMap Expr) IO) ()
-execute g rev = \case
-  [] -> pure ()
-  (Join nt:rest) -> interpolateNT' g rev nt >> execute g rev rest
-  (TreeLike t:rest) -> interpolateTree g rev t >> execute g rev rest
-
-interpolateTree ::
-     Grammar Expr
-  -> Context Expr
-  -> Tree NT
-  -> ExceptT Model (StateT (IntMap Expr) IO) ()
-interpolateTree g rev ntTree = do
-  let root = Tree.rootLabel ntTree
-  m <- get
-  let context = case root of
[…]

### 2018-05-22 src/Shara/Shara.hs

@@ -14,23 +14,16 @@ import           Formula
 import qualified Formula.Z3                  as Z3
 import           Shara.CDD
 import           Shara.Interpolate
-import           Shara.Plan
-
-data SolveKind
-  = Topological
-  | LicketySplit LicketySplitOptions
-  deriving (Show, Read, Eq, Ord)
 
 shara ::
      MonadIO m
-  => SolveKind
-  -> IntMap [Var]
+  => IntMap [Var]
   -> SG.Grammar [Var] Expr
   -> m (Either Model (IntMap Expr))
-shara sk vocab sg = evalStateT (go =<< cdd sg) (emptyCDDState vocab)
+shara vocab sg = evalStateT (go =<< cdd sg) (emptyCDDState vocab)
   where
     go g =
-      sharaStep sk g >>= \case
+      sharaStep g >>= \case
         Left m -> pure (Left m)
         Right m ->
           inductive vocab m sg >>= \case
@@ -41,33 +34,23 @@ shara sk vocab sg = evalStateT (go =<< cdd sg) (emptyCDDState vocab)
 -- unwind the grammar: Used for nontercursive problems.
 sharaNonrecursive ::
      MonadIO m
-  => SolveKind
-  -> IntMap [Var]
+  => IntMap [Var]
   -> SG.Grammar [Var] Expr
   -> m (Either Model (IntMap Expr))
-sharaNonrecursive sk vocab sg =
-  evalStateT (sharaStep sk =<< cdd sg) (emptyCDDState vocab)
+sharaNonrecursive vocab sg =
+  evalStateT (sharaStep =<< cdd sg) (emptyCDDState vocab)
 
 -- | Perform a single step -- solving a particular unwinding of the grammar.
 sharaStep ::
      (MonadIO m, MonadState CDDState m)
-  => SolveKind
-  -> IGrammar m' Expr
+  => IGrammar m' Expr
   -> m (Either Model (IntMap Expr))
-sharaStep sk g =
-  liftIO (interp (finitePrefix g)) >>= \case
-  -- solveDirect sk (finitePrefix g) >>= \case
+sharaStep g =
+  liftIO (interpolate (finitePrefix g)) >>= \case
     Left m -> pure (Left m)
     Right m -> do
       cs <- use clones
       pure (Right $ collapse cs m)
-
-solveDirect ::
-     MonadIO m => SolveKind -> Grammar Expr -> m (Either Model (IntMap Expr))
-solveDirect =
-  \case
[…]
